### 2021-05-23 paperwork/Research Plan/Research Plan.tex
lets see did this work?

\documentclass{article}
\usepackage{adjustbox}
\usepackage[round]{natbib}

\title{The Effect of Natural and Anthropogenic Forcing on ENSO Amplitude in the CESM Large Ensemble}
\author{Benjamin Goldman}

\begin{document}

\maketitle

\section{Rationale}

El Ni\~{n}o/Southern Oscillation (ENSO) is a cyclical change in the temperature of the equatorial Pacific Ocean. It is accompanied by disturbances to the Walker Circulation and reduced upwelling in the eastern Pacific \citep{bjerknes1969atmospheric}. ENSO has extreme effects on global climate, affecting the intensity of rainfall in the Americas, the location of Rossby Waves, the paths of storms, and much more \citep{liu2007atmospheric}. Because of these critical effects, determining how climate change will affect ENSO is crucial.

Scientists have begun to research the effects of natural and anthropogenic forcing on ENSO. For example, \citet{zhu2017reduced} showed that during the Last Glacial Maximum, lower CO$_{2}$ levels reduced ENSO variability. \citet{levine2017impact} suggested that the Atlantic Multidecadal Oscillation (AMO) may also play a role in determining ENSO amplitude and demonstrated the difficulty in predicting long-term changes to ENSO. Researchers have also shown the effect of natural forcing on ENSO variability. \citet{emile2007nino} suggested the role of ENSO in the effect of solar and orbital forcing on the Earth's climate, and \citet{liu2007atmospheric} determined that ozone changes dampen CO$_{2}$-forced reduction of ENSO amplitude. The conflicting relationship of ozone and CO$_{2}$ has also been shown for ENSO diversity, as \citet{stevenson2019forced} showed that the effects of ozone and greenhouse gasses on the development of Eastern and Central Pacific ENSO events tend to cancel out. \citet{zheng2018response} showed that ENSO variability is highly dependent on internal variability, and therefore a model with a large ensemble is necessary to measure a robust signal.

\section{Research Goals}

\begin{itemize}

	\item Determine the effect of a range of climate factors on Nin\~{o} amplitude.
	\item Examine mechanisms responsible for observed effect(s)
	\item The following factors will be considered:
	\begin{itemize}
		\item Greenhouse gas emissions
		\item Aerosol emissions
		\item Biomass burning
		\item Land use/cover
		\item Ozone
	\end{itemize}

\end{itemize}

\section{Research Methods}

\subsection{Procedures}

\begin{enumerate}

	\item Data source: Community Earth System Model (CESM LME) output, publicly available but compiled by mentor
	\begin{itemize}
		\item Full-forcing and single-forcing ensembles using recorded radiative forcing from 1920-2005, and RCP 8.5 projections from 2005-2100
		\item 2000-year control run with pre-industrial forcing
	\end{itemize}
	\item The Python programming language with the Scipy package will be used to perform all calculations.
	\item Running centered 20-year Ni\~{n}o 3.4 variance will be calculated for each ensemble member and the PI control.

\end{enumerate}

\subsection{Data Analysis}

\begin{enumerate}

	\item Mean and standard error will be calculated for each ensemble.
	\item Probability Distribution Function (PDF) will be estimated for the control run.
	\item Ni\~{n}o 3.4 index for single-forcing ensembles will be subtracted from full-forcing ensemble to isolate changes associated with individual factors.
	\item Mean and standard error for full-forcing and single-forcing cases will be compared to that of the control.
	\item The bootstrap method will be used to compare the single and full-forcing distributions.
	\item Control Ni\~{n}o 3.4 variance PDF's will be calculated for high and low AMO and AMOC index using CVDP (Climate Variability Diagnostics Package) provided by mentor.
	\item Correlation between Ni\~{n}o variance and equatorial Pacific ocean tempecature will be calculated for the full-forcing ensemble.
	\item A mixed-layer heat budget analysis may be calculated to examine changes to ocean structure.

\end{enumerate}

\subsection{Risk and Safety}

No risk/safety issues relevant as all methods are digital.

\subsection{Role of Mentor and Student}

\begin{adjustbox}{width=1.5\textwidth, center=\textwidth}
\begin{tabular}{| l | c | r |}

\hline
Student & Shared & Mentor\\
\hline

Perform all calculations & Discuss and interpret results and data & Review results and methods\\
Produce plots and documentation &       & Suggest changes and future calculations\\
 & & Provide/compile datasets\\

\hline

\end{tabular}
\end{adjustbox}

\bibliographystyle{apa}
\bibliography{refs.bib}

\end{document}

### 2021-05-23 paperwork/ResearchPlan2/plan.tex
lets see did this work?

% Created 2021-05-21 Fri 08:44
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{natbib}
\author{Benjamin Goldman}
\date{\today}
\title{Research Plan for the Impact of Anthropogenic Forcing on ENSO Amplitude}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Research Plan for the Impact of Anthropogenic Forcing on ENSO Amplitude},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{Rationale}
\label{sec:org38aeab9}

El Niño/Southern Oscillation (ENSO) is a cyclical change in the temperature of the equatorial Pacific Ocean. It is accompanied by disturbances to the Walker Circulation and reduced upwelling in the eastern Pacific \citep{bjerknes1969atmospheric}. ENSO has extreme effects on global climate, affecting the intensity of rainfall in the Americas, the location of Rossby Waves, the paths of storms, and much more \citep{liu2007atmospheric}. Because of these critical effects, determining how climate change will affect ENSO is crucial.

Scientists have begun to research the effects of natural and anthropogenic forcing on ENSO. For example, \citet{zhu2017reduced} showed that during the Last Glacial Maximum, lower CO\(_2\) levels reduced ENSO variability. \cite{levine2017impact} suggested that the Atlantic Multidecadal Oscillation (AMO) may also play a role in determining ENSO amplitude and demonstrated the difficulty in predicting long-term changes to ENSO. Researchers have also shown the effect of natural forcing on ENSO variability. \citet{zheng2018response} showed that ENSO variability is highly dependent on internal variability, and therefore a model with a large ensemble is necessary to measure a robust signal. Despite the copious noise in the simulated ENSO amplitude record, progress has been made in determining the sign of changes to ENSO amplitude over the 21st century. Recent studies \citep{maher2018enso}, \citep{cai2018increased} suggest that ENSO is likely to intensify, using records of increased ENSO amplitude in large model ensembles as evidence.

These past studies provide a background of predicted increased ENSO amplitude. With this background as a starting point, this study aims to assess probable causes for this predicted increasing amplitude. Using 2 recent derivatives of the National Center for Atmospheric Research's (NCAR) Community Earth System Model version 1 (CESM1), the CESM1 Large Ensemble and the CESM1 Single Forcing Ensembles. Using these tools, significant increases to ENSO amplitude over the 21st century were attributed to the combined influence of greenhouse gases and aerosol emissions on ocean stratification and mixed layer heat transfer. However, some questions remain, and new data sources beckon. Firstly, past results conducted indicate a predicted decrease in ENSO amplitude after approximately 2060, which the researchers aim to investigate. Additionally, the sign of changes to ENSO amplitude under greenhouse forcing agree with those under aerosol forcing, where other studies indicated that they should be the opposite due to the differing nature of greenhouse and aerosol emissions \citep{deser2020isolating}. Moreover, it is still unclear specifically how greenhouse and aerosol emissions affect mixed layer heat transfer. Finally, Large Ensemble and Single Forcing runs are nearly complete for NCAR's next generation of the CESM, CESM2. Past methods will be repeated on this dataset and results will be compared to those from the CESM1, with the hope that the CESM2 results will provide insight into those descrepancies that appear in the CESM1 output. Additionally, deeper analysis of the role of air currents and ocean structure will be conducted for each ensemble to further clarify the processes that mediate the aformentioned forced effect on ENSO amplitude.

\section{Research Goals}
\label{sec:orgc5ea83c}
\begin{itemize}
\item Estimate future changes to ENSO amplitude in the NCAR's Community Earth System Model Version 2 (CESM2)
\item Compare these changes to those previously found in the CESM1.
\item Examine contributions of greenhouse gasses, aerosols, and natural forcing in both models.
\item Explore physical mechanism that mediates emissions' effect on ENSO amplitude.
\end{itemize}

\section{Methods}
\label{sec:org42d33f3}
\subsection{Materials and Data}
\label{sec:orgb6a9935}
\begin{itemize}
\item NCAR's CESM1 and CESM2 will be used.
\begin{itemize}
\item Fully forced historical and projected simulations.
\item Single forcing simulations that are forced by all but one factor.
\end{itemize}
\item The Python and R programming language will be used with third party packages as needed.
\end{itemize}

\subsection{Procedures}
\label{sec:org1d14759}
\emph{Methods are subject to change in response to results}
\begin{itemize}
\item Prodedures from previous version of this study will be repeated using the Python and R languages.
\item ENSO amplitude will be measured in each simulation by calculating the 20-year running variance of the Niño 3.4 index.
\item A bootstrap process will be used to isolate the influence of single forcings, comparing their contributions in the context of global warming.
\item Correlation coefficient between Niño 3.4 variance and sea temperature will be calculated for each simulation.
\item A mixed-layer heat budget analysis will be used to further clarify the role of heat transfer in determining changes to ENSO under global warming.
\end{itemize}

\subsection{Risk and Safety}
\label{sec:org7d700d5}
No risk/safety issues relevant as all methods are digital.

\section{Role of Mentor and Student}
\label{sec:org841d5d2}
\begin{center}
\begin{tabular}{lll}
Student & Shared & Mentor\\
\hline
Perform all calculations & Discuss and interperet results & Review results and methods\\
Produce plots and documents &  & Provide large datasets\\
\end{tabular}
\end{center}

\bibliographystyle{apa}
\bibliography{references}
\end{document}

### 2021-05-23 paperwork/abstract/abstract.tex
lets see did this work?

% Created 2021-05-10 Mon 19:08
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\date{\today}
\title{Abstract}
\hypersetup{
 pdfauthor={John Doe},
 pdftitle={Abstract},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\begin{center}

{\bf The Impact of Industrial Emissions on ENSO Amplitude}
{\bf Benjamin Goldman}

\end{center}

The El Niño/Southern Oscillation (ENSO) is the dominant mode of interannual climate variability, with substantial associated global socio-economic impacts. Due to their significance, shifts in ENSO under climate change also have the potential to substantial impact human society and natural ecosystems. However, it is currently unclear what effect greenhouse gas (GHG) and industrial aerosol (AER) emissions will have on ENSO and even less clear what effect a combination of these factors might have when changing in tandem. This study examines transient changes to ENSO variance under a variety of forcing scenarios using the CESM1 Large and Single-Forcing Ensembles. These multi-member ensembles span the historical record (1920-2005) and much of the 21st C (2006-2080 for GHG/AER). A 2000-year pre-industrial (PI) control simulation is used to account for model drift and 20-year running variance of the Niño 3.4 SST index is used as a proxy for ENSO variance. The ensemble mean and standard error of each ensemble is calculated, to estimate the statistical significance of simulated changes. We identify significant increases in variance under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come. Computing detrended and smoothed correlation between 20-year variance of the Niño 3.4 index and ocean temperature in the tropical Pacific Ocean reveals that stratification caused by global warming is connected with increased ENSO amplitude.
\end{document}

### 2021-05-23 paperwork/paperAndRefsList/agu_archive/paper.tex
lets see did this work?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
%
% This file includes commands and instructions
% given in the order necessary to produce a final output that will
% satisfy AGU requirements, including customized APA reference formatting.
%
% You may copy this file and give it your
% article name, and enter your text.
%
%
% Step 1: Set the \documentclass
%
%

%% To submit your paper:
\documentclass[draft]{agujournal2019}
\usepackage{url} %this package should fix any errors with URLs in refs.
\synctex=1

\usepackage{lineno}
\usepackage[finalnew]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers

\addeditor{bengoldman}
\addeditor{fasullo}

%%%%%%%
% As of 2018 we recommend use of the TrackChanges package to mark revisions.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/
%%%%%%%

%\draftfalse

%% Enter journal name below.
%% Choose from this list of Journals:
%
% JGR: Atmospheres
% JGR: Biogeosciences
% JGR: Earth Surface
% JGR: Oceans
% JGR: Planets
% JGR: Solid Earth
% JGR: Space Physics
% Global Biogeochemical Cycles
% Geophysical Research Letters
% Paleoceanography and Paleoclimatology
% Radio Science
% Reviews of Geophysics
% Tectonics
% Space Weather
% Water Resources Research
% Geochemistry, Geophysics, Geosystems
% Journal of Advances in Modeling Earth Systems (JAMES)
% Earth's Future
% Earth and Space Science
% Geohealth
%
% ie, \journalname{Water Resources Research}

\journalname{Geophysical Research Letters}

\begin{document}

%% ------------------------------------------------------------------------ %%
%  Title
%
% (A title should be specific, informative, and brief. Use
% abbreviations only if they are defined in the abstract. Titles that
% start with general keywords then specific terms are optimized in
% searches)
%
%% ------------------------------------------------------------------------ %%

% Example: \title{This is a test title}

\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}

%% ------------------------------------------------------------------------ %%
%
%  AUTHORS AND AFFILIATIONS
%
%% ------------------------------------------------------------------------ %%

% Authors are individuals who have significantly contributed to the
% research and preparation of the article. Group authors are allowed, if
% each author in the group is separately identified in an appendix.)

% List authors by first name or initial followed by last name and
% separated by commas. Use \affil{} to number affiliations, and
% \thanks{} for author notes.
% Additional author notes should be indicated with \thanks{} (for
% example, for current addresses).

% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
% Author\affil{3,4}\thanks{Also funded by Monsanto.}}

\authors{John Fasullo\affil{1}, Benjamin Goldman\affil{2}}

% \affiliation{1}{First Affiliation}
% \affiliation{2}{Second Affiliation}
% \affiliation{3}{Third Affiliation}
% \affiliation{4}{Fourth Affiliation}

\affiliation{1}{Climate and Global Dynamics Division, National Center for Atmospheric Research, 1850 Table Mesa Dr., Boulder, CO, USA}

\affiliation{2}{Science Research Program, White Plains High School, 550 North Street, White Plains, NY, USA}

%(repeat as many times as is necessary)

%% Corresponding Author:
% Corresponding author mailing address and e-mail address:

% (include name and email addresses of the corresponding author.  More
% than one corresponding author is allowed in this LaTeX file and for
% publication; but only one corresponding author is allowed in our
% editorial system.)

% Example: \correspondingauthor{First and Last Name}{email@address.edu}

\correspondingauthor{Benjamin Goldman}{bg502257@live.wpcsd.k12.ny.us}

%% Keypoints, final entry on title page.

%  List up to three key points (at least one is required)
%  Key Points summarize the main points and conclusions of the article
%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences

% Example:
% \begin{keypoints}
% \item	List up to three key points (at least one is required)
% \item	Key Points summarize the main points and conclusions of the article
% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
% \end{keypoints}

\begin{keypoints}
\item NCAR's CESM Large Ensemble predicts increase to Ni\~{n}o 3.4 variance over the 21$^{st}$ century.
\item Increase can mainly be attributed to greenhouse gas and aerosol emissions.
\item Changes to ENSO amplitude are linked to changes to equatorial Pacific ocean stratification.
\end{keypoints}

%% ------------------------------------------------------------------------ %%
%
%  ABSTRACT and PLAIN LANGUAGE SUMMARY
%
% A good Abstract will begin with a short description of the problem
% being addressed, briefly describe the new data or analyses, then
% briefly states the main conclusion(s) and how they are supported and
% uncertainties.

% The Plain Language Summary should be written for a broad audience,
% including journalists and the science-interested public, that will not have 
% a background in your field.
%
% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
% see http://sharingscience.agu.org/creating-plain-language-summary/)
%
%% ------------------------------------------------------------------------ %%

%% \begin{abstract} starts the second page

\begin{abstract}

    The El Niño/Southern Oscillation (ENSO) is the dominant mode of interannual climate variability, with substantial associated global socio-economic impacts. Due to their significance, shifts in ENSO under climate change also have the potential to substantially impact human society and natural ecosystems. However, it is currently unclear what effect greenhouse gas (GHG) and industrial aerosol (AER) emissions have on ENSO, as well as what effect these factors have when combined. This study examined transient changes to ENSO variance under a variety of forcing scenarios using the CESM1 Large and Single-Forcing Ensembles. These multi-member ensembles span the historical record (1920-2005) and much of the 21st C (2006-2080 for GHG/AER). A 2000-year pre-industrial (PI) control simulation is used to account for model drift and 20-year running variance of the Niño 3.4 SST index is used as a proxy for ENSO variance. The ensemble mean and standard error of each ensemble was calculated, while the Probability Density Function (PDF) is computed for the PI control simulation to estimate the statistical significance of simulated changes. We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Ni\~{n}o 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude. We identifed significant increases in variance of the Ni\~{n}o 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.

\end{abstract}

%\section*{Plain Language Summary}

%[ enter your Plain Language Summary here or delete this section]

%% ------------------------------------------------------------------------ %%
%
%  TEXT
%
%% ------------------------------------------------------------------------ %%

%%% Suggested section heads:
% \section{Introduction}
%
% The main text should start with an introduction. Except for short
% manuscripts (such as comments and replies), the text should be divided
% into sections, each with its own heading.

% Headings should be sentence fragments and do not begin with a
% lowercase letter or number. Examples of good headings are:

% \section{Materials and Methods}
% Here is text on Materials and Methods.
%
% \subsection{A descriptive heading about methods}
% More about Methods.
%
% \section{Data} (Or section title might be a descriptive heading about data)
%
% \section{Results} (Or section title might be a descriptive heading about the
% results)
%
% \section{Conclusions}

\section{Introduction}

\note[bengoldman]{problem/entity}El Ni\~{n}o is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \cite{bjerknes1969atmospheric}. The reasons for studying ENSO are clear, as ENSO has drastically affected climate patterns worldwide, modulating rainfall and temperature in nearly every continent \cite{ropelewski1987global}. For example, the recent 2015-2016 El Ni\~{n}o event contributed to record-breaking high temperatures and droughts in South America \cite{jimenez2016record}. At the same time, long-term anthropogenic greenhouse gas emissions are causing global temperatures to increase through a greenhouse effect. The effect of greenhouse emissions and other factors on ENSO intensity remains unclear.

The two major ways in which the earth's climate varies are climate change and climate variability. Climate change is defined as climate changes caused by external factors, most notably greenhouse gasses, natural and artificial aerosol emissions, land use changes, and stratospheric ozone changes. Greenhouse gas emissions have a clear impact on the earth's climate, global warming. Climate change is usually long-term. In contrast, internal variability is defined as changes to the earth's climate originating from natural climatic processes, such as ENSO, Pacific Decadal Oscillation (PDO), Atlantic Multidecadal Oscillation (AMO), and others. Climate variability occurs on much shorter timescales, and is usually cyclical and driven by feedback loops.

\note[bengoldman]{focus on future experiments, describe results of literature, not methods} \note[bengoldman]{past research} Research on the effect of external forcing on ENSO remains inconclusive, as results from similar studies conflict. \citeA{nowack2017role} predicted an overall increase in Ni\~{n}o 3.4 standard deviation under a combination of 4xCO$_2$ and interactive ozone forcing using single-model simulations, showing that greenhouse gasses increase the frequency of extreme ENSO by favoring a more Ni\~{n}o-like in the tropical Pacific, while ozone dampens this effect. In contrast, a few studies have found that ENSO amplitude decreases under global warming in certain coupled models \cite{kohyama2018weakening}.

However, other studies have failed to find any statistically significant ENSO response to external forcing \cite{stevenson2012significant}. Analysis using NCAR's CESM Large Ensemble shows an ensemble size of at least 15 models is necessary to attribute changes to Ni\~{n}o variance to external forcing and reject the null hypothesis that internal variability is responsible for changes to ENSO \cite{zheng2018response}. Several modes of internal variability have been shown to modulate ENSO, including the AMO \cite{levine2017impact} and Tropical Pacific Decadal Variability (TPDV) \cite{zheng2018response}. An analysis of the Max Plank Institute's Grand Ensemble as well as NCAR's CESM Large Ensemble suggests that 80\% of changes to ENSO amplitude can be attributed to internal variability, but given a large enough ensemble, significant changes in ENSO amplitude due to climate change can be detected \cite{maher2018enso}.

In this study, we show that NCAR's CESM predicts significant increase in ENSO amplitude in the 21st century, and that greenhouse gasses and aerosol emissions play important roles in causing this increase. As in previous studies, the role of internal variability in conjunction with forcing was examined. We hypothesized that increased stratification in the future plays a large role in causing this predicted increase.

\section{Data}

The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE). The Large Ensemble contains 40 simulations of the CESM1 coupled model, forced with historic radiative forcing from 1920 to 2005 and according to the RCP 8.5 protocol from 2006 to 2100 \cite{kay2015community}. The single forcing ensemble is a collection of sub-ensembles for various climate factors (greenhouse gas, aerosols, biomass burning, land use, ozone). Each simulation is forced by a combination of all factors except for one, allowing the impact of a single factor to be deduced by subtracting the ensemble mean from the fully-forced ensemble mean \cite{deser2020isolating}. For example, the xghg (greenhouse gas) ensemble is forced by aerosol emissions, biomass burning, land use, and ozone. There is also a preindustrial control simulation, with all radiative forcing fixed at 1850 levels.

\section{Methods and Results}
%% yay

\subsection{Ni\~{n}o 3.4 Variance}
We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Ni\~{n}o 3.4 region of the Pacific ocean (5N-5S, 170W-120W). We measure variance on 20-year centered sliding windows. We calculate the multi-model mean for each ensemble, as well as the ensemble standard error. The results of this calculation are shown in figure \ref{fig:variance_1}.

The fully-forced ensemble exhibits moderate increase in variance, with the other ensembles showing less meaningful changes. In the fully-forced ensemble, the variance of the Ni\~{n}o 3.4 region increases beyond 2 standard errors of the control, increasing until the mid-21st century, and then decreasing gradually. The excluded greenhouse, aerosol, and biomass-burning ensembles also contain a slight increase, which may not be statistically significant. Although the excluded land use ensemble mean has a strong increase in variance, this result is unlikely to be meaningful due to the xlulc ensemble's low sample size. Additionally, all ensembles exhibit considerable noise, in line with previous studies \cite{maher2018enso}.

\begin{figure}
    \includegraphics{figures/variance_1.pdf}
    \caption{20-year windowed variance of the Ni\~{n}o 3.4 region SST, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc, f) fixedO3. Grey bars are PI control mean and 1x, 2x, 3x standard error. Colored bars represent ensemble standard error.}
    \label{fig:variance_1}
\end{figure}

The fully-forced ensemble exhibits reduced variance in the mid-late 20th century, below that of the PI control. The most likely explanation for this phenomenon is internal variability. To test the likelihood of this explanation, we test the influence of the Atlantic Multidecadal Oscillation (AMO) and Atlantic Meridional Overturning Current (AMOC) on ENSO in the control simulation. The AMO has been shown to have an influence on ENSO strength and seasonal growth rate \cite{levine2017impact}. We filter the control Ni\~{n}o 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \cite{phillips2014evaluating}. The Probability Distribution Function of Ni\~{n}o 3.4 variance is estimated for AMO/C > 1/2$\sigma$ and < -1/-2$\sigma$ using a Kernel Density Estimation. No meaningful differences were found in the distribution of ENSO intensity under any of these conditions (See Suplementary information figure 1).

\subsection{Single Forcing Scenarios and the Bootstrap Process}
We analyze the role of individual factors using the CESM Single Forcing Ensembles. To separate the influence of a single factor from the fully-forced ensemble, we employ a bootstrap test. For each single-forcing ensemble, a single simulation is randomly selected, and the Ni\~{n}o 3.4 20-year variance record is subtracted from that from a randomly selected fully-forced simulation. We repeat this process 1000 times for each ensemble, and then calculate the mean and standard error for each ensemble. These results are shown in figure \ref{fig:bootstrap_1}. The greenhouse-only ensemble as well as the aerosol-only ensemble exhibit increased variance, signaling that greenhouse and aerosol emissions likely both play a significant role in ENSO's forced response in the full-forcing ensemble. Interestingly, the influence of greenhouse gasses and aerosols are non-conflicting, in contrast with previous studies that show opposite effects of greenhouse gas and aerosol forcing on ENSO \cite{stevenson2019forced}. All the other single forcing ensembles exhibit insignificant differences from the fully-forced ensemble. The biomass burning case shows very small deviations from the fully-forced case, while the ozone ensemble's period of recording is too small to draw meaningful conclusions. However, \citeA{nowack2017role} showed that ozone forcing may dampen the effect of greenhouse-forced increases to ENSO amplitude by reducing changes to Pacific sea temperature and the Walker Circulation. The land use/cover case, while it does show large deviations from the fully-forced case, has an ensemble size (5 members) too small to lend any credibility to these changes.

\begin{figure}
    \includegraphics{figures/bootstrap_1.pdf}
    \caption{Difference between fully forced and single forced ensembles derived from the bootstrap process for a) greenhouse gas, b) aerosol emissions c) biomass burning, d) land use/cover, e) ozone; f) compares the sum of the bootstrapped ensemble means for the single forcing ensembles (blue) and the ensemble mean for the full forcing ensemble, detrended by a constant to center on zero (orange).}
    \label{fig:bootstrap_1}
\end{figure}

In both the fully-forced scenario, and the greenhouse and aerosol only simulations, there is noticeably reduced Ni\~{n}o 3.4 variance in the mid-20\^{th} century, below 2 standard errors of the control. We hypothesize that this discrepancy may be the result of anomalous initial conditions caused by internal variability of the control, as the control conditions are used to initialize all the forced runs. To test this hypothesis, we analyzed the impact of the Atlantic Multidecadal Oscillation (AMO) and the Atlantic Meridional Overturning Current (AMOC) on Ni\~{n}o 3.4 variance in the control simulation using records of the AMO and AMOC from the Climate Variability Diagnostics Package (CVDP) \cite{phillips2014evaluating}. We filtered the 20-year variance of the Ni\~{n}o 3.4 sea surface temperature in the control based on the strength of the AMO/AMOC, separating ENSO variance into groups where AMO < -1$\sigma$, AMO < -2$\sigma$, AMO > 1$\sigma$, AMO >2$\sigma$, and the same for AMOC. We then estimated the probability density function for each group using a kernel density estimator. We observed no consistent difference in the distribution of Ni\~{n}o 3.4 variance between any group. So far the question of reduced variance is unanswered, and it will be addressed in further depth later in the study.

\subsection{Correlation With Ocean Temperature}
Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles. To do this we used 4 slices of ocean temperature from the fully-forced, xghg, and xaer ensembles, including a slice averaged along the equator and slices through the western, central, and eastern Pacific basins. We linearly detrend and smooth with a 30-year windowed mean the timeseries at each gridpoint and the Ni\~{n}o 3.4 variance. Next, we calculated the Pearson's correlation coefficient between each gridpoint and the Ni\~{n}o 3.4 variance timeseries. The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt_1}, and for the central slice in figure \ref{fig:tempcep_1}.

Overall, the majority of the Pacific basin exhibits negative correlation with Ni\~{n}o 3.4 variance when linearly detrended and smoothed. The fully-forced simulation contains strongly negative correlation between the subsurface layer of the equatorial Pacific and ENSO amplitude, leading to the hypothesis that forced changes to Pacific Ocean stratification may be connected to ENSO intensity. However, it is unclear whether there is also a causal relationship between the two. Additionally, it is unclear whether the overall warming trend of the upper Pacific ocean is connected to changes to ENSO intensity. We notice that the high levels of correlation between the subsurface layer and Ni\~{n}o 3.4 variance are present in the xGHG and xAER ensembles, as well as the fully forced ensemble, suggesting that there is a relationship between ENSO amplitude and subsurface Pacific Ocean temperature in a variety of forcing scenarios.

\begin{figure}
    \includegraphics{figures/tempdt.pdf}
    \caption{Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Ni\~{n}o 3.4 variance. Grey color indicates points where p<0.05. The white regions at 25 and 280 degrees East are Africa and South America.}
    \label{fig:tempdt_1}
\end{figure}

\begin{figure}
    \includegraphics{figures/tempcep.pdf}
    \caption{Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Ni\~{n}o 3.4 variance. Grey color indicates points where p<0.05.}
    \label{fig:tempcep_1}
\end{figure}

\section{Conclusions and Discussion}
%% yay

In this study, predicted changes to ENSO amplitude are examined in the CESM1 Large Ensemble and Single Forcing Ensemble. The fully-forced large ensemble exhibits increased ENSO amplitude, as measured by calculating the 20-year variance of the Ni\~{n}o 3.4 index. These changes appear to be statistically significant, but need to be scrutinized further, as they are in disagreement with results of some previous studies \cite{stevenson2012significant}. Additionally, the fully-forced ensemble, as well as all the others contains considerable noise, as the individual members cover a wide range of variability. Greenhouse gas and aerosol emissions are likely the most significant contributors to these forced changes, as shown by the fact that the ensemble mean for the xGHG and xAER ensembles are the most different from the fully-forced mean. All the other single-forcing ensembles had insignificant differences from the fully forced ensemble, or were lacking a sample size large enough to lend credibility to their results. Analysis of the correlation coefficient between ENSO amplitude and Pacific ocean temperature reveals that changes to ENSO intensity are tied to changes to the subsurface layer of the tropical Pacific Ocean. It is still unclear what is driving this connection, as well as how external forcing such as global warming affects it. There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \cite{dewitte2013reinterpreting}.

This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \cite{maher2018enso}. Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citeA{stevenson2019forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity. In contrast, this study observes that the impact on ENSO amplitude of greenhouse gas and aerosols has the same sign. This result is also surprising given that greenhouse and aerosol emissions have been shown to have opposite effects on sea surface temperature and global circulation, with greenhouse emissions favoring a more Ni\~{n}o-like state with weakened Walker circulation and higher SST, and aerosol emissions counteracting these effects \cite{boer2000transient}.

This study has a number of goals for future continuation. The most pressing of these goals is a deeper analysis of changes to Pacific Ocean structure, which will be done by examining correlation between ENSO amplitude and potential density, as well as comparisons of changes to stratification in the Pacific. Additionally, the methods of this study will be repeated on the CESM2 large ensemble, which has a larger ensemble size and a longer record period, as compared to CESM1. Analysis of the CMIP6 models are also necessary to verify these results across variations in model physics. Deeper statistical analyses of changes to ENSO amplitude should also be tone to further examine the likelihood and intensity of future changes to ENSO amplitude. Lastly, correlation between ENSO amplitude and internal variability should be further investigated using the records provided by the CVDP. Doing these further analyses will help to verify the results of this study, while more precisely analyzing future changes to ENSO.

The results of this study will help to direct future studies on the impact of climate change on ENSO by providing one further data point supporting the conclusion that ENSO will intensify due to global warming. This study and others like it will help societies impacted by ENSO to prepare for intensification in the future.

%%

%  Numbered lines in equations:
%  To add line numbers to lines in equations,
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

%% Enter Figures and Tables near as possible to where they are first mentioned:
%
% DO NOT USE \psfrag or \subfigure commands.
%
% Figure captions go below the figure.
% Table titles go above tables;  other caption information
%  should be placed in last line of the table, using
% \multicolumn2l{$^a$ This is a table note.}
%
%----------------
% EXAMPLE FIGURES
%
% \begin{figure}
% \includegraphics{example.png}
% \caption{caption}
% \end{figure}
%
% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
% \begin{figure}
% \noindent\includegraphics[width=\textwidth]{anothersample.png}
%\caption{caption}
%\label{pngfiguresample}
%\end{figure}
%
%
% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options. This is common when trying to add a PDF figure without pdflatex.
% \begin{figure}
% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
%\caption{caption}
%\label{pdffiguresample}
%\end{figure}
%
%
% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
%

%
% ---------------
% EXAMPLE TABLE
%
% \begin{table}
% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
% \centering
% \begin{tabular}{l c}
% \hline
%  Run  & Time (min)  \\
% \hline
%   $l1$  & 260   \\
%   $l2$  & 300   \\
%   $l3$  & 340   \\
%   $h1$  & 270   \\
%   $h2$  & 250   \\
%   $h3$  & 380   \\
%   $r1$  & 370   \\
%   $r2$  & 390   \\
% \hline
% \multicolumn{2}{l}{$^{a}$Footnote text here.}
% \end{tabular}
% \end{table}

%% SIDEWAYS FIGURE and TABLE
% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
%
% \begin{sidewaysfigure}
% \includegraphics[width=20pc]{figsamp}
% \caption{caption here}
% \label{newfig}
% \end{sidewaysfigure}
%
%  \begin{sidewaystable}
%  \caption{Caption here}
% \label{tab:signif_gap_clos}
%  \begin{tabular}{ccc}
% one&two&three\\
% four&five&six
%  \end{tabular}
%  \end{sidewaystable}

%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
%\begin{linenomath*}
%\begin{equation}
%y|{f} \sim g(m, \sigma),
%\end{equation}
%\end{linenomath*}

%%% End of body of article

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional Appendix goes here
%
% The \appendix command resets counters and redefines section heads
%
% After typing \appendix
%
%\section{Here Is Appendix Title}
% will show
% A: Here Is Appendix Title
%
%\appendix
%\section{Here is a sample appendix}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Optional Glossary, Notation or Acronym section goes here:
%
%%%%%%%%%%%%%%
% Glossary is only allowed in Reviews of Geophysics
%  \begin{glossary}
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \term{Term}
%   Term Definition here
%  \end{glossary}

%
%%%%%%%%%%%%%%
% Acronyms
%   \begin{acronyms}
%   \acro{Acronym}
%   Definition here
%   \acro{EMOS}
%   Ensemble model output statistics
%   \acro{ECMWF}
%   Centre for Medium-Range Weather Forecasts
%   \end{acronyms}

%
%%%%%%%%%%%%%%
% Notation
%   \begin{notation}
%   \notation{$a+b$} Notation Definition here
%   \notation{$e=mc^2$}
%   Equation in German-born physicist Albert Einstein's theory of special
%  relativity that showed that the increased relativistic mass ($m$) of a
%  body comes from the energy of motion of the body—that is, its kinetic
%  energy ($E$)—divided by the speed of light squared ($c^2$).
%   \end{notation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  ACKNOWLEDGMENTS
%
% The acknowledgments must list:
%
% >>>>	A statement that indicates to the reader where the data
% 	supporting the conclusions can be obtained (for example, in the
% 	references, tables, supporting information, and other databases).
%
% 	All funding sources related to this work from all authors
%
% 	Any real or perceived financial conflicts of interests for any
%	author
%
% 	Other affiliations for any author that may be perceived as
% 	having a conflict of interest with respect to the results of this
% 	paper.
%
%
% It is also the appropriate place to thank colleagues and other contributors.
% AGU does not normally allow dedications.

\acknowledgments

%% ------------------------------------------------------------------------ %%
%% References and Citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% \bibliography{<name of your .bib file>} don't specify the file extension
%
% don't specify bibliographystyle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\bibliography{references.bib}
This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No. 1852977. Computing resources (doi:10.5065/D6RX99HX) were provided by the Climate Simulation Laboratory at NCAR's Computational and Information Systems Laboratory, sponsored by the National Science Foundation and other agencies. I would like to thank Ms. Kimberly Fleming and the White Plains High School Science Research Program for providing support in this study.

%Reference citation instructions and examples:
%
% Please use ONLY \cite and \citeA for reference citations.
% \cite for parenthetical references
% ...as shown in recent studies (Simpson et al., 2019)
% \citeA for in-text citations
% ...Simpson et al. (2019) have shown...
%
%
%...as shown by \citeA{jskilby}.
%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
%...has been shown \cite{jskilbye}.
%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
%...has been shown by \cite <e.g.,>[and others]{lewin76}.
%
% apacite uses < > for prenotes and [ ] for postnotes
% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
%

\end{document}

\section*{extra stuff}

More Information and Advice:

%% ------------------------------------------------------------------------ %%
%
%  SECTION HEADS
%
%% ------------------------------------------------------------------------ %%

% Capitalize the first letter of each word (except for
% prepositions, conjunctions, and articles that are
% three or fewer letters).

% AGU follows standard outline style; therefore, there cannot be a section 1 without
% a section 2, or a section 2.3.1 without a section 2.3.2.
% Please make sure your section numbers are balanced.
% ---------------
% Level 1 head
%
% Use the \section{} command to identify level 1 heads;
% type the appropriate head wording between the curly
% brackets, as shown below.
%
%An example:
%\section{Level 1 Head: Introduction}
%
% ---------------
% Level 2 head
%
% Use the \subsection{} command to identify level 2 heads.
%An example:
%\subsection{Level 2 Head}
%
% ---------------
% Level 3 head
%
% Use the \subsubsection{} command to identify level 3 heads
%An example:
%\subsubsection{Level 3 Head}
%
%---------------
% Level 4 head
%
% Use the \subsubsubsection{} command to identify level 3 heads
% An example:
%\subsubsubsection{Level 4 Head} An example.
%
%% ------------------------------------------------------------------------ %%
%
%  IN-TEXT LISTS
%
%% ------------------------------------------------------------------------ %%
%
% Do not use bulleted lists; enumerated lists are okay.
% \begin{enumerate}
% \item
% \item
% \item
% \end{enumerate}
%
%% ------------------------------------------------------------------------ %%
%
%  EQUATIONS
%
%% ------------------------------------------------------------------------ %%

% Single-line equations are centered.
% Equation arrays will appear left-aligned.

Math coded inside display math mode \[ ...\]
 will not be numbered, e.g.,:
 \[ x^2=y^2 + z^2\]

 Math coded inside \begin{equation} and \end{equation} will
 be automatically numbered, e.g.,:
 \begin{equation}
 x^2=y^2 + z^2
 \end{equation}

% To create multiline equations, use the
% \begin{eqnarray} and \end{eqnarray} environment
% as demonstrated below.
\begin{eqnarray}
  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
        && + (y - y_{0}) \sin \Theta  \nonumber \\
  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
        && + (y - y_{0}) \cos \Theta.
\end{eqnarray}

%If you don't want an equation number, use the star form:
%\begin{eqnarray*}...\end{eqnarray*}

% Break each line at a sign of operation
% (+, -, etc.) if possible, with the sign of operation
% on the new line.

% Indent second and subsequent lines to align with
% the first character following the equal sign on the
% first line.

% Use an \hspace{} command to insert horizontal space
% into your equation if necessary. Place an appropriate
% unit of measure between the curly braces, e.g.
% \hspace{1in}; you may have to experiment to achieve
% the correct amount of space.

%% ------------------------------------------------------------------------ %%
%
%  EQUATION NUMBERING: COUNTER
%
%% ------------------------------------------------------------------------ %%

% You may change equation numbering by resetting
% the equation counter or by explicitly numbering
% an equation.

% To explicitly number an equation, type \eqnum{}
% (with the desired number between the brackets)
% after the \begin{equation} or \begin{eqnarray}
% command.  The \eqnum{} command will affect only
% the equation it appears with; LaTeX will number
% any equations appearing later in the manuscript
% according to the equation counter.
%

% If you have a multiline equation that needs only
% one equation number, use a \nonumber command in
% front of the double backslashes (\\) as shown in
% the multiline equation above.

% If you are using line numbers, remember to surround
% equations with \begin{linenomath*}...\end{linenomath*}

%  To add line numbers to lines in equations:
%  \begin{linenomath*}
%  \begin{equation}
%  \end{equation}
%  \end{linenomath*}

### 2021-05-23 paperwork/paperAndRefsList/papernew.tex
lets see did this work?

--- 2021-05-23 paperwork/paperAndRefsList/agu_archive/paper.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% AGUJournalTemplate.tex: this template file is for articles formatted with LaTeX
-%
-% This file includes commands and instructions
-% given in the order necessary to produce a final output that will
-% satisfy AGU requirements, including customized APA reference formatting.
-%
-% You may copy this file and give it your
-% article name, and enter your text.
-%
-%
-% Step 1: Set the \documentclass
-%
-%
-
-%% To submit your paper:
-\documentclass[draft]{agujournal2019}
-\usepackage{url} %this package should fix any errors with URLs in refs.
-\synctex=1
-
-\usepackage{lineno}
-\usepackage[finalnew]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
-\linenumbers
-
-\addeditor{bengoldman}
-\addeditor{fasullo}
-
-%%%%%%%
-% As of 2018 we recommend use of the TrackChanges package to mark revisions.
-% The trackchanges package adds five new LaTeX commands:
-%
-%  \note[editor]{The note}
-%  \annote[editor]{Text to annotate}{The note}
-%  \add[editor]{Text to add}
-%  \remove[editor]{Text to remove}
-%  \change[editor]{Text to remove}{Text to add}
-%
-% complete documentation is here: http://trackchanges.sourceforge.net/
-%%%%%%%
-
-%\draftfalse
-
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-% ie, \journalname{Water Resources Research}
-
-\journalname{Geophysical Research Letters}
-
+% Created 2021-05-21 Fri 09:44
+% Intended LaTeX compiler: pdflatex
+\documentclass[11pt]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{graphicx}
+\usepackage{grffile}
+\usepackage{longtable}
+\usepackage{wrapfig}
+\usepackage{rotating}
+\usepackage[normalem]{ulem}
+\usepackage{amsmath}
+\usepackage{textcomp}
+\usepackage{amssymb}
+\usepackage{capt-of}
+\usepackage{hyperref}
+\usepackage{natbib}
+\usepackage[letterpaper, margins=1in]{geometry}
+\setlength{\parskip}{0.35em}
+\author{Benjamin Goldman}
+\date{\today}
+\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
+\hypersetup{
+ pdfauthor={Benjamin Goldman},
+ pdftitle={The Impact of Anthropogenic Forcing on ENSO Amplitude},
+ pdfkeywords={},
+ pdfsubject={},
+ pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
+ pdflang={English}}
 \begin{document}
 
-%% ------------------------------------------------------------------------ %%
-%  Title
-%
-% (A title should be specific, informative, and brief. Use
-% abbreviations only if they are defined in the abstract. Titles that
-% start with general keywords then specific terms are optimized in
-% searches)
-%
-%% ------------------------------------------------------------------------ %%
-
-% Example: \title{This is a test title}
-
-\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  AUTHORS AND AFFILIATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
-
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-
-\authors{John Fasullo\affil{1}, Benjamin Goldman\affil{2}}
-
-% \affiliation{1}{First Affiliation}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
-
-\affiliation{1}{Climate and Global Dynamics Division, National Center for Atmospheric Research, 1850 Table Mesa Dr., Boulder, CO, USA}
-
-\affiliation{2}{Science Research Program, White Plains High School, 550 North Street, White Plains, NY, USA}
-
-%(repeat as many times as is necessary)
-
-%% Corresponding Author:
-% Corresponding author mailing address and e-mail address:
-
-% (include name and email addresses of the corresponding author.  More
-% than one corresponding author is allowed in this LaTeX file and for
-% publication; but only one corresponding author is allowed in our
-% editorial system.)
-
-% Example: \correspondingauthor{First and Last Name}{email@address.edu}
-
-\correspondingauthor{Benjamin Goldman}{bg502257@live.wpcsd.k12.ny.us}
-
-%% Keypoints, final entry on title page.
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-
-% Example:
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-% \end{keypoints}
-
-\begin{keypoints}
-\item NCAR's CESM Large Ensemble predicts increase to Ni\~{n}o 3.4 variance over the 21$^{st}$ century.
-\item Increase can mainly be attributed to greenhouse gas and aerosol emissions.
-\item Changes to ENSO amplitude are linked to changes to equatorial Pacific ocean stratification.
-\end{keypoints}
-
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT and PLAIN LANGUAGE SUMMARY
-%
-% A good Abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
-
-% The Plain Language Summary should be written for a broad audience,
-% including journalists and the science-interested public, that will not have 
-% a background in your field.
-%
-% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
-% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
-% see http://sharingscience.agu.org/creating-plain-language-summary/)
-%
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
-
+\maketitle
 \begin{abstract}
@@
 The ensemble mean and standard error of each ensemble was calculated, while the Probability Density Function (PDF) is computed for the PI control simulation to estimate the statistical significance of simulated changes.
-We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Ni\~{n}o 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
-We identifed significant increases in variance of the Ni\~{n}o 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
+We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
+We identifed significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
 
@@
 
-%\section*{Plain Language Summary}
-
-%[ enter your Plain Language Summary here or delete this section]
-
-%% ------------------------------------------------------------------------ %%
-%
-%  TEXT
-%
-%% ------------------------------------------------------------------------ %%
-
-%%% Suggested section heads:
-% \section{Introduction}
-%
-% The main text should start with an introduction. Except for short
-% manuscripts (such as comments and replies), the text should be divided
-% into sections, each with its own heading.
-
-% Headings should be sentence fragments and do not begin with a
-% lowercase letter or number. Examples of good headings are:
-
-% \section{Materials and Methods}
-% Here is text on Materials and Methods.
-%
-% \subsection{A descriptive heading about methods}
-% More about Methods.
-%
-% \section{Data} (Or section title might be a descriptive heading about data)
-%
-% \section{Results} (Or section title might be a descriptive heading about the
-% results)
-%
-% \section{Conclusions}
-
 \section{Introduction}
-
-\note[bengoldman]{problem/entity}El Ni\~{n}o is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \cite{bjerknes1969atmospheric}.
-The reasons for studying ENSO are clear, as ENSO has drastically affected climate patterns worldwide, modulating rainfall and temperature in nearly every continent \cite{ropelewski1987global}.
-For example, the recent 2015-2016 El Ni\~{n}o event contributed to record-breaking high temperatures and droughts in South America \cite{jimenez2016record}.
+\label{sec:org7fb1b29}
+El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
+The reasons for studying ENSO are clear, as ENSO has drastically affected climate patterns worldwide, modulating rainfall and temperature in nearly every continent \citep{ropelewski1987global}.
+For example, the recent 2015-2016 El Niño event contributed to record-breaking high temperatures and droughts in South America \citep{jimenez2016record}.
 At the same time, long-term anthropogenic greenhouse gas emissions are causing global temperatures to increase through a greenhouse effect.
@@
 
-\note[bengoldman]{focus on future experiments, describe results of literature, not methods} \note[bengoldman]{past research} Research on the effect of external forcing on ENSO remains inconclusive, as results from similar studies conflict.
-\citeA{nowack2017role} predicted an overall increase in Ni\~{n}o 3.4 standard deviation under a combination of 4xCO$_2$ and interactive ozone forcing using single-model simulations, showing that greenhouse gasses increase the frequency of extreme ENSO by favoring a more Ni\~{n}o-like in the tropical Pacific, while ozone dampens this effect.
-In contrast, a few studies have found that ENSO amplitude decreases under global warming in certain coupled models \cite{kohyama2018weakening}.
+Research on the effect of external forcing on ENSO remains inconclusive, as results from similar studies conflict.
+\citet{nowack2017role} predicted an overall increase in Niño 3.4 standard deviation under a combination of 4xCO\(_2\) and interactive ozone forcing using single-model simulations, showing that greenhouse gasses increase the frequency of extreme ENSO by favoring a more Niño-like in the tropical Pacific, while ozone dampens this effect.
+In contrast, a few studies have found that ENSO amplitude decreases under global warming in certain coupled models \citep{kohyama2018weakening}.
 
-However, other studies have failed to find any statistically significant ENSO response to external forcing \cite{stevenson2012significant}.
-Analysis using NCAR's CESM Large Ensemble shows an ensemble size of at least 15 models is necessary to attribute changes to Ni\~{n}o variance to external forcing and reject the null hypothesis that internal variability is responsible for changes to ENSO \cite{zheng2018response}.
-Several modes of internal variability have been shown to modulate ENSO, including the AMO \cite{levine2017impact} and Tropical Pacific Decadal Variability (TPDV) \cite{zheng2018response}.
-An analysis of the Max Plank Institute's Grand Ensemble as well as NCAR's CESM Large Ensemble suggests that 80\% of changes to ENSO amplitude can be attributed to internal variability, but given a large enough ensemble, significant changes in ENSO amplitude due to climate change can be detected \cite{maher2018enso}.
+However, other studies have failed to find any statistically significant ENSO response to external forcing \citep{stevenson2012significant}.
+Analysis using NCAR's CESM Large Ensemble shows an ensemble size of at least 15 models is necessary to attribute changes to Niño variance to external forcing and reject the null hypothesis that internal variability is responsible for changes to ENSO \citep{zheng2018response}.
+Several modes of internal variability have been shown to modulate ENSO, including the AMO \citep{levine2017impact} and Tropical Pacific Decadal Variability (TPDV) \citep{zheng2018response}.
+An analysis of the Max Plank Institute's Grand Ensemble as well as NCAR's CESM Large Ensemble suggests that 80$\backslash$% of changes to ENSO amplitude can be attributed to internal variability, but given a large enough ensemble, significant changes in ENSO amplitude due to climate change can be detected \citep{maher2018enso}.
 
@@
 \section{Data}
-
+\label{sec:org91f7f64}
 The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
-The Large Ensemble contains 40 simulations of the CESM1 coupled model, forced with historic radiative forcing from 1920 to 2005 and according to the RCP 8.5 protocol from 2006 to 2100 \cite{kay2015community}.
+The Large Ensemble contains 40 simulations of the CESM1 coupled model, forced with historic radiative forcing from 1920 to 2005 and according to the RCP 8.5 protocol from 2006 to 2100 \citep{kay2015community}.
 The single forcing ensemble is a collection of sub-ensembles for various climate factors (greenhouse gas, aerosols, biomass burning, land use, ozone).
-Each simulation is forced by a combination of all factors except for one, allowing the impact of a single factor to be deduced by subtracting the ensemble mean from the fully-forced ensemble mean \cite{deser2020isolating}.
+Each simulation is forced by a combination of all factors except for one, allowing the impact of a single factor to be deduced by subtracting the ensemble mean from the fully-forced ensemble mean \citep{deser2020isolating}\}.
 For example, the xghg (greenhouse gas) ensemble is forced by aerosol emissions, biomass burning, land use, and ozone.
@@
 \section{Methods and Results}
-%% yay
+\label{sec:orgd9fdb5e}
 
-\subsection{Ni\~{n}o 3.4 Variance}
-We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Ni\~{n}o 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
+\subsection{Niño 3.4 Variance}
+\label{sec:orge2424fe}
+We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
 We measure variance on 20-year centered sliding windows.
@@
 The fully-forced ensemble exhibits moderate increase in variance, with the other ensembles showing less meaningful changes.
-In the fully-forced ensemble, the variance of the Ni\~{n}o 3.4 region increases beyond 2 standard errors of the control, increasing until the mid-21st century, and then decreasing gradually.
+In the fully-forced ensemble, the variance of the Niño 3.4 region increases beyond 2 standard errors of the control, increasing until the mid-21st century, and then decreasing gradually.
 The excluded greenhouse, aerosol, and biomass-burning ensembles also contain a slight increase, which may not be statistically significant.
 Although the excluded land use ensemble mean has a strong increase in variance, this result is unlikely to be meaningful due to the xlulc ensemble's low sample size.
-Additionally, all ensembles exhibit considerable noise, in line with previous studies \cite{maher2018enso}.
+Additionally, all ensembles exhibit considerable noise, in line with previous studies \citep{maher2018enso}.
 
-\begin{figure}
-    \includegraphics{figures/variance_1.pdf}
-    \caption{20-year windowed variance of the Ni\~{n}o 3.4 region SST, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc, f) fixedO3.
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./figures/variance_1.pdf}
+\caption{\label{fig:variance_1}20-year windowed variance of the Niño 3.4 region SST, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc, f) fixedO3.
 Grey bars are PI control mean and 1x, 2x, 3x standard error.
 Colored bars represent ensemble standard error.}
-    \label{fig:variance_1}
 \end{figure}
@@
 To test the likelihood of this explanation, we test the influence of the Atlantic Multidecadal Oscillation (AMO) and Atlantic Meridional Overturning Current (AMOC) on ENSO in the control simulation.
-The AMO has been shown to have an influence on ENSO strength and seasonal growth rate \cite{levine2017impact}.
-We filter the control Ni\~{n}o 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \cite{phillips2014evaluating}.
-The Probability Distribution Function of Ni\~{n}o 3.4 variance is estimated for AMO/C > 1/2$\sigma$ and < -1/-2$\sigma$ using a Kernel Density Estimation.
+The AMO has been shown to have an influence on ENSO strength and seasonal growth rate \citep{levine2017impact}.
+We filter the control Niño 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \citep{phillips2014evaluating}\}.
+The Probability Distribution Function of Niño 3.4 variance is estimated for AMO/C > 1/2\(\sigma\) and < -1/-2\(\sigma\) using a Kernel Density Estimation.
 No meaningful differences were found in the distribution of ENSO intensity under any of these conditions (See Suplementary information figure 1).
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
+\label{sec:org8837c57}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
 To separate the influence of a single factor from the fully-forced ensemble, we employ a bootstrap test.
-For each single-forcing ensemble, a single simulation is randomly selected, and the Ni\~{n}o 3.4 20-year variance record is subtracted from that from a randomly selected fully-forced simulation.
+For each single-forcing ensemble, a single simulation is randomly selected, and the Niño 3.4 20-year variance record is subtracted from that from a randomly selected fully-forced simulation.
 We repeat this process 1000 times for each ensemble, and then calculate the mean and standard error for each ensemble.
@@
 The greenhouse-only ensemble as well as the aerosol-only ensemble exhibit increased variance, signaling that greenhouse and aerosol emissions likely both play a significant role in ENSO's forced response in the full-forcing ensemble.
-Interestingly, the influence of greenhouse gasses and aerosols are non-conflicting, in contrast with previous studies that show opposite effects of greenhouse gas and aerosol forcing on ENSO \cite{stevenson2019forced}.
+Interestingly, the influence of greenhouse gasses and aerosols are non-conflicting, in contrast with previous studies that show opposite effects of greenhouse gas and aerosol forcing on ENSO \citep{stevenson2019forced}.
 All the other single forcing ensembles exhibit insignificant differences from the fully-forced ensemble.
 The biomass burning case shows very small deviations from the fully-forced case, while the ozone ensemble's period of recording is too small to draw meaningful conclusions.
-However, \citeA{nowack2017role} showed that ozone forcing may dampen the effect of greenhouse-forced increases to ENSO amplitude by reducing changes to Pacific sea temperature and the Walker Circulation.
+However, \citet{nowack2017role} showed that ozone forcing may dampen the effect of greenhouse-forced increases to ENSO amplitude by reducing changes to Pacific sea temperature and the Walker Circulation.
 The land use/cover case, while it does show large deviations from the fully-forced case, has an ensemble size (5 members) too small to lend any credibility to these changes.
 
-\begin{figure}
-    \includegraphics{figures/bootstrap_1.pdf}
-    \caption{Difference between fully forced and single forced ensembles derived from the bootstrap process for a) greenhouse gas, b) aerosol emissions c) biomass burning, d) land use/cover, e) ozone; f) compares the sum of the bootstrapped ensemble means for the single forcing ensembles (blue) and the ensemble mean for the full forcing ensemble, detrended by a constant to center on zero (orange).}
-    \label{fig:bootstrap_1}
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./figures/bootstrap_1.pdf}
+\caption{\label{fig:bootstrap_1}Difference between fully forced and single forced ensembles derived from the bootstrap process for a) greenhouse gas, b) aerosol emissions c) biomass burning, d) land use/cover, e) ozone; f) compares the sum of the bootstrapped ensemble means for the single forcing ensembles (blue) and the ensemble mean for the full forcing ensemble, detrended by a constant to center on zero (orange).}
 \end{figure}
 
-In both the fully-forced scenario, and the greenhouse and aerosol only simulations, there is noticeably reduced Ni\~{n}o 3.4 variance in the mid-20\^{th} century, below 2 standard errors of the control.
+In both the fully-forced scenario, and the greenhouse and aerosol only simulations, there is noticeably reduced Niño 3.4 variance in the mid-20$\backslash$\textsuperscript{th} century, below 2 standard errors of the control.
 We hypothesize that this discrepancy may be the result of anomalous initial conditions caused by internal variability of the control, as the control conditions are used to initialize all the forced runs.
-To test this hypothesis, we analyzed the impact of the Atlantic Multidecadal Oscillation (AMO) and the Atlantic Meridional Overturning Current (AMOC) on Ni\~{n}o 3.4 variance in the control simulation using records of the AMO and AMOC from the Climate Variability Diagnostics Package (CVDP) \cite{phillips2014evaluating}.
-We filtered the 20-year variance of the Ni\~{n}o 3.4 sea surface temperature in the control based on the strength of the AMO/AMOC, separating ENSO variance into groups where AMO < -1$\sigma$, AMO < -2$\sigma$, AMO > 1$\sigma$, AMO >2$\sigma$, and the same for AMOC.
+To test this hypothesis, we analyzed the impact of the Atlantic Multidecadal Oscillation (AMO) and the Atlantic Meridional Overturning Current (AMOC) on Niño 3.4 variance in the control simulation using records of the AMO and AMOC from the Climate Variability Diagnostics Package (CVDP) \citep{phillips2014evaluating}.
+We filtered the 20-year variance of the Niño 3.4 sea surface temperature in the control based on the strength of the AMO/AMOC, separating ENSO variance into groups where AMO < -1\(\sigma\), AMO < -2\(\sigma\), AMO > 1\(\sigma\), AMO >2\(\sigma\), and the same for AMOC.
 We then estimated the probability density function for each group using a kernel density estimator.
-We observed no consistent difference in the distribution of Ni\~{n}o 3.4 variance between any group.
+We observed no consistent difference in the distribution of Niño 3.4 variance between any group.
 So far the question of reduced variance is unanswered, and it will be addressed in further depth later in the study.
@@
 \subsection{Correlation With Ocean Temperature}
+\label{sec:org9556be7}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
 To do this we used 4 slices of ocean temperature from the fully-forced, xghg, and xaer ensembles, including a slice averaged along the equator and slices through the western, central, and eastern Pacific basins.
-We linearly detrend and smooth with a 30-year windowed mean the timeseries at each gridpoint and the Ni\~{n}o 3.4 variance.
-Next, we calculated the Pearson's correlation coefficient between each gridpoint and the Ni\~{n}o 3.4 variance timeseries.
+We linearly detrend and smooth with a 30-year windowed mean the timeseries at each gridpoint and the Niño 3.4 variance.
+Next, we calculated the Pearson's correlation coefficient between each gridpoint and the Niño 3.4 variance timeseries.
 The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt_1}, and for the central slice in figure \ref{fig:tempcep_1}.
 
-Overall, the majority of the Pacific basin exhibits negative correlation with Ni\~{n}o 3.4 variance when linearly detrended and smoothed.
+Overall, the majority of the Pacific basin exhibits negative correlation with Niño 3.4 variance when linearly detrended and smoothed.
 The fully-forced simulation contains strongly negative correlation between the subsurface layer of the equatorial Pacific and ENSO amplitude, leading to the hypothesis that forced changes to Pacific Ocean stratification may be connected to ENSO intensity.
@@
 Additionally, it is unclear whether the overall warming trend of the upper Pacific ocean is connected to changes to ENSO intensity.
-We notice that the high levels of correlation between the subsurface layer and Ni\~{n}o 3.4 variance are present in the xGHG and xAER ensembles, as well as the fully forced ensemble, suggesting that there is a relationship between ENSO amplitude and subsurface Pacific Ocean temperature in a variety of forcing scenarios.
+We notice that the high levels of correlation between the subsurface layer and Niño 3.4 variance are present in the xGHG and xAER ensembles, as well as the fully forced ensemble, suggesting that there is a relationship between ENSO amplitude and subsurface Pacific Ocean temperature in a variety of forcing scenarios.
 
-\begin{figure}
-    \includegraphics{figures/tempdt.pdf}
-    \caption{Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Ni\~{n}o 3.4 variance.
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./figures/tempdt.pdf}
+\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Niño 3.4 variance.
 Grey color indicates points where p<0.05.
 The white regions at 25 and 280 degrees East are Africa and South America.}
-    \label{fig:tempdt_1}
 \end{figure}
 
-\begin{figure}
-    \includegraphics{figures/tempcep.pdf}
-    \caption{Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Ni\~{n}o 3.4 variance.
-Grey color indicates points where p<0.05.}
-    \label{fig:tempcep_1}
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{./figures/tempcep.pdf}
+\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.
+Grey color indicates points where p>0.05.}
 \end{figure}
@@
 \section{Conclusions and Discussion}
-%% yay
+\label{sec:org1483582}
 
 In this study, predicted changes to ENSO amplitude are examined in the CESM1 Large Ensemble and Single Forcing Ensemble.
-The fully-forced large ensemble exhibits increased ENSO amplitude, as measured by calculating the 20-year variance of the Ni\~{n}o 3.4 index.
-These changes appear to be statistically significant, but need to be scrutinized further, as they are in disagreement with results of some previous studies \cite{stevenson2012significant}.
+The fully-forced large ensemble exhibits increased ENSO amplitude, as measured by calculating the 20-year variance of the Niño 3.4 index.
+These changes appear to be statistically significant, but need to be scrutinized further, as they are in disagreement with results of some previous studies \citep{stevenson2012significant}.
 Additionally, the fully-forced ensemble, as well as all the others contains considerable noise, as the individual members cover a wide range of variability.
@@
 It is still unclear what is driving this connection, as well as how external forcing such as global warming affects it.
-There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \cite{dewitte2013reinterpreting}.
+There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \citep{dewitte2013reinterpreting}.
 
-This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \cite{maher2018enso}.
-Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citeA{stevenson2019forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
+This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \citep{maher2018enso}.
+Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citet{stevenson2019forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
 In contrast, this study observes that the impact on ENSO amplitude of greenhouse gas and aerosols has the same sign.
-This result is also surprising given that greenhouse and aerosol emissions have been shown to have opposite effects on sea surface temperature and global circulation, with greenhouse emissions favoring a more Ni\~{n}o-like state with weakened Walker circulation and higher SST, and aerosol emissions counteracting these effects \cite{boer2000transient}.
+This result is also surprising given that greenhouse and aerosol emissions have been shown to have opposite effects on sea surface temperature and global circulation, with greenhouse emissions favoring a more Niño-like state with weakened Walker circulation and higher SST, and aerosol emissions counteracting these effects \citep{boer2000transient}.
 
@@
 
-%%
-
-%  Numbered lines in equations:
-%  To add line numbers to lines in equations,
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}
-
-%% Enter Figures and Tables near as possible to where they are first mentioned:
-%
-% DO NOT USE \psfrag or \subfigure commands.
-%
-% Figure captions go below the figure.
-% Table titles go above tables;  other caption information
-%  should be placed in last line of the table, using
-% \multicolumn2l{$^a$ This is a table note.}
-%
-%----------------
-% EXAMPLE FIGURES
-%
-% \begin{figure}
-% \includegraphics{example.png}
-% \caption{caption}
-% \end{figure}
-%
-% Giving latex a width will help it to scale the figure properly. A simple trick is to use \textwidth. Try this if large figures run off the side of the page.
-% \begin{figure}
-% \noindent\includegraphics[width=\textwidth]{anothersample.png}
-%\caption{caption}
-%\label{pngfiguresample}
-%\end{figure}
-%
-%
-% If you get an error about an unknown bounding box, try specifying the width and height of the figure with the natwidth and natheight options.
-This is common when trying to add a PDF figure without pdflatex.
-% \begin{figure}
-% \noindent\includegraphics[natwidth=800px,natheight=600px]{samplefigure.pdf}
-%\caption{caption}
-%\label{pdffiguresample}
-%\end{figure}
-%
-%
-% PDFLatex does not seem to be able to process EPS figures. You may want to try the epstopdf package.
-%
-
-%
-% ---------------
-% EXAMPLE TABLE
-%
-% \begin{table}
-% \caption{Time of the Transition Between Phase 1 and Phase 2$^{a}$}
-% \centering
-% \begin{tabular}{l c}
-% \hline
-%  Run  & Time (min)  \\
-% \hline
-%   $l1$  & 260   \\
-%   $l2$  & 300   \\
-%   $l3$  & 340   \\
-%   $h1$  & 270   \\
-%   $h2$  & 250   \\
-%   $h3$  & 380   \\
-%   $r1$  & 370   \\
-%   $r2$  & 390   \\
-% \hline
-% \multicolumn{2}{l}{$^{a}$Footnote text here.}
-% \end{tabular}
-% \end{table}
-
-%% SIDEWAYS FIGURE and TABLE
-% AGU prefers the use of {sidewaystable} over {landscapetable} as it causes fewer problems.
-%
-% \begin{sidewaysfigure}
-% \includegraphics[width=20pc]{figsamp}
-% \caption{caption here}
-% \label{newfig}
-% \end{sidewaysfigure}
-%
-%  \begin{sidewaystable}
-%  \caption{Caption here}
-% \label{tab:signif_gap_clos}
-%  \begin{tabular}{ccc}
-% one&two&three\\
-% four&five&six
-%  \end{tabular}
-%  \end{sidewaystable}
-
-%% If using numbered lines, please surround equations with \begin{linenomath*}...\end{linenomath*}
-%\begin{linenomath*}
-%\begin{equation}
-%y|{f} \sim g(m, \sigma),
-%\end{equation}
-%\end{linenomath*}
-
-%%% End of body of article
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional Appendix goes here
-%
-% The \appendix command resets counters and redefines section heads
-%
-% After typing \appendix
-%
-%\section{Here Is Appendix Title}
-% will show
-% A: Here Is Appendix Title
-%
-%\appendix
-%\section{Here is a sample appendix}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% Optional Glossary, Notation or Acronym section goes here:
-%
-%%%%%%%%%%%%%%
-% Glossary is only allowed in Reviews of Geophysics
-%  \begin{glossary}
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \term{Term}
-%   Term Definition here
-%  \end{glossary}
-
-%
-%%%%%%%%%%%%%%
-% Acronyms
-%   \begin{acronyms}
-%   \acro{Acronym}
-%   Definition here
-%   \acro{EMOS}
-%   Ensemble model output statistics
-%   \acro{ECMWF}
-%   Centre for Medium-Range Weather Forecasts
-%   \end{acronyms}
-
-%
-%%%%%%%%%%%%%%
-% Notation
-%   \begin{notation}
-%   \notation{$a+b$} Notation Definition here
-%   \notation{$e=mc^2$}
-%   Equation in German-born physicist Albert Einstein's theory of special
-%  relativity that showed that the increased relativistic mass ($m$) of a
-%  body comes from the energy of motion of the body—that is, its kinetic
-%  energy ($E$)—divided by the speed of light squared ($c^2$).
-%   \end{notation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%  ACKNOWLEDGMENTS
-%
-% The acknowledgments must list:
-%
-% >>>>	A statement that indicates to the reader where the data
-% 	supporting the conclusions can be obtained (for example, in the
-% 	references, tables, supporting information, and other databases).
-%
-% 	All funding sources related to this work from all authors
-%
-% 	Any real or perceived financial conflicts of interests for any
-%	author
-%
-% 	Other affiliations for any author that may be perceived as
-% 	having a conflict of interest with respect to the results of this
-% 	paper.
-%
-%
-% It is also the appropriate place to thank colleagues and other contributors.
-% AGU does not normally allow dedications.
-
-\acknowledgments
-
-%% ------------------------------------------------------------------------ %%
-%% References and Citations
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-% \bibliography{<name of your .bib file>} don't specify the file extension
-%
-% don't specify bibliographystyle
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\bibliography{references.bib}
+\section{Acknowledgements}
+\label{sec:org9a83329}
 This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
 1852977.
-Computing resources (doi:10.5065/D6RX99HX) were provided by the Climate Simulation Laboratory at NCAR's Computational and Information Systems Laboratory, sponsored by the National Science Foundation and other agencies.
+Computing resources (\href{https://doi.org/10.5065/D6RX99HX}{doi:10.5065/D6RX99HX}) were provided by the Climate Simulation Laboratory at NCAR's Computational and Information Systems Laboratory, sponsored by the National Science Foundation and other agencies.
 I would like to thank Ms.
@@
 
-%Reference citation instructions and examples:
-%
-% Please use ONLY \cite and \citeA for reference citations.
-% \cite for parenthetical references
-% ...as shown in recent studies (Simpson et al., 2019)
-% \citeA for in-text citations
-% ...Simpson et al. (2019) have shown...
-%
-%
-%...as shown by \citeA{jskilby}.
-%...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
-%...has been shown \cite{jskilbye}.
-%...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
-%... \cite <i.e.>[]{lewin76,carson86,bartoldy02,rinaldi03}.
-%...has been shown by \cite <e.g.,>[and others]{lewin76}.
-%
-% apacite uses < > for prenotes and [ ] for postnotes
-% DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
-%
-
+\bibliographystyle{apa}
+\bibliography{references}
 \end{document}
-
-\section*{extra stuff}
-
-More Information and Advice:
-
-%% ------------------------------------------------------------------------ %%
-%
-%  SECTION HEADS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Capitalize the first letter of each word (except for
-% prepositions, conjunctions, and articles that are
-% three or fewer letters).
-
-% AGU follows standard outline style; therefore, there cannot be a section 1 without
-% a section 2, or a section 2.3.1 without a section 2.3.2.
-% Please make sure your section numbers are balanced.
-% ---------------
-% Level 1 head
-%
-% Use the \section{} command to identify level 1 heads;
-% type the appropriate head wording between the curly
-% brackets, as shown below.
-%
-%An example:
-%\section{Level 1 Head: Introduction}
-%
-% ---------------
-% Level 2 head
-%
-% Use the \subsection{} command to identify level 2 heads.
-%An example:
-%\subsection{Level 2 Head}
-%
-% ---------------
-% Level 3 head
-%
-% Use the \subsubsection{} command to identify level 3 heads
-%An example:
-%\subsubsection{Level 3 Head}
-%
-%---------------
-% Level 4 head
-%
-% Use the \subsubsubsection{} command to identify level 3 heads
-% An example:
-%\subsubsubsection{Level 4 Head} An example.
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  IN-TEXT LISTS
-%
-%% ------------------------------------------------------------------------ %%
-%
-% Do not use bulleted lists; enumerated lists are okay.
-% \begin{enumerate}
-% \item
-% \item
-% \item
-% \end{enumerate}
-%
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATIONS
-%
-%% ------------------------------------------------------------------------ %%
-
-% Single-line equations are centered.
-% Equation arrays will appear left-aligned.
-
-Math coded inside display math mode \[ ...\]
- will not be numbered, e.g.,:
- \[ x^2=y^2 + z^2\]
-
- Math coded inside \begin{equation} and \end{equation} will
- be automatically numbered, e.g.,:
- \begin{equation}
- x^2=y^2 + z^2
- \end{equation}
-
-% To create multiline equations, use the
-% \begin{eqnarray} and \end{eqnarray} environment
-% as demonstrated below.
-\begin{eqnarray}
-  x_{1} & = & (x - x_{0}) \cos \Theta \nonumber \\
-        && + (y - y_{0}) \sin \Theta  \nonumber \\
-  y_{1} & = & -(x - x_{0}) \sin \Theta \nonumber \\
-        && + (y - y_{0}) \cos \Theta.
-\end{eqnarray}
-
-%If you don't want an equation number, use the star form:
-%\begin{eqnarray*}...\end{eqnarray*}
-
-% Break each line at a sign of operation
-% (+, -, etc.) if possible, with the sign of operation
-% on the new line.
-
-% Indent second and subsequent lines to align with
-% the first character following the equal sign on the
-% first line.
-
-% Use an \hspace{} command to insert horizontal space
-% into your equation if necessary. Place an appropriate
-% unit of measure between the curly braces, e.g.
-% \hspace{1in}; you may have to experiment to achieve
-% the correct amount of space.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  EQUATION NUMBERING: COUNTER
-%
-%% ------------------------------------------------------------------------ %%
-
-% You may change equation numbering by resetting
-% the equation counter or by explicitly numbering
-% an equation.
-
-% To explicitly number an equation, type \eqnum{}
-% (with the desired number between the brackets)
-% after the \begin{equation} or \begin{eqnarray}
-% command.  The \eqnum{} command will affect only
-% the equation it appears with; LaTeX will number
-% any equations appearing later in the manuscript
-% according to the equation counter.
-%
-
-% If you have a multiline equation that needs only
-% one equation number, use a \nonumber command in
-% front of the double backslashes (\\) as shown in
-% the multiline equation above.
-
-% If you are using line numbers, remember to surround
-% equations with \begin{linenomath*}...\end{linenomath*}
-
-%  To add line numbers to lines in equations:
-%  \begin{linenomath*}
-%  \begin{equation}
-%  \end{equation}
-%  \end{linenomath*}

### 2021-05-23 presentations/presentation_current/presentation.tex
lets see did this work?

% Created 2021-05-23 Sun 12:21
% Intended LaTeX compiler: pdflatex
\documentclass[presentation]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{natbib}
\renewcommand{\bibsection}{}
\usetheme{metropolis}
\author{Benjamin Goldman}
\date{\today}
\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={The Impact of Anthropogenic Forcing on ENSO Amplitude},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\begin{frame}[label={sec:org91e1c08}]{What is El Niño/Southern Oscillation (ENSO)?}
\begin{columns}
\begin{column}{0.6\columnwidth}
\begin{itemize}
\item Drives extreme weather around the world
\item Oscillation between warm and cold temperature in the Pacific Ocean
\item Some events are more strong than others
\item Significant effect on people: 2015-2016 event
\item Major issue is prediction
\end{itemize}
\end{column}

\begin{column}{0.4\columnwidth}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./images/el_nino.jpg}
\caption{Changes to tropical Pacific climate during El Niño. \url{https://www.esa.int/ESA\_Multimedia/Images/2018/08/El\_Nino}}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label={sec:org5f7e4ec}]{Long-Term vs Short-Term Change}
\begin{columns}
\begin{column}{0.4\columnwidth}
\begin{itemize}
\item Long-term change: climate change/global warming
\begin{itemize}
\item Causes: greenhouse gasses, aerosols (smoke), land use, etc.
\end{itemize}
\item Short-term change: climate variability
\begin{itemize}
\item ENSO, seasons, AMO (Atlantic Multidecadal Oscillation), etc.
\end{itemize}
\end{itemize}
\end{column}

\begin{column}{0.6\columnwidth}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./images/climate_change.pdf}
\caption{Global average temperature changes since 1880. Red line: smoothed average, black line: unsmoothed average. \url{https://data.giss.nasa.gov/gistemp/graphs\_v4}}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label={sec:org3adee30}]{Review of literature}
\begin{itemize}
\item \citet{chen2017possible}
\begin{itemize}
\item Past studies disagree about whether ENSO will strengthen or weaken.
\item Simulation discrepancy caused by modeling of ENSO mechanics.
\end{itemize}
\item \citet{maher2018enso}
\begin{itemize}
\item Used a large dataset of climate predictions.
\item ENSO may become stronger in the future.
\end{itemize}
\item \citet{cai2018increased}
\begin{itemize}
\item Found that models agree by using a more flexible way of defining ENSO events
\item ENSO is strengthening because global warming is leading to higher stratification.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={sec:org1492cdc}]{Research Goals}
\begin{itemize}
\item Overall changes to ENSO amplitude
\begin{itemize}
\item Estimate future changes to ENSO amplitude using the CESM1 dataset.
\end{itemize}
\item Role of individual factors
\begin{itemize}
\item Compare contributions of greenhouse gasses, aerosols, land use, biomass burning, and ozone to ENSO intensity.
\end{itemize}
\item Changes to ocean structure
\begin{itemize}
\item Examine changes to correlation coefficient between ENSO intensity and ocean temperature for each simulation.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={sec:orgc837d3d}]{Data: the CESM1 Large Ensemble}
\begin{itemize}
\item Explore hypothetical scenarios with a computer model \citet{kay2015community}.
\item Estimation of how the earth's climate actually works.
\item Experimental group: Receives input of rising greenhouse gas and/or aerosol levels.
\item Control group: Emissions fixed at levels before industrial revolution.
\end{itemize}
\end{frame}

\begin{frame}[label={sec:org5c44c75}]{Niño 3.4 Variance}
\begin{columns}
\begin{column}{0.6\columnwidth}
\begin{itemize}
\item How to calculate ENSO intensity in the model output?
\item Step 1: Calculate sea temperature in Niño 3.4 region of tropical Pacific Ocean.
\item Step 2: Convert temperature dataset to dataset representing change in temperature variation over time.
\item Calculate variance around one point, move point forward slightly, repeat.
\end{itemize}
\end{column}

\begin{column}{0.4\columnwidth}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./images/nino3_4.pdf}
\caption{Niño 3.4 region is the shaded box.}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label={sec:org8bfbd4b}]{Butterfly Effect: The Need for a Large Ensemble}
\begin{columns}
\begin{column}{0.5\columnwidth}
\begin{itemize}
\item Butterfly effect: Small differences in initial conditions can become big differences in end result \citep{lorenz1963deterministic}.
\item Each simulation by itself is inaccurate.
\item Repeat simulation with slightly different initial conditions.
\item Due to larger sample size, noise can be filtered out by calculating the mean.
\end{itemize}
\end{column}

\begin{column}{0.5\columnwidth}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./images/variance_3.pdf}
\caption{Niño 3.4 20-year variance for individual members in full forcing ensemble.}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label={sec:org7c52937}]{Model Predictions: ENSO in the Future}
\begin{columns}
\begin{column}{.5\columnwidth}
\begin{itemize}
\item Calculate mean and standard error of ENSO intensity in ensemble and control.
\item ENSO is predicted to intensify in the 21st century!
\item Statistically significant: exceeds 2 standard errors.
\item Decreasing variance after 2060: still under investigation.
\end{itemize}
\end{column}

\begin{column}{0.5\columnwidth}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./images/variance_2.pdf}
\caption{20-year variance of Niño 3.4 index for fully-forced ensemble. Grey bar shows control mean and standard errors}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label={sec:orge7e784d}]{Role of Greenhouse and Aerosol Emissions}
\begin{columns}
\begin{column}{0.5\columnwidth}
\begin{itemize}
\item Greenhouse gasses and aerosols contribute to increase in variance.
\item Aerosols and greenhouse gasses have same sign: disagree with previous studies \citep{deser2020isolating}
\item Greenhouse gasses and aerosols are both human-produced.
\end{itemize}
\end{column}

\begin{column}{0.5\columnwidth}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./images/bootstrap_2.pdf}
\caption{Influence of individual human factors. Yellow is greenhouse gasses, green is aerosols.}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label={sec:orga9b609d}]{Correlation With Changes in Ocean Temperature}
\begin{itemize}
\item Examine relationship between ocean temperature and ENSO intensity in each simulation.
\item Calculate correlation coefficient between ENSO intensity and ocean temperature.
\item Find correlation coefficient at each grid-point.
\end{itemize}
\end{frame}

\begin{frame}[label={sec:org81efb26}]{Physical Mediator: Heating Difference}
\begin{columns}
\begin{column}{.62\columnwidth}
\begin{itemize}
\item Strong negative correlation in fully forced ensemble below surface.
\item Positive correlation in greenhouse ensemble and weak/zero correlation in aerosols ensemble
\item Rising temperatures heat different layers of ocean at different rates, modifying heat transfer.
\end{itemize}
\end{column}

\begin{column}{.50\columnwidth}
\begin{figure}[htbp]
\centering
\includegraphics[width=.9\linewidth]{./images/diff_tempdt.pdf}
\caption{Correlation between ENSO intensity and ocean temperature in 3 major ensembles}
\end{figure}
\end{column}
\end{columns}
\end{frame}

\begin{frame}[label={sec:orgeac6a2d}]{Conclusion and Discussion}
\begin{itemize}
\item Predicted increase in variance
\begin{itemize}
\item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
\end{itemize}
\item Greenhouse gasses and aerosols
\begin{itemize}
\item Increase is likely caused by the combined influence of greenhouse gasses and aerosols.
\end{itemize}
\item Heat transfer
\begin{itemize}
\item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
\end{itemize}
\item Notable disagreement
\begin{itemize}
\item Greenhouse gasses and aerosols both increase ENSO amplitude, in contrast to \citet{deser2020isolating}
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={sec:orgb0a7373}]{Applications, Next Steps, Limitations}
\begin{itemize}
\item Improve prediction ability to help people prepare for increased likelihood of extreme weather.
\item Reduce danger by switching to renewable energy.
\item Limitations:
\begin{itemize}
\item Only used one climate model.
\item Niño 3.4 index may not be fully accurate for various models (Cai et. al. 2018).
\end{itemize}
\item Next steps:
\begin{itemize}
\item Work with other datasets, such as the new CESM2.
\item Examine other variables to further analyze mediator process.
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={sec:orgbef5f98}]{Acknowledgements}
\begin{itemize}
\item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No. 1852977.
\item Thank you to my teacher, my family, and my mentor!
\item Role of mentor:
\begin{itemize}
\item Provide raw data from his facility
\item Suggest methods and interpretations
\item Provide feedback on results
\item Make similar calculations to check student's results
\end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[label={sec:org99c254a}]{References}
\bibliographystyle{apa}
\fontsize{7pt}{7.2}\selectfont
\bibliography{references}
\end{frame}
\end{document}

### 2021-05-23 presentations/script.tex
lets see did this work?

% Created 2021-05-23 Sun 12:23
% Intended LaTeX compiler: pdflatex
\documentclass{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\class{Science Research}
\professor{Ms. Fleming}
\author{Benjamin Goldman}
\date{\today}
\title{Script}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Script},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle
Hello, my name is Ben Goldman, and I am presenting my study called ``The impact of industrial emissions on El Nino intensity.'' You've probably heard about some kind of extreme weather in the news, or experienced an unusually hot or cold year. This phenomenon is most likely caused by El Nino.

El Nino, which researchers also refer to as El Nino Southern Oscillation, or ENSO, is a crucial element of the climate system. El Nino, or ENSO, is a change in the temperature of the equatorial Pacific Ocean. This change in temperature leads to disruption of air currents and changes in rainfall. ENSO is often responsible for extreme weather, such as the catastrophic flooding in South America during the devastating 2016 event. Due to this extreme impact, it is important for people to predict ENSO, which is what my study is focused on.

The earth's climate is very complex because it changes in a variety of ways, namely long-term and short-term. The red line on this graph of global air temperature from 1880 to the present depicts a long-term change, in this case global warming. It is caused by human greenhouse gas emissions. However, climate also changes on short timescales, over the course of years instead of decades, as shown by the jagged black line. ENSO is one example of this short-term variability. Studying climate is often difficult because it may be unclear whether certain events can be attributed to short-term effects, long-term effects, or both. My study is on how short and long-term variability interact with each other, or more specifically, how do long term changes in global temperature affect the intensity of short-term changes, in this case, ENSO?

Scientists have attempted to answer this question of global warming's influence on ENSO, mainly using computer simulations of the earth's climate. Many researchers who ran these simulations found that ENSO is predicted to become more intense, while others predicted a weakening. In 2017, Chen et. al. found that differences between simulations's modeling of air movement above the Pacific Ocean was largely responsible. However, in recent years, it's become clear that ENSO is more likely to intensify rather than weaken. One interesting study showed that by using a more flexible method of measuring ENSO events, a majority of simulations predict increased ENSO intensity. It has also become clear that ENSO is intensifying due to increasing ocean stratification, or the temperature difference between the upper and central layers of the pacific ocean. However, it is still unclear what human activities are causing ENSO to intensify, so the aim of my study is to assess the role of greenhouse gasses in comparison to aerosols and natural changes.

My study has 3 main goals. The first goal is to measure predicted changes to ENSO intensity, judging whether they agree with those documented in past research. The next goal is to analyze the relative role of greenhouse gasses, like carbon dioxide, and aerosols, such as smoke and dust, in affecting ENSO intensity in the future, to figure out what human activities are contributing most to ENSO intensification. Finally, I intended to examine changes to stratification and heat transfer under greenhouse and aerosol forcing.

All of my data so far comes from a computer simulation of the climate, an estimation of the Earth's climate using mathematical equations. These equations allow researchers like me to make predictions about what will happen in the future if humans continue polluting the environment. Then, I used a control group that simulates future changes to the entire Earth's climate with predicted increases in greenhouse and aerosol emissions, and a control group that uses fixed preindustrial levels of pollutants.

My first step was to measure ENSO intensity in the simulation output. To do this, I averaged sea surface temperature in the Niño 3.4 box of the Pacific ocean, shown in this blue box. Then, I converted this temperature record into a record of ENSO intensity with a 20-year sliding windowed variance calculation. By finding variance around one point, and then moving that point forward one month, I derived a dataset capturing ENSO intensity changing over time.

This plot shows the results of my calculation. Each line represents ENSO intensity in each member of the dataset. Notice that they start out close together, and then gradually diverge over time. This is an example of the butterfly effect, where small differences in initial conditions evolve into significant differences. This chaotic property of climate means that by itself, each simulation run is quite inaccurate, but by taking the average of running a simulation multiple times, one can filter out a great deal of this noise.

ENSO is predicted to intensify in the future, as shown in this graph. The blue bar shows variance in the experimental simulations, while the grey bar shows that of the control. There is a significant increase over time, meaning that ENSO is becoming stronger in the models. Notice the decrease after 2060, which I am still investigating.

The next step is to determine how large of a role greenhouse emissions play versus aerosol emissions and natural factors. To do this, I had to separate the influence of each factor from the model output, by using simulations of the Earth's climate that used all but one category of influence. For example, the greenhouse gas dataset, or ensemble, was forced by aerosols and natural factors. Next, by subtracting this dataset from the original dataset which received all factors as input, I derived changes to ENSO intensity under only one single factor.

Here are the results for greenhouse gasses and aerosols, where the graph shows the portion of changes to ENSO intensity that are attributed only to these factors. Both the greenhouse gas and aerosol lines are above zero, meaning that they are contributing to ENSO intensity increasing. Natural factors were left out of this graph because their influences were all statistically insignificant. Interestingly, changes to ENSO intensity under aerosol emissions have the same sign as that from greenhouse gas emissions, where other studies predicted that aerosols should have an opposite effect from greenhouse gasses. Now both greenhouse gasses and aerosols are human-produced, meaning that the majority of changes to ENSO intensity are humanity's fault.

Finally, I looked at what changes to ocean structure occur under greenhouse and aerosol emissions, to determine what physical processes are mediating their influence. To do this, I calculated the correlation coefficient between ENSO intensity and sea temperature at the equator, after detrending and smoothing both inputs.

Shown here are the results of this process. Each plot represents a different forcing scenario. The fully-forced scenario exhibits a strongly negative correlation coefficient in the subsurface layer of the pacific, meaning that the temperature in that region is connected to ENSO intensity. This correlation coefficient is positive in the greenhouse gas-only dataset in the middle, and close to zero in the aerosol dataset in the third row. We believe that this region of the ocean is connected to ENSO intensity because rising air temperatures from the greenhouse effect lead to different layers of the ocean heating at different rates, affecting how heat is transferred during an ENSO event.

We can conclude from these results that ENSO will likely become more intense in the future, and that this increase can be attributed to a combination of greenhouse and aerosol emissions. These emissions affect ENSO intensity through modifying how heat is transferred between different layers of the ocean. These broad conclusions agree with past research such as Cai et. al. 2018, and more analysis is being done on those that do not, such as the sign of the changes to ENSO intensity in the aerosol-only simulation, conflicting with Deser et. al. 2020.

My research contributes to a growing body of evidence demonstrating that global warming causes extreme weather, which in turn, harms humans and natural ecosystems. Climate research like mine will continue to encourage people to limit their greenhouse gas emissions. This study has a few limitations, the most important of which being the fact that only one out of many existing models is used, and that the Niño 3.4 index has been shown to be inaccurate for certain models. The next steps for my research are to conduct similar analysis on the output from different models and to examine other possible mediators to ENSO intensity under global warming, such as air currents.

Thank you to the scientists at the National Center for Atmospheric Research for running the simulations that my research uses. Thank you to my family for their encouragement and enthusiasm for my project, and to my teacher for teaching me how to conduct research at this level. Thank you to my mentor for providing and compiling the raw data, for helping to interpret my results select methods, and for reviewing my work.

Thank you for listening.
\end{document}

### 2021-05-23 miscellaneous/Summer2020/UHSFiles/Code/20year_variance.py

@@ -1,0 +1,726 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Apr 27 14:01:34 2020
+
+@author: bengoldman
+"""
+# %%
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats, signal
+from matplotlib import gridspec
+from math import sqrt
+from random import randint
+import os
+from matplotlib import cm
+
+from helpers import *
+
+# %%
+
+os.chdir('/Users/bengoldman/ENSO-amplitude/Data')
+ctl = np.loadtxt("ctl.csv")
+ff_set = np.loadtxt("ff_set.csv")
+ff_1850 = np.loadtxt("ff_1850.csv")
+ghg_set = np.loadtxt("ghg_set.csv")
+aer_set = np.loadtxt("aer_set.csv")
+bmb_set = np.loadtxt("bmb_set.csv")
+luc_set = np.loadtxt("luc_set.csv")
+ozo_set = np.loadtxt("ozo_set.csv")
+oic_set = np.loadtxt("oic_set.csv")
+oic_1850 = np.loadtxt("oic_1850.csv")
+ctl_amoc = np.loadtxt("ctl_amoc.csv")
+ctl_amo = np.loadtxt("ctl_amo.csv")
+ff_tempdt_avg = np.load("/Users/bengoldman/ENSO-amplitude/Data/ff_tempdt_average.npy")
+lat = np.loadtxt("lat.csv")
+lon = np.loadtxt("lon.csv")
+lon1 = np.loadtxt("lon1.csv")
+depth = np.loadtxt("depth.csv")
+
+os.chdir('/Users/bengoldman/ENSO-amplitude/Figures')
+
+# %%
+ctl_var = pd.DataFrame(ctl).rolling(240, center=True).var().to_numpy()[:, 0]
+ff_var_set = pd.DataFrame(ff_set).rolling(240, center=True).var().to_numpy()
+ff_1850_var = pd.DataFrame(ff_1850).rolling(240, center=True).var().to_numpy()
+ghg_var_set = pd.DataFrame(ghg_set).rolling(240, center=True).var().to_numpy()
+aer_var_set = pd.DataFrame(aer_set).rolling(240, center=True).var().to_numpy()
[…]

### 2021-05-23 miscellaneous/Summer2020/UHSFiles/Code/control_variance.py

@@ -1,0 +1,62 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats, signal
+from matplotlib import gridspec
+from math import sqrt
+from random import randint
+import os
+from matplotlib import cm
+
+from helpers import *
+
+ctl_var_nan = ctl_var[~np.isnan(ctl_var)]
+
+x = np.linspace(0.7, 2, 500)
+
+ctl_var_div_20 = np.zeros((1000, 20))
+for i in range(20):
+    ctl_var_div_20[:, i] = ctl_var_nan[i*1000:(i+1)*1000]
+
+ctl_var_div_10 = np.zeros((2000, 10))
+for i in range(10):
+    ctl_var_div_10[:, i] = ctl_var_nan[i*2000:(i+1)*2000]
+
+ctl_var_div_5 = np.zeros((4000, 5))
+for i in range(5):
+    ctl_var_div_5[:, i] = ctl_var_nan[i*4000:(i+1)*4000]
+
+ctl_var_div_20_pdf = np.zeros((500, 20))
+for i in range(20):
+    ctl_var_div_20_pdf[:, i] = stats.gaussian_kde(ctl_var_div_20[:, i]).evaluate(x)
+
+ctl_var_div_10_pdf = np.zeros((500, 10))
+for i in range(10):
+    ctl_var_div_10_pdf[:, i] = stats.gaussian_kde(ctl_var_div_10[:, i]).evaluate(x)
+
+ctl_var_div_5_pdf = np.zeros((500, 5))
+for i in range(5):
+    ctl_var_div_5_pdf[:, i] = stats.gaussian_kde(ctl_var_div_5[:, i]).evaluate(x)
+
+ctl_var_pdf = stats.gaussian_kde(ctl_var_nan).evaluate(x)
+
+# %%
+gs = gridspec.GridSpec(20, 3)
+
+fig = plt.figure(figsize=(10, 10))
+
+for i in range(20):
+    ax = fig.add_subplot(gs[i, 0])
+    ax.plot(x, ctl_var_div_20_pdf[:, i])
+#    ax.set_xticklabels([])
+
+for i in range(10):
+    ax = fig.add_subplot(gs[2*i:2*i+2, 1])
+    ax.plot(x, ctl_var_div_10_pdf[:, i])
+
+for i in range(5):
+    ax = fig.add_subplot(gs[4*i:4*i+4, 2])
[…]

### 2021-05-23 miscellaneous/Summer2020/UHSFiles/Code/correlation_xsections.py

@@ -1,0 +1,79 @@
+from helpers import *
+from variance import ff_var_mean, ghg_var_mean, aer_var_mean
+
+# %%
+print("ff")
+ff_tempdt = np.load("../Data/ff_tempdt_avg.npy")
+ff_tempdt_rs, ff_tempdt_ps = correlation_detrended(ff_tempdt, ff_var_mean)
+
+# %%
+print("ff")
+ff_tempcep = np.load("../Data/ff_tempcep_avg.npy")
+ff_tempcep_rs, ff_tempcep_ps = correlation_detrended(ff_tempcep, ff_var_mean)
+
+# %%
+print("ff")
+ff_tempwep = np.load("../Data/ff_tempwep_avg.npy")
+ff_tempwep_rs, ff_tempwep_ps = correlation_detrended(ff_tempwep, ff_var_mean)
+
+# %%
+print("ff")
+ff_tempeep = np.load("../Data/ff_tempeep_avg.npy")
+ff_tempeep_rs, ff_tempeep_ps = correlation_detrended(ff_tempeep, ff_var_mean)
+
+# %%
+print("ghg")
+ghg_tempdt = np.load("../Data/ghg_tempdt_avg.npy")
+ghg_tempdt_rs, ghg_tempdt_ps = correlation_detrended(ghg_tempdt, ghg_var_mean)
+
+# %%
+print("ghg")
+ghg_tempcep = np.load("../Data/ghg_tempcep_avg.npy")
+ghg_tempcep_rs, ghg_tempcep_ps = correlation_detrended(ghg_tempcep, ghg_var_mean)
+
+# %%
+print("ghg")
+ghg_tempwep = np.load("../Data/ghg_tempwep_avg.npy")
+ghg_tempwep_rs, ghg_tempwep_ps = correlation_detrended(ghg_tempwep, ghg_var_mean)
+
+# %%
+print("ghg")
+ghg_tempeep = np.load("../Data/ghg_tempeep_avg.npy")
+ghg_tempeep_rs, ghg_tempeep_ps = correlation_detrended(ghg_tempeep, ghg_var_mean)
+
+# %%
+print("aer")
+aer_tempdt = np.load("../Data/aer_tempdt_avg.npy")
+aer_tempdt_rs, aer_tempdt_ps = correlation_detrended(aer_tempdt, aer_var_mean)
+
+# %%
+print("aer")
[…]

### 2021-05-23 miscellaneous/Summer2020/UHSFiles/Code/correlations.py

@@ -1,0 +1,10 @@
+from helpers import *
+from variance import ff_var_set
+
+# %%
+os.chdir("/Volumes/Extreme SSD/LE/TEMPDT/ff")
+ff = get_ensemble(0, "TEMPDT", 2172)
+ff_correlation = np.zeros((34, 60, 180))
+
+# %%
+correlation = list(map(lambda x, y, z: correlation_detrended(ff_var_set[x, y, z], ff_var_set), ff))

### 2021-05-23 miscellaneous/Summer2020/UHSFiles/Code/helpers.py

@@ -1,0 +1,223 @@
+import netCDF4 as nc
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats, signal
+from matplotlib import gridspec
+from math import sqrt
+from random import randint
+import os
+from matplotlib import cm, colors
+import re
+
+lat = np.loadtxt("../Data/lat.csv")
+lon = np.loadtxt("../Data/lon.csv")
+lon1 = np.loadtxt("../Data/lon1.csv")
+lat1 = np.loadtxt("../Data/lat1.csv")
+depth = np.loadtxt("../Data/depth.csv")/100
+
+date_2172 = np.linspace(1920, 2100, 2172)
+date_1872 = np.linspace(1850, 2005, 1872)
+date_1032 = np.linspace(1920, 2005, 1032)
+date_1932 = np.linspace(1920, 2080, 1932)
+date_1320 = np.linspace(1920, 2029, 1320)
+date_612 = np.linspace(1955, 2005, 612)
+date_960 = np.linspace(1920, 1999, 960)
+date_1800 = np.linspace(1850, 1999, 1800)
+date_80 = np.linspace(1920, 1999, 80)
+date_ctl_20600 = np.linspace(400, 2200, 21600)
+date_ctl_1800 = np.linspace(400, 2200, 1800)
+
+
+def envelope(ax, x, y, err, col):
+    ax.plot(x, y, color=col)
+    ax.fill_between(x, y-err, y+err, alpha=.2, color=col)
+
+
+def figlett(ax, a):
+    ax.text(0.01, 0.95, a, transform=ax.transAxes, va='top')
+
+size1 = (7, 4)
+size2 = (7, 8)
+mp = np.ones((400, 3))
+mp[21:40] = [.6, .6, .6]
+mp[-40:-21] = [.6, .6, .6]
+mp[0:20, 0] = np.linspace(0, 1, 20)
+mp[0:20, 1] = np.linspace(0, 1, 20)
+mp[0:20, 2] = 1
+mp[-21:-1, 0] =1
+mp[-21:-1, 1] = np.linspace(1, 0, 20)
+mp[-21:-1, 2] = np.linspace(1, 0, 20)
+cmap = colors.ListedColormap(mp)
+
+mp = np.zeros((2, 4))
+mp[1, :] = [1, 1, 1, 1]
+mp[0, :] = [1, 1, 1, 0]
[…]

### 2021-05-23 miscellaneous/Summer2020/UHSFiles/Code/open_xsections.py

@@ -1,0 +1,98 @@
+from helpers import *
+import feather as ft
+
+# %%
+print("ff tempdt")
+ff_tempdt = get_ensemble("/Volumes/Extreme SSD/LE/TEMPDT/ff", 0, "TEMPDT")
+ff_tempdt_avg = np.average(ff_tempdt, axis=0)
+del ff_tempdt
+os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
+np.save("ff_tempdt_avg", ff_tempdt_avg)
+
+# %%
+print("ff tempcep")
+ff_tempcep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPCEP/ff", 0, "TEMPCEP")
+ff_tempcep_avg = np.average(ff_tempcep, axis=0)
+del ff_tempcep
+os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
+np.save("ff_tempcep_avg", ff_tempcep_avg)
+
+# %%
+print("ff tempwep")
+ff_tempwep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPWEP/ff", 0, "TEMPWEP")
+ff_tempwep_avg = np.average(ff_tempwep, axis=0)
+del ff_tempwep
+os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
+np.save("ff_tempwep_avg", ff_tempwep_avg)
+
+# %%
+print("ff tempeep")
+ff_tempeep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPEEP/ff", 0, "TEMPEEP")
+ff_tempeep_avg = np.average(ff_tempeep, axis=0)
+del ff_tempeep
+os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
+np.save("ff_tempeep_avg", ff_tempeep_avg)
+
+# %%
+print("ghg tempdt")
+ghg_tempdt = get_ensemble("/Volumes/Extreme SSD/LE/TEMPDT/xghg", 1, "TEMPDT")
+ghg_tempdt_avg = np.average(ghg_tempdt, axis=0)
+del ghg_tempdt
+os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
+np.save("ghg_tempdt_avg", ghg_tempdt_avg)
+
+# %%
+print("ghg tempcep")
[…]

### 2021-05-23 miscellaneous/Summer2020/UHSFiles/Code/tempeep.py

@@ -1,0 +1,190 @@
+
+import sys
+
+sys.path.append("./")
+from helpers import *
+from variance import ff_var_mean, ghg_var_mean, aer_var_mean
+
+ff_tempeep_avg = np.load("../Data/ff_tempeep_average.npy")
+ff_tempeep_avg_det = ff_tempeep_avg
+
+ff_correlation = np.zeros(ff_tempeep_avg_det[0].shape)
+ff_p_vals = np.zeros(ff_tempeep_avg_det[0].shape)
+ff_var_mean_det = ff_var_mean
+signal.detrend(ff_var_mean_det[~np.isnan(ff_var_mean_det)], overwrite_data=True)
+ff_nino_ts = pd.DataFrame(ff_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
+for i in range(len(ff_tempeep_avg_det[0, :, 0])):
+    for j in range(len(ff_tempeep_avg_det[0, i, :])):
+        print(str(i) + ":" + str(j) )
+        if np.isnan(ff_tempeep_avg_det[200, i, j]):
+            r, p = (np.nan, np.nan)
+        else:
+            signal.detrend(ff_tempeep_avg_det[:, i, j], overwrite_data=True)
+            ff_tempeep_avg_det[:, i, j] = pd.DataFrame(ff_tempeep_avg_det[:, i, j]).rolling(360, center=True).mean().to_numpy()[:, 0]
+            r, p = stats.pearsonr(ff_tempeep_avg_det[300:-300, i, j], ff_nino_ts[300:-300])
+        ff_correlation[i, j], ff_p_vals[i, j] = r, p
+
+ghg_tempeep_avg = np.load("../Data/ghg_tempeep_average.npy")
+ghg_tempeep_avg_det = ghg_tempeep_avg
+
+ghg_correlation = np.zeros(ghg_tempeep_avg_det[0].shape)
+ghg_p_vals = np.zeros(ghg_tempeep_avg_det[0].shape)
+ghg_var_mean_det = ghg_var_mean
+signal.detrend(ghg_var_mean_det[~np.isnan(ghg_var_mean_det)], overwrite_data=True)
+ghg_nino_ts = pd.DataFrame(ghg_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
+for i in range(len(ghg_tempeep_avg_det[0, :, 0])):
[…]

### 2021-05-23 miscellaneous/Summer2020/UHSFiles/Code/tempwep.py

@@ -1,0 +1,192 @@
+import sys
+
+sys.path.append("./")
+from helpers import *
+from variance import ff_var_mean, ghg_var_mean, aer_var_mean
+
+ff_tempwep_avg = np.load("../Data/ff_tempwep_average.npy")
+ff_tempwep_avg_det = ff_tempwep_avg
+
+ff_correlation = np.zeros(ff_tempwep_avg_det[0].shape)
+ff_p_vals = np.zeros(ff_tempwep_avg_det[0].shape)
+ff_var_mean_det = ff_var_mean
+signal.detrend(ff_var_mean_det[~np.isnan(ff_var_mean_det)], overwrite_data=True)
+ff_nino_ts = pd.DataFrame(ff_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
+for i in range(len(ff_tempwep_avg_det[0, :, 0])):
+    for j in range(len(ff_tempwep_avg_det[0, i, :])):
+        print(str(i) + ":" + str(j) )
+        if np.isnan(ff_tempwep_avg_det[200, i, j]):
+            r, p = (np.nan, np.nan)
+        else:
+            signal.detrend(ff_tempwep_avg_det[:, i, j], overwrite_data=True)
+            ff_tempwep_avg_det[:, i, j] = pd.DataFrame(ff_tempwep_avg_det[:, i, j]).rolling(360, center=True).mean().to_numpy()[:, 0]
+            r, p = stats.pearsonr(ff_tempwep_avg_det[300:-300, i, j], ff_nino_ts[300:-300])
+        ff_correlation[i, j], ff_p_vals[i, j] = r, p
+
+ghg_tempwep_avg = np.load("../Data/ghg_tempwep_average.npy")
+ghg_tempwep_avg_det = ghg_tempwep_avg
+
+ghg_correlation = np.zeros(ghg_tempwep_avg_det[0].shape)
+ghg_p_vals = np.zeros(ghg_tempwep_avg_det[0].shape)
+ghg_var_mean_det = ghg_var_mean
+signal.detrend(ghg_var_mean_det[~np.isnan(ghg_var_mean_det)], overwrite_data=True)
+ghg_nino_ts = pd.DataFrame(ghg_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
+for i in range(len(ghg_tempwep_avg_det[0, :, 0])):
[…]

### 2021-05-23 miscellaneous/Summer2020/UHSFiles/Code/variance.py

@@ -1,0 +1,71 @@
+
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats, signal
+from math import sqrt
+from random import randint
+import os
+
+# %%
+
+os.chdir('/Users/bengoldman/ENSO-amplitude/Data')
+ctl = np.loadtxt("ctl.csv")
+ff_set = np.loadtxt("ff_set.csv")
+ff_1850 = np.loadtxt("ff_1850.csv")
+ghg_set = np.loadtxt("ghg_set.csv")
+aer_set = np.loadtxt("aer_set.csv")
+bmb_set = np.loadtxt("bmb_set.csv")
+luc_set = np.loadtxt("luc_set.csv")
+ozo_set = np.loadtxt("ozo_set.csv")
+oic_set = np.loadtxt("oic_set.csv")
+oic_1850 = np.loadtxt("oic_1850.csv")
+ctl_amoc = np.loadtxt("ctl_amoc.csv")
+ctl_amo = np.loadtxt("ctl_amo.csv")
+ff_tempdt_avg = np.load("/Users/bengoldman/ENSO-amplitude/Data/ff_tempdt_average.npy")
+lat = np.loadtxt("lat.csv")
+lon = np.loadtxt("lon.csv")
+lon1 = np.loadtxt("lon1.csv")
+depth = np.loadtxt("depth.csv")
+
+os.chdir('/Users/bengoldman/ENSO-amplitude/Figures')
+
+# %%
+ctl_var = pd.DataFrame(ctl).rolling(240, center=True).var().to_numpy()[:, 0]
+ff_var_set = pd.DataFrame(ff_set).rolling(240, center=True).var().to_numpy()
+ff_1850_var = pd.DataFrame(ff_1850).rolling(240, center=True).var().to_numpy()
+ghg_var_set = pd.DataFrame(ghg_set).rolling(240, center=True).var().to_numpy()
+aer_var_set = pd.DataFrame(aer_set).rolling(240, center=True).var().to_numpy()
+bmb_var_set = pd.DataFrame(bmb_set).rolling(240, center=True).var().to_numpy()
+luc_var_set = pd.DataFrame(luc_set).rolling(240, center=True).var().to_numpy()
+ozo_var_set = pd.DataFrame(ozo_set).rolling(240, center=True).var().to_numpy()
[…]

### 2021-05-23 data/Code/20year_variance.py
moved in data analysis code/figs/data

@@ -1,0 +1,663 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Mon Apr 27 14:01:34 2020
+
+@author: bengoldman
+"""
+# %%
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats, signal
+from matplotlib import gridspec
+from math import sqrt
+from random import randint
+import os
+from matplotlib import cm
+import sys
+from helpers import *
+
+# %%
+from variance import *
+
+# %%
+os.chdir("../Data")
+
+ctl = np.loadtxt("ctl.csv")
+ff_set = np.loadtxt("ff_set.csv")
+ff_1850 = np.loadtxt("ff_1850.csv")
+ghg_set = np.loadtxt("ghg_set.csv")
+aer_set = np.loadtxt("aer_set.csv")
+bmb_set = np.loadtxt("bmb_set.csv")
+luc_set = np.loadtxt("luc_set.csv")
+ozo_set = np.loadtxt("ozo_set.csv")
+oic_set = np.loadtxt("oic_set.csv")
+oic_1850 = np.loadtxt("oic_1850.csv")
+ctl_amoc = np.loadtxt("ctl_amoc.csv")
+ctl_amo = np.loadtxt("ctl_amo.csv")
+lat = np.loadtxt("lat.csv")
+lon = np.loadtxt("lon.csv")
+lon1 = np.loadtxt("lon1.csv")
+depth = np.loadtxt("depth.csv")
+
+# %%
+os.chdir('/Users/bengoldman/ENSO-amplitude/Figures')
+
+# %%
+ctl_var = pd.DataFrame(ctl).rolling(240, center=True).var().to_numpy()[:, 0]
+ff_var_set = pd.DataFrame(ff_set).rolling(240, center=True).var().to_numpy()
+ff_1850_var = pd.DataFrame(ff_1850).rolling(240, center=True).var().to_numpy()
+ghg_var_set = pd.DataFrame(ghg_set).rolling(240, center=True).var().to_numpy()
+aer_var_set = pd.DataFrame(aer_set).rolling(240, center=True).var().to_numpy()
[…]

### 2021-05-23 data/Code/control_variance.py

@@ -1,0 +1,147 @@
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats, signal
+from matplotlib import gridspec
+from math import sqrt
+from random import randint
+import os
+from matplotlib import cm
+
+from helpers import *
+from variance import *
+
+ctl_var_nan = ctl_var[~np.isnan(ctl_var)]
+
+x = np.linspace(0.7, 2, 500)
+
+ctl_var_div_20 = np.zeros((1000, 20))
+for i in range(20):
+    ctl_var_div_20[:, i] = ctl_var_nan[i*1000:(i+1)*1000]
+
+ctl_var_div_10 = np.zeros((2000, 10))
+for i in range(10):
+    ctl_var_div_10[:, i] = ctl_var_nan[i*2000:(i+1)*2000]
+
+ctl_var_div_5 = np.zeros((4000, 5))
+for i in range(5):
+    ctl_var_div_5[:, i] = ctl_var_nan[i*4000:(i+1)*4000]
+
+ctl_var_div_20_pdf = np.zeros((500, 20))
+for i in range(20):
+    ctl_var_div_20_pdf[:, i] = stats.gaussian_kde(ctl_var_div_20[:, i]).evaluate(x)
+
+ctl_var_div_10_pdf = np.zeros((500, 10))
+for i in range(10):
+    ctl_var_div_10_pdf[:, i] = stats.gaussian_kde(ctl_var_div_10[:, i]).evaluate(x)
+
+ctl_var_div_5_pdf = np.zeros((500, 5))
+for i in range(5):
+    ctl_var_div_5_pdf[:, i] = stats.gaussian_kde(ctl_var_div_5[:, i]).evaluate(x)
+
+ctl_var_pdf = stats.gaussian_kde(ctl_var_nan).evaluate(x)
+
+# %%
+gs = gridspec.GridSpec(20, 3)
+
+fig = plt.figure(figsize=(10, 10))
+
+for i in range(20):
+    ax = fig.add_subplot(gs[i, 0])
+    ax.plot(x, ctl_var_div_20_pdf[:, i])
+#    ax.set_xticklabels([])
+
+for i in range(10):
[…]

### 2021-05-23 data/Code/correlation_xsections.py

@@ -1,0 +1,231 @@
+# %%
+from helpers import *
+from variance import ff_var_mean, ghg_var_mean, aer_var_mean, ghg_bs_mean, aer_bs_mean
+
+# %%
+ff_tempdt = np.load("/Volumes/Extreme SSD/LE/compiled/ff_tempdt_avg.npy")
+ff_tempcep = np.load("/Volumes/Extreme SSD/LE/compiled/ff_tempcep_avg.npy")
+ff_tempwep = np.load("/Volumes/Extreme SSD/LE/compiled/ff_tempwep_avg.npy")
+ff_tempeep = np.load("/Volumes/Extreme SSD/LE/compiled/ff_tempeep_avg.npy")
+ghg_tempdt = np.load("/Volumes/Extreme SSD/LE/compiled/ghg_tempdt_avg.npy")
+ghg_tempcep = np.load("/Volumes/Extreme SSD/LE/compiled/ghg_tempcep_avg.npy")
+ghg_tempwep = np.load("/Volumes/Extreme SSD/LE/compiled/ghg_tempwep_avg.npy")
+ghg_tempeep = np.load("/Volumes/Extreme SSD/LE/compiled/ghg_tempeep_avg.npy")
+aer_tempdt = np.load("/Volumes/Extreme SSD/LE/compiled/aer_tempdt_avg.npy")
+aer_tempcep = np.load("/Volumes/Extreme SSD/LE/compiled/aer_tempcep_avg.npy")
+aer_tempwep = np.load("/Volumes/Extreme SSD/LE/compiled/aer_tempwep_avg.npy")
+aer_tempeep = np.load("/Volumes/Extreme SSD/LE/compiled/aer_tempeep_avg.npy")
+
+# %%
+print("ghg")
+diff_ghg_tempdt_rs, diff_ghg_tempdt_ps = correlation_detrended(-(ghg_tempdt-ff_tempdt[:1932]), ghg_bs_mean)
+print("ghg")
+diff_ghg_tempcep_rs, diff_ghg_tempcep_ps = correlation_detrended(-(ghg_tempcep-ff_tempcep[:1932]), ghg_bs_mean)
+print("ghg")
+diff_ghg_tempwep_rs, diff_ghg_tempwep_ps = correlation_detrended(-(ghg_tempwep-ff_tempwep[:1932]), ghg_bs_mean)
+print("ghg")
[…]

### 2021-05-23 data/Code/nino_contour.py

@@ -1,0 +1,48 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Sun May 17 11:51:06 2020
+
+@author: bengoldman
+"""
+import netCDF4 as nc
+import numpy as np
+import matplotlib.pyplot as plt
+from eofs.standard import Eof
+from matplotlib import cm
+import cartopy.crs as ccrs
+import cartopy
+
+# %%
+f1 = nc.Dataset("/Volumes/Extreme SSD/science_research/obs_data/HadISST_sst.nc")
+
+sst1 = f1.variables["sst"][:, 59:120, 300:359]
+sst2 = f1.variables["sst"][:, 59:120, 0:100]
+sst = np.concatenate((sst1, sst2), axis=2)
+
+lats = np.flip(f1.variables["latitude"][59:120])
+lons1 = f1.variables["longitude"][300:359]
+lons2 = f1.variables["longitude"][0:100] + 360
+lons = np.concatenate((lons1, lons2), axis=0)
+
+# %%
+sst[sst == - 1e+30] = np.nan
+
+solver = Eof(sst)
+eof1 = np.flip(solver.eofsAsCorrelation(neofs=2)[1], axis=0)
+
+# %%
+
+# %%
+proj = ccrs.PlateCarree(central_longitude=180)
+
+fig = plt.figure(figsize=(15, 5))
+ax = fig.add_subplot(111, projection=proj)
+
+ax.contourf(lons, lats, eof1, cmap="coolwarm", transform=ccrs.PlateCarree(central_longitude=0))
+
+ax.coastlines()
+ax.add_feature(cartopy.feature.LAND, edgecolor='black')
+
+fig.colorbar(cm.ScalarMappable(cmap="coolwarm"), ax=ax)
+fig.suptitle("ENSO Temperature Anomaly Pattern", y=.91)

### 2021-05-23 data/Code/openCVDP.py

@@ -1,0 +1,29 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Fri May 22 11:00:38 2020
+
+@author: bengoldman
+"""
+
+import numpy as np
+import netCDF4 as nc
+import os
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats
+from math import sqrt
+from random import randint
+
+
+# %%
+
+
+os.chdir('/Users/bengoldman/ENSO-amplitude/Data')
+
+ctl_cvdp = nc.Dataset("/Volumes/Extreme SSD/science_research/cvdp_data/CESM1_LENS_Coupled_Control.cvdp_data.401-2200.nc")
+ctl_amoc = ctl_cvdp.variables["amoc_timeseries_ann"][:]
+ctl_amo = ctl_cvdp.variables["amo_timeseries_mon"][:]
+
+np.savetxt("ctl_amoc.csv", ctl_amoc)
+np.savetxt("ctl_amo.csv", ctl_amo)

### 2021-05-23 data/Code/open_nino.py

@@ -1,0 +1,209 @@
+#!/usr/bin/env python3
+# -*- coding: utf-8 -*-
+"""
+Created on Sun Apr 26 16:12:51 2020
+
+@author: bengoldman
+"""
+
+# %%
+import numpy as np
+import netCDF4 as nc
+import os
+
+
+lat1 = 90
+lat2 = 101
+lon1 = 152
+lon2 = 192
+
+# %%
+os.chdir("/Volumes/Extreme SSD/science_research/large_ensemble/CESM_LE/TREFHT/")
+
+# Full forcing set
+ff_set = np.zeros((2172, 34))
+print("")
+print("Full forcing")
+for i in range(2, 36):
+    if i < 10:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.cam.h0.TREFHT.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.cam.h0.TREFHT.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.cam.h0.TREFHT.208101-210012.nc" % i)
+
+        ff_set[:, i-2] = np.concatenate((
+            np.average(f1.variables["TREFHT"][:, lat1:lat2, lon1:lon2], axis=(1, 2)),
+            np.average(f2.variables["TREFHT"][:, lat1:lat2, lon1:lon2], axis=(1, 2)),
+            np.average(f3.variables["TREFHT"][:, lat1:lat2, lon1:lon2], axis=(1, 2))
+            )) - 273
+        f1.close()
+        f2.close()
+        f3.close()
+    elif i < 34:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.cam.h0.TREFHT.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.cam.h0.TREFHT.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.cam.h0.TREFHT.208101-210012.nc" % i)
+
+        ff_set[:, i-2] = np.concatenate((
+            np.average(f1.variables["TREFHT"][:, lat1:lat2, lon1:lon2], axis=(1, 2)),
[…]

### 2021-05-23 data/Code/open_xsections.py

@@ -1,0 +1,98 @@
+from helpers import *
+import feather as ft
+
+# %%
+print("ff tempdt")
+ff_tempdt = get_ensemble("/Volumes/Extreme SSD/LE/TEMPDT/ff", 0, "TEMPDT")
+ff_tempdt_avg = np.average(ff_tempdt, axis=0)
+del ff_tempdt
+os.chdir("/Volumes/Extreme SSD/LE/compiled/")
+np.save("ff_tempdt_avg", ff_tempdt_avg)
+
+# %%
+print("ff tempcep")
+ff_tempcep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPCEP/ff", 0, "TEMPCEP")
+ff_tempcep_avg = np.average(ff_tempcep, axis=0)
+del ff_tempcep
+os.chdir("/Volumes/Extreme SSD/LE/compiled/")
+np.save("ff_tempcep_avg", ff_tempcep_avg)
+
+# %%
+print("ff tempwep")
+ff_tempwep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPWEP/ff", 0, "TEMPWEP")
+ff_tempwep_avg = np.average(ff_tempwep, axis=0)
+del ff_tempwep
+os.chdir("/Volumes/Extreme SSD/LE/compiled/")
+np.save("ff_tempwep_avg", ff_tempwep_avg)
+
+# %%
+print("ff tempeep")
+ff_tempeep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPEEP/ff", 0, "TEMPEEP")
+ff_tempeep_avg = np.average(ff_tempeep, axis=0)
+del ff_tempeep
+os.chdir("/Volumes/Extreme SSD/LE/compiled/")
+np.save("ff_tempeep_avg", ff_tempeep_avg)
+
+# %%
+print("ghg tempdt")
+ghg_tempdt = get_ensemble("/Volumes/Extreme SSD/LE/TEMPDT/xghg", 1, "TEMPDT")
+ghg_tempdt_avg = np.average(ghg_tempdt, axis=0)
+del ghg_tempdt
+os.chdir("/Volumes/Extreme SSD/LE/compiled/")
+np.save("ghg_tempdt_avg", ghg_tempdt_avg)
+
+# %%
+print("ghg tempcep")
+ghg_tempcep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPCEP/xghg", 1, "TEMPCEP")
[…]

### 2021-05-23 data/Code/pptfigs.py

@@ -1,0 +1,16 @@
+from helpers import *
+
+import cartopy.crs as ccrs
+import cartopy
+import matplotlib.patches as mpatches
+
+fig = plt.figure(figsize=(4, 4))
+ax = fig.add_subplot(projection=ccrs.PlateCarree(central_longitude=180))
+ax.coastlines()
+ax.gridlines()
+ax.add_patch(mpatches.Rectangle(xy=[-170, -5], width=50, height=10, color="C0", transform=ccrs.PlateCarree()))
+ax.set_xlim((-75, 100))
+ax.set_ylim((-80, 80))
+plt.tight_layout()
+plt.savefig("../Figures/nino.pdf")
+

### 2021-05-23 data/Code/tempdt.py

@@ -1,0 +1,164 @@
+from helpers import *
+from variance import ff_var_mean, ghg_var_mean, aer_var_mean
+
+ff_tempdt_avg = np.load("../Data/ff_tempdt_average.npy")
+ff_tempdt_avg_det = ff_tempdt_avg
+
+ff_correlation = np.zeros(ff_tempdt_avg_det[0].shape)
+ff_p_vals = np.zeros(ff_tempdt_avg_det[0].shape)
+ff_var_mean_det = ff_var_mean
+signal.detrend(ff_var_mean_det[~np.isnan(ff_var_mean_det)], overwrite_data=True)
+ff_nino_ts = pd.DataFrame(ff_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
+for i in range(len(ff_tempdt_avg_det[0, :, 0])):
+    for j in range(len(ff_tempdt_avg_det[0, i, :])):
+        print(str(i) + ":" + str(j) )
+        if np.isnan(ff_tempdt_avg_det[200, i, j]):
+            r, p = (np.nan, np.nan)
+        else:
+            signal.detrend(ff_tempdt_avg_det[:, i, j], overwrite_data=True)
+            ff_tempdt_avg_det[:, i, j] = pd.DataFrame(ff_tempdt_avg_det[:, i, j]).rolling(360, center=True).mean().to_numpy()[:, 0]
+            r, p = stats.pearsonr(ff_tempdt_avg_det[300:-300, i, j], ff_nino_ts[300:-300])
+        ff_correlation[i, j], ff_p_vals[i, j] = r, p
+
+ghg_tempdt_avg = np.load("../Data/ghg_tempdt_average.npy")
+ghg_tempdt_avg_det = ghg_tempdt_avg
+
+ghg_correlation = np.zeros(ghg_tempdt_avg_det[0].shape)
+ghg_p_vals = np.zeros(ghg_tempdt_avg_det[0].shape)
+ghg_var_mean_det = ghg_var_mean
+signal.detrend(ghg_var_mean_det[~np.isnan(ghg_var_mean_det)], overwrite_data=True)
+ghg_nino_ts = pd.DataFrame(ghg_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
+for i in range(len(ghg_tempdt_avg_det[0, :, 0])):
+    for j in range(len(ghg_tempdt_avg_det[0, i, :])):
[…]

### 2021-05-23 data/Code/variance.py

@@ -1,0 +1,127 @@
+
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats, signal
+from math import sqrt
+from random import randint
+import os
+
+# %%
+
+os.chdir('/Users/bengoldman/ENSO-amplitude/Data')
+ctl = np.loadtxt("ctl.csv")
+ff_set = np.loadtxt("ff_set.csv")
+ff_1850 = np.loadtxt("ff_1850.csv")
+ghg_set = np.loadtxt("ghg_set.csv")
+aer_set = np.loadtxt("aer_set.csv")
+bmb_set = np.loadtxt("bmb_set.csv")
+luc_set = np.loadtxt("luc_set.csv")
+ozo_set = np.loadtxt("ozo_set.csv")
+oic_set = np.loadtxt("oic_set.csv")
+oic_1850 = np.loadtxt("oic_1850.csv")
+ctl_amoc = np.loadtxt("ctl_amoc.csv")
+ctl_amo = np.loadtxt("ctl_amo.csv")
+lat = np.loadtxt("lat.csv")
+lon = np.loadtxt("lon.csv")
+lon1 = np.loadtxt("lon1.csv")
+depth = np.loadtxt("depth.csv")
+
+os.chdir('/Users/bengoldman/ENSO-amplitude/Figures')
+
+# %%
+ctl_var = pd.DataFrame(ctl).rolling(240, center=True).var().to_numpy()[:, 0]
+ff_var_set = pd.DataFrame(ff_set).rolling(240, center=True).var().to_numpy()
+ff_1850_var = pd.DataFrame(ff_1850).rolling(240, center=True).var().to_numpy()
+ghg_var_set = pd.DataFrame(ghg_set).rolling(240, center=True).var().to_numpy()
+aer_var_set = pd.DataFrame(aer_set).rolling(240, center=True).var().to_numpy()
+bmb_var_set = pd.DataFrame(bmb_set).rolling(240, center=True).var().to_numpy()
+luc_var_set = pd.DataFrame(luc_set).rolling(240, center=True).var().to_numpy()
+ozo_var_set = pd.DataFrame(ozo_set).rolling(240, center=True).var().to_numpy()
[…]

### 2021-05-23 data/Code/variance.r

@@ -1,0 +1,119 @@
+library(zoo)
+library(ncdf4)
+library(ggplot2)
+library(dplyr)
+library(gridExtra)
+library(grid)
+library(boot)
+
+
+
+
+rollvar <- function(x) {
+  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
+  return(out)
+}
+
+se <- function(x, ...) sqrt(var(x, ...)/length(x))
+
+extract_nino3.4 <- function(ncfile, xvar) {
+  ncin <- nc_open(ncfile)
+  out <- apply(
+    ncvar_get(ncin, xvar, start = c(153, 91, 1, 1), count = c(40, 11, -1, -1)), 
+    MARGIN = c(3, 4),
+    FUN = mean
+  )
+  return(out)
+  nc_close(ncin)
+}
+
+
+ncin <- nc_open("/Volumes/Extreme SSD/CESM/CESM1/CESM1/TREFHT/b.e11.BOTH.f09_g16.ensemble.cam.h0.TREFHT.192001-210012.nc")
+date <- as.Date(ncvar_get(ncin, "time"), origin = "1920-01-01")
+date <- seq(1920, 2100, length.out = 2172)
+
+path_ff <- "/Volumes/Extreme\ SSD/CESM/CESM1/CESM1/TREFHT/"
+path_sf <- "/Volumes/Extreme\ SSD/CESM/CESM1/CESM1SF/TREFHT/"
+
+file_ff <- "b.e11.BOTH.f09_g16.ensemble.cam.h0.TREFHT.192001-210012.nc"
+files_sf <- c(
+  "b.e11.B20TRLENS_RCP85.f09_g16.xghg.ensemble.cam.h0.TREFHT.192001-208012.nc",
+  "b.e11.B20TRLENS_RCP85.f09_g16.xaer.ensemble.cam.h0.TREFHT.192001-208012.nc",
+  "b.e11.B20TRLENS_RCP85.f09_g16.xbmb.ensemble.cam.h0.TREFHT.192001-202912.nc",
+  "b.e11.B20TRLENS_RCP85.f09_g16.xlulc.ensemble.cam.h0.TREFHT.192001-202912.nc",
+  "b.e11.B20LE_fixedO3_ensemble.cam.h0.TREFHT.195501-200512.nc"
+  )
[…]

### 2021-05-23 presentations/presentation_current/presentation.tex
stuff

@@
-% Created 2021-05-23 Sun 12:21
+% Created 2021-05-23 Sun 17:17
 % Intended LaTeX compiler: pdflatex
@@
 
-\begin{frame}[label={sec:org91e1c08}]{What is El Niño/Southern Oscillation (ENSO)?}
+\begin{frame}[label={sec:orgb06ffc8}]{What is El Niño/Southern Oscillation (ENSO)?}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org5f7e4ec}]{Long-Term vs Short-Term Change}
+\begin{frame}[label={sec:org8030abb}]{Long-Term vs Short-Term Change}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org3adee30}]{Review of literature}
+\begin{frame}[label={sec:org5552ddc}]{Review of literature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org1492cdc}]{Research Goals}
+\begin{frame}[label={sec:orgc04d095}]{Research Goals}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgc837d3d}]{Data: the CESM1 Large Ensemble}
+\begin{frame}[label={sec:orgf0d7e18}]{Data: the CESM1 Large Ensemble}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org5c44c75}]{Niño 3.4 Variance}
+\begin{frame}[label={sec:orgd2d90e6}]{Niño 3.4 Variance}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org8bfbd4b}]{Butterfly Effect: The Need for a Large Ensemble}
+\begin{frame}[label={sec:orgfb6c2f9}]{Butterfly Effect: The Need for a Large Ensemble}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org7c52937}]{Model Predictions: ENSO in the Future}
+\begin{frame}[label={sec:orgc7880f2}]{Model Predictions: ENSO in the Future}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orge7e784d}]{Role of Greenhouse and Aerosol Emissions}
+\begin{frame}[label={sec:org07ac6e1}]{Analysis of Individual Factors}
+\begin{itemize}
+\item Why is ENSO predicted to intensify? What human impacts play the largest role?
+\begin{itemize}
+\item Factors include: Greenhouse gasses, aerosols, natural factors.
+\end{itemize}
+\item Separate out individual influences in model output.
+\item Single forcing ensembles: forced by all factors except for 1.
+\item Subtract “all-but-one” ensembles from original “full-forcing” ensemble.
+\item Resulting data represents influence of only one factor.
+\end{itemize}
+\end{frame}
+
+\begin{frame}[label={sec:org50ac92b}]{Role of Greenhouse and Aerosol Emissions}
 \begin{columns}
@@
 \item Greenhouse gasses and aerosols contribute to increase in variance.
-\item Aerosols and greenhouse gasses have same sign: disagree with previous studies \citep{deser2020isolating}
+\item Aerosols and greenhouse gasses have same sign: disagree with previous studies \citep{deser2020isolating}.
 \item Greenhouse gasses and aerosols are both human-produced.
@@
 
-\begin{frame}[label={sec:orga9b609d}]{Correlation With Changes in Ocean Temperature}
+\begin{frame}[label={sec:orgdf0933d}]{Correlation With Changes in Ocean Temperature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org81efb26}]{Physical Mediator: Heating Difference}
+\begin{frame}[label={sec:orgc26b911}]{Physical Mediator: Heating Difference}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgeac6a2d}]{Conclusion and Discussion}
+\begin{frame}[label={sec:org85cbf31}]{Conclusion and Discussion}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgb0a7373}]{Applications, Next Steps, Limitations}
+\begin{frame}[label={sec:orgbda44eb}]{Applications, Next Steps, Limitations}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgbef5f98}]{Acknowledgements}
+\begin{frame}[label={sec:org1c1bea6}]{Acknowledgements}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org99c254a}]{References}
+\begin{frame}[label={sec:org1cb487a}]{References}
 \bibliographystyle{apa}

### 2021-05-23 presentations/script.tex
stuff

@@
-% Created 2021-05-23 Sun 12:23
+% Created 2021-05-23 Sun 17:01
 % Intended LaTeX compiler: pdflatex
@@
 \maketitle
-Hello, my name is Ben Goldman, and I am presenting my study called ``The impact of industrial emissions on El Nino intensity.'' You've probably heard about some kind of extreme weather in the news, or experienced an unusually hot or cold year.
+Hello, my name is Ben Goldman, and I am presenting my study called ``The impact of anthropogenic forcing on ENSO amplitude.'' You've probably heard about some kind of extreme weather in the news, or experienced an unusually hot or cold year.
 This phenomenon is most likely caused by El Nino.
@@
 This change in temperature leads to disruption of air currents and changes in rainfall.
-ENSO is often responsible for extreme weather, such as the catastrophic flooding in South America during the devastating 2016 event.
+ENSO is often responsible for extreme weather, such as the catastrophic flooding in South America during the 2016 event.
 Due to this extreme impact, it is important for people to predict ENSO, which is what my study is focused on.
@@
 The earth's climate is very complex because it changes in a variety of ways, namely long-term and short-term.
-The red line on this graph of global air temperature from 1880 to the present depicts a long-term change, in this case global warming.
-It is caused by human greenhouse gas emissions.
-However, climate also changes on short timescales, over the course of years instead of decades, as shown by the jagged black line.
+The red line on this graph of global air temperature from 1880 to the present depicts a long-term change, in this case global warming, which is caused by human greenhouse gas emissions.
+However, the climate also changes on short timescales, over the course of years instead of decades, as shown by the jagged black line.
 ENSO is one example of this short-term variability.
@@
 found that differences between simulations's modeling of air movement above the Pacific Ocean was largely responsible.
-However, in recent years, it's become clear that ENSO is more likely to intensify rather than weaken.
-One interesting study showed that by using a more flexible method of measuring ENSO events, a majority of simulations predict increased ENSO intensity.
+However, in recent years, it's become clear that ENSO is more likely to intensify rather than weaken, as studies like Maher et.
+al.
+(2018) showed that in a large group of simulations, a majority of them showed increased ENSO strength.
+Also, Cai et.
+al.
+(2018) showed that by using a more flexible method of measuring ENSO events, a majority of simulations predict increased ENSO intensity.
 It has also become clear that ENSO is intensifying due to increasing ocean stratification, or the temperature difference between the upper and central layers of the pacific ocean.
-However, it is still unclear what human activities are causing ENSO to intensify, so the aim of my study is to assess the role of greenhouse gasses in comparison to aerosols and natural changes.
+However, it is still unclear what human activities are causing ENSO to intensify, so the aim of my study is to assess the role of greenhouse gasses in comparison to aerosols and natural changes in forcing increased ENSO amplitude.
 
-My study has 3 main goals.
-The first goal is to measure predicted changes to ENSO intensity, judging whether they agree with those documented in past research.
-The next goal is to analyze the relative role of greenhouse gasses, like carbon dioxide, and aerosols, such as smoke and dust, in affecting ENSO intensity in the future, to figure out what human activities are contributing most to ENSO intensification.
-Finally, I intended to examine changes to stratification and heat transfer under greenhouse and aerosol forcing.
+So my first goal of this study is to measure predicted changes to ENSO intensity, judging whether they agree with those documented in past research.
+The next goal is to analyze the relative role of greenhouse gasses, like carbon dioxide, and aerosols, such as smoke and dust, in affecting ENSO intensity in the future.
+This will enable me to figure out what human activities are contributing most to ENSO intensification.
+Finally, I intended to examine changes to stratification and heat transfer under greenhouse and aerosol forcing, to examine what is actually happening on the planet that translates increased temperature into strengthened ENSO.
 
@@
 These equations allow researchers like me to make predictions about what will happen in the future if humans continue polluting the environment.
-Then, I used a control group that simulates future changes to the entire Earth's climate with predicted increases in greenhouse and aerosol emissions, and a control group that uses fixed preindustrial levels of pollutants.
+I used an experimental group that simulates future changes to the entire Earth's climate with predicted increases in greenhouse and aerosol emissions, and a control group that uses fixed preindustrial levels of pollutants.
 
@@
 There is a significant increase over time, meaning that ENSO is becoming stronger in the models.
-Notice the decrease after 2060, which I am still investigating.
+There is a decrease after 2060, which I am still investigating.
 

### 2021-06-01 paperwork/classmates papers/celia_feedback.tex
presentation edits

% Created 2021-05-27 Thu 09:03
% Intended LaTeX compiler: pdflatex
\documentclass{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{Benjamin Goldman}
\date{14 May 2021}
\title{Celia Paper Feedback}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Celia Paper Feedback},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{Abstract}
\label{sec:org9eff226}
\begin{itemize}
\item You might want to put everything in past tense or future tense, since the paper will be submitted after you have results
\item Otherwise, it is very clear and concise
\item Also, sometimes they say not to use first person. It seems fine as it is but Ms. Fleming might have a preference
\end{itemize}

\section{Introduction}
\label{sec:org2841516}
\begin{itemize}
\item Paragraph 1 does a good job introducing the topic, I especially liked the last sentence about how light pollution is as of yet mostly unquantified.
\item In the sentence with ``\ldots{}can easily accumulate over and light pollution is a problem\ldots{}'' I would turn the and into a sentence break so it is easier to read.
\item For paragraph 2 when you cite a paper, you should use parenthetical citations
\begin{itemize}
\item So like say ``According to Longcore and Rich (2004), light pollution could \ldots{}'' or ``Light pollution could alter \ldots{} (Longcore and Rich 2004).''
\end{itemize}
\item Expand a little bit on paragraph 2, maybe cite a few more recent papers, and say how they affect your hypothesis by creating a gap.
\item I would only include the figure if it directly supports the ideas in the introduction. Its a cool picture, but I'm not sure its necessary. Maybe you could use like a picture of sources of light pollution or something.
\item Maybe in the 3rd paragraph elaborate a little bit on the dependent variable, so like say what the implications of a high light level are, etc.
\item Are you examining how light pollution is changing over time, or on average in a certain time span, or only in the present?
\item Similar citation thing for p. 4.
\item Add a sentence or 2 after each paper citation to add more for how it impacts your study or other studies.
\item For the last intro paragraph, you mention your hypothesis being supported, but it's not clear what your hypothesis actually is.
\item The part about the designation being worth it or not is interesting, and because it would be an implication of your results, you should clarify what you mean about it.
\item You could also mention ecosystems and human health in addition to tourism
\item Not clear what the gap in literature is
\item Overall, you did a good job introducing each paper so that they all contribute to your study, but what's missing is an explanation of the importance of each paper, and the gap.
\end{itemize}

\section{Methodology}
\label{sec:org9ba9d76}
\begin{itemize}
\item Do a separate paragraph on data where you talk about where the raw lighting data is coming from. Is it satellites, human recordings, some kind of meter, etc?
\item Median
\item I think past or present tense would work fine here too.
\item You mentioned using ANOVA?
\end{itemize}
\end{document}

### 2021-06-01 presentations/presentation_current/presentation.tex
presentation edits

@@
-% Created 2021-05-23 Sun 17:17
+% Created 2021-06-01 Tue 09:02
 % Intended LaTeX compiler: pdflatex
@@
 
-\begin{frame}[label={sec:orgb06ffc8}]{What is El Niño/Southern Oscillation (ENSO)?}
+\begin{frame}[label={sec:org94c92b5}]{What is El Niño/Southern Oscillation (ENSO)?}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org8030abb}]{Long-Term vs Short-Term Change}
+\begin{frame}[label={sec:orga6e57de}]{Long-Term vs Short-Term Change}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org5552ddc}]{Review of literature}
+\begin{frame}[label={sec:org98f4b82}]{Review of literature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgc04d095}]{Research Goals}
+\begin{frame}[label={sec:orgeb64aee}]{Research Goals}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgf0d7e18}]{Data: the CESM1 Large Ensemble}
+\begin{frame}[label={sec:orgced4cac}]{Data: the CESM1 Large Ensemble}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgd2d90e6}]{Niño 3.4 Variance}
+\begin{frame}[label={sec:orgc2d187b}]{Niño 3.4 Variance}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgfb6c2f9}]{Butterfly Effect: The Need for a Large Ensemble}
+\begin{frame}[label={sec:org2b84dfc}]{Butterfly Effect: The Need for a Large Ensemble}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgc7880f2}]{Model Predictions: ENSO in the Future}
+\begin{frame}[label={sec:org82b4406}]{Model Predictions: ENSO in the Future}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org07ac6e1}]{Analysis of Individual Factors}
+\begin{frame}[label={sec:orgc997c2e}]{Analysis of Individual Factors}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org50ac92b}]{Role of Greenhouse and Aerosol Emissions}
+\begin{frame}[label={sec:org948ad63}]{Role of Greenhouse and Aerosol Emissions}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgdf0933d}]{Correlation With Changes in Ocean Temperature}
+\begin{frame}[label={sec:orga1b0a8e}]{Correlation With Changes in Ocean Temperature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgc26b911}]{Physical Mediator: Heating Difference}
+\begin{frame}[label={sec:org93082a0}]{Physical Mediator: Heating Difference}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org85cbf31}]{Conclusion and Discussion}
+\begin{frame}[label={sec:orgff36aaf}]{Conclusion and Discussion}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgbda44eb}]{Applications, Next Steps, Limitations}
+\begin{frame}[label={sec:org6ab9458}]{Applications, Next Steps, Limitations}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org1c1bea6}]{Acknowledgements}
+\begin{frame}[label={sec:orgccfbd1e}]{Acknowledgements}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org1cb487a}]{References}
+\begin{frame}[label={sec:org45ad4cf}]{References}
 \bibliographystyle{apa}

### 2021-06-02 presentations/presentation_current/presentation.tex
reloaded pres

@@
-% Created 2021-06-01 Tue 09:02
+% Created 2021-06-02 Wed 08:17
 % Intended LaTeX compiler: pdflatex
@@
 
-\begin{frame}[label={sec:org94c92b5}]{What is El Niño/Southern Oscillation (ENSO)?}
+\begin{frame}[label={sec:org3a2b290}]{What is El Niño/Southern Oscillation (ENSO)?}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orga6e57de}]{Long-Term vs Short-Term Change}
+\begin{frame}[label={sec:org82e6eda}]{Long-Term vs Short-Term Change}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org98f4b82}]{Review of literature}
+\begin{frame}[label={sec:orgc6e3570}]{Review of literature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgeb64aee}]{Research Goals}
+\begin{frame}[label={sec:org514c83e}]{Research Goals}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgced4cac}]{Data: the CESM1 Large Ensemble}
+\begin{frame}[label={sec:orge5e22ee}]{Data: the CESM1 Large Ensemble}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgc2d187b}]{Niño 3.4 Variance}
+\begin{frame}[label={sec:orgc40518a}]{Niño 3.4 Variance}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org2b84dfc}]{Butterfly Effect: The Need for a Large Ensemble}
+\begin{frame}[label={sec:orgae620b1}]{Butterfly Effect: The Need for a Large Ensemble}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org82b4406}]{Model Predictions: ENSO in the Future}
+\begin{frame}[label={sec:orgd198f85}]{Model Predictions: ENSO in the Future}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgc997c2e}]{Analysis of Individual Factors}
+\begin{frame}[label={sec:org81c8216}]{Analysis of Individual Factors}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org948ad63}]{Role of Greenhouse and Aerosol Emissions}
+\begin{frame}[label={sec:orgc1945dc}]{Role of Greenhouse and Aerosol Emissions}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orga1b0a8e}]{Correlation With Changes in Ocean Temperature}
+\begin{frame}[label={sec:orgeb8bfbe}]{Correlation With Changes in Ocean Temperature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org93082a0}]{Physical Mediator: Heating Difference}
+\begin{frame}[label={sec:org80b8911}]{Physical Mediator: Heating Difference}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgff36aaf}]{Conclusion and Discussion}
+\begin{frame}[label={sec:orgf45e4aa}]{Conclusion and Discussion}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org6ab9458}]{Applications, Next Steps, Limitations}
+\begin{frame}[label={sec:orgb0efbd3}]{Applications, Next Steps, Limitations}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgccfbd1e}]{Acknowledgements}
+\begin{frame}[label={sec:org30b35b8}]{Acknowledgements}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org45ad4cf}]{References}
+\begin{frame}[label={sec:org37198dc}]{References}
 \bibliographystyle{apa}

### 2021-06-02 presentations/presentation_current/presentation.tex
deleted unnessecary images

@@
-% Created 2021-06-02 Wed 08:17
+% Created 2021-06-02 Wed 08:23
 % Intended LaTeX compiler: pdflatex
@@
 
-\begin{frame}[label={sec:org3a2b290}]{What is El Niño/Southern Oscillation (ENSO)?}
+\begin{frame}[label={sec:org6324502}]{What is El Niño/Southern Oscillation (ENSO)?}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org82e6eda}]{Long-Term vs Short-Term Change}
+\begin{frame}[label={sec:orga9c1cb1}]{Long-Term vs Short-Term Change}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgc6e3570}]{Review of literature}
+\begin{frame}[label={sec:orge51aa65}]{Review of literature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org514c83e}]{Research Goals}
+\begin{frame}[label={sec:org0833bc2}]{Research Goals}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orge5e22ee}]{Data: the CESM1 Large Ensemble}
+\begin{frame}[label={sec:orgb114c8a}]{Data: the CESM1 Large Ensemble}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgc40518a}]{Niño 3.4 Variance}
+\begin{frame}[label={sec:orgb922573}]{Niño 3.4 Variance}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgae620b1}]{Butterfly Effect: The Need for a Large Ensemble}
+\begin{frame}[label={sec:orgfda0878}]{Butterfly Effect: The Need for a Large Ensemble}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgd198f85}]{Model Predictions: ENSO in the Future}
+\begin{frame}[label={sec:org54ceb45}]{Model Predictions: ENSO in the Future}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org81c8216}]{Analysis of Individual Factors}
+\begin{frame}[label={sec:org09b2b52}]{Analysis of Individual Factors}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgc1945dc}]{Role of Greenhouse and Aerosol Emissions}
+\begin{frame}[label={sec:org64fbfaf}]{Role of Greenhouse and Aerosol Emissions}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgeb8bfbe}]{Correlation With Changes in Ocean Temperature}
+\begin{frame}[label={sec:orgb590ee5}]{Correlation With Changes in Ocean Temperature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org80b8911}]{Physical Mediator: Heating Difference}
+\begin{frame}[label={sec:org420b553}]{Physical Mediator: Heating Difference}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgf45e4aa}]{Conclusion and Discussion}
+\begin{frame}[label={sec:orgb66a216}]{Conclusion and Discussion}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgb0efbd3}]{Applications, Next Steps, Limitations}
+\begin{frame}[label={sec:org527ca8d}]{Applications, Next Steps, Limitations}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org30b35b8}]{Acknowledgements}
+\begin{frame}[label={sec:orgdaf6ace}]{Acknowledgements}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org37198dc}]{References}
+\begin{frame}[label={sec:orgd4caeb0}]{References}
 \bibliographystyle{apa}

### 2021-06-07 paperwork/paperAndRefsList/papernew.tex
reorganized

@@
-% Created 2021-05-21 Fri 09:44
+% Created 2021-06-03 Thu 13:31
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:org7fb1b29}
+\label{sec:orgff39771}
 El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
@@
 \section{Data}
-\label{sec:org91f7f64}
+\label{sec:org7e5e5d2}
 The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
@@
 \section{Methods and Results}
-\label{sec:orgd9fdb5e}
+\label{sec:org8ed47a9}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:orge2424fe}
+\label{sec:org61ed09e}
 We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org8837c57}
+\label{sec:org5cc43db}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:org9556be7}
+\label{sec:orge288389}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 \section{Conclusions and Discussion}
-\label{sec:org1483582}
+\label{sec:org33563ec}
 
@@
 \section{Acknowledgements}
-\label{sec:org9a83329}
+\label{sec:org493e249}
 This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.

### 2021-06-07 presentations/presentation_current/board_test.tex
reorganized

--- 2021-06-02 presentations/presentation_current/presentation.tex
@@
-% Created 2021-06-02 Wed 08:23
+% Created 2021-06-03 Thu 13:30
 % Intended LaTeX compiler: pdflatex
@@
 \usepackage{natbib}
+\usepackage[orientation=landscape,size=b1,scale=1.0, debug]{beamerposter}
 \renewcommand{\bibsection}{}
-\usetheme{metropolis}
+\setbeamertemplate{caption}[numbered]
+\usepackage[margin=1in]{geometry}
+\usetheme{Madrid}
 \author{Benjamin Goldman}
@@
 
-\maketitle
-
-\begin{frame}[label={sec:org6324502}]{What is El Niño/Southern Oscillation (ENSO)?}
+\begin{frame}[label={sec:orgb7a0c80}]{Board}
 \begin{columns}
-\begin{column}{0.6\columnwidth}
+\begin{column}{0.01\columnwidth}
+\end{column}
+\begin{column}{0.24\columnwidth}
+\begin{block}{Introduction}
+\begin{block}{What is El Niño/Southern Oscillation (ENSO)?}
 \begin{itemize}
@@
 \end{itemize}
-\end{column}
 
-\begin{column}{0.4\columnwidth}
 \begin{figure}[htbp]
 \centering
-\includegraphics[width=.9\linewidth]{./images/el_nino.jpg}
+\includegraphics[width=.9\linewidth]{./images/el_nino_reduced.jpg}
 \caption{Changes to tropical Pacific climate during El Niño. \url{https://www.esa.int/ESA\_Multimedia/Images/2018/08/El\_Nino}}
 \end{figure}
-\end{column}
-\end{columns}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:orga9c1cb1}]{Long-Term vs Short-Term Change}
-\begin{columns}
-\begin{column}{0.4\columnwidth}
+\begin{block}{Long-Term vs Short-Term Chang}
 \begin{itemize}
@@
 \end{itemize}
-\end{column}
 
-\begin{column}{0.6\columnwidth}
 \begin{figure}[htbp]
@@
 \end{figure}
-\end{column}
-\end{columns}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:orge51aa65}]{Review of literature}
+\begin{block}{Review of literature}
 \begin{itemize}
@@
 \end{itemize}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:org0833bc2}]{Research Goals}
+\begin{block}{Research Goals}
 \begin{itemize}
@@
 \end{itemize}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:orgb114c8a}]{Data: the CESM1 Large Ensemble}
+\begin{block}{Data: the CESM1 Large Ensemble}
 \begin{itemize}
@@
 \end{itemize}
-\end{frame}
+\end{block}
+\end{block}
+\end{column}
 
-\begin{frame}[label={sec:orgb922573}]{Niño 3.4 Variance}
+\begin{column}{0.01\columnwidth}
+\end{column}
+\begin{column}{0.48\columnwidth}
+\begin{block}{Methods}
 \begin{columns}
-\begin{column}{0.6\columnwidth}
+\begin{column}{0.48\columnwidth}
+\begin{block}{Niño 3.4 Variance}
 \begin{itemize}
 \item How to calculate ENSO intensity in the model output?
-\item Step 1: Calculate sea temperature in Niño 3.4 region of tropical Pacific Ocean.
+\item Step 1: Calculate sea temperature in Niño 3.4 region (5°N - 5°S, 120°-170°W) of tropical Pacific Ocean.
 \item Step 2: Convert temperature dataset to dataset representing change in temperature variation over time.
@@
 \end{itemize}
-\end{column}
+\end{block}
 
-\begin{column}{0.4\columnwidth}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./images/nino3_4.pdf}
-\caption{Niño 3.4 region is the shaded box.}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}[label={sec:orgfda0878}]{Butterfly Effect: The Need for a Large Ensemble}
-\begin{columns}
-\begin{column}{0.5\columnwidth}
+\begin{block}{Butterfly Effect: The Need for a Large Ensemble}
 \begin{itemize}
@@
 \end{itemize}
-\end{column}
 
-\begin{column}{0.5\columnwidth}
 \begin{figure}[htbp]
@@
 \end{figure}
-\end{column}
-\end{columns}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:org54ceb45}]{Model Predictions: ENSO in the Future}
-\begin{columns}
-\begin{column}{.5\columnwidth}
+\begin{block}{Model Predictions: ENSO in the Future}
 \begin{itemize}
@@
 \end{figure}
+\end{block}
 \end{column}
-\end{columns}
-\end{frame}
 
-\begin{frame}[label={sec:org09b2b52}]{Analysis of Individual Factors}
+\begin{column}{0.01\columnwidth}
+\end{column}
+\begin{column}{0.48\columnwidth}
+\begin{block}{Analysis of Individual Factors}
 \begin{itemize}
@@
 \end{itemize}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:org64fbfaf}]{Role of Greenhouse and Aerosol Emissions}
-\begin{columns}
-\begin{column}{0.5\columnwidth}
+\begin{block}{Role of Greenhouse and Aerosol Emissions}
 \begin{itemize}
@@
 \end{figure}
-\end{column}
-\end{columns}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:orgb590ee5}]{Correlation With Changes in Ocean Temperature}
+\begin{block}{Correlation With Changes in Ocean Temperature}
 \begin{itemize}
@@
 \end{itemize}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:org420b553}]{Physical Mediator: Heating Difference}
-\begin{columns}
-\begin{column}{.62\columnwidth}
+\begin{block}{Physical Mediator: Heating Difference}
 \begin{itemize}
@@
 \end{itemize}
-\end{column}
 
-\begin{column}{.50\columnwidth}
 \begin{figure}[htbp]
@@
 \end{figure}
+\end{block}
 \end{column}
 \end{columns}
-\end{frame}
+\end{block}
+\end{column}
 
-\begin{frame}[label={sec:orgb66a216}]{Conclusion and Discussion}
+\begin{column}{0.01\columnwidth}
+\end{column}
+\begin{column}{0.24\columnwidth}
+\begin{block}{Conclusion and Discussion}
 \begin{itemize}
@@
 \end{itemize}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:org527ca8d}]{Applications, Next Steps, Limitations}
+\begin{block}{Applications, Next Steps, Limitations}
 \begin{itemize}
@@
 \end{itemize}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:orgdaf6ace}]{Acknowledgements}
+\begin{block}{Acknowledgements}
 \begin{itemize}
@@
 \end{itemize}
-\end{frame}
+\end{block}
 
-\begin{frame}[label={sec:orgd4caeb0}]{References}
+\begin{block}{References}
 \bibliographystyle{apa}
@@
 \bibliography{references}
+\end{block}
+\end{column}
+
+\begin{column}{0.01\columnwidth}
+\end{column}
+\end{columns}
 \end{frame}

### 2021-06-10 presentations/presentation_current/presentation.tex
symposium and hours

@@
-% Created 2021-06-02 Wed 08:23
+% Created 2021-06-09 Wed 14:40
 % Intended LaTeX compiler: pdflatex
@@
 
-\begin{frame}[label={sec:org6324502}]{What is El Niño/Southern Oscillation (ENSO)?}
+\begin{frame}[label={sec:org6e6eccf}]{What is El Niño/Southern Oscillation (ENSO)?}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orga9c1cb1}]{Long-Term vs Short-Term Change}
+\begin{frame}[label={sec:org6b9d33a}]{Long-Term vs Short-Term Change}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orge51aa65}]{Review of literature}
+\begin{frame}[label={sec:org24e313e}]{Review of literature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org0833bc2}]{Research Goals}
+\begin{frame}[label={sec:org77e0249}]{Research Goals}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgb114c8a}]{Data: the CESM1 Large Ensemble}
+\begin{frame}[label={sec:orgd6cf415}]{Data: the CESM1 Large Ensemble}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgb922573}]{Niño 3.4 Variance}
+\begin{frame}[label={sec:org2cd69f1}]{Niño 3.4 Variance}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgfda0878}]{Butterfly Effect: The Need for a Large Ensemble}
+\begin{frame}[label={sec:org8562e6e}]{Butterfly Effect: The Need for a Large Ensemble}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org54ceb45}]{Model Predictions: ENSO in the Future}
+\begin{frame}[label={sec:orgfda20ee}]{Model Predictions: ENSO in the Future}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org09b2b52}]{Analysis of Individual Factors}
+\begin{frame}[label={sec:org5795ab7}]{Analysis of Individual Factors}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org64fbfaf}]{Role of Greenhouse and Aerosol Emissions}
+\begin{frame}[label={sec:orgd092c4d}]{Role of Greenhouse and Aerosol Emissions}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgb590ee5}]{Correlation With Changes in Ocean Temperature}
+\begin{frame}[label={sec:org4409680}]{Correlation With Changes in Ocean Temperature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org420b553}]{Physical Mediator: Heating Difference}
+\begin{frame}[label={sec:org52a5575}]{Physical Mediator: Heating Difference}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgb66a216}]{Conclusion and Discussion}
+\begin{frame}[label={sec:org3c6bb69}]{Conclusion and Discussion}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org527ca8d}]{Applications, Next Steps, Limitations}
+\begin{frame}[label={sec:orgdf5ae8f}]{Applications, Next Steps, Limitations}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgdaf6ace}]{Acknowledgements}
+\begin{frame}[label={sec:org8ed07de}]{Acknowledgements}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgd4caeb0}]{References}
+\begin{frame}[label={sec:orgc3b51d1}]{References}
 \bibliographystyle{apa}

### 2021-06-16 paperwork/paper_archive/paper.tex
latex mixup?

--- 2021-06-07 paperwork/paperAndRefsList/papernew.tex
@@
-% Created 2021-06-03 Thu 13:31
-% Intended LaTeX compiler: pdflatex
-\documentclass[11pt]{article}
-\usepackage[utf8]{inputenc}
+\documentclass[12pt]{article}
 \usepackage[T1]{fontenc}
 \usepackage{graphicx}
-\usepackage{grffile}
-\usepackage{longtable}
-\usepackage{wrapfig}
-\usepackage{rotating}
-\usepackage[normalem]{ulem}
-\usepackage{amsmath}
-\usepackage{textcomp}
 \usepackage{amssymb}
-\usepackage{capt-of}
 \usepackage{hyperref}
-\usepackage{natbib}
-\usepackage[letterpaper, margins=1in]{geometry}
+\usepackage[
+backend=biber,
+style=apa,
+sorting=ynt
+]{biblatex}
+\addbibresource{references.bib}
+\usepackage[letterpaper, margin=1in]{geometry}
 \setlength{\parskip}{0.35em}
@@
 \title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
-\hypersetup{
- pdfauthor={Benjamin Goldman},
- pdftitle={The Impact of Anthropogenic Forcing on ENSO Amplitude},
- pdfkeywords={},
- pdfsubject={},
- pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
- pdflang={English}}
+
 \begin{document}
@@
 \section{Introduction}
-\label{sec:orgff39771}
-El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
-The reasons for studying ENSO are clear, as ENSO has drastically affected climate patterns worldwide, modulating rainfall and temperature in nearly every continent \citep{ropelewski1987global}.
-For example, the recent 2015-2016 El Niño event contributed to record-breaking high temperatures and droughts in South America \citep{jimenez2016record}.
+El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \parencite{bjerknes1969atmospheric}.
+The reasons for studying ENSO are clear, as ENSO has drastically affected climate patterns worldwide, modulating rainfall and temperature in nearly every continent \parencite{ropelewski1987global}.
+For example, the recent 2015-2016 El Niño event contributed to record-breaking high temperatures and droughts in South America \parencite{jimenez2016record}.
 At the same time, long-term anthropogenic greenhouse gas emissions are causing global temperatures to increase through a greenhouse effect.
@@
 Research on the effect of external forcing on ENSO remains inconclusive, as results from similar studies conflict.
-\citet{nowack2017role} predicted an overall increase in Niño 3.4 standard deviation under a combination of 4xCO\(_2\) and interactive ozone forcing using single-model simulations, showing that greenhouse gasses increase the frequency of extreme ENSO by favoring a more Niño-like in the tropical Pacific, while ozone dampens this effect.
-In contrast, a few studies have found that ENSO amplitude decreases under global warming in certain coupled models \citep{kohyama2018weakening}.
+\cite{nowack2017role} predicted an overall increase in Niño 3.4 standard deviation under a combination of 4xCO$_2$ and interactive ozone forcing using single-model simulations, showing that greenhouse gasses increase the frequency of extreme ENSO by favoring a more Niño-like in the tropical Pacific, while ozone dampens this effect.
+In contrast, a few studies have found that ENSO amplitude decreases under global warming in certain coupled models \parencite{kohyama2018weakening}.
 
-However, other studies have failed to find any statistically significant ENSO response to external forcing \citep{stevenson2012significant}.
-Analysis using NCAR's CESM Large Ensemble shows an ensemble size of at least 15 models is necessary to attribute changes to Niño variance to external forcing and reject the null hypothesis that internal variability is responsible for changes to ENSO \citep{zheng2018response}.
-Several modes of internal variability have been shown to modulate ENSO, including the AMO \citep{levine2017impact} and Tropical Pacific Decadal Variability (TPDV) \citep{zheng2018response}.
-An analysis of the Max Plank Institute's Grand Ensemble as well as NCAR's CESM Large Ensemble suggests that 80$\backslash$% of changes to ENSO amplitude can be attributed to internal variability, but given a large enough ensemble, significant changes in ENSO amplitude due to climate change can be detected \citep{maher2018enso}.
+However, other studies have failed to find any statistically significant ENSO response to external forcing \parencite{stevenson2012significant}.
+Analysis using NCAR's CESM Large Ensemble shows an ensemble size of at least 15 models is necessary to attribute changes to Niño variance to external forcing and reject the null hypothesis that internal variability is responsible for changes to ENSO \parencite{zheng2018response}.
+Several modes of internal variability have been shown to modulate ENSO, including the AMO \parencite{levine2017impact} and Tropical Pacific Decadal Variability (TPDV) \parencite{zheng2018response}.
+An analysis of the Max Plank Institute's Grand Ensemble as well as NCAR's CESM Large Ensemble suggests that 80\% of changes to ENSO amplitude can be attributed to internal variability, but given a large enough ensemble, significant changes in ENSO amplitude due to climate change can be detected \parencite{maher2018enso}.
 
@@
 \section{Data}
-\label{sec:org7e5e5d2}
 The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
-The Large Ensemble contains 40 simulations of the CESM1 coupled model, forced with historic radiative forcing from 1920 to 2005 and according to the RCP 8.5 protocol from 2006 to 2100 \citep{kay2015community}.
+The Large Ensemble contains 40 simulations of the CESM1 coupled model, forced with historic radiative forcing from 1920 to 2005 and according to the RCP 8.5 protocol from 2006 to 2100 \parencite{kay2015community}.
 The single forcing ensemble is a collection of sub-ensembles for various climate factors (greenhouse gas, aerosols, biomass burning, land use, ozone).
-Each simulation is forced by a combination of all factors except for one, allowing the impact of a single factor to be deduced by subtracting the ensemble mean from the fully-forced ensemble mean \citep{deser2020isolating}\}.
+Each simulation is forced by a combination of all factors except for one, allowing the impact of a single factor to be deduced by subtracting the ensemble mean from the fully-forced ensemble mean \parencite{deser2020isolating}\}.
 For example, the xghg (greenhouse gas) ensemble is forced by aerosol emissions, biomass burning, land use, and ozone.
@@
 \section{Methods and Results}
-\label{sec:org8ed47a9}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org61ed09e}
 We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 Although the excluded land use ensemble mean has a strong increase in variance, this result is unlikely to be meaningful due to the xlulc ensemble's low sample size.
-Additionally, all ensembles exhibit considerable noise, in line with previous studies \citep{maher2018enso}.
+Additionally, all ensembles exhibit considerable noise, in line with previous studies \parencite{maher2018enso}.
 
@@
 To test the likelihood of this explanation, we test the influence of the Atlantic Multidecadal Oscillation (AMO) and Atlantic Meridional Overturning Current (AMOC) on ENSO in the control simulation.
-The AMO has been shown to have an influence on ENSO strength and seasonal growth rate \citep{levine2017impact}.
-We filter the control Niño 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \citep{phillips2014evaluating}\}.
+The AMO has been shown to have an influence on ENSO strength and seasonal growth rate \parencite{levine2017impact}.
+We filter the control Niño 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \parencite{phillips2014evaluating}\}.
 The Probability Distribution Function of Niño 3.4 variance is estimated for AMO/C > 1/2\(\sigma\) and < -1/-2\(\sigma\) using a Kernel Density Estimation.
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org5cc43db}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 The greenhouse-only ensemble as well as the aerosol-only ensemble exhibit increased variance, signaling that greenhouse and aerosol emissions likely both play a significant role in ENSO's forced response in the full-forcing ensemble.
-Interestingly, the influence of greenhouse gasses and aerosols are non-conflicting, in contrast with previous studies that show opposite effects of greenhouse gas and aerosol forcing on ENSO \citep{stevenson2019forced}.
+Interestingly, the influence of greenhouse gasses and aerosols are non-conflicting, in contrast with previous studies that show opposite effects of greenhouse gas and aerosol forcing on ENSO \parencite{stevenson2019forced}.
 All the other single forcing ensembles exhibit insignificant differences from the fully-forced ensemble.
 The biomass burning case shows very small deviations from the fully-forced case, while the ozone ensemble's period of recording is too small to draw meaningful conclusions.
-However, \citet{nowack2017role} showed that ozone forcing may dampen the effect of greenhouse-forced increases to ENSO amplitude by reducing changes to Pacific sea temperature and the Walker Circulation.
+However, \cite{nowack2017role} showed that ozone forcing may dampen the effect of greenhouse-forced increases to ENSO amplitude by reducing changes to Pacific sea temperature and the Walker Circulation.
 The land use/cover case, while it does show large deviations from the fully-forced case, has an ensemble size (5 members) too small to lend any credibility to these changes.
@@
 We hypothesize that this discrepancy may be the result of anomalous initial conditions caused by internal variability of the control, as the control conditions are used to initialize all the forced runs.
-To test this hypothesis, we analyzed the impact of the Atlantic Multidecadal Oscillation (AMO) and the Atlantic Meridional Overturning Current (AMOC) on Niño 3.4 variance in the control simulation using records of the AMO and AMOC from the Climate Variability Diagnostics Package (CVDP) \citep{phillips2014evaluating}.
+To test this hypothesis, we analyzed the impact of the Atlantic Multidecadal Oscillation (AMO) and the Atlantic Meridional Overturning Current (AMOC) on Niño 3.4 variance in the control simulation using records of the AMO and AMOC from the Climate Variability Diagnostics Package (CVDP) \parencite{phillips2014evaluating}.
 We filtered the 20-year variance of the Niño 3.4 sea surface temperature in the control based on the strength of the AMO/AMOC, separating ENSO variance into groups where AMO < -1\(\sigma\), AMO < -2\(\sigma\), AMO > 1\(\sigma\), AMO >2\(\sigma\), and the same for AMOC.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:orge288389}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 \includegraphics[width=.9\linewidth]{./figures/tempdt.pdf}
-\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Niño 3.4 variance.
+\caption{\label{fig:tempdt_1}Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Niño 3.4 variance.
 Grey color indicates points where p<0.05.
@@
 \includegraphics[width=.9\linewidth]{./figures/tempcep.pdf}
-\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.
+\caption{\label{fig:tempcep_1}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.
 Grey color indicates points where p>0.05.}
@@
 \section{Conclusions and Discussion}
-\label{sec:org33563ec}
 
@@
 The fully-forced large ensemble exhibits increased ENSO amplitude, as measured by calculating the 20-year variance of the Niño 3.4 index.
-These changes appear to be statistically significant, but need to be scrutinized further, as they are in disagreement with results of some previous studies \citep{stevenson2012significant}.
+These changes appear to be statistically significant, but need to be scrutinized further, as they are in disagreement with results of some previous studies \parencite{stevenson2012significant}.
 Additionally, the fully-forced ensemble, as well as all the others contains considerable noise, as the individual members cover a wide range of variability.
@@
 It is still unclear what is driving this connection, as well as how external forcing such as global warming affects it.
-There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \citep{dewitte2013reinterpreting}.
+There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \parencite{dewitte2013reinterpreting}.
 
-This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \citep{maher2018enso}.
-Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citet{stevenson2019forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
+This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \parencite{maher2018enso}.
+Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \cite{stevenson2019forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
 In contrast, this study observes that the impact on ENSO amplitude of greenhouse gas and aerosols has the same sign.
-This result is also surprising given that greenhouse and aerosol emissions have been shown to have opposite effects on sea surface temperature and global circulation, with greenhouse emissions favoring a more Niño-like state with weakened Walker circulation and higher SST, and aerosol emissions counteracting these effects \citep{boer2000transient}.
+This result is also surprising given that greenhouse and aerosol emissions have been shown to have opposite effects on sea surface temperature and global circulation, with greenhouse emissions favoring a more Niño-like state with weakened Walker circulation and higher SST, and aerosol emissions counteracting these effects \parencite{boer2000transient}.
 
@@
 \section{Acknowledgements}
-\label{sec:org493e249}
 This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
@@
 
-\bibliographystyle{apa}
-\bibliography{references}
+\printbibliography
+
 \end{document}

### 2021-06-16 paperwork/papernew/papernew.tex
latex mixup?

--- 2021-06-07 paperwork/paperAndRefsList/papernew.tex
@@
-% Created 2021-06-03 Thu 13:31
+% Created 2021-06-16 Wed 08:13
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:orgff39771}
+\label{sec:org14be98b}
 El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
@@
 \section{Data}
-\label{sec:org7e5e5d2}
+\label{sec:org15697ab}
 The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
@@
 \section{Methods and Results}
-\label{sec:org8ed47a9}
+\label{sec:org469de48}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org61ed09e}
+\label{sec:org56d15ab}
 We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org5cc43db}
+\label{sec:org582da32}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:orge288389}
+\label{sec:orga46145a}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 \section{Conclusions and Discussion}
-\label{sec:org33563ec}
+\label{sec:org787504a}
 
@@
 \section{Acknowledgements}
-\label{sec:org493e249}
+\label{sec:orgaff78f0}
 This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.

### 2021-06-16 presentations/presentation_current/presentation.tex
latex mixup?

@@
-% Created 2021-06-09 Wed 14:40
+% Created 2021-06-16 Wed 08:12
 % Intended LaTeX compiler: pdflatex
@@
 
-\begin{frame}[label={sec:org6e6eccf}]{What is El Niño/Southern Oscillation (ENSO)?}
+\begin{frame}[label={sec:orgb1df2b4}]{What is El Niño/Southern Oscillation (ENSO)?}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org6b9d33a}]{Long-Term vs Short-Term Change}
+\begin{frame}[label={sec:orga0cf916}]{Long-Term vs Short-Term Change}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org24e313e}]{Review of literature}
+\begin{frame}[label={sec:org8f2e28a}]{Review of literature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org77e0249}]{Research Goals}
+\begin{frame}[label={sec:org26b0634}]{Research Goals}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgd6cf415}]{Data: the CESM1 Large Ensemble}
+\begin{frame}[label={sec:org389fba7}]{Data: the CESM1 Large Ensemble}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org2cd69f1}]{Niño 3.4 Variance}
+\begin{frame}[label={sec:org80dd814}]{Niño 3.4 Variance}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org8562e6e}]{Butterfly Effect: The Need for a Large Ensemble}
+\begin{frame}[label={sec:org8592647}]{Butterfly Effect: The Need for a Large Ensemble}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgfda20ee}]{Model Predictions: ENSO in the Future}
+\begin{frame}[label={sec:orgf5c3604}]{Model Predictions: ENSO in the Future}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org5795ab7}]{Analysis of Individual Factors}
+\begin{frame}[label={sec:org9a47952}]{Analysis of Individual Factors}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgd092c4d}]{Role of Greenhouse and Aerosol Emissions}
+\begin{frame}[label={sec:org0a966de}]{Role of Greenhouse and Aerosol Emissions}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org4409680}]{Correlation With Changes in Ocean Temperature}
+\begin{frame}[label={sec:org288c45c}]{Correlation With Changes in Ocean Temperature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org52a5575}]{Physical Mediator: Heating Difference}
+\begin{frame}[label={sec:org8bce193}]{Physical Mediator: Heating Difference}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org3c6bb69}]{Conclusion and Discussion}
+\begin{frame}[label={sec:orge890f4d}]{Conclusion and Discussion}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgdf5ae8f}]{Applications, Next Steps, Limitations}
+\begin{frame}[label={sec:orge9c36c0}]{Applications, Next Steps, Limitations}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org8ed07de}]{Acknowledgements}
+\begin{frame}[label={sec:org4cda3b8}]{Acknowledgements}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgc3b51d1}]{References}
+\begin{frame}[label={sec:org3d198c1}]{References}
 \bibliographystyle{apa}

### 2021-06-20 paperwork/paperAndRefsList/papernew.tex
ran paper sts stuff

@@
-% Created 2021-06-03 Thu 13:31
-% Intended LaTeX compiler: pdflatex
+% Created 2021-06-18 Fri 09:09
+% Intended LaTeX compiler: xelatex
 \documentclass[11pt]{article}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
 \usepackage{graphicx}
@@
 \section{Introduction}
-\label{sec:orgff39771}
+\label{sec:org644a82a}
 El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
@@
 \section{Data}
-\label{sec:org7e5e5d2}
+\label{sec:org03d34a1}
 The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
@@
 \section{Methods and Results}
-\label{sec:org8ed47a9}
+\label{sec:orgd811f4f}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org61ed09e}
+\label{sec:org61a5214}
 We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org5cc43db}
+\label{sec:orga982838}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:orge288389}
+\label{sec:org546253d}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 \section{Conclusions and Discussion}
-\label{sec:org33563ec}
+\label{sec:org66b548e}
 
@@
 \section{Acknowledgements}
-\label{sec:org493e249}
+\label{sec:orgc00f71c}
 This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
@@
 
-\bibliographystyle{apa}
+\bibliographystyle{apalike}
 \bibliography{references}

### 2021-06-20 presentations/presentation_current/presentation.tex
ran paper sts stuff

@@
-% Created 2021-06-16 Wed 08:12
+% Created 2021-06-18 Fri 09:11
 % Intended LaTeX compiler: pdflatex
@@
 \renewcommand{\bibsection}{}
+\DeclareUnicodeCharacter{0303}{\~{n}}
 \usetheme{metropolis}
@@
 
-\begin{frame}[label={sec:orgb1df2b4}]{What is El Niño/Southern Oscillation (ENSO)?}
+\begin{frame}[label={sec:org8153378}]{What is El Niño/Southern Oscillation (ENSO)?}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orga0cf916}]{Long-Term vs Short-Term Change}
+\begin{frame}[label={sec:org30dba5e}]{Long-Term vs Short-Term Change}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org8f2e28a}]{Review of literature}
+\begin{frame}[label={sec:orge54f468}]{Review of literature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org26b0634}]{Research Goals}
+\begin{frame}[label={sec:org85f5f49}]{Research Goals}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org389fba7}]{Data: the CESM1 Large Ensemble}
+\begin{frame}[label={sec:org6749e99}]{Data: the CESM1 Large Ensemble}
 \begin{itemize}
-\item Explore hypothetical scenarios with a computer model \citet{kay2015community}.
+\item Explore hypothetical scenarios with a computer model \citep{kay2015community}.
 \item Estimation of how the earth's climate actually works.
@@
 
-\begin{frame}[label={sec:org80dd814}]{Niño 3.4 Variance}
+\begin{frame}[label={sec:org650aae3}]{Niño 3.4 Variance}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org8592647}]{Butterfly Effect: The Need for a Large Ensemble}
+\begin{frame}[label={sec:orgdfcbcd5}]{Butterfly Effect: The Need for a Large Ensemble}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgf5c3604}]{Model Predictions: ENSO in the Future}
+\begin{frame}[label={sec:org7d549e7}]{Model Predictions: ENSO in the Future}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org9a47952}]{Analysis of Individual Factors}
+\begin{frame}[label={sec:org3a535ea}]{Analysis of Individual Factors}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org0a966de}]{Role of Greenhouse and Aerosol Emissions}
+\begin{frame}[label={sec:org87d91c6}]{Role of Greenhouse and Aerosol Emissions}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org288c45c}]{Correlation With Changes in Ocean Temperature}
+\begin{frame}[label={sec:org1369c26}]{Correlation With Changes in Ocean Temperature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org8bce193}]{Physical Mediator: Heating Difference}
+\begin{frame}[label={sec:org10d0dfb}]{Physical Mediator: Heating Difference}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orge890f4d}]{Conclusion and Discussion}
+\begin{frame}[label={sec:orga972c2c}]{Conclusion and Discussion}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orge9c36c0}]{Applications, Next Steps, Limitations}
+\begin{frame}[label={sec:orgb31e23e}]{Applications, Next Steps, Limitations}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org4cda3b8}]{Acknowledgements}
+\begin{frame}[label={sec:org35dd772}]{Acknowledgements}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org3d198c1}]{References}
-\bibliographystyle{apa}
+\begin{frame}[label={sec:org98d3ee4}]{References}
+\bibliographystyle{apalike}
 \fontsize{7pt}{7.2}\selectfont

### 2021-07-07 data/Code/analyzeVariance.r
r stuff

@@ -1,0 +1,50 @@
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+
+ff <- read.table("ff_var.csv", sep = ",") %>% as.matrix()
+ghg <- read.table("ghg_var.csv", sep = ",") %>% as.matrix()
+aer <- read.table("aer_var.csv", sep = ",") %>% as.matrix()
+bmb <- read.table("bmb_var.csv", sep = ",") %>% as.matrix()
+luc <- read.table("luc_var.csv", sep = ",") %>% as.matrix()
+
+ff_mean <- apply(X = ff, MARGIN = 1, FUN = mean)
+ghg_mean <- apply(X = ghg, MARGIN = 1, FUN = mean)
+aer_mean <- apply(X = aer, MARGIN = 1, FUN = mean)
+bmb_mean <- apply(X = bmb, MARGIN = 1, FUN = mean)
+luc_mean <- apply(X = luc, MARGIN = 1, FUN = mean)
+
+se <- function(x, ...) sd(x, ...)/sqrt(length(x))
+
+ff_se <- apply(X = ff, MARGIN = 1, FUN = se)
+ghg_se <- apply(X = ghg, MARGIN = 1, FUN = se)
+aer_se <- apply(X = aer, MARGIN = 1, FUN = se)
+bmb_se <- apply(X = bmb, MARGIN = 1, FUN = se)
+luc_se <- apply(X = luc, MARGIN = 1, FUN = se)
+
+mydate <- read.table("mydate.csv", sep = ",") %>% as.vector()
+
+ff_df <- data.frame(time = mydate, mean = ff_mean, sterr = ff_se)
+ghg_df <- data.frame(time = mydate, mean = ghg_mean, sterr = ghg_se)
+aer_df <- data.frame(time = mydate, mean = aer_mean, sterr = aer_se)
+bmb_df <- data.frame(time = mydate, mean = bmb_mean, sterr = bmb_se)
+luc_df <- data.frame(time = mydate, mean = luc_mean, sterr = luc_se)
+
+
+nino <- ff_df %>% mutate(case = factor("Full Forcing")) %>%
+  bind_rows(ghg_df %>%
+              mutate(case = factor("Greenhouse"))) %>%
+  bind_rows(aer_df %>%
+              mutate(case = factor("Aerosols"))) %>%
+  bind_rows(bmb_df %>%
+              mutate(case = factor("Biomass")))  %>%
[…]

### 2021-07-07 data/Code/extractVariance.r

@@ -1,6 +1,7 @@
 library(ncdf4)
 library(zoo)
 library(dplyr)
+library(ggplot2)
 
 rollvar <- function(x) {
   out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
@@ -34,19 +35,28 @@ files_sf <- c("b.e11.B20TRLENS_RCP85.f09_g16.xghg.all.cam.h0.TREFHT.192001-20801
 
 ff_var <- extract_nino3.4(paste(path_ff, file_ff, sep = ""), "TREFHT") %>%
   apply(2, rollvar)
+write.table(ff_var, file = "ff_var.csv", sep = ",")
 
 ghg_var <- extract_nino3.4(paste(path_sf, files_sf[1], sep = ""), "TREFHT") %>%
     apply(2, rollvar) %>%
     rbind(matrix(data=NA, nrow = (2172-1932), ncol = 20))
+write.table(ghg_var, file = "ghg_var.csv", sep = ",")
 
 aer_var <- extract_nino3.4(paste(path_sf, files_sf[2], sep = ""), "TREFHT") %>%
-  apply(2, rollvar)
-  #rbind(matrix(data = NA, nrow = (2172-1932), ncol = 20))
+  apply(2, rollvar) %>%
+  rbind(matrix(data = NA, nrow = (2172-1932), ncol = 20))
+write.table(aer_var, file = "aer_var.csv", sep = ",")
 
 bmb_var <- extract_nino3.4(paste(path_sf, files_sf[3], sep = ""), "TREFHT") %>%
   apply(2, rollvar) %>%
   rbind(matrix(data = NA, nrow = (2172-1320), ncol = 15))
+write.table(bmb_var, file = "bmb_var.csv", sep = ",")
 
 luc_var <- extract_nino3.4(paste(path_sf, files_sf[4], sep = ""), "TREFHT") %>%
   apply(2, rollvar) %>%
   rbind(matrix(data = NA, nrow = (2172-1320), ncol = 5))
+write.table(luc_var, file = "luc_var.csv", sep = ",")
+
+ncin <- nc_open("/Volumes/Extreme SSD/DATA/stacked/CESM1/TREFHT/b.e11.BRCP85C5CNBDRD.f09_g16.001.cam.h0.TREFHT.192001-210012.nc")
+mydate <- as.Date(as.character(ncvar_get(ncin, "date")), "%Y%m%d")
+write.table(mydate, file = "mydate.csv", sep = ",")

### 2021-07-07 paperwork/paper/papernew.tex
new reference list bib

--- 2021-06-20 paperwork/paperAndRefsList/papernew.tex
@@
-% Created 2021-06-18 Fri 09:09
+% Created 2021-07-07 Wed 18:07
 % Intended LaTeX compiler: xelatex
@@
 \section{Introduction}
-\label{sec:org644a82a}
+\label{sec:orgdd1c479}
 El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
@@
 \section{Data}
-\label{sec:org03d34a1}
+\label{sec:org4518aec}
 The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
@@
 \section{Methods and Results}
-\label{sec:orgd811f4f}
+\label{sec:org3c41840}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org61a5214}
+\label{sec:org20d4dc8}
 We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:orga982838}
+\label{sec:orge5f0b01}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:org546253d}
+\label{sec:org9cb6717}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 \section{Conclusions and Discussion}
-\label{sec:org66b548e}
+\label{sec:org3b1ac08}
 
@@
 \section{Acknowledgements}
-\label{sec:orgc00f71c}
+\label{sec:org8775e78}
 This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.

### 2021-07-07 paperwork/paper/paper.tex
paper folder renamed because why not

--- 2021-07-07 paperwork/paper/papernew.tex
@@
-% Created 2021-07-07 Wed 18:07
+% Created 2021-07-07 Wed 18:17
 % Intended LaTeX compiler: xelatex
@@
 \section{Introduction}
-\label{sec:orgdd1c479}
+\label{sec:orgd83aafa}
 El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
@@
 \section{Data}
-\label{sec:org4518aec}
+\label{sec:org4103a64}
 The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
@@
 \section{Methods and Results}
-\label{sec:org3c41840}
+\label{sec:org458007c}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org20d4dc8}
+\label{sec:org831a275}
 We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:orge5f0b01}
+\label{sec:org4438ebb}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:org9cb6717}
+\label{sec:org6571ab0}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 \section{Conclusions and Discussion}
-\label{sec:org3b1ac08}
+\label{sec:org4393892}
 
@@
 \section{Acknowledgements}
-\label{sec:org8775e78}
+\label{sec:orgc33b858}
 This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.

### 2021-07-08 paperwork/paper/paper.tex
fix some refs

@@
-% Created 2021-07-07 Wed 18:17
+% Created 2021-07-07 Wed 20:51
 % Intended LaTeX compiler: xelatex
@@
 \section{Introduction}
-\label{sec:orgd83aafa}
+\label{sec:org3fc8d9e}
 El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
@@
 However, other studies have failed to find any statistically significant ENSO response to external forcing \citep{stevenson2012significant}.
-Analysis using NCAR's CESM Large Ensemble shows an ensemble size of at least 15 models is necessary to attribute changes to Niño variance to external forcing and reject the null hypothesis that internal variability is responsible for changes to ENSO \citep{zheng2018response}.
-Several modes of internal variability have been shown to modulate ENSO, including the AMO \citep{levine2017impact} and Tropical Pacific Decadal Variability (TPDV) \citep{zheng2018response}.
+Analysis using NCAR's CESM Large Ensemble shows an ensemble size of at least 15 models is necessary to attribute changes to Niño variance to external forcing and reject the null hypothesis that internal variability is responsible for changes to ENSO \citep{zheng2017response}.
+Several modes of internal variability have been shown to modulate ENSO, including the AMO \citep{levine2017impact} and Tropical Pacific Decadal Variability (TPDV) \citep{zheng2017response}.
 An analysis of the Max Plank Institute's Grand Ensemble as well as NCAR's CESM Large Ensemble suggests that 80$\backslash$% of changes to ENSO amplitude can be attributed to internal variability, but given a large enough ensemble, significant changes in ENSO amplitude due to climate change can be detected \citep{maher2018enso}.
@@
 \section{Data}
-\label{sec:org4103a64}
+\label{sec:org0534843}
 The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
@@
 \section{Methods and Results}
-\label{sec:org458007c}
+\label{sec:org6ea3bdd}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org831a275}
+\label{sec:org907e898}
 We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org4438ebb}
+\label{sec:orge41456d}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 The greenhouse-only ensemble as well as the aerosol-only ensemble exhibit increased variance, signaling that greenhouse and aerosol emissions likely both play a significant role in ENSO's forced response in the full-forcing ensemble.
-Interestingly, the influence of greenhouse gasses and aerosols are non-conflicting, in contrast with previous studies that show opposite effects of greenhouse gas and aerosol forcing on ENSO \citep{stevenson2019forced}.
+Interestingly, the influence of greenhouse gasses and aerosols are non-conflicting, in contrast with previous studies that show opposite effects of greenhouse gas and aerosol forcing on ENSO \citep{stevenson2017forced}.
 All the other single forcing ensembles exhibit insignificant differences from the fully-forced ensemble.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:org6571ab0}
+\label{sec:orgabb8356}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 Next, we calculated the Pearson's correlation coefficient between each gridpoint and the Niño 3.4 variance timeseries.
-The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt_1}, and for the central slice in figure \ref{fig:tempcep_1}.
+The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt}, and for the central slice in figure \ref{fig:tempcep}.
 
@@
 \includegraphics[width=.9\linewidth]{./figures/tempcep.pdf}
-\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.
+\caption{\label{fig:tempcep}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.
 Grey color indicates points where p>0.05.}
@@
 \section{Conclusions and Discussion}
-\label{sec:org4393892}
+\label{sec:org811a462}
 
@@
 It is still unclear what is driving this connection, as well as how external forcing such as global warming affects it.
-There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \citep{dewitte2013reinterpreting}.
+There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \citep{dewitte2012reinterpreting}.
 
 This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \citep{maher2018enso}.
-Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citet{stevenson2019forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
+Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citet{stevenson2017forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
 In contrast, this study observes that the impact on ENSO amplitude of greenhouse gas and aerosols has the same sign.
@@
 \section{Acknowledgements}
-\label{sec:orgc33b858}
+\label{sec:org25124e8}
 This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.

### 2021-07-08 articles/notes.tex
notes on all my articles yay very nice

% Created 2021-07-08 Thu 18:46
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{enumitem}
\setlist{noitemsep}
\author{Benjamin Goldman}
\date{\today}
\title{Notes}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Notes},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents

\section{\cite{bjerknes1969atmospheric}}
\label{sec:org2ba3d44}

\begin{itemize}
\item First big paper on ENSO having a big impact
\item connected changes in ocean currents to Walker Circulation
\item ENSO phase affects behavior of the Indian Ocean monsoon.
\end{itemize}

\section{\cite{an2017feedback}}
\label{sec:org301fc0b}

\begin{itemize}
\item Used SVD (Singular Value Decomposition) together with the Mixed Layer Heat Budget Analysis to look at which feedbacks contributed most to ENSO's variation between models
\item Influence of thermocline feedback is determined by how strongly equatorial horizontal winds affect the slope of the thermocline.
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{boer2000transient}}
\label{sec:org0861beb}

\section{\cite{cai2018increased}}
\label{sec:org4e917df}

\begin{itemize}
\item Increased ENSO variance in most CMIP5 models in EP ENSO center.
\item Likely caused by greenhouse gases
\item Higher ocean stratification allows for stronger communication between atmospheric and oceanic temperatures.
\item Used EOF analysis.
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{chen2015causes}}
\label{sec:org1579516}

\section{\cite{chen2017possible}}
\label{sec:org0513968}

\begin{itemize}
\item Models are disagreeing on ENSO in the future because they have different representations of the mechanics and mean state of the Pacific subtropical cell
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{deser2020isolating}}
\label{sec:org1861fc1}

\begin{itemize}
\item Main documentation for CESM1 Single Forcing Ensemble
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{dewitte2012reinterpreting}}
\label{sec:org0dc6c59}

\section{\cite{emile2007nino}}
\label{sec:org66963df}

\begin{itemize}
\item Analyzed wavelet power spectrum of ENSO variability in models forced by sunspot and orbital changes
\item Orbital changes increase long-term ENSO variability
\item It is possible that ENSO was the mechanic that allowed prehistoric solar/orbital changes to control the earth's climate
\end{itemize}

\section{\cite{graham2014effectiveness}}
\label{sec:orgbe0857b}

\begin{itemize}
\item tested how accurate the Bjerknes Stability Index is at measuring the mechanics of ENSO in a couple models
\item BJ index overestimates the importance of the Thermocline feedback.
\item BJ index assumes that terms should be linear when combined, but they actually aren't.
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{torrence1998practical}}
\label{sec:org96cb080}

\begin{itemize}
\item How to use wavelets to estimate power spectrum in timeseries.
\item Uses ENSO data \emph{very niiceee}
\item Windowed Fourier Transform sucks butt because it is dependent on a time step parameter that can muck with the results depending on which value you choose.
\item A wavelet is a short \textbf{\emph{blirp}} of a wave with a mean of zero and finite amplitude/frequency and limited time domain.
\end{itemize}

\bibliographystyle{apalike}
\bibliography{references}
\end{document}

### 2021-07-20 presentations/presentation_2021/presentation.tex
renamed stuff will make a new presentation

--- 2021-06-20 presentations/presentation_current/presentation.tex
@@
-% Created 2021-06-18 Fri 09:11
+% Created 2021-07-07 Wed 18:06
 % Intended LaTeX compiler: pdflatex
@@
 
-\begin{frame}[label={sec:org8153378}]{What is El Niño/Southern Oscillation (ENSO)?}
+\begin{frame}[label={sec:org4e471fb}]{What is El Niño/Southern Oscillation (ENSO)?}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org30dba5e}]{Long-Term vs Short-Term Change}
+\begin{frame}[label={sec:orgbda0410}]{Long-Term vs Short-Term Change}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orge54f468}]{Review of literature}
+\begin{frame}[label={sec:org86b3819}]{Review of literature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org85f5f49}]{Research Goals}
+\begin{frame}[label={sec:orga816529}]{Research Goals}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org6749e99}]{Data: the CESM1 Large Ensemble}
+\begin{frame}[label={sec:org6382c25}]{Data: the CESM1 Large Ensemble}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org650aae3}]{Niño 3.4 Variance}
+\begin{frame}[label={sec:orgdd97104}]{Niño 3.4 Variance}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orgdfcbcd5}]{Butterfly Effect: The Need for a Large Ensemble}
+\begin{frame}[label={sec:org95b465d}]{Butterfly Effect: The Need for a Large Ensemble}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org7d549e7}]{Model Predictions: ENSO in the Future}
+\begin{frame}[label={sec:org9e668e7}]{Model Predictions: ENSO in the Future}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org3a535ea}]{Analysis of Individual Factors}
+\begin{frame}[label={sec:orgdeea30d}]{Analysis of Individual Factors}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org87d91c6}]{Role of Greenhouse and Aerosol Emissions}
+\begin{frame}[label={sec:org2a9fa36}]{Role of Greenhouse and Aerosol Emissions}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:org1369c26}]{Correlation With Changes in Ocean Temperature}
+\begin{frame}[label={sec:org56f69a6}]{Correlation With Changes in Ocean Temperature}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org10d0dfb}]{Physical Mediator: Heating Difference}
+\begin{frame}[label={sec:org67801d2}]{Physical Mediator: Heating Difference}
 \begin{columns}
@@
 
-\begin{frame}[label={sec:orga972c2c}]{Conclusion and Discussion}
+\begin{frame}[label={sec:org8f5cccb}]{Conclusion and Discussion}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:orgb31e23e}]{Applications, Next Steps, Limitations}
+\begin{frame}[label={sec:org4e06df6}]{Applications, Next Steps, Limitations}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org35dd772}]{Acknowledgements}
+\begin{frame}[label={sec:orgda59867}]{Acknowledgements}
 \begin{itemize}
@@
 
-\begin{frame}[label={sec:org98d3ee4}]{References}
+\begin{frame}[label={sec:orgcc97b05}]{References}
 \bibliographystyle{apalike}

### 2021-07-20 presentations/presentation_current/presentation.tex
renamed stuff will make a new presentation

--- 2021-06-16 presentations/presentation_current/presentation.tex
@@
-% Created 2021-06-16 Wed 08:12
+% Created 2021-07-20 Tue 18:30
 % Intended LaTeX compiler: pdflatex
@@
 \date{\today}
-\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
+\title{Presentation}
 \hypersetup{
  pdfauthor={Benjamin Goldman},
- pdftitle={The Impact of Anthropogenic Forcing on ENSO Amplitude},
+ pdftitle={Presentation},
  pdfkeywords={},
@@
 
-\begin{frame}[label={sec:orgb1df2b4}]{What is El Niño/Southern Oscillation (ENSO)?}
-\begin{columns}
-\begin{column}{0.6\columnwidth}
-\begin{itemize}
-\item Drives extreme weather around the world
-\item Oscillation between warm and cold temperature in the Pacific Ocean
-\item Some events are more strong than others
-\item Significant effect on people: 2015-2016 event
-\item Major issue is prediction
-\end{itemize}
-\end{column}
-
-\begin{column}{0.4\columnwidth}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./images/el_nino.jpg}
-\caption{Changes to tropical Pacific climate during El Niño. \url{https://www.esa.int/ESA\_Multimedia/Images/2018/08/El\_Nino}}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}[label={sec:orga0cf916}]{Long-Term vs Short-Term Change}
-\begin{columns}
-\begin{column}{0.4\columnwidth}
-\begin{itemize}
-\item Long-term change: climate change/global warming
-\begin{itemize}
-\item Causes: greenhouse gasses, aerosols (smoke), land use, etc.
-\end{itemize}
-\item Short-term change: climate variability
-\begin{itemize}
-\item ENSO, seasons, AMO (Atlantic Multidecadal Oscillation), etc.
-\end{itemize}
-\end{itemize}
-\end{column}
-
-\begin{column}{0.6\columnwidth}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./images/climate_change.pdf}
-\caption{Global average temperature changes since 1880. Red line: smoothed average, black line: unsmoothed average. \url{https://data.giss.nasa.gov/gistemp/graphs\_v4}}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}[label={sec:org8f2e28a}]{Review of literature}
-\begin{itemize}
-\item \citet{chen2017possible}
-\begin{itemize}
-\item Past studies disagree about whether ENSO will strengthen or weaken.
-\item Simulation discrepancy caused by modeling of ENSO mechanics.
-\end{itemize}
-\item \citet{maher2018enso}
-\begin{itemize}
-\item Used a large dataset of climate predictions.
-\item ENSO may become stronger in the future.
-\end{itemize}
-\item \citet{cai2018increased}
-\begin{itemize}
-\item Found that models agree by using a more flexible way of defining ENSO events
-\item ENSO is strengthening because global warming is leading to higher stratification.
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label={sec:org26b0634}]{Research Goals}
-\begin{itemize}
-\item Overall changes to ENSO amplitude
-\begin{itemize}
-\item Estimate future changes to ENSO amplitude using the CESM1 dataset.
-\end{itemize}
-\item Role of individual factors
-\begin{itemize}
-\item Compare contributions of greenhouse gasses, aerosols, land use, biomass burning, and ozone to ENSO intensity.
-\end{itemize}
-\item Changes to ocean structure
-\begin{itemize}
-\item Examine changes to correlation coefficient between ENSO intensity and ocean temperature for each simulation.
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label={sec:org389fba7}]{Data: the CESM1 Large Ensemble}
-\begin{itemize}
-\item Explore hypothetical scenarios with a computer model \citet{kay2015community}.
-\item Estimation of how the earth's climate actually works.
-\item Experimental group: Receives input of rising greenhouse gas and/or aerosol levels.
-\item Control group: Emissions fixed at levels before industrial revolution.
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label={sec:org80dd814}]{Niño 3.4 Variance}
-\begin{columns}
-\begin{column}{0.6\columnwidth}
-\begin{itemize}
-\item How to calculate ENSO intensity in the model output?
-\item Step 1: Calculate sea temperature in Niño 3.4 region of tropical Pacific Ocean.
-\item Step 2: Convert temperature dataset to dataset representing change in temperature variation over time.
-\item Calculate variance around one point, move point forward slightly, repeat.
-\end{itemize}
-\end{column}
-
-\begin{column}{0.4\columnwidth}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./images/nino3_4.pdf}
-\caption{Niño 3.4 region is the shaded box.}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}[label={sec:org8592647}]{Butterfly Effect: The Need for a Large Ensemble}
-\begin{columns}
-\begin{column}{0.5\columnwidth}
-\begin{itemize}
-\item Butterfly effect: Small differences in initial conditions can become big differences in end result \citep{lorenz1963deterministic}.
-\item Each simulation by itself is inaccurate.
-\item Repeat simulation with slightly different initial conditions.
-\item Due to larger sample size, noise can be filtered out by calculating the mean.
-\end{itemize}
-\end{column}
-
-\begin{column}{0.5\columnwidth}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./images/variance_3.pdf}
-\caption{Niño 3.4 20-year variance for individual members in full forcing ensemble.}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}[label={sec:orgf5c3604}]{Model Predictions: ENSO in the Future}
-\begin{columns}
-\begin{column}{.5\columnwidth}
-\begin{itemize}
-\item Calculate mean and standard error of ENSO intensity in ensemble and control.
-\item ENSO is predicted to intensify in the 21st century!
-\item Statistically significant: exceeds 2 standard errors.
-\item Decreasing variance after 2060: still under investigation.
-\end{itemize}
-\end{column}
-
-\begin{column}{0.5\columnwidth}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./images/variance_2.pdf}
-\caption{20-year variance of Niño 3.4 index for fully-forced ensemble. Grey bar shows control mean and standard errors}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}[label={sec:org9a47952}]{Analysis of Individual Factors}
-\begin{itemize}
-\item Why is ENSO predicted to intensify? What human impacts play the largest role?
-\begin{itemize}
-\item Factors include: Greenhouse gasses, aerosols, natural factors.
-\end{itemize}
-\item Separate out individual influences in model output.
-\item Single forcing ensembles: forced by all factors except for 1.
-\item Subtract “all-but-one” ensembles from original “full-forcing” ensemble.
-\item Resulting data represents influence of only one factor.
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label={sec:org0a966de}]{Role of Greenhouse and Aerosol Emissions}
-\begin{columns}
-\begin{column}{0.5\columnwidth}
-\begin{itemize}
-\item Greenhouse gasses and aerosols contribute to increase in variance.
-\item Aerosols and greenhouse gasses have same sign: disagree with previous studies \citep{deser2020isolating}.
-\item Greenhouse gasses and aerosols are both human-produced.
-\end{itemize}
-\end{column}
-
-\begin{column}{0.5\columnwidth}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./images/bootstrap_2.pdf}
-\caption{Influence of individual human factors. Yellow is greenhouse gasses, green is aerosols.}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}[label={sec:org288c45c}]{Correlation With Changes in Ocean Temperature}
-\begin{itemize}
-\item Examine relationship between ocean temperature and ENSO intensity in each simulation.
-\item Calculate correlation coefficient between ENSO intensity and ocean temperature.
-\item Find correlation coefficient at each grid-point.
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label={sec:org8bce193}]{Physical Mediator: Heating Difference}
-\begin{columns}
-\begin{column}{.62\columnwidth}
-\begin{itemize}
-\item Strong negative correlation in fully forced ensemble below surface.
-\item Positive correlation in greenhouse ensemble and weak/zero correlation in aerosols ensemble
-\item Rising temperatures heat different layers of ocean at different rates, modifying heat transfer.
-\end{itemize}
-\end{column}
-
-\begin{column}{.50\columnwidth}
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./images/diff_tempdt.pdf}
-\caption{Correlation between ENSO intensity and ocean temperature in 3 major ensembles}
-\end{figure}
-\end{column}
-\end{columns}
-\end{frame}
-
-\begin{frame}[label={sec:orge890f4d}]{Conclusion and Discussion}
-\begin{itemize}
-\item Predicted increase in variance
-\begin{itemize}
-\item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
-\end{itemize}
-\item Greenhouse gasses and aerosols
-\begin{itemize}
-\item Increase is likely caused by the combined influence of greenhouse gasses and aerosols.
-\end{itemize}
-\item Heat transfer
-\begin{itemize}
-\item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
-\end{itemize}
-\item Notable disagreement
-\begin{itemize}
-\item Greenhouse gasses and aerosols both increase ENSO amplitude, in contrast to \citet{deser2020isolating}
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label={sec:orge9c36c0}]{Applications, Next Steps, Limitations}
-\begin{itemize}
-\item Improve prediction ability to help people prepare for increased likelihood of extreme weather.
-\item Reduce danger by switching to renewable energy.
-\item Limitations:
-\begin{itemize}
-\item Only used one climate model.
-\item Niño 3.4 index may not be fully accurate for various models (Cai et. al. 2018).
-\end{itemize}
-\item Next steps:
-\begin{itemize}
-\item Work with other datasets, such as the new CESM2.
-\item Examine other variables to further analyze mediator process.
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label={sec:org4cda3b8}]{Acknowledgements}
-\begin{itemize}
-\item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
-1852977.
-\item Thank you to my teacher, my family, and my mentor!
-\item Role of mentor:
-\begin{itemize}
-\item Provide raw data from his facility
-\item Suggest methods and interpretations
-\item Provide feedback on results
-\item Make similar calculations to check student's results
-\end{itemize}
-\end{itemize}
-\end{frame}
-
-\begin{frame}[label={sec:org3d198c1}]{References}
-\bibliographystyle{apa}
-\fontsize{7pt}{7.2}\selectfont
-\bibliography{references}
+\begin{frame}[label={sec:org6b28ed4}]{Test}
 \end{frame}

### 2021-07-20 data/Code/analyzeVariance.r

@@ -3,6 +3,8 @@ library(ncdf4)
 library(zoo)
 library(dplyr)
 library(ggplot2)
+library(WaveletComp)
+library(reshape2)
 
 # CESM1 LENS and SF
 ff <- read.table("ff_var.csv", sep = ",") %>% as.matrix()
@@ -10,6 +12,8 @@ ghg <- read.table("ghg_var.csv", sep = ",") %>% as.matrix()
 aer <- read.table("aer_var.csv", sep = ",") %>% as.matrix()
 bmb <- read.table("bmb_var.csv", sep = ",") %>% as.matrix()
 luc <- read.table("luc_var.csv", sep = ",") %>% as.matrix()
+
+ff_raw_mean <- read.table("ff_raw.csv", sep = ",") %>% as.matrix() %>% apply(1, mean)
 
 ff_mean <- apply(X = ff, MARGIN = 1, FUN = mean)
 ghg_mean <- apply(X = ghg, MARGIN = 1, FUN = mean)
@@ -71,6 +75,74 @@ ff_compare <-
     cbind(mydate_2)
 colnames(ff_compare) <- c("Mean", "SE", "Case", "Date")
 
+# Bootstrap process to generate sf timeseries
+
+subtract.bootstrap <- function(data1, data2, reps) {
+    len <- dim(data1)[1]
+    num1 <- dim(data1)[2]
+    num2 <- dim(data2)[2]
+    out <- matrix(NA, ncol = reps, nrow = len)
+#    writeLines(as.character(dim(out)))
+#    writeLines(as.character(len))
+    choice1 <- sample(x = 1:num1, size = reps, replace = TRUE)
+    choice2 <- sample(x = 1:num2, size = reps, replace = TRUE)
+    for(i in 1:reps){
+        writeLines(as.character(i))
+#        writeLines(as.character(choice1[i]))
+#        writeLines(as.character(choice2[i]))
+        answer <- data1[, choice1[i]] - data2[, choice2[i]]
+#        writeLines(as.character(length(answer)))
+        out[, i] <- answer
+    }
+    return(out)
+}
+
+reps <- 1000
+se_bs <- function(x, n) {return(sd(x)/sqrt(n))}
+
+ghg_sf <- subtract.bootstrap(ff, ghg, reps)
+ghg_sf_mean <- ghg_sf %>% apply(1, mean)
[…]

### 2021-07-20 data/Code/extractVariance.r

@@ -60,3 +60,6 @@ write.table(luc_var, file = "luc_var.csv", sep = ",")
 ncin <- nc_open("/Volumes/Extreme SSD/DATA/stacked/CESM1/TREFHT/b.e11.BRCP85C5CNBDRD.f09_g16.001.cam.h0.TREFHT.192001-210012.nc")
 mydate <- as.Date(as.character(ncvar_get(ncin, "date")), "%Y%m%d")
 write.table(mydate, file = "mydate.csv", sep = ",")
+
+ff_raw <- extract_nino3.4(paste(path_ff, file_ff, sep = ""), "TREFHT")
+write.table(ff_raw, file = "ff_raw.csv", sep = ",")

### 2021-07-21 presentations/presentation_current/presentation.tex
new r stuff outlined presentation yay

@@
-% Created 2021-07-20 Tue 18:30
+% Created 2021-07-20 Tue 20:15
 % Intended LaTeX compiler: pdflatex
@@
 \renewcommand{\bibsection}{}
-\usetheme{metropolis}
+\usetheme{default}
+\usecolortheme{owl}
 \author{Benjamin Goldman}
@@
 
-\begin{frame}[label={sec:org6b28ed4}]{Test}
+\section{Introduction}
+\label{sec:orgc44fc10}
+
+\begin{frame}[label={sec:orgf174f87}]{El Nino}
 \end{frame}
+\begin{frame}[label={sec:orgbb4e216}]{Climate Change}
+\end{frame}
+\begin{frame}[label={sec:org9a7241e}]{Chaos/Timeseries analysis}
+\end{frame}
+\begin{frame}[label={sec:org0558743}]{Gap/rationale}
+\end{frame}
+\begin{frame}[label={sec:orgaa26d89}]{Hypothesis}
+\end{frame}
+
+\section{Materials, Methods, and Results}
+\label{sec:org45d9343}
+\begin{frame}[label={sec:orgfdd47ac}]{Climate simulation}
+\end{frame}
+\begin{frame}[label={sec:org59b9f58}]{CESM1 and CESM2}
+\end{frame}
+\begin{frame}[label={sec:org8e1af0f}]{R and Python tools}
+**
+\end{frame}
+
+\section{Discussion and Conclusion}
+\label{sec:orgc16fbba}
 \end{document}

### 2021-07-21 data/Code/test.r

@@ -1,0 +1,17 @@
+# a little bit of exploratory stuff on what the pacific looks like in different seasons at different times yay
+
+library(dplyr)
+library(ncdf4)
+library(ggplot2)
+library(reshape2)
+
+ncin <- nc_open("/Volumes/Extreme SSD/DATA/BEN/CESM1/TEMP/b.e11.BRCP85C5CNBDRD.f09_g16.001.pop.h.TEMPDT.192001-210012.nc")
+
+tempdt <- ncvar_get(ncin, "TEMPDT")
+
+tempdt_melt <- tempdt[100,,seq(9, 2172, by = 12)] %>% melt
+# tempdt_melt <- tempdt[,,1] %>% melt
+ggplot(tempdt_melt, aes(Var2, Var1, z = value)) +
+    geom_contour_filled(bins = 16)
+
+plot(tempdt[1,1,])

### 2021-07-21 data/Code_old/extract.sh

@@ -7,18 +7,6 @@ set -euo pipefail
 
 nces -d time,840,1871 "b.e11.B20TRC5CNBDRD.f09_g16.001.cam.h0.TREFHT.185001-200512.nc" "b.e11.B20TRC5CNBDRD.f09_g16.001.cam.h0.TREFHT.192001-200512.nc"
 
-# ensemble concatenate cesm1 trefht members *DEPRECATED*
-
-ls | grep "b.e11.B20TRC5CNBDRD.f09_g16.\d\{3\}.cam.h0.TREFHT.192001-200512.nc" | ncecat -D 2 -o "b.e11.B20TRC5CNBDRD.f09_g16.all.cam.h0.TREFHT.192001-200512.nc"
-ls | grep "b.e11.BRCP85C5CNBDRD.f09_g16.\d\{3\}.cam.h0.TREFHT.200601-208012.nc" | ncecat -D 2 -o "b.e11.BRCP85C5CNBDRD.f09_g16.001-033.cam.h0.TREFHT.200601-208012.nc"
-ls | grep "b.e11.BRCP85C5CNBDRD.f09_g16.\d\{3\}.cam.h0.TREFHT.208101-210012.nc" | ncecat -D 2 -o "b.e11.BRCP85C5CNBDRD.f09_g16.001-033.cam.h0.TREFHT.208101-210012.nc"
-ls | grep "b.e11.BRCP85C5CNBDRD.f09_g16.\d\{3\}.cam.h0.TREFHT.200601-210012.nc" | ncecat -D 2 -o "b.e11.BRCP85C5CNBDRD.f09_g16.034-105.cam.h0.TREFHT.200601-210012.nc"
-
-# record concatenate 2006-2080 and 2080-2100 files *DEPRECATED*
-
-ls | grep "b.e11.BRCP85C5CNBDRD.f09_g16.001-033.cam.h0.TREFHT.\d\{6\}-\d\{6\}.nc" | ncrcat -D 2 -o "b.e11.BRCP85C5CNBDRD.f09_g16.001-033.cam.h0.TREFHT.200601-210012.nc"
[…]

### 2021-07-22 data/Code/analyzeVariance.r
continued to plan presentation, little r stuff

@@ -39,6 +39,8 @@ aer_df <- data.frame(date = mydate, mean = aer_mean, se = aer_se)
 bmb_df <- data.frame(date = mydate, mean = bmb_mean, se = bmb_se)
 luc_df <- data.frame(date = mydate, mean = luc_mean, se = luc_se)
 
+ff_members <- ff %>% melt() %>% mutate(date = (Var1/12 + 1920))
+
 cesm_1 <- ff_df %>% mutate(case = factor("Full Forcing")) %>%
   bind_rows(ghg_df %>%
               mutate(case = factor("Greenhouse"))) %>%
@@ -59,6 +61,8 @@ ff_2_se <- apply(X = ff_2, MARGIN = 1, FUN = se)
 mydate_2 <- seq(1850, 2100, length.out = 3012)
 
 ff_2_df <- data.frame(date = mydate_2, mean = ff_2_mean, se = ff_2_se)
+
+ff_2_members <- ff_2 %>% melt() %>% mutate(date = (Var1/12 + 1850))
 
 #ff_compare <- ff_df %>% mutate(case = factor("CESM1")) %>%
 #    bind_rows(ff_2_df %>%
@@ -155,6 +159,16 @@ ggplot(ff_df, aes(y = mean, x = date, ymin = (mean - se), ymax = (mean + se))) +
     geom_line(color = "Blue") +
     geom_ribbon(alpha = .5, fill = "Blue", color = NA)
 
+ggplot(ff_2_df, aes(y = mean, x = date, ymin = (mean - se), ymax = (mean + se))) +
+    geom_line() +
+    geom_ribbon(alpha = .5)
+
+ggplot(ff_members, aes(x = date, y = value, group = Var2)) +
+    geom_line(alpha = .1) +
+
+ggplot(ff_2_members, aes(x = date, y = value, group = Var2)) +
+    geom_line(alpha = .1)
+
 ggplot(ff_compare, aes(y = Mean, x = Date, color = Case, fill = Case, ymin = (Mean - SE), ymax = (Mean + SE))) +
     geom_line() +
     geom_ribbon(alpha = 0.5, color = NA)
@@ -164,10 +178,6 @@ ggplot(cesm_1_sf, aes(y = Mean, x = Date, color = Case, fill = Case, ymin = (Mea
     facet_grid(Case ~ .) +
     geom_line() +
     geom_ribbon(alpha = 0.5, color = NA)
-
-
[…]

### 2021-07-22 presentations/presentation_current/presentation.tex
presentation skeleton

--- 2021-07-20 presentations/presentation_current/presentation.tex
@@
-% Created 2021-07-20 Tue 18:30
-% Intended LaTeX compiler: pdflatex
-\documentclass[presentation]{beamer}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{graphicx}
-\usepackage{grffile}
-\usepackage{longtable}
-\usepackage{wrapfig}
-\usepackage{rotating}
-\usepackage[normalem]{ulem}
-\usepackage{amsmath}
-\usepackage{textcomp}
-\usepackage{amssymb}
-\usepackage{capt-of}
-\usepackage{hyperref}
+\documentclass{beamer}
+\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
+\author{Ben Goldman}
+\date{\today}
 \usepackage{natbib}
+\usepackage{tikz}
+\usetheme{metropolis}
 \renewcommand{\bibsection}{}
-\usetheme{metropolis}
-\author{Benjamin Goldman}
-\date{\today}
-\title{Presentation}
-\hypersetup{
- pdfauthor={Benjamin Goldman},
- pdftitle={Presentation},
- pdfkeywords={},
- pdfsubject={},
- pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
- pdflang={English}}
+
 \begin{document}
@@
 
-\begin{frame}[label={sec:org6b28ed4}]{Test}
+\begin{frame}
+  \frametitle{Climate Change}
+  \citet{bjerknes1969atmospheric}
+
 \end{frame}
+
+\begin{frame}
+  \frametitle{El Niño}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Climate Simulation}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{ENSO in the Future}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Gap and Goal}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Research Questions}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{signal and Noise}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Ensembles: CESM1 and CESM2}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{R and Python tools}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Role of Mentor and Student}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Measuring ENSO}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Measuring ENSO Intensity}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{ENSO is Becoming Stronger}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{It's not That Simple}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Single Forcing Ensembles}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Influence of Aerosols and Greenhouse Gasses}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Correlation With Ocean Temperature}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stratification}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{Stratification in Other Ensembles}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{References}
+  \bibliographystyle{apalike}
+  \bibliography{references.bib}
+\end{frame}
+
+\maketitle
+
 \end{document}

### 2021-07-23 data/Code/extractVariance.r
hours and some other stuff

@@ -33,33 +33,30 @@ files_sf <- c("b.e11.B20TRLENS_RCP85.f09_g16.xghg.all.cam.h0.TREFHT.192001-20801
               "b.e11.B20TRLENS_RCP85.f09_g16.xbmb.all.cam.h0.TREFHT.192001-202912.nc",
               "b.e11.B20TRLENS_RCP85.f09_g16.xlulc.all.cam.h0.TREFHT.192001-202912.nc")
 
-ff_var <- extract_nino3.4(paste(path_ff, file_ff, sep = ""), "TREFHT") %>%
-  apply(2, rollvar)
-write.table(ff_var, file = "ff_var.csv", sep = ",")
+ff <- extract_nino3.4(paste(path_ff, file_ff, sep = ""), "TREFHT")
+write.table(ff %>% apply(2, rollvar), file = "ff_var.csv", sep = ",")
+write.table(ff, file = "ff.csv", sep = ",")
 
-ghg_var <- extract_nino3.4(paste(path_sf, files_sf[1], sep = ""), "TREFHT") %>%
-    apply(2, rollvar) %>%
+ghg <- extract_nino3.4(paste(path_sf, files_sf[1], sep = ""), "TREFHT") %>%
     rbind(matrix(data=NA, nrow = (2172-1932), ncol = 20))
-write.table(ghg_var, file = "ghg_var.csv", sep = ",")
+write.table(ghg %>% apply(2, rollvar), file = "ghg_var.csv", sep = ",")
+write.table(ghg, file = "ghg.csv", sep = ",")
 
-aer_var <- extract_nino3.4(paste(path_sf, files_sf[2], sep = ""), "TREFHT") %>%
-  apply(2, rollvar) %>%
+aer <- extract_nino3.4(paste(path_sf, files_sf[2], sep = ""), "TREFHT") %>%
   rbind(matrix(data = NA, nrow = (2172-1932), ncol = 20))
-write.table(aer_var, file = "aer_var.csv", sep = ",")
+write.table(aer %>% apply(2, rollvar), file = "aer_var.csv", sep = ",")
+write.table(aer, file = "aer.csv", sep = ",")
 
-bmb_var <- extract_nino3.4(paste(path_sf, files_sf[3], sep = ""), "TREFHT") %>%
-  apply(2, rollvar) %>%
+bmb <- extract_nino3.4(paste(path_sf, files_sf[3], sep = ""), "TREFHT") %>%
[…]

### 2021-08-01 data/Code/analyzeVariance.r
lotsa refactoring

@@ -1,188 +1,0 @@
-# Imports
-library(ncdf4)
-library(zoo)
-library(dplyr)
-library(ggplot2)
-library(WaveletComp)
-library(reshape2)
-
-# CESM1 LENS and SF
-ff <- read.table("ff_var.csv", sep = ",") %>% as.matrix()
-ghg <- read.table("ghg_var.csv", sep = ",") %>% as.matrix()
-aer <- read.table("aer_var.csv", sep = ",") %>% as.matrix()
-bmb <- read.table("bmb_var.csv", sep = ",") %>% as.matrix()
-luc <- read.table("luc_var.csv", sep = ",") %>% as.matrix()
-
-ff_raw_mean <- read.table("ff_raw.csv", sep = ",") %>% as.matrix() %>% apply(1, mean)
-
-ff_mean <- apply(X = ff, MARGIN = 1, FUN = mean)
-ghg_mean <- apply(X = ghg, MARGIN = 1, FUN = mean)
-aer_mean <- apply(X = aer, MARGIN = 1, FUN = mean)
-bmb_mean <- apply(X = bmb, MARGIN = 1, FUN = mean)
-luc_mean <- apply(X = luc, MARGIN = 1, FUN = mean)
-
-se <- function(x, ...) sd(x, ...)/sqrt(length(x))
-
-ff_se <- apply(X = ff, MARGIN = 1, FUN = se)
-ghg_se <- apply(X = ghg, MARGIN = 1, FUN = se)
-aer_se <- apply(X = aer, MARGIN = 1, FUN = se)
-bmb_se <- apply(X = bmb, MARGIN = 1, FUN = se)
-luc_se <- apply(X = luc, MARGIN = 1, FUN = se)
-
-date.num <- seq(1920, 2100, length.out = 2172)
-date.str <- read.table("mydate.csv", sep = ",")$x
-mydate <- date.num
-
-ff_df <- data.frame(date = mydate, mean = ff_mean, se = ff_se)
-ghg_df <- data.frame(date = mydate, mean = ghg_mean, se = ghg_se)
-aer_df <- data.frame(date = mydate, mean = aer_mean, se = aer_se)
-bmb_df <- data.frame(date = mydate, mean = bmb_mean, se = bmb_se)
-luc_df <- data.frame(date = mydate, mean = luc_mean, se = luc_se)
-
-ff_members <- ff %>% melt() %>% mutate(date = (Var1/12 + 1920))
-
[…]

### 2021-08-01 data/Code/analyzeVarianceCESM2.r

@@ -1,18 +1,0 @@
-library(ncdf4)
-library(zoo)
-library(dplyr)
-library(ggplot2)
-
-ff <- read.table("ff_var_CESM2.csv", sep = ",") %>% as.matrix()
-
-ff_mean <- apply(X = ff, MARGIN = 1, FUN = mean)
-
-se <- function(x, ...) sd(x, ...)/sqrt(length(x))
-
-ff_se <- apply(X = ff, MARGIN = 1, FUN = se)
-
-mydate <- seq(1850, 2100, length.out = 3012)
-
-ggplot(ff_df, aes(y = mean, x = date, ymin = (mean - se), ymax = (mean + se))) +
-    geom_line(color = "Blue") +
-    geom_ribbon(alpha = .5, fill = "Blue", color = NA)

### 2021-08-01 data/Code/extractVariance.r

@@ -1,62 +1,0 @@
-library(ncdf4)
-library(zoo)
-library(dplyr)
-library(ggplot2)
-
-rollvar <- function(x) {
-  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
-  return(out)
-}
-
-
-se <- function(x, ...) sqrt(var(x, ...)/length(x))
-
-extract_nino3.4 <- function(ncfile, xvar) {
-  ncin <- nc_open(ncfile)
-  out <- apply(
-    ncvar_get(ncin, xvar, start = c(153, 91, 1, 1), count = c(40, 11, -1, -1)),
-    MARGIN = c(3, 4),
-    FUN = mean
-  )
-  return(out)
-  nc_close(ncin)
-}
-
-path_ff <- "/Volumes/Extreme\ SSD/DATA/stacked/CESM1/TREFHT/"
-path_sf <- "/Volumes/Extreme\ SSD/DATA/stacked/CESM1SF/TREFHT/"
-
-file_ff <- "b.e11.BRCP85C5CNBDRD.f09_g16.all.cam.h0.TREFHT.192001-210012.nc"
-
-
-files_sf <- c("b.e11.B20TRLENS_RCP85.f09_g16.xghg.all.cam.h0.TREFHT.192001-208012.nc",
-              "b.e11.B20TRLENS_RCP85.f09_g16.xaer.all.cam.h0.TREFHT.192001-208012.nc",
-              "b.e11.B20TRLENS_RCP85.f09_g16.xbmb.all.cam.h0.TREFHT.192001-202912.nc",
-              "b.e11.B20TRLENS_RCP85.f09_g16.xlulc.all.cam.h0.TREFHT.192001-202912.nc")
-
-ff <- extract_nino3.4(paste(path_ff, file_ff, sep = ""), "TREFHT")
-write.table(ff %>% apply(2, rollvar), file = "ff_var.csv", sep = ",")
-write.table(ff, file = "ff.csv", sep = ",")
-
-ghg <- extract_nino3.4(paste(path_sf, files_sf[1], sep = ""), "TREFHT") %>%
-    rbind(matrix(data=NA, nrow = (2172-1932), ncol = 20))
-write.table(ghg %>% apply(2, rollvar), file = "ghg_var.csv", sep = ",")
-write.table(ghg, file = "ghg.csv", sep = ",")
-
-aer <- extract_nino3.4(paste(path_sf, files_sf[2], sep = ""), "TREFHT") %>%
[…]

### 2021-08-01 data/Code/extractVarianceCESM2.r

@@ -1,36 +1,0 @@
-library(dplyr)
-library(ncdf4)
-library(zoo)
-
-rollvar <- function(x) {
-  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
-  return(out)
-}
-
-file_ff <- "b.e21.BOTH.f09_g17.LE2-XXXX.all.cam.h0.TREFHT.185001-210012.nc"
-path_ff <- "/Volumes/Extreme SSD/DATA/stacked/CESM2/TREFHT/"
-
-ncin <- nc_open(paste(path_ff, file_ff, sep = ""))
-
-#nino <- ncvar_get(ncin, "TREFHT", start = c(153, 91, 1, 1), count = c(40, 11, 3012, 20))
-
-
-extract_nino3.4 <- function(ncfile, xvar) {
-    ncin <- nc_open(ncfile)
-    out <- matrix(0, nrow = 3012, ncol = 98)
-    for(t in 1:3012){
-        slice <- apply(
-            ncvar_get(ncin, xvar, start = c(153, 91, t, 1), count = c(40, 11, 1, -1)),
-            MARGIN = 3,
-            FUN = mean)
-        out[t,] <- slice
-        cat(paste(t, "\n"))
-    }
-    return(out)
-}
-
-nino <- extract_nino3.4(paste(path_ff, file_ff, sep = ""), "TREFHT")
-
-nino_var <- apply(nino, FUN = rollvar, MARGIN = 2)
-
-write.table(nino_var, file = "ff_var_CESM2.csv", sep = ",")

### 2021-08-01 data/Code/test.r

@@ -1,17 +1,0 @@
-# a little bit of exploratory stuff on what the pacific looks like in different seasons at different times yay
-
-library(dplyr)
-library(ncdf4)
-library(ggplot2)
-library(reshape2)
-
-ncin <- nc_open("/Volumes/Extreme SSD/DATA/BEN/CESM1/TEMP/b.e11.BRCP85C5CNBDRD.f09_g16.001.pop.h.TEMPDT.192001-210012.nc")
-
-tempdt <- ncvar_get(ncin, "TEMPDT")
-
-tempdt_melt <- tempdt[100,,seq(9, 2172, by = 12)] %>% melt
-# tempdt_melt <- tempdt[,,1] %>% melt
-ggplot(tempdt_melt, aes(Var2, Var1, z = value)) +
-    geom_contour_filled(bins = 16)
-
-plot(tempdt[1,1,])

### 2021-08-01 data/Code/wavelets.r

@@ -1,116 +1,0 @@
-# Imports
-library(ncdf4)
-library(zoo)
-library(dplyr)
-library(ggplot2)
-library(WaveletComp)
-library(reshape2)
-
-# get dato from file, apply 1 year rolling average to remove seasonality and calculate ensemble means
-ff <- read.table("ff.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
-ghg <- read.table("ghg.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
-aer <- read.table("aer.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
-bmb <- read.table("bmb.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
-luc <- read.table("luc.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
-
-# mydate <- read.table("mydate.csv", sep = ",")$x
-
-mydate_1 <- seq(1920, 2100, length.out = 2172)
-mydate_2 <- seq(1920, 2099, length.out = 2161)
-mydate_3 <- seq(1920, 2079, length.out = 1921)
-
-## ff_raw_mean <- read.table("ff.csv", sep = ",") %>% apply(1, mean)
-
-## ff_rw_ts_dec <- ts(ff_raw_mean, frequency = 12) %>% decompose()
-## ff_resid <- ff_rw_ts_dec$random
-## ff_resid_wvlt_power <- WaveletTransform(ff_resid[7:2166])$Power %>% melt()
-
-## ff_reg_wvlt_power <- WaveletTransform(ff_raw_mean)$Power %>% melt
-
-## ff_resid_wvlt_power$Var1 <- (2172 / ff_resid_wvlt_power)$Var1 / 12
-## ff_reg_wvlt_power$Var1 <- (2172 / ff_reg_wvlt_power$Var1) / 12
-
-## ggplot(ff_resid_wvlt_power, aes(Var2, Var1, z = value)) +
-##     geom_contour_filled() +
-##     scale_y_log10()
-
-## ggplot(ff_reg_wvlt_power, aes(Var2, Var1, z = value)) +
-##     geom_contour_filled() +
-##     scale_y_log10()
-
[…]

### 2021-08-01 data/Code_old/20year_variance.py

@@ -1,663 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Apr 27 14:01:34 2020
-
-@author: bengoldman
-"""
-# %%
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy import stats, signal
-from matplotlib import gridspec
-from math import sqrt
-from random import randint
-import os
-from matplotlib import cm
-import sys
-from helpers import *
-
-# %%
-from variance import *
-
-# %%
-os.chdir("../Data")
-
-ctl = np.loadtxt("ctl.csv")
-ff_set = np.loadtxt("ff_set.csv")
-ff_1850 = np.loadtxt("ff_1850.csv")
-ghg_set = np.loadtxt("ghg_set.csv")
-aer_set = np.loadtxt("aer_set.csv")
-bmb_set = np.loadtxt("bmb_set.csv")
-luc_set = np.loadtxt("luc_set.csv")
-ozo_set = np.loadtxt("ozo_set.csv")
-oic_set = np.loadtxt("oic_set.csv")
-oic_1850 = np.loadtxt("oic_1850.csv")
-ctl_amoc = np.loadtxt("ctl_amoc.csv")
-ctl_amo = np.loadtxt("ctl_amo.csv")
-lat = np.loadtxt("lat.csv")
-lon = np.loadtxt("lon.csv")
-lon1 = np.loadtxt("lon1.csv")
-depth = np.loadtxt("depth.csv")
-
-# %%
-os.chdir('/Users/bengoldman/ENSO-amplitude/Figures')
-
-# %%
-ctl_var = pd.DataFrame(ctl).rolling(240, center=True).var().to_numpy()[:, 0]
-ff_var_set = pd.DataFrame(ff_set).rolling(240, center=True).var().to_numpy()
-ff_1850_var = pd.DataFrame(ff_1850).rolling(240, center=True).var().to_numpy()
-ghg_var_set = pd.DataFrame(ghg_set).rolling(240, center=True).var().to_numpy()
-aer_var_set = pd.DataFrame(aer_set).rolling(240, center=True).var().to_numpy()
[…]

### 2021-08-01 data/Code_old/control_variance.py

@@ -1,147 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy import stats, signal
-from matplotlib import gridspec
-from math import sqrt
-from random import randint
-import os
-from matplotlib import cm
-
-from helpers import *
-from variance import *
-
-ctl_var_nan = ctl_var[~np.isnan(ctl_var)]
-
-x = np.linspace(0.7, 2, 500)
-
-ctl_var_div_20 = np.zeros((1000, 20))
-for i in range(20):
-    ctl_var_div_20[:, i] = ctl_var_nan[i*1000:(i+1)*1000]
-
-ctl_var_div_10 = np.zeros((2000, 10))
-for i in range(10):
-    ctl_var_div_10[:, i] = ctl_var_nan[i*2000:(i+1)*2000]
-
-ctl_var_div_5 = np.zeros((4000, 5))
-for i in range(5):
-    ctl_var_div_5[:, i] = ctl_var_nan[i*4000:(i+1)*4000]
-
-ctl_var_div_20_pdf = np.zeros((500, 20))
-for i in range(20):
-    ctl_var_div_20_pdf[:, i] = stats.gaussian_kde(ctl_var_div_20[:, i]).evaluate(x)
-
-ctl_var_div_10_pdf = np.zeros((500, 10))
-for i in range(10):
-    ctl_var_div_10_pdf[:, i] = stats.gaussian_kde(ctl_var_div_10[:, i]).evaluate(x)
-
-ctl_var_div_5_pdf = np.zeros((500, 5))
-for i in range(5):
-    ctl_var_div_5_pdf[:, i] = stats.gaussian_kde(ctl_var_div_5[:, i]).evaluate(x)
-
-ctl_var_pdf = stats.gaussian_kde(ctl_var_nan).evaluate(x)
-
-# %%
-gs = gridspec.GridSpec(20, 3)
-
-fig = plt.figure(figsize=(10, 10))
-
-for i in range(20):
-    ax = fig.add_subplot(gs[i, 0])
-    ax.plot(x, ctl_var_div_20_pdf[:, i])
-#    ax.set_xticklabels([])
-
-for i in range(10):
[…]

### 2021-08-01 data/Code_old/correlation_xsections.py

@@ -1,231 +1,0 @@
-# %%
-from helpers import *
-from variance import ff_var_mean, ghg_var_mean, aer_var_mean, ghg_bs_mean, aer_bs_mean
-
-# %%
-ff_tempdt = np.load("/Volumes/Extreme SSD/LE/compiled/ff_tempdt_avg.npy")
-ff_tempcep = np.load("/Volumes/Extreme SSD/LE/compiled/ff_tempcep_avg.npy")
-ff_tempwep = np.load("/Volumes/Extreme SSD/LE/compiled/ff_tempwep_avg.npy")
-ff_tempeep = np.load("/Volumes/Extreme SSD/LE/compiled/ff_tempeep_avg.npy")
-ghg_tempdt = np.load("/Volumes/Extreme SSD/LE/compiled/ghg_tempdt_avg.npy")
-ghg_tempcep = np.load("/Volumes/Extreme SSD/LE/compiled/ghg_tempcep_avg.npy")
-ghg_tempwep = np.load("/Volumes/Extreme SSD/LE/compiled/ghg_tempwep_avg.npy")
-ghg_tempeep = np.load("/Volumes/Extreme SSD/LE/compiled/ghg_tempeep_avg.npy")
-aer_tempdt = np.load("/Volumes/Extreme SSD/LE/compiled/aer_tempdt_avg.npy")
-aer_tempcep = np.load("/Volumes/Extreme SSD/LE/compiled/aer_tempcep_avg.npy")
-aer_tempwep = np.load("/Volumes/Extreme SSD/LE/compiled/aer_tempwep_avg.npy")
-aer_tempeep = np.load("/Volumes/Extreme SSD/LE/compiled/aer_tempeep_avg.npy")
-
-# %%
-print("ghg")
-diff_ghg_tempdt_rs, diff_ghg_tempdt_ps = correlation_detrended(-(ghg_tempdt-ff_tempdt[:1932]), ghg_bs_mean)
-print("ghg")
-diff_ghg_tempcep_rs, diff_ghg_tempcep_ps = correlation_detrended(-(ghg_tempcep-ff_tempcep[:1932]), ghg_bs_mean)
-print("ghg")
-diff_ghg_tempwep_rs, diff_ghg_tempwep_ps = correlation_detrended(-(ghg_tempwep-ff_tempwep[:1932]), ghg_bs_mean)
-print("ghg")
[…]

### 2021-08-01 data/Code_old/extract.sh

@@ -1,216 +1,0 @@
-#!/usr/bin/env bash
-set -euo pipefail
-
-# You probably shouldn't run this file, just copy and paste from it into the shell.
-
-# clip 1850-2005 member at 1920
-
-nces -d time,840,1871 "b.e11.B20TRC5CNBDRD.f09_g16.001.cam.h0.TREFHT.185001-200512.nc" "b.e11.B20TRC5CNBDRD.f09_g16.001.cam.h0.TREFHT.192001-200512.nc"
-
-# Take 2
-
-for i in {001..033};
-do
-    f1="b.e11.B20TRC5CNBDRD.f09_g16.$i.cam.h0.TREFHT.192001-200512.nc";
-    f2="b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.200601-208012.nc";
-    f3="b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.208101-210012.nc";
-    ncrcat $f1 $f2 $f3 -D 2 -o "b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.192001-210012.nc";
-done
-
-for i in {034,035,101,102,103,104,105};
-do
-    f1="b.e11.B20TRC5CNBDRD.f09_g16.$i.cam.h0.TREFHT.192001-200512.nc";
-    f2="b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.200601-210012.nc";
-    ncrcat $f1 $f2 -D 2 -o "b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.192001-210012.nc";
-done
-
-ncrcat b.e11.B20TRC5CNBDRD.f09_g16.035.cam.h0.TREFHT.192001-200512.nc b.e11.BRCP85C5CNBDRD.f09_g16.035.cam.h0.TREFHT.200601-210012.nc -D 3 -o b.e11.BRCP85C5CNBDRD.f09_g16.035.cam.h0.TREFHT.192001-210012.nc
-
-# Try this to check stuff
-for f in $(ls | grep "192001-210012.nc$"); do echo $f; ncdump -h $f | grep "currently"; echo ""; done
[…]

### 2021-08-01 data/Code_old/helpers.py

@@ -1,226 +1,0 @@
-#! /usr/bin/env python3
-
-import netCDF4 as nc
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy import stats, signal
-from matplotlib import gridspec
-from math import sqrt
-from random import randint
-import os
-from matplotlib import cm, colors
-import re
-
-lat = np.loadtxt("../Data/lat.csv")
-lon = np.loadtxt("../Data/lon.csv")
-lon1 = np.loadtxt("../Data/lon1.csv")
-lat1 = np.loadtxt("../Data/lat1.csv")
-depth = np.loadtxt("../Data/depth.csv")/100
-
-date_2172 = np.linspace(1920, 2100, 2172)
-date_1872 = np.linspace(1850, 2005, 1872)
-date_1032 = np.linspace(1920, 2005, 1032)
-date_1932 = np.linspace(1920, 2080, 1932)
-date_1320 = np.linspace(1920, 2029, 1320)
-date_612 = np.linspace(1955, 2005, 612)
-date_960 = np.linspace(1920, 1999, 960)
-date_1800 = np.linspace(1850, 1999, 1800)
-date_80 = np.linspace(1920, 1999, 80)
-date_ctl_20600 = np.linspace(400, 2200, 21600)
-date_ctl_1800 = np.linspace(400, 2200, 1800)
-
-
-def envelope(ax, x, y, err, col):
-    ax.plot(x, y, color=col)
-    ax.fill_between(x, y-err, y+err, alpha=.2, color=col)
-
-
-def figlett(ax, a):
-    ax.text(0.01, 0.99, a, transform=ax.transAxes, va='top')
-
-size1 = (4, 3)
-size2 = (4, 5)
-size3 = (6, 5)
-mp = np.ones((400, 3))
-mp[21:40] = [.6, .6, .6]
-mp[-40:-21] = [.6, .6, .6]
-mp[0:20, 0] = np.linspace(0, 1, 20)
-mp[0:20, 1] = np.linspace(0, 1, 20)
-mp[0:20, 2] = 1
-mp[-21:-1, 0] =1
-mp[-21:-1, 1] = np.linspace(1, 0, 20)
-mp[-21:-1, 2] = np.linspace(1, 0, 20)
-cmap = colors.ListedColormap(mp)
-
-mp = np.zeros((2, 4))
-mp[1, :] = [1, 1, 1, 1]
[…]

### 2021-08-01 data/Code_old/open_tempdt.py

@@ -1,151 +1,0 @@
-# %%
-import numpy as np
-import netCDF4 as nc
-import os
-
-os.chdir("/Volumes/Extreme SSD/LE/TEMPDT/ff/")
-
-# %%
-ff_tempdt_set = np.zeros((34, 2172, 60, 180))
-for i in range(2, 36):
-    print(i)
-    if i < 10:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.pop.h.TEMPDT.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPDT.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPDT.208101-210012.nc" % i)
-
-        v1 = f1.variables["TEMPDT"][:, :, :]
-        v2 = f2.variables["TEMPDT"][:, :, :]
-        v3 = f3.variables["TEMPDT"][:, :, :]
-
-        v1[v1.mask] = np.nan
-        v2[v2.mask] = np.nan
-        v3[v3.mask] = np.nan
-        ff_tempdt_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
-
-        f1.close()
-        f2.close()
-        f3.close()
-
-    elif i < 34:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.pop.h.TEMPDT.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPDT.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPDT.208101-210012.nc" % i)
-
-
-        v1 = f1.variables["TEMPDT"][:, :, :]
-        v2 = f2.variables["TEMPDT"][:, :, :]
-        v3 = f3.variables["TEMPDT"][:, :, :]
-        
-        v1[v1.mask] = np.nan
-        v2[v2.mask] = np.nan
-        v3[v3.mask] = np.nan
-        ff_tempdt_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
-
-        f1.close()
-        f2.close()
[…]

### 2021-08-01 data/Code_old/open_tempeep.py

@@ -1,143 +1,0 @@
-# %%
-import numpy as np
-import netCDF4 as nc
-import os
-
-os.chdir("/Volumes/Extreme SSD/LE/TEMPEEP/ff/")
-
-# %%
-ff_tempeep_set = np.zeros((34, 2172, 60, 41))
-for i in range(2, 36):
-    print(i, end="|")
-    if i < 10:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.pop.h.TEMPEEP.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPEEP.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPEEP.208101-210012.nc" % i)
-
-        v1 = f1.variables["TEMPEEP"][:, :, :]
-        v2 = f2.variables["TEMPEEP"][:, :, :]
-        v3 = f3.variables["TEMPEEP"][:, :, :]
-
-        v1[v1.mask] = np.nan
-        v2[v2.mask] = np.nan
-        v3[v3.mask] = np.nan
-        ff_tempeep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
-
-        f1.close()
-        f2.close()
-        f3.close()
-
-    elif i < 34:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.pop.h.TEMPEEP.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPEEP.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPEEP.208101-210012.nc" % i)
-
-
-        v1 = f1.variables["TEMPEEP"][:, :, :]
-        v2 = f2.variables["TEMPEEP"][:, :, :]
-        v3 = f3.variables["TEMPEEP"][:, :, :]
-        
-        v1[v1.mask] = np.nan
-        v2[v2.mask] = np.nan
-        v3[v3.mask] = np.nan
-        ff_tempeep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
-
-        f1.close()
[…]

### 2021-08-01 data/Code_old/open_tempwep.py

@@ -1,143 +1,0 @@
-# %%
-import numpy as np
-import netCDF4 as nc
-import os
-
-os.chdir("/Volumes/Extreme SSD/LE/TEMPWEP/ff/")
-
-# %%
-ff_tempwep_set = np.zeros((34, 2172, 60, 41))
-for i in range(2, 36):
-    print(i, end="|")
-    if i < 10:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.pop.h.TEMPWEP.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPWEP.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPWEP.208101-210012.nc" % i)
-
-        v1 = f1.variables["TEMPWEP"][:, :, :]
-        v2 = f2.variables["TEMPWEP"][:, :, :]
-        v3 = f3.variables["TEMPWEP"][:, :, :]
-
-        v1[v1.mask] = np.nan
-        v2[v2.mask] = np.nan
-        v3[v3.mask] = np.nan
-        ff_tempwep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
-
-        f1.close()
-        f2.close()
-        f3.close()
-
-    elif i < 34:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.pop.h.TEMPWEP.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPWEP.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPWEP.208101-210012.nc" % i)
-
-
-        v1 = f1.variables["TEMPWEP"][:, :, :]
-        v2 = f2.variables["TEMPWEP"][:, :, :]
-        v3 = f3.variables["TEMPWEP"][:, :, :]
-        
-        v1[v1.mask] = np.nan
-        v2[v2.mask] = np.nan
-        v3[v3.mask] = np.nan
-        ff_tempwep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
-
-        f1.close()
[…]

### 2021-08-01 data/Code_old/variance.py

@@ -1,127 +1,0 @@
-
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy import stats, signal
-from math import sqrt
-from random import randint
-import os
-
-# %%
-
-os.chdir('/Users/bengoldman/ENSO-amplitude/Data')
-ctl = np.loadtxt("ctl.csv")
-ff_set = np.loadtxt("ff_set.csv")
-ff_1850 = np.loadtxt("ff_1850.csv")
-ghg_set = np.loadtxt("ghg_set.csv")
-aer_set = np.loadtxt("aer_set.csv")
-bmb_set = np.loadtxt("bmb_set.csv")
-luc_set = np.loadtxt("luc_set.csv")
-ozo_set = np.loadtxt("ozo_set.csv")
-oic_set = np.loadtxt("oic_set.csv")
-oic_1850 = np.loadtxt("oic_1850.csv")
-ctl_amoc = np.loadtxt("ctl_amoc.csv")
-ctl_amo = np.loadtxt("ctl_amo.csv")
-lat = np.loadtxt("lat.csv")
-lon = np.loadtxt("lon.csv")
-lon1 = np.loadtxt("lon1.csv")
-depth = np.loadtxt("depth.csv")
-
-os.chdir('/Users/bengoldman/ENSO-amplitude/Figures')
-
-# %%
-ctl_var = pd.DataFrame(ctl).rolling(240, center=True).var().to_numpy()[:, 0]
-ff_var_set = pd.DataFrame(ff_set).rolling(240, center=True).var().to_numpy()
-ff_1850_var = pd.DataFrame(ff_1850).rolling(240, center=True).var().to_numpy()
-ghg_var_set = pd.DataFrame(ghg_set).rolling(240, center=True).var().to_numpy()
-aer_var_set = pd.DataFrame(aer_set).rolling(240, center=True).var().to_numpy()
-bmb_var_set = pd.DataFrame(bmb_set).rolling(240, center=True).var().to_numpy()
-luc_var_set = pd.DataFrame(luc_set).rolling(240, center=True).var().to_numpy()
-ozo_var_set = pd.DataFrame(ozo_set).rolling(240, center=True).var().to_numpy()
[…]

### 2021-08-01 data/Code_old/variance.r

@@ -1,119 +1,0 @@
-library(zoo)
-library(ncdf4)
-library(ggplot2)
-library(dplyr)
-library(gridExtra)
-library(grid)
-library(boot)
-
-
-
-
-rollvar <- function(x) {
-  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
-  return(out)
-}
-
-se <- function(x, ...) sqrt(var(x, ...)/length(x))
-
-extract_nino3.4 <- function(ncfile, xvar) {
-  ncin <- nc_open(ncfile)
-  out <- apply(
-    ncvar_get(ncin, xvar, start = c(153, 91, 1, 1), count = c(40, 11, -1, -1)), 
-    MARGIN = c(3, 4),
-    FUN = mean
-  )
-  return(out)
-  nc_close(ncin)
-}
-
-
-ncin <- nc_open("/Volumes/Extreme SSD/CESM/CESM1/CESM1/TREFHT/b.e11.BOTH.f09_g16.ensemble.cam.h0.TREFHT.192001-210012.nc")
-date <- as.Date(ncvar_get(ncin, "time"), origin = "1920-01-01")
-date <- seq(1920, 2100, length.out = 2172)
-
-path_ff <- "/Volumes/Extreme\ SSD/CESM/CESM1/CESM1/TREFHT/"
-path_sf <- "/Volumes/Extreme\ SSD/CESM/CESM1/CESM1SF/TREFHT/"
-
-file_ff <- "b.e11.BOTH.f09_g16.ensemble.cam.h0.TREFHT.192001-210012.nc"
-files_sf <- c(
-  "b.e11.B20TRLENS_RCP85.f09_g16.xghg.ensemble.cam.h0.TREFHT.192001-208012.nc",
-  "b.e11.B20TRLENS_RCP85.f09_g16.xaer.ensemble.cam.h0.TREFHT.192001-208012.nc",
-  "b.e11.B20TRLENS_RCP85.f09_g16.xbmb.ensemble.cam.h0.TREFHT.192001-202912.nc",
-  "b.e11.B20TRLENS_RCP85.f09_g16.xlulc.ensemble.cam.h0.TREFHT.192001-202912.nc",
-  "b.e11.B20LE_fixedO3_ensemble.cam.h0.TREFHT.195501-200512.nc"
-  )
[…]

### 2021-08-01 data/analyzeVariance.r

@@ -1,0 +1,171 @@
+# Imports
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+library(WaveletComp)
+library(reshape2)
+
+prepare <- function(x) {
+    x.mean <- x %>% apply(1, mean)
+    return(x - x.mean)
+}
+
+rollvar <- function(x) {
+  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
+  return(out)
+}
+
+load_variance <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% prepare %>% apply(2, rollvar))
+
+se <- function(x, ...) sd(x, ...)/sqrt(length(x))
+
+bind_stats <- function(this.date, this.mean, this.se, date.start = NA, date.end = NA) {
+    if(length(this.mean) == length(this.date)) {
+        this.df <- data.frame(date = this.date, mean = this.mean, se = this.se)
+    } else {
+        if(is.na(date.end)){
+            date.end <- length(this.date)
+        }
+        if(is.na(date.start)){
+            date.start <- 0
+        }
+        this.df <- data.frame(date = this.date, mean = c(
+                                                    rep(NA, date.start),
+                                                    this.mean,
+                                                    rep(NA, (length(this.date) - date.end))),
+                              se =
+                                  c(rep(NA, date.start),
+                                    this.se,
+                                    rep(NA, (length(this.date) - date.end))))
+        writeLines(as.character(date.start))
+        writeLines(as.character(date.end))
+        return(this.df)
+    }
+}
+
+make_df <- function(this, this.date, date.start = NA, date.end = NA) {
+    this.mean <- this %>% apply(1, mean)
+    this.se <- this %>% apply(1, se)
+    writeLines(as.character(length(this.mean)))
+    this.df <- bind_stats(this.date, this.mean, this.se, date.start, date.end)
+    return(this.df)
+}
+
+date1 <- seq(1920, 2100, length.out = 2172)
[…]

### 2021-08-01 data/analyzeVarianceCESM2.r

@@ -1,0 +1,18 @@
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+
+ff <- read.table("ff_var_CESM2.csv", sep = ",") %>% as.matrix()
+
+ff_mean <- apply(X = ff, MARGIN = 1, FUN = mean)
+
+se <- function(x, ...) sd(x, ...)/sqrt(length(x))
+
+ff_se <- apply(X = ff, MARGIN = 1, FUN = se)
+
+mydate <- seq(1850, 2100, length.out = 3012)
+
+ggplot(ff_df, aes(y = mean, x = date, ymin = (mean - se), ymax = (mean + se))) +
+    geom_line(color = "Blue") +
+    geom_ribbon(alpha = .5, fill = "Blue", color = NA)

### 2021-08-01 data/archive/Code_old/extract.sh

@@ -1,0 +1,216 @@
+#!/usr/bin/env bash
+set -euo pipefail
+
+# You probably shouldn't run this file, just copy and paste from it into the shell.
+
+# clip 1850-2005 member at 1920
+
+nces -d time,840,1871 "b.e11.B20TRC5CNBDRD.f09_g16.001.cam.h0.TREFHT.185001-200512.nc" "b.e11.B20TRC5CNBDRD.f09_g16.001.cam.h0.TREFHT.192001-200512.nc"
+
+# Take 2
+
+for i in {001..033};
+do
+    f1="b.e11.B20TRC5CNBDRD.f09_g16.$i.cam.h0.TREFHT.192001-200512.nc";
+    f2="b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.200601-208012.nc";
+    f3="b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.208101-210012.nc";
+    ncrcat $f1 $f2 $f3 -D 2 -o "b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.192001-210012.nc";
+done
+
+for i in {034,035,101,102,103,104,105};
+do
+    f1="b.e11.B20TRC5CNBDRD.f09_g16.$i.cam.h0.TREFHT.192001-200512.nc";
+    f2="b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.200601-210012.nc";
+    ncrcat $f1 $f2 -D 2 -o "b.e11.BRCP85C5CNBDRD.f09_g16.$i.cam.h0.TREFHT.192001-210012.nc";
+done
+
+ncrcat b.e11.B20TRC5CNBDRD.f09_g16.035.cam.h0.TREFHT.192001-200512.nc b.e11.BRCP85C5CNBDRD.f09_g16.035.cam.h0.TREFHT.200601-210012.nc -D 3 -o b.e11.BRCP85C5CNBDRD.f09_g16.035.cam.h0.TREFHT.192001-210012.nc
+
+# Try this to check stuff
+for f in $(ls | grep "192001-210012.nc$"); do echo $f; ncdump -h $f | grep "currently"; echo ""; done
[…]

### 2021-08-01 data/archive/Code_old/fix_data.py

@@ -1,0 +1,10 @@
+import netCDF4 as nc
+import numpy as np
+import os
+
+os.chdir("/Volumes/Extreme SSD/LE/TEMPWEP/ff")
+for file in os.listdir():
+    f = nc.Dataset(file, mode="a")
+    del f.variables["TEMPWEP"].valid_range
+    print(f.variables["TEMPWEP"])
+    f.close()

### 2021-08-01 data/archive/Code_old/helpers.py

@@ -1,0 +1,226 @@
+#! /usr/bin/env python3
+
+import netCDF4 as nc
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+from scipy import stats, signal
+from matplotlib import gridspec
+from math import sqrt
+from random import randint
+import os
+from matplotlib import cm, colors
+import re
+
+lat = np.loadtxt("../Data/lat.csv")
+lon = np.loadtxt("../Data/lon.csv")
+lon1 = np.loadtxt("../Data/lon1.csv")
+lat1 = np.loadtxt("../Data/lat1.csv")
+depth = np.loadtxt("../Data/depth.csv")/100
+
+date_2172 = np.linspace(1920, 2100, 2172)
+date_1872 = np.linspace(1850, 2005, 1872)
+date_1032 = np.linspace(1920, 2005, 1032)
+date_1932 = np.linspace(1920, 2080, 1932)
+date_1320 = np.linspace(1920, 2029, 1320)
+date_612 = np.linspace(1955, 2005, 612)
+date_960 = np.linspace(1920, 1999, 960)
+date_1800 = np.linspace(1850, 1999, 1800)
+date_80 = np.linspace(1920, 1999, 80)
+date_ctl_20600 = np.linspace(400, 2200, 21600)
+date_ctl_1800 = np.linspace(400, 2200, 1800)
+
+
+def envelope(ax, x, y, err, col):
+    ax.plot(x, y, color=col)
+    ax.fill_between(x, y-err, y+err, alpha=.2, color=col)
+
+
+def figlett(ax, a):
+    ax.text(0.01, 0.99, a, transform=ax.transAxes, va='top')
+
+size1 = (4, 3)
+size2 = (4, 5)
+size3 = (6, 5)
+mp = np.ones((400, 3))
+mp[21:40] = [.6, .6, .6]
+mp[-40:-21] = [.6, .6, .6]
+mp[0:20, 0] = np.linspace(0, 1, 20)
+mp[0:20, 1] = np.linspace(0, 1, 20)
+mp[0:20, 2] = 1
+mp[-21:-1, 0] =1
+mp[-21:-1, 1] = np.linspace(1, 0, 20)
+mp[-21:-1, 2] = np.linspace(1, 0, 20)
+cmap = colors.ListedColormap(mp)
+
+mp = np.zeros((2, 4))
+mp[1, :] = [1, 1, 1, 1]
[…]

### 2021-08-01 data/archive/Code_old/open_tempcep.py

@@ -1,0 +1,143 @@
+# %%
+import numpy as np
+import netCDF4 as nc
+import os
+
+os.chdir("/Volumes/Extreme SSD/LE/TEMPCEP/ff/")
+
+# %%
+ff_tempcep_set = np.zeros((34, 2172, 60, 41))
+for i in range(2, 36):
+    print(i, end="|")
+    if i < 10:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.pop.h.TEMPCEP.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPCEP.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPCEP.208101-210012.nc" % i)
+
+        v1 = f1.variables["TEMPCEP"][:, :, :]
+        v2 = f2.variables["TEMPCEP"][:, :, :]
+        v3 = f3.variables["TEMPCEP"][:, :, :]
+
+        v1[v1.mask] = np.nan
+        v2[v2.mask] = np.nan
+        v3[v3.mask] = np.nan
+        ff_tempcep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
+
+        f1.close()
+        f2.close()
+        f3.close()
+
+    elif i < 34:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.pop.h.TEMPCEP.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPCEP.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPCEP.208101-210012.nc" % i)
+
+
+        v1 = f1.variables["TEMPCEP"][:, :, :]
+        v2 = f2.variables["TEMPCEP"][:, :, :]
+        v3 = f3.variables["TEMPCEP"][:, :, :]
+        
+        v1[v1.mask] = np.nan
+        v2[v2.mask] = np.nan
+        v3[v3.mask] = np.nan
+        ff_tempcep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
+
+        f1.close()
[…]

### 2021-08-01 data/archive/Code_old/open_tempdt.py

@@ -1,0 +1,151 @@
+# %%
+import numpy as np
+import netCDF4 as nc
+import os
+
+os.chdir("/Volumes/Extreme SSD/LE/TEMPDT/ff/")
+
+# %%
+ff_tempdt_set = np.zeros((34, 2172, 60, 180))
+for i in range(2, 36):
+    print(i)
+    if i < 10:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.pop.h.TEMPDT.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPDT.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPDT.208101-210012.nc" % i)
+
+        v1 = f1.variables["TEMPDT"][:, :, :]
+        v2 = f2.variables["TEMPDT"][:, :, :]
+        v3 = f3.variables["TEMPDT"][:, :, :]
+
+        v1[v1.mask] = np.nan
+        v2[v2.mask] = np.nan
+        v3[v3.mask] = np.nan
+        ff_tempdt_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
+
+        f1.close()
+        f2.close()
+        f3.close()
+
+    elif i < 34:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.pop.h.TEMPDT.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPDT.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPDT.208101-210012.nc" % i)
+
+
+        v1 = f1.variables["TEMPDT"][:, :, :]
+        v2 = f2.variables["TEMPDT"][:, :, :]
+        v3 = f3.variables["TEMPDT"][:, :, :]
+        
+        v1[v1.mask] = np.nan
+        v2[v2.mask] = np.nan
+        v3[v3.mask] = np.nan
+        ff_tempdt_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
+
+        f1.close()
+        f2.close()
[…]

### 2021-08-01 data/archive/Code_old/open_tempeep.py

@@ -1,0 +1,143 @@
+# %%
+import numpy as np
+import netCDF4 as nc
+import os
+
+os.chdir("/Volumes/Extreme SSD/LE/TEMPEEP/ff/")
+
+# %%
+ff_tempeep_set = np.zeros((34, 2172, 60, 41))
+for i in range(2, 36):
+    print(i, end="|")
+    if i < 10:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.pop.h.TEMPEEP.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPEEP.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPEEP.208101-210012.nc" % i)
+
+        v1 = f1.variables["TEMPEEP"][:, :, :]
+        v2 = f2.variables["TEMPEEP"][:, :, :]
+        v3 = f3.variables["TEMPEEP"][:, :, :]
+
+        v1[v1.mask] = np.nan
+        v2[v2.mask] = np.nan
+        v3[v3.mask] = np.nan
+        ff_tempeep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
+
+        f1.close()
+        f2.close()
+        f3.close()
+
+    elif i < 34:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.pop.h.TEMPEEP.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPEEP.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPEEP.208101-210012.nc" % i)
+
+
+        v1 = f1.variables["TEMPEEP"][:, :, :]
+        v2 = f2.variables["TEMPEEP"][:, :, :]
+        v3 = f3.variables["TEMPEEP"][:, :, :]
+        
+        v1[v1.mask] = np.nan
+        v2[v2.mask] = np.nan
+        v3[v3.mask] = np.nan
+        ff_tempeep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
+
+        f1.close()
[…]

### 2021-08-01 data/archive/Code_old/open_tempwep.py

@@ -1,0 +1,143 @@
+# %%
+import numpy as np
+import netCDF4 as nc
+import os
+
+os.chdir("/Volumes/Extreme SSD/LE/TEMPWEP/ff/")
+
+# %%
+ff_tempwep_set = np.zeros((34, 2172, 60, 41))
+for i in range(2, 36):
+    print(i, end="|")
+    if i < 10:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.pop.h.TEMPWEP.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPWEP.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPWEP.208101-210012.nc" % i)
+
+        v1 = f1.variables["TEMPWEP"][:, :, :]
+        v2 = f2.variables["TEMPWEP"][:, :, :]
+        v3 = f3.variables["TEMPWEP"][:, :, :]
+
+        v1[v1.mask] = np.nan
+        v2[v2.mask] = np.nan
+        v3[v3.mask] = np.nan
+        ff_tempwep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
+
+        f1.close()
+        f2.close()
+        f3.close()
+
+    elif i < 34:
+        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.pop.h.TEMPWEP.192001-200512.nc" % i)
+        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPWEP.200601-208012.nc" % i)
+        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPWEP.208101-210012.nc" % i)
+
+
+        v1 = f1.variables["TEMPWEP"][:, :, :]
+        v2 = f2.variables["TEMPWEP"][:, :, :]
+        v3 = f3.variables["TEMPWEP"][:, :, :]
+        
+        v1[v1.mask] = np.nan
+        v2[v2.mask] = np.nan
+        v3[v3.mask] = np.nan
+        ff_tempwep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
+
+        f1.close()
[…]

### 2021-08-01 data/archive/Code_old/tempcep.py

@@ -1,0 +1,191 @@
+import sys
+
+sys.path.append("./")
+from helpers import *
+from variance import ff_var_mean, ghg_var_mean, aer_var_mean
+
+# %%
+ff_tempcep_avg = np.load("../Data/ff_tempcep_average.npy")
+ff_tempcep_avg_det = ff_tempcep_avg
+
+ff_correlation = np.zeros(ff_tempcep_avg_det[0].shape)
+ff_p_vals = np.zeros(ff_tempcep_avg_det[0].shape)
+ff_var_mean_det = ff_var_mean
+signal.detrend(ff_var_mean_det[~np.isnan(ff_var_mean_det)], overwrite_data=True)
+ff_nino_ts = pd.DataFrame(ff_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
+for i in range(len(ff_tempcep_avg_det[0, :, 0])):
+    for j in range(len(ff_tempcep_avg_det[0, i, :])):
+        print(str(i) + ":" + str(j) )
+        if np.isnan(ff_tempcep_avg_det[200, i, j]):
+            r, p = (np.nan, np.nan)
+        else:
+            signal.detrend(ff_tempcep_avg_det[:, i, j], overwrite_data=True)
+            ff_tempcep_avg_det[:, i, j] = pd.DataFrame(ff_tempcep_avg_det[:, i, j]).rolling(360, center=True).mean().to_numpy()[:, 0]
+            r, p = stats.pearsonr(ff_tempcep_avg_det[300:-300, i, j], ff_nino_ts[300:-300])
+        ff_correlation[i, j], ff_p_vals[i, j] = r, p
+
+ghg_tempcep_avg = np.load("../Data/ghg_tempcep_average.npy")
+ghg_tempcep_avg_det = ghg_tempcep_avg
+
+ghg_correlation = np.zeros(ghg_tempcep_avg_det[0].shape)
+ghg_p_vals = np.zeros(ghg_tempcep_avg_det[0].shape)
+ghg_var_mean_det = ghg_var_mean
+signal.detrend(ghg_var_mean_det[~np.isnan(ghg_var_mean_det)], overwrite_data=True)
+ghg_nino_ts = pd.DataFrame(ghg_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
+for i in range(len(ghg_tempcep_avg_det[0, :, 0])):
[…]

### 2021-08-01 data/extractVariance.r

@@ -1,0 +1,62 @@
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+
+rollvar <- function(x) {
+  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
+  return(out)
+}
+
+
+se <- function(x, ...) sqrt(var(x, ...)/length(x))
+
+extract_nino3.4 <- function(ncfile, xvar) {
+  ncin <- nc_open(ncfile)
+  out <- apply(
+    ncvar_get(ncin, xvar, start = c(153, 91, 1, 1), count = c(40, 11, -1, -1)),
+    MARGIN = c(3, 4),
+    FUN = mean
+  )
+  return(out)
+  nc_close(ncin)
+}
+
+path_ff <- "/Volumes/Extreme\ SSD/DATA/stacked/CESM1/TREFHT/"
+path_sf <- "/Volumes/Extreme\ SSD/DATA/stacked/CESM1SF/TREFHT/"
+
+file_ff <- "b.e11.BRCP85C5CNBDRD.f09_g16.all.cam.h0.TREFHT.192001-210012.nc"
+
+
+files_sf <- c("b.e11.B20TRLENS_RCP85.f09_g16.xghg.all.cam.h0.TREFHT.192001-208012.nc",
+              "b.e11.B20TRLENS_RCP85.f09_g16.xaer.all.cam.h0.TREFHT.192001-208012.nc",
+              "b.e11.B20TRLENS_RCP85.f09_g16.xbmb.all.cam.h0.TREFHT.192001-202912.nc",
+              "b.e11.B20TRLENS_RCP85.f09_g16.xlulc.all.cam.h0.TREFHT.192001-202912.nc")
+
+ff <- extract_nino3.4(paste(path_ff, file_ff, sep = ""), "TREFHT")
+write.table(ff %>% apply(2, rollvar), file = "ff_var.csv", sep = ",")
+write.table(ff, file = "./data/CESM1/ff.csv", sep = ",")
+
+ghg <- extract_nino3.4(paste(path_sf, files_sf[1], sep = ""), "TREFHT") %>%
+    rbind(matrix(data=NA, nrow = (2172-1932), ncol = 20))
+write.table(ghg %>% apply(2, rollvar), file = "ghg_var.csv", sep = ",")
+write.table(ghg, file = "./data/CESM1/ghg.csv", sep = ",")
+
[…]

### 2021-08-01 data/extractVarianceCESM2.r

@@ -1,0 +1,36 @@
+library(dplyr)
+library(ncdf4)
+library(zoo)
+
+rollvar <- function(x) {
+  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
+  return(out)
+}
+
+file_ff <- "b.e21.BOTH.f09_g17.LE2-XXXX.all.cam.h0.TREFHT.185001-210012.nc"
+path_ff <- "/Volumes/Extreme SSD/DATA/stacked/CESM2/TREFHT/"
+
+ncin <- nc_open(paste(path_ff, file_ff, sep = ""))
+
+#nino <- ncvar_get(ncin, "TREFHT", start = c(153, 91, 1, 1), count = c(40, 11, 3012, 20))
+
+
+extract_nino3.4 <- function(ncfile, xvar) {
+    ncin <- nc_open(ncfile)
+    out <- matrix(0, nrow = 3012, ncol = 98)
+    for(t in 1:3012){
+        slice <- apply(
+            ncvar_get(ncin, xvar, start = c(153, 91, t, 1), count = c(40, 11, 1, -1)),
+            MARGIN = 3,
+            FUN = mean)
+        out[t,] <- slice
+        cat(paste(t, "\n"))
+    }
+    return(out)
+}
+
+nino <- extract_nino3.4(paste(path_ff, file_ff, sep = ""), "TREFHT")
+
+nino_var <- apply(nino, FUN = rollvar, MARGIN = 2)
+
+write.table(nino, file = "./data/CESM2/ff.csv", sep = ",")

### 2021-08-01 data/wavelets.r

@@ -1,0 +1,116 @@
+# Imports
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+library(WaveletComp)
+library(reshape2)
+
+# get dato from file, apply 1 year rolling average to remove seasonality and calculate ensemble means
+ff <- read.table("ff.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
+ghg <- read.table("ghg.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
+aer <- read.table("aer.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
+bmb <- read.table("bmb.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
+luc <- read.table("luc.csv", sep = ",") %>% apply(1, mean) %>% rollapply(width = 12, FUN = mean, align = "center")
+
+# mydate <- read.table("mydate.csv", sep = ",")$x
+
+mydate_1 <- seq(1920, 2100, length.out = 2172)
+mydate_2 <- seq(1920, 2099, length.out = 2161)
+mydate_3 <- seq(1920, 2079, length.out = 1921)
+
+## ff_raw_mean <- read.table("ff.csv", sep = ",") %>% apply(1, mean)
+
+## ff_rw_ts_dec <- ts(ff_raw_mean, frequency = 12) %>% decompose()
+## ff_resid <- ff_rw_ts_dec$random
+## ff_resid_wvlt_power <- WaveletTransform(ff_resid[7:2166])$Power %>% melt()
+
+## ff_reg_wvlt_power <- WaveletTransform(ff_raw_mean)$Power %>% melt
+
+## ff_resid_wvlt_power$Var1 <- (2172 / ff_resid_wvlt_power)$Var1 / 12
+## ff_reg_wvlt_power$Var1 <- (2172 / ff_reg_wvlt_power$Var1) / 12
+
+## ggplot(ff_resid_wvlt_power, aes(Var2, Var1, z = value)) +
+##     geom_contour_filled() +
+##     scale_y_log10()
+
+## ggplot(ff_reg_wvlt_power, aes(Var2, Var1, z = value)) +
+##     geom_contour_filled() +
+##     scale_y_log10()
+
[…]

### 2021-08-03 data/analyzeVariance.r
figure stuff!

@@ -52,50 +52,6 @@ make_df <- function(this, this.date, date.start = NA, date.end = NA) {
     return(this.df)
 }
 
-date1 <- seq(1920, 2100, length.out = 2172)
-date2 <- seq(1850, 2100, length.out = 3012)
-
-# CESM1 LENS and SF
-ff1 <- load_variance("./data/CESM1/ff.csv")
-ghg1 <- load_variance("./data/CESM1/ghg.csv")
-aer1 <- load_variance("./data/CESM1/aer.csv")
-bmb1 <- load_variance("./data/CESM1/bmb.csv")
-luc1 <- load_variance("./data/CESM1/luc.csv")
-
-ff1_df <- make_df(ff1, date2, date.start = 840)
-ghg1_df <- make_df(ghg1, date2, date.start = 840)
-aer1_df <- make_df(aer1, date2, date.start = 840)
-bmb1_df <- make_df(bmb1, date2, date.start = 840)
-luc1_df <- make_df(luc1, date2, date.start = 840)
-
-ff2 <- load_variance("./data/CESM2/ff.csv")
-ff2_df <- make_df(ff2, date2)
-
-ff_members <- ff %>% melt() %>% mutate(date = (Var1/12 + 1920))
-
-cesm_1 <- ff1_df %>% mutate(case = factor("Full Forcing")) %>%
-  bind_rows(ghg1_df %>%
-              mutate(case = factor("Greenhouse"))) %>%
-  bind_rows(aer1_df %>%
-              mutate(case = factor("Aerosols"))) %>%
-  bind_rows(bmb1_df %>%
-              mutate(case = factor("Biomass")))  %>%
-  bind_rows(luc1_df %>%
-              mutate(case = factor("Land Use")))
-
-ff_2_members <- ff_2 %>% melt() %>% mutate(date = (Var1/12 + 1850))
-
-#ff_compare <- ff_df %>% mutate(case = factor("CESM1")) %>%
-#    bind_rows(ff_2_df %>%
-#              mutate(case = factor("CESM2")) %>%
-#              rbind)
-
-ff_compare <- ff1_df %>% mutate(case = factor("CESM1")) %>%
-    bind_rows(ff2_df %>% mutate(case = factor("CESM2")))
-
-# Bootstrap process to generate sf timeseries
-
-reps <- 1000
 se_bs <- function(x, n) {return(sd(x)/sqrt(n))}
 
[…]

### 2021-08-03 competitions/sts/regeneron_essay_qs.tex
moved stuff

% Created 2021-07-20 Tue 17:46
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{Benjamin Goldman}
\date{\today}
\title{Regeneron STS Application Essay Questions}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Regeneron STS Application Essay Questions},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{Research Project “Layperson’s Summary” (maximum 200 words)}
\label{sec:org3b34335}
Summarize your project in layperson’s terms, while maintaining scientific accuracy. Your explanation should be easily understandable and include background, procedures, conclusions and relevance. This summary will aid readers, including evaluators, journalists and the public.

\section{Project Benefits and Impact (maximum 200 words)}
\label{sec:org8665bc4}
What benefits do you think your research will bring to the world, and/or to your field? What additional steps, and by whom, might be needed for this benefit to be realized?

\section{Your Potential as a Scientist, Mathematician or Engineer (maximum 250 words)}
\label{sec:orgd6d85d8}
Address through specific and concrete examples what characteristics you have that best demonstrate your affinity and aptitude for being a good scientist.
\begin{itemize}
\item What have you done that illustrates scientific aptitude, leadership, curiosity, inventiveness and/or initiative?
\item After completing your research, how has your interest in science, engineering, and/or math been clarified? What are your other STEM-related interests besides your project?
\item How does your experience suggest future success as a scientist, mathematician or engineer?
\item What do you plan to study in post-secondary education and what occupation do you plan to pursue?  What do you hope to be doing 10 years from now?
\end{itemize}

\section{Major Scientific Question (maximum 250 words)}
\label{sec:org3b74cef}
What is a major scientific question in your field whose answer you believe will have a significant impact on the world in the next 20 years, and why? Using examples from your own experience or research, explain how you might envision addressing the question over the next 20 years.

\section{``Tweet'' about your project! Tell us about your project in 280 characters or less.}
\label{sec:org9998841}
The Society might share this response if you are named a scholar or finalist.

\section{Greatest Accomplishment or Challenge (COMMON APP)}
\label{sec:org228c064}
Please answer ONE of the prompts below from the Common App; we prefer that you think beyond your research project in this essay. (maximum 200 words)
Discuss an accomplishment, event, or realization that sparked a period of personal growth and a new understanding of yourself or others.
OR
The lessons we take from obstacles we encounter can be fundamental to later success. Recount a time when you faced a challenge, setback, or failure. How did it affect you, and what did you learn from the experience?

\section{About You (COMMON APP) (optional)}
\label{sec:orgbad95ff}
Some students have a background, identity, interest, or talent that is so meaningful they believe their application would be incomplete without it. If this sounds like you, then please share your story. (maximum 200 words)

\section{COVID-19 Pandemic (adapted from COMMON APP) (optional)}
\label{sec:org2636bf9}

The COVID-19 pandemic has been experienced differently throughout the country and world. Share how it has impacted your life, especially its impact on your learning and as applicable on your submitted project and ability to work on that project. Examples include:
\begin{itemize}
\item Illness or loss within your family or support network
\item Employment or housing disruptions within your family
\item Food insecurity
\item Toll on mental and emotional health
\item New obligations such as part-time work or care for siblings or family members
\item Availability of computer or internet access required to continue your studies
\item Access to a safe and quiet study space
\item A new direction for your major or career interests
\end{itemize}
\end{document}

### 2021-08-03 presentations/presentation/presentation.tex
moved stuff

\documentclass{beamer}
\title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
\author{Ben Goldman}
\date{\today}
\usepackage{natbib}
\usepackage{tikz}
\usetheme{metropolis}
\renewcommand{\bibsection}{}

% \begin{frame}{Example Frame With Text and Image}
%   \begin{columns}
%     \column{0.5\textwidth}
%     \begin{itemize}
%     \item Here we go...
%     \end{itemize}
%     \column{0.5\textwidth}
%     \begin{figure}
%       \centering
%       \includegraphics[width=\textwidth]{figures/example.jpg}
%       \caption{This is a very nice figure}
%       \label{fig:this}
%     \end{figure}
%   \end{columns}
% \end{frame}

\begin{document}

\maketitle

\section{Introduction}

\begin{frame}{Climate Change}

  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
    \item Here we go...
    \item \cite{bjerknes1969atmospheric}
    \end{itemize}
    \column{0.5\textwidth}
    \begin{figure}
      \centering
      \includegraphics[width=\textwidth]{figures/example.jpg}
      \caption{This is a very nice figure}
      \label{fig:this}
    \end{figure}
  \end{columns}
\end{frame}

\begin{frame}{El Niño}

\end{frame}

\begin{frame}{Climate Simulation}

\end{frame}

\begin{frame}{ENSO in the Future}
  Problem: What effect does global warming have on ENSO? Can we expect ENSO to become stronger, weaker, or neither?
\end{frame}

\begin{frame}{Gap and Goal}

\end{frame}

\begin{frame}{Research Questions}

\end{frame}

\section{Data and Methods}

\begin{frame}{Ensembles: CESM1 and CESM2}

\end{frame}

\begin{frame}{Analysis Tools}
  \begin{columns}[t]
    \column{.33\textwidth}
    R:
    \begin{itemize}
    \item ncdf4
    \item zoo
    \item dplyr
    \item ggplot2
    \item WaveletComp
    \item reshape2
    \end{itemize}
    \column{.33\textwidth}
    Python:
    \begin{itemize}
    \item numpy
    \item pandas
    \item scipy
    \item matplotlib
    \item netCDF4
    \end{itemize}
    \column{.33\textwidth}
    Other:
    \begin{itemize}
    \item nco
    \end{itemize}
  \end{columns}

\end{frame}

\begin{frame}{Role of Mentor and Student}
  \begin{columns}[t]
    \column{.5\textwidth}
    Mentor:
    \begin{itemize}
    \item Suggest future methods
    \item Conduct parallel analysis to complement student work
    \item Provide raw precollected data
    \item Interpret data produced by student
    \item Review student writing
    \end{itemize}
    \column{.5\textwidth}
    Student:
    \begin{itemize}
    \item Analyze data on computer
    \item Produce graphics for analysis and publication
    \item Write documentation
    \item Suggest interpretations of data
    \end{itemize}
  \end{columns}

\end{frame}

\begin{frame}{Measuring ENSO}

\end{frame}

\begin{frame}{Measuring ENSO Intensity}

\end{frame}

\begin{frame}{Signal and Noise}

\end{frame}

\begin{frame}{ENSO is Becoming Stronger}

\end{frame}

\begin{frame}{It's not That Simple}

\end{frame}

\begin{frame}{Single Forcing Ensembles}

\end{frame}

\begin{frame}{Influence of Aerosols and Greenhouse Gasses}

\end{frame}

\begin{frame}{Correlation With Ocean Temperature}

\end{frame}

\begin{frame}{Stratification}

\end{frame}

\begin{frame}{Stratification in Other Ensembles}

\end{frame}

\section{Conclusion}

\begin{frame}{Conclusions}

\end{frame}

\begin{frame}{Discussion}

\end{frame}

\begin{frame}{Acknowledgments}

\end{frame}

\begin{frame}{References}
  \bibliographystyle{apalike}
  \bibliography{references.bib}
\end{frame}

\maketitle

\end{document}

### 2021-08-05 data/extract.sh
smbb CESM2 input only

@@ -66,36 +66,15 @@ ls | grep "b.e11.B20TRLENS_RCP85.f09_g16.xlulc.\d\{3\}.cam.h0.TREFHT.192001-2029
 ls | grep "b.e11.B1850C5CN.f09_g16.005.cam.h0.TREFHT.\d\{6\}-\d\{6\}.nc" | ncrcat -D 2 -o "b.e11.B1850C5CN.f09_g16.005.cam.h0.TREFHT.040001-220012.nc"
 
 # CESM2 TREFHT
-bits=("1001.001"
+bits=(
 "1011.001"
-"1021.002"
 "1031.002"
-"1041.003"
 "1051.003"
-"1061.004"
 "1071.004"
-"1081.005"
 "1091.005"
-"1101.006"
 "1111.006"
-"1121.007"
 "1131.007"
-"1141.008"
 "1151.008"
-"1161.009"
-"1171.009"
-"1181.010"
-"1191.010"
-"1231.001"
-"1231.002"
-"1231.003"
-"1231.004"
-"1231.005"
-"1231.006"
-"1231.007"
-"1231.008"
-"1231.009"
-"1231.010"
 "1231.011"
 "1231.012"
 "1231.013"
@@ -106,16 +85,6 @@ bits=("1001.001"
 "1231.018"
 "1231.019"
 "1231.020"
-"1251.001"
-"1251.002"
-"1251.003"
-"1251.004"
-"1251.005"
-"1251.006"
-"1251.007"
-"1251.008"
-"1251.009"
-"1251.010"
 "1251.011"
 "1251.012"
 "1251.013"
@@ -126,16 +95,6 @@ bits=("1001.001"
 "1251.018"
 "1251.019"
 "1251.020"
-"1281.001"
-"1281.002"
-"1281.003"
-"1281.004"
-"1281.005"
-"1281.006"
-"1281.007"
-"1281.008"
-"1281.009"
-"1281.010"
 "1281.011"
 "1281.012"
 "1281.013"
@@ -146,16 +105,6 @@ bits=("1001.001"
 "1281.018"
 "1281.019"
 "1281.020"
-"1301.001"
-"1301.002"
-"1301.003"
-"1301.004"
-"1301.005"
-"1301.006"
-"1301.007"
-"1301.008"
-"1301.009"
-"1301.010"
 "1301.011"
 "1301.012"
 "1301.013"
@@ -165,13 +114,14 @@ bits=("1001.001"
 "1301.017"
[…]

### 2021-08-06 data/eof_fig.r
tried to make a figure for presentation

@@ -1,0 +1,40 @@
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+
+ncin <- nc_open("/Volumes/Extreme SSD/DATA/HadISST_sst.nc")
+
+sst <- ncvar_get(ncin, "sst")
+sst[sst == -1000] <- NA
+
+lat <- ncvar_get(ncin, "latitude")
+lon <- ncvar_get(ncin, "longitude")
+
+rownames(sst) <- lon
+colnames(sst) <- lat
+
+sst_wrap <- array(NA, dim = dim(sst))
+## lon_wrap <- rep(NA, times = length(lon))
+## lon_wrap[0:211] <- lon[150:360]
+## lon_wrap[212:360] <- lon[1:149]
+lon_wrap <- lon + 180
+sst_wrap[0:180,,] <- sst[181:360,,]
+sst_wrap[181:360,,] <- sst[1:180,,]
+rownames(sst_wrap) <- lon_wrap
+colnames(sst_wrap) <- lat
+
+## NEED TO MAKE ANOMALY PLOT!
+
+sst_example <- apply(sst_wrap[,,1356:1360], c(1, 2), mean)
+sst_example2 <- apply(sst_wrap[,,1359 + 24 :1360 + 24], c(1, 2), mean)
+
+sst_melt <- sst_example2 %>% melt(varnames = c("longitude", "latitude"))
+ggplot(sst_melt %>% filter(latitude < 50 &
+                           latitude > -50 &
+                           longitude > 50 &
+                           longitude < 300), aes(longitude, latitude, z = value)) +
+    geom_contour_filled(bins = 16)
+
+#  %>% filter(latitude < 20 & latitude > -20),

### 2021-08-09 paperwork/archive/paper_2020/_region_.tex
saved intermediate data

--- 2021-06-16 paperwork/papernew/papernew.tex
@@
-% Created 2021-06-16 Wed 08:13
-% Intended LaTeX compiler: pdflatex
-\documentclass[11pt]{article}
+\message{ !name(papernew.tex)}\documentclass[11pt]{article}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{natbib}
-\usepackage[letterpaper, margins=1in]{geometry}
+\usepackage[letterpaper, margin=1in]{geometry}
 \setlength{\parskip}{0.35em}
@@
 
-\maketitle
-\begin{abstract}
-
-    The El Niño/Southern Oscillation (ENSO) is the dominant mode of interannual climate variability, with substantial associated global socio-economic impacts.
-Due to their significance, shifts in ENSO under climate change also have the potential to substantially impact human society and natural ecosystems.
-However, it is currently unclear what effect greenhouse gas (GHG) and industrial aerosol (AER) emissions have on ENSO, as well as what effect these factors have when combined.
-This study examined transient changes to ENSO variance under a variety of forcing scenarios using the CESM1 Large and Single-Forcing Ensembles.
-These multi-member ensembles span the historical record (1920-2005) and much of the 21st C (2006-2080 for GHG/AER).
-A 2000-year pre-industrial (PI) control simulation is used to account for model drift and 20-year running variance of the Niño 3.4 SST index is used as a proxy for ENSO variance.
-The ensemble mean and standard error of each ensemble was calculated, while the Probability Density Function (PDF) is computed for the PI control simulation to estimate the statistical significance of simulated changes.
-We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
-We identifed significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
-
-\end{abstract}
-
+\message{ !name(papernew.tex) !offset(6) }
 \section{Introduction}
-\label{sec:org14be98b}
+\label{sec:org72159d6}
 El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
@@
 
-\section{Data}
-\label{sec:org15697ab}
-The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
-The Large Ensemble contains 40 simulations of the CESM1 coupled model, forced with historic radiative forcing from 1920 to 2005 and according to the RCP 8.5 protocol from 2006 to 2100 \citep{kay2015community}.
-The single forcing ensemble is a collection of sub-ensembles for various climate factors (greenhouse gas, aerosols, biomass burning, land use, ozone).
-Each simulation is forced by a combination of all factors except for one, allowing the impact of a single factor to be deduced by subtracting the ensemble mean from the fully-forced ensemble mean \citep{deser2020isolating}\}.
-For example, the xghg (greenhouse gas) ensemble is forced by aerosol emissions, biomass burning, land use, and ozone.
-There is also a preindustrial control simulation, with all radiative forcing fixed at 1850 levels.
+\message{ !name(papernew.tex) !offset(71) }
 
-\section{Methods and Results}
-\label{sec:org469de48}
-
-\subsection{Niño 3.4 Variance}
-\label{sec:org56d15ab}
-We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
-We measure variance on 20-year centered sliding windows.
-We calculate the multi-model mean for each ensemble, as well as the ensemble standard error.
-The results of this calculation are shown in figure \ref{fig:variance_1}.
-
-The fully-forced ensemble exhibits moderate increase in variance, with the other ensembles showing less meaningful changes.
-In the fully-forced ensemble, the variance of the Niño 3.4 region increases beyond 2 standard errors of the control, increasing until the mid-21st century, and then decreasing gradually.
-The excluded greenhouse, aerosol, and biomass-burning ensembles also contain a slight increase, which may not be statistically significant.
-Although the excluded land use ensemble mean has a strong increase in variance, this result is unlikely to be meaningful due to the xlulc ensemble's low sample size.
-Additionally, all ensembles exhibit considerable noise, in line with previous studies \citep{maher2018enso}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/variance_1.pdf}
-\caption{\label{fig:variance_1}20-year windowed variance of the Niño 3.4 region SST, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc, f) fixedO3.
-Grey bars are PI control mean and 1x, 2x, 3x standard error.
-Colored bars represent ensemble standard error.}
-\end{figure}
-
-The fully-forced ensemble exhibits reduced variance in the mid-late 20th century, below that of the PI control.
-The most likely explanation for this phenomenon is internal variability.
-To test the likelihood of this explanation, we test the influence of the Atlantic Multidecadal Oscillation (AMO) and Atlantic Meridional Overturning Current (AMOC) on ENSO in the control simulation.
-The AMO has been shown to have an influence on ENSO strength and seasonal growth rate \citep{levine2017impact}.
-We filter the control Niño 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \citep{phillips2014evaluating}\}.
-The Probability Distribution Function of Niño 3.4 variance is estimated for AMO/C > 1/2\(\sigma\) and < -1/-2\(\sigma\) using a Kernel Density Estimation.
-No meaningful differences were found in the distribution of ENSO intensity under any of these conditions (See Suplementary information figure 1).
-
-\subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org582da32}
-We analyze the role of individual factors using the CESM Single Forcing Ensembles.
-To separate the influence of a single factor from the fully-forced ensemble, we employ a bootstrap test.
-For each single-forcing ensemble, a single simulation is randomly selected, and the Niño 3.4 20-year variance record is subtracted from that from a randomly selected fully-forced simulation.
-We repeat this process 1000 times for each ensemble, and then calculate the mean and standard error for each ensemble.
-These results are shown in figure \ref{fig:bootstrap_1}.
-The greenhouse-only ensemble as well as the aerosol-only ensemble exhibit increased variance, signaling that greenhouse and aerosol emissions likely both play a significant role in ENSO's forced response in the full-forcing ensemble.
-Interestingly, the influence of greenhouse gasses and aerosols are non-conflicting, in contrast with previous studies that show opposite effects of greenhouse gas and aerosol forcing on ENSO \citep{stevenson2019forced}.
-All the other single forcing ensembles exhibit insignificant differences from the fully-forced ensemble.
-The biomass burning case shows very small deviations from the fully-forced case, while the ozone ensemble's period of recording is too small to draw meaningful conclusions.
-However, \citet{nowack2017role} showed that ozone forcing may dampen the effect of greenhouse-forced increases to ENSO amplitude by reducing changes to Pacific sea temperature and the Walker Circulation.
-The land use/cover case, while it does show large deviations from the fully-forced case, has an ensemble size (5 members) too small to lend any credibility to these changes.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/bootstrap_1.pdf}
-\caption{\label{fig:bootstrap_1}Difference between fully forced and single forced ensembles derived from the bootstrap process for a) greenhouse gas, b) aerosol emissions c) biomass burning, d) land use/cover, e) ozone; f) compares the sum of the bootstrapped ensemble means for the single forcing ensembles (blue) and the ensemble mean for the full forcing ensemble, detrended by a constant to center on zero (orange).}
-\end{figure}
-
-In both the fully-forced scenario, and the greenhouse and aerosol only simulations, there is noticeably reduced Niño 3.4 variance in the mid-20$\backslash$\textsuperscript{th} century, below 2 standard errors of the control.
-We hypothesize that this discrepancy may be the result of anomalous initial conditions caused by internal variability of the control, as the control conditions are used to initialize all the forced runs.
-To test this hypothesis, we analyzed the impact of the Atlantic Multidecadal Oscillation (AMO) and the Atlantic Meridional Overturning Current (AMOC) on Niño 3.4 variance in the control simulation using records of the AMO and AMOC from the Climate Variability Diagnostics Package (CVDP) \citep{phillips2014evaluating}.
-We filtered the 20-year variance of the Niño 3.4 sea surface temperature in the control based on the strength of the AMO/AMOC, separating ENSO variance into groups where AMO < -1\(\sigma\), AMO < -2\(\sigma\), AMO > 1\(\sigma\), AMO >2\(\sigma\), and the same for AMOC.
-We then estimated the probability density function for each group using a kernel density estimator.
-We observed no consistent difference in the distribution of Niño 3.4 variance between any group.
-So far the question of reduced variance is unanswered, and it will be addressed in further depth later in the study.
-
-\subsection{Correlation With Ocean Temperature}
-\label{sec:orga46145a}
-Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
-To do this we used 4 slices of ocean temperature from the fully-forced, xghg, and xaer ensembles, including a slice averaged along the equator and slices through the western, central, and eastern Pacific basins.
-We linearly detrend and smooth with a 30-year windowed mean the timeseries at each gridpoint and the Niño 3.4 variance.
-Next, we calculated the Pearson's correlation coefficient between each gridpoint and the Niño 3.4 variance timeseries.
-The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt_1}, and for the central slice in figure \ref{fig:tempcep_1}.
-
-Overall, the majority of the Pacific basin exhibits negative correlation with Niño 3.4 variance when linearly detrended and smoothed.
-The fully-forced simulation contains strongly negative correlation between the subsurface layer of the equatorial Pacific and ENSO amplitude, leading to the hypothesis that forced changes to Pacific Ocean stratification may be connected to ENSO intensity.
-However, it is unclear whether there is also a causal relationship between the two.
-Additionally, it is unclear whether the overall warming trend of the upper Pacific ocean is connected to changes to ENSO intensity.
-We notice that the high levels of correlation between the subsurface layer and Niño 3.4 variance are present in the xGHG and xAER ensembles, as well as the fully forced ensemble, suggesting that there is a relationship between ENSO amplitude and subsurface Pacific Ocean temperature in a variety of forcing scenarios.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/tempdt.pdf}
-\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Niño 3.4 variance.
-Grey color indicates points where p<0.05.
-The white regions at 25 and 280 degrees East are Africa and South America.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/tempcep.pdf}
-\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.
-Grey color indicates points where p>0.05.}
-\end{figure}
-
-\section{Conclusions and Discussion}
-\label{sec:org787504a}
-
-In this study, predicted changes to ENSO amplitude are examined in the CESM1 Large Ensemble and Single Forcing Ensemble.
-The fully-forced large ensemble exhibits increased ENSO amplitude, as measured by calculating the 20-year variance of the Niño 3.4 index.
-These changes appear to be statistically significant, but need to be scrutinized further, as they are in disagreement with results of some previous studies \citep{stevenson2012significant}.
-Additionally, the fully-forced ensemble, as well as all the others contains considerable noise, as the individual members cover a wide range of variability.
-Greenhouse gas and aerosol emissions are likely the most significant contributors to these forced changes, as shown by the fact that the ensemble mean for the xGHG and xAER ensembles are the most different from the fully-forced mean.
-All the other single-forcing ensembles had insignificant differences from the fully forced ensemble, or were lacking a sample size large enough to lend credibility to their results.
-Analysis of the correlation coefficient between ENSO amplitude and Pacific ocean temperature reveals that changes to ENSO intensity are tied to changes to the subsurface layer of the tropical Pacific Ocean.
-It is still unclear what is driving this connection, as well as how external forcing such as global warming affects it.
-There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \citep{dewitte2013reinterpreting}.
-
-This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \citep{maher2018enso}.
-Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citet{stevenson2019forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
-In contrast, this study observes that the impact on ENSO amplitude of greenhouse gas and aerosols has the same sign.
-This result is also surprising given that greenhouse and aerosol emissions have been shown to have opposite effects on sea surface temperature and global circulation, with greenhouse emissions favoring a more Niño-like state with weakened Walker circulation and higher SST, and aerosol emissions counteracting these effects \citep{boer2000transient}.
-
-This study has a number of goals for future continuation.
-The most pressing of these goals is a deeper analysis of changes to Pacific Ocean structure, which will be done by examining correlation between ENSO amplitude and potential density, as well as comparisons of changes to stratification in the Pacific.
-Additionally, the methods of this study will be repeated on the CESM2 large ensemble, which has a larger ensemble size and a longer record period, as compared to CESM1.
-Analysis of the CMIP6 models are also necessary to verify these results across variations in model physics.
-Deeper statistical analyses of changes to ENSO amplitude should also be tone to further examine the likelihood and intensity of future changes to ENSO amplitude.
-Lastly, correlation between ENSO amplitude and internal variability should be further investigated using the records provided by the CVDP.
-Doing these further analyses will help to verify the results of this study, while more precisely analyzing future changes to ENSO.
-
-The results of this study will help to direct future studies on the impact of climate change on ENSO by providing one further data point supporting the conclusion that ENSO will intensify due to global warming.
-This study and others like it will help societies impacted by ENSO to prepare for intensification in the future.
-
-\section{Acknowledgements}
-\label{sec:orgaff78f0}
-This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
-1852977.
-Computing resources (\href{https://doi.org/10.5065/D6RX99HX}{doi:10.5065/D6RX99HX}) were provided by the Climate Simulation Laboratory at NCAR's Computational and Information Systems Laboratory, sponsored by the National Science Foundation and other agencies.
-I would like to thank Ms.
-Kimberly Fleming and the White Plains High School Science Research Program for providing support in this study.
-
-\bibliographystyle{apa}
-\bibliography{references}
 \end{document}

### 2021-08-09 paperwork/paper/paper.tex
saved intermediate data

@@
-% Created 2021-07-07 Wed 20:51
-% Intended LaTeX compiler: xelatex
+% Created 2021-08-08 Sun 15:05
+% Intended LaTeX compiler: pdflatex
 \documentclass[11pt]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 \usepackage{graphicx}
@@
 \usepackage{hyperref}
-\usepackage{natbib}
 \usepackage[letterpaper, margins=1in]{geometry}
-\setlength{\parskip}{0.35em}
 \author{Benjamin Goldman}
@@
 
-The El Niño/Southern Oscillation (ENSO) is the dominant mode of interannual climate variability, with substantial associated global socio-economic impacts.
-Due to their significance, shifts in ENSO under climate change also have the potential to substantially impact human society and natural ecosystems.
-However, it is currently unclear what effect greenhouse gas (GHG) and industrial aerosol (AER) emissions have on ENSO, as well as what effect these factors have when combined.
-This study examined transient changes to ENSO variance under a variety of forcing scenarios using the CESM1 Large and Single-Forcing Ensembles.
-These multi-member ensembles span the historical record (1920-2005) and much of the 21st C (2006-2080 for GHG/AER).
-A 2000-year pre-industrial (PI) control simulation is used to account for model drift and 20-year running variance of the Niño 3.4 SST index is used as a proxy for ENSO variance.
-The ensemble mean and standard error of each ensemble was calculated, while the Probability Density Function (PDF) is computed for the PI control simulation to estimate the statistical significance of simulated changes.
-We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
-We identifed significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
-
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:org3fc8d9e}
-El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
-The reasons for studying ENSO are clear, as ENSO has drastically affected climate patterns worldwide, modulating rainfall and temperature in nearly every continent \citep{ropelewski1987global}.
-For example, the recent 2015-2016 El Niño event contributed to record-breaking high temperatures and droughts in South America \citep{jimenez2016record}.
-At the same time, long-term anthropogenic greenhouse gas emissions are causing global temperatures to increase through a greenhouse effect.
-The effect of greenhouse emissions and other factors on ENSO intensity remains unclear.
-
-The two major ways in which the earth's climate varies are climate change and climate variability.
-Climate change is defined as climate changes caused by external factors, most notably greenhouse gasses, natural and artificial aerosol emissions, land use changes, and stratospheric ozone changes.
-Greenhouse gas emissions have a clear impact on the earth's climate, global warming.
-Climate change is usually long-term.
-In contrast, internal variability is defined as changes to the earth's climate originating from natural climatic processes, such as ENSO, Pacific Decadal Oscillation (PDO), Atlantic Multidecadal Oscillation (AMO), and others.
-Climate variability occurs on much shorter timescales, and is usually cyclical and driven by feedback loops.
-
-Research on the effect of external forcing on ENSO remains inconclusive, as results from similar studies conflict.
-\citet{nowack2017role} predicted an overall increase in Niño 3.4 standard deviation under a combination of 4xCO\(_2\) and interactive ozone forcing using single-model simulations, showing that greenhouse gasses increase the frequency of extreme ENSO by favoring a more Niño-like in the tropical Pacific, while ozone dampens this effect.
-In contrast, a few studies have found that ENSO amplitude decreases under global warming in certain coupled models \citep{kohyama2018weakening}.
-
-However, other studies have failed to find any statistically significant ENSO response to external forcing \citep{stevenson2012significant}.
-Analysis using NCAR's CESM Large Ensemble shows an ensemble size of at least 15 models is necessary to attribute changes to Niño variance to external forcing and reject the null hypothesis that internal variability is responsible for changes to ENSO \citep{zheng2017response}.
-Several modes of internal variability have been shown to modulate ENSO, including the AMO \citep{levine2017impact} and Tropical Pacific Decadal Variability (TPDV) \citep{zheng2017response}.
-An analysis of the Max Plank Institute's Grand Ensemble as well as NCAR's CESM Large Ensemble suggests that 80$\backslash$% of changes to ENSO amplitude can be attributed to internal variability, but given a large enough ensemble, significant changes in ENSO amplitude due to climate change can be detected \citep{maher2018enso}.
-
-In this study, we show that NCAR's CESM predicts significant increase in ENSO amplitude in the 21st century, and that greenhouse gasses and aerosol emissions play important roles in causing this increase.
-As in previous studies, the role of internal variability in conjunction with forcing was examined.
-We hypothesized that increased stratification in the future plays a large role in causing this predicted increase.
-
+\label{sec:org1b56736}
 \section{Data}
-\label{sec:org0534843}
-The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
-The Large Ensemble contains 40 simulations of the CESM1 coupled model, forced with historic radiative forcing from 1920 to 2005 and according to the RCP 8.5 protocol from 2006 to 2100 \citep{kay2015community}.
-The single forcing ensemble is a collection of sub-ensembles for various climate factors (greenhouse gas, aerosols, biomass burning, land use, ozone).
-Each simulation is forced by a combination of all factors except for one, allowing the impact of a single factor to be deduced by subtracting the ensemble mean from the fully-forced ensemble mean \citep{deser2020isolating}\}.
-For example, the xghg (greenhouse gas) ensemble is forced by aerosol emissions, biomass burning, land use, and ozone.
-There is also a preindustrial control simulation, with all radiative forcing fixed at 1850 levels.
-
+\label{sec:orgaaab045}
 \section{Methods and Results}
-\label{sec:org6ea3bdd}
-
-\subsection{Niño 3.4 Variance}
-\label{sec:org907e898}
-We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
-We measure variance on 20-year centered sliding windows.
-We calculate the multi-model mean for each ensemble, as well as the ensemble standard error.
-The results of this calculation are shown in figure \ref{fig:variance_1}.
-
-The fully-forced ensemble exhibits moderate increase in variance, with the other ensembles showing less meaningful changes.
-In the fully-forced ensemble, the variance of the Niño 3.4 region increases beyond 2 standard errors of the control, increasing until the mid-21st century, and then decreasing gradually.
-The excluded greenhouse, aerosol, and biomass-burning ensembles also contain a slight increase, which may not be statistically significant.
-Although the excluded land use ensemble mean has a strong increase in variance, this result is unlikely to be meaningful due to the xlulc ensemble's low sample size.
-Additionally, all ensembles exhibit considerable noise, in line with previous studies \citep{maher2018enso}.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/variance_1.pdf}
-\caption{\label{fig:variance_1}20-year windowed variance of the Niño 3.4 region SST, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc, f) fixedO3.
-Grey bars are PI control mean and 1x, 2x, 3x standard error.
-Colored bars represent ensemble standard error.}
-\end{figure}
-
-The fully-forced ensemble exhibits reduced variance in the mid-late 20th century, below that of the PI control.
-The most likely explanation for this phenomenon is internal variability.
-To test the likelihood of this explanation, we test the influence of the Atlantic Multidecadal Oscillation (AMO) and Atlantic Meridional Overturning Current (AMOC) on ENSO in the control simulation.
-The AMO has been shown to have an influence on ENSO strength and seasonal growth rate \citep{levine2017impact}.
-We filter the control Niño 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \citep{phillips2014evaluating}\}.
-The Probability Distribution Function of Niño 3.4 variance is estimated for AMO/C > 1/2\(\sigma\) and < -1/-2\(\sigma\) using a Kernel Density Estimation.
-No meaningful differences were found in the distribution of ENSO intensity under any of these conditions (See Suplementary information figure 1).
-
-\subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:orge41456d}
-We analyze the role of individual factors using the CESM Single Forcing Ensembles.
-To separate the influence of a single factor from the fully-forced ensemble, we employ a bootstrap test.
-For each single-forcing ensemble, a single simulation is randomly selected, and the Niño 3.4 20-year variance record is subtracted from that from a randomly selected fully-forced simulation.
-We repeat this process 1000 times for each ensemble, and then calculate the mean and standard error for each ensemble.
-These results are shown in figure \ref{fig:bootstrap_1}.
-The greenhouse-only ensemble as well as the aerosol-only ensemble exhibit increased variance, signaling that greenhouse and aerosol emissions likely both play a significant role in ENSO's forced response in the full-forcing ensemble.
-Interestingly, the influence of greenhouse gasses and aerosols are non-conflicting, in contrast with previous studies that show opposite effects of greenhouse gas and aerosol forcing on ENSO \citep{stevenson2017forced}.
-All the other single forcing ensembles exhibit insignificant differences from the fully-forced ensemble.
-The biomass burning case shows very small deviations from the fully-forced case, while the ozone ensemble's period of recording is too small to draw meaningful conclusions.
-However, \citet{nowack2017role} showed that ozone forcing may dampen the effect of greenhouse-forced increases to ENSO amplitude by reducing changes to Pacific sea temperature and the Walker Circulation.
-The land use/cover case, while it does show large deviations from the fully-forced case, has an ensemble size (5 members) too small to lend any credibility to these changes.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/bootstrap_1.pdf}
-\caption{\label{fig:bootstrap_1}Difference between fully forced and single forced ensembles derived from the bootstrap process for a) greenhouse gas, b) aerosol emissions c) biomass burning, d) land use/cover, e) ozone; f) compares the sum of the bootstrapped ensemble means for the single forcing ensembles (blue) and the ensemble mean for the full forcing ensemble, detrended by a constant to center on zero (orange).}
-\end{figure}
-
-In both the fully-forced scenario, and the greenhouse and aerosol only simulations, there is noticeably reduced Niño 3.4 variance in the mid-20$\backslash$\textsuperscript{th} century, below 2 standard errors of the control.
-We hypothesize that this discrepancy may be the result of anomalous initial conditions caused by internal variability of the control, as the control conditions are used to initialize all the forced runs.
-To test this hypothesis, we analyzed the impact of the Atlantic Multidecadal Oscillation (AMO) and the Atlantic Meridional Overturning Current (AMOC) on Niño 3.4 variance in the control simulation using records of the AMO and AMOC from the Climate Variability Diagnostics Package (CVDP) \citep{phillips2014evaluating}.
-We filtered the 20-year variance of the Niño 3.4 sea surface temperature in the control based on the strength of the AMO/AMOC, separating ENSO variance into groups where AMO < -1\(\sigma\), AMO < -2\(\sigma\), AMO > 1\(\sigma\), AMO >2\(\sigma\), and the same for AMOC.
-We then estimated the probability density function for each group using a kernel density estimator.
-We observed no consistent difference in the distribution of Niño 3.4 variance between any group.
-So far the question of reduced variance is unanswered, and it will be addressed in further depth later in the study.
-
-\subsection{Correlation With Ocean Temperature}
-\label{sec:orgabb8356}
-Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
-To do this we used 4 slices of ocean temperature from the fully-forced, xghg, and xaer ensembles, including a slice averaged along the equator and slices through the western, central, and eastern Pacific basins.
-We linearly detrend and smooth with a 30-year windowed mean the timeseries at each gridpoint and the Niño 3.4 variance.
-Next, we calculated the Pearson's correlation coefficient between each gridpoint and the Niño 3.4 variance timeseries.
-The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt}, and for the central slice in figure \ref{fig:tempcep}.
-
-Overall, the majority of the Pacific basin exhibits negative correlation with Niño 3.4 variance when linearly detrended and smoothed.
-The fully-forced simulation contains strongly negative correlation between the subsurface layer of the equatorial Pacific and ENSO amplitude, leading to the hypothesis that forced changes to Pacific Ocean stratification may be connected to ENSO intensity.
-However, it is unclear whether there is also a causal relationship between the two.
-Additionally, it is unclear whether the overall warming trend of the upper Pacific ocean is connected to changes to ENSO intensity.
-We notice that the high levels of correlation between the subsurface layer and Niño 3.4 variance are present in the xGHG and xAER ensembles, as well as the fully forced ensemble, suggesting that there is a relationship between ENSO amplitude and subsurface Pacific Ocean temperature in a variety of forcing scenarios.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/tempdt.pdf}
-\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Niño 3.4 variance.
-Grey color indicates points where p<0.05.
-The white regions at 25 and 280 degrees East are Africa and South America.}
-\end{figure}
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{./figures/tempcep.pdf}
-\caption{\label{fig:tempcep}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.
-Grey color indicates points where p>0.05.}
-\end{figure}
-
-\section{Conclusions and Discussion}
-\label{sec:org811a462}
-
-In this study, predicted changes to ENSO amplitude are examined in the CESM1 Large Ensemble and Single Forcing Ensemble.
-The fully-forced large ensemble exhibits increased ENSO amplitude, as measured by calculating the 20-year variance of the Niño 3.4 index.
-These changes appear to be statistically significant, but need to be scrutinized further, as they are in disagreement with results of some previous studies \citep{stevenson2012significant}.
-Additionally, the fully-forced ensemble, as well as all the others contains considerable noise, as the individual members cover a wide range of variability.
-Greenhouse gas and aerosol emissions are likely the most significant contributors to these forced changes, as shown by the fact that the ensemble mean for the xGHG and xAER ensembles are the most different from the fully-forced mean.
-All the other single-forcing ensembles had insignificant differences from the fully forced ensemble, or were lacking a sample size large enough to lend credibility to their results.
-Analysis of the correlation coefficient between ENSO amplitude and Pacific ocean temperature reveals that changes to ENSO intensity are tied to changes to the subsurface layer of the tropical Pacific Ocean.
-It is still unclear what is driving this connection, as well as how external forcing such as global warming affects it.
-There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \citep{dewitte2012reinterpreting}.
-
-This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \citep{maher2018enso}.
-Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citet{stevenson2017forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
-In contrast, this study observes that the impact on ENSO amplitude of greenhouse gas and aerosols has the same sign.
-This result is also surprising given that greenhouse and aerosol emissions have been shown to have opposite effects on sea surface temperature and global circulation, with greenhouse emissions favoring a more Niño-like state with weakened Walker circulation and higher SST, and aerosol emissions counteracting these effects \citep{boer2000transient}.
-
-This study has a number of goals for future continuation.
-The most pressing of these goals is a deeper analysis of changes to Pacific Ocean structure, which will be done by examining correlation between ENSO amplitude and potential density, as well as comparisons of changes to stratification in the Pacific.
-Additionally, the methods of this study will be repeated on the CESM2 large ensemble, which has a larger ensemble size and a longer record period, as compared to CESM1.
-Analysis of the CMIP6 models are also necessary to verify these results across variations in model physics.
-Deeper statistical analyses of changes to ENSO amplitude should also be tone to further examine the likelihood and intensity of future changes to ENSO amplitude.
-Lastly, correlation between ENSO amplitude and internal variability should be further investigated using the records provided by the CVDP.
-Doing these further analyses will help to verify the results of this study, while more precisely analyzing future changes to ENSO.
-
-The results of this study will help to direct future studies on the impact of climate change on ENSO by providing one further data point supporting the conclusion that ENSO will intensify due to global warming.
-This study and others like it will help societies impacted by ENSO to prepare for intensification in the future.
-
-\section{Acknowledgements}
-\label{sec:org25124e8}
-This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
-1852977.
-Computing resources (\href{https://doi.org/10.5065/D6RX99HX}{doi:10.5065/D6RX99HX}) were provided by the Climate Simulation Laboratory at NCAR's Computational and Information Systems Laboratory, sponsored by the National Science Foundation and other agencies.
-I would like to thank Ms.
-Kimberly Fleming and the White Plains High School Science Research Program for providing support in this study.
+\label{sec:org3f97e45}
+\section{Conclusion}
+\label{sec:org0e6ec22}
+\section{Acknowledgments}
+\label{sec:orgaaa58bc}
 

### 2021-08-09 data/analyzeVariance.r

@@ -6,127 +6,10 @@ library(ggplot2)
 library(WaveletComp)
 library(reshape2)
 
-prepare <- function(x) {
-    x.mean <- x %>% apply(1, mean)
-    return(x - x.mean)
-}
-
-rollvar <- function(x) {
-  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
-  return(out)
-}
-
-load_variance <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% prepare %>% apply(2, rollvar))
-
-se <- function(x, ...) sd(x, ...)/sqrt(length(x))
-
-bind_stats <- function(this.date, this.mean, this.se, date.start = NA, date.end = NA) {
-    if(length(this.mean) == length(this.date)) {
-        this.df <- data.frame(date = this.date, mean = this.mean, se = this.se)
-    } else {
-        if(is.na(date.end)){
-            date.end <- length(this.date)
-        }
-        if(is.na(date.start)){
-            date.start <- 0
-        }
-        this.df <- data.frame(date = this.date, mean = c(
-                                                    rep(NA, date.start),
-                                                    this.mean,
-                                                    rep(NA, (length(this.date) - date.end))),
-                              se =
-                                  c(rep(NA, date.start),
-                                    this.se,
-                                    rep(NA, (length(this.date) - date.end))))
-        writeLines(as.character(date.start))
-        writeLines(as.character(date.end))
-        return(this.df)
-    }
-}
-
-make_df <- function(this, this.date, date.start = NA, date.end = NA) {
-    this.mean <- this %>% apply(1, mean)
-    this.se <- this %>% apply(1, se)
-    writeLines(as.character(length(this.mean)))
-    this.df <- bind_stats(this.date, this.mean, this.se, date.start, date.end)
-    return(this.df)
-}
-
-se_bs <- function(x, n) {return(sd(x)/sqrt(n))}
-
[…]

### 2021-08-09 data/correlation.r

@@ -1,0 +1,6 @@
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+library(WaveletComp)
+library(reshape2)

### 2021-08-09 data/helpers.r

@@ -1,0 +1,83 @@
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+library(WaveletComp)
+library(reshape2)
+
+# subtract ensemble mean from member
+prepare <- function(x) {
+    x.mean <- x %>% apply(1, mean)
+    return(x - x.mean)
+}
+
+# calculate 20-year rolling variance
+rollvar <- function(x) {
+  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
+  return(out)
+}
+
+# load variance from ensemble file
+load_variance <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% prepare %>% apply(2, rollvar))
+
+se <- function(x, ...) sd(x, ...)/sqrt(length(x))
+
+# bind ensemble statistics into dataframe while preserving date offset
+bind_stats <- function(this.date, this.mean, this.se, date.start = NA, date.end = NA) {
+    if(length(this.mean) == length(this.date)) {
+        this.df <- data.frame(date = this.date, mean = this.mean, se = this.se)
+    } else {
+        if(is.na(date.end)){
+            date.end <- length(this.date)
+        }
+        if(is.na(date.start)){
+            date.start <- 0
+        }
+        this.df <- data.frame(date = this.date, mean = c(
+                                                    rep(NA, date.start),
+                                                    this.mean,
+                                                    rep(NA, (length(this.date) - date.end))),
+                              se =
+                                  c(rep(NA, date.start),
+                                    this.se,
+                                    rep(NA, (length(this.date) - date.end))))
+        writeLines(as.character(date.start))
+        writeLines(as.character(date.end))
+        return(this.df)
+    }
+}
+
+# Produce ensemble stats dataframe
+make_df <- function(this, this.date, date.start = NA, date.end = NA) {
+    this.mean <- this %>% apply(1, mean)
+    this.se <- this %>% apply(1, se)
+    writeLines(as.character(length(this.mean)))
+    this.df <- bind_stats(this.date, this.mean, this.se, date.start, date.end)
+    return(this.df)
+}
+
+# calculate standard error in bootstrap process setting
+se_bs <- function(x, n) {return(sd(x)/sqrt(n))}
+
+# calculate single forcing impact through repeated random sampling
[…]

### 2021-08-09 data/varcalc.r

@@ -1,0 +1,143 @@
+# Imports
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+library(WaveletComp)
+library(reshape2)
+
+# subtract ensemble mean from member
+prepare <- function(x) {
+    x.mean <- x %>% apply(1, mean)
+    return(x - x.mean)
+}
+
+# calculate 20-year rolling variance
+rollvar <- function(x) {
+  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
+  return(out)
+}
+
+# load variance from ensemble file
+load_variance <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% prepare %>% apply(2, rollvar))
+
+se <- function(x, ...) sd(x, ...)/sqrt(length(x))
+
+# bind ensemble statistics into dataframe while preserving date offset
+bind_stats <- function(this.date, this.mean, this.se, date.start = NA, date.end = NA) {
+    if(length(this.mean) == length(this.date)) {
+        this.df <- data.frame(date = this.date, mean = this.mean, se = this.se)
+    } else {
+        if(is.na(date.end)){
+            date.end <- length(this.date)
+        }
+        if(is.na(date.start)){
+            date.start <- 0
+        }
+        this.df <- data.frame(date = this.date, mean = c(
+                                                    rep(NA, date.start),
+                                                    this.mean,
+                                                    rep(NA, (length(this.date) - date.end))),
+                              se =
+                                  c(rep(NA, date.start),
+                                    this.se,
+                                    rep(NA, (length(this.date) - date.end))))
+        writeLines(as.character(date.start))
+        writeLines(as.character(date.end))
+        return(this.df)
+    }
+}
+
+# Produce ensemble stats dataframe
+make_df <- function(this, this.date, date.start = NA, date.end = NA) {
+    this.mean <- this %>% apply(1, mean)
+    this.se <- this %>% apply(1, se)
[…]

### 2021-08-09 articles/notes/notes.tex
removed archive to separate database, reorganized some stuff

% Created 2021-07-13 Tue 22:21
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\usepackage{enumitem}
\usepackage{natbib}
\setlist{noitemsep}
\author{Benjamin Goldman}
\date{\today}
\title{Notes}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Notes},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle
\tableofcontents

\section{\cite{an2017feedback}}
\label{sec:org3655487}

\begin{itemize}
\item Used SVD (Singular Value Decomposition) together with the Mixed Layer Heat Budget Analysis to look at which feedbacks contributed most to ENSO's variation between models
\item Influence of thermocline feedback is determined by how strongly equatorial horizontal winds affect the slope of the thermocline.
\end{itemize}

\section{\cite{bjerknes1969atmospheric}}
\label{sec:org136b470}

\begin{itemize}
\item First big paper on ENSO having a big impact
\item connected changes in ocean currents to Walker Circulation
\item ENSO phase affects behavior of the Indian Ocean monsoon.
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{boer2000transient}}
\label{sec:org04662d3}

\section{\cite{cai2018increased}}
\label{sec:org328012a}

\begin{itemize}
\item Increased ENSO variance in most CMIP5 models in EP ENSO center.
\item Likely caused by greenhouse gases
\item Higher ocean stratification allows for stronger communication between atmospheric and oceanic temperatures.
\item Used EOF analysis.
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{chen2015causes}}
\label{sec:orgd3d49d0}

\section{\cite{chen2017possible}}
\label{sec:org21b124c}

\begin{itemize}
\item Models are disagreeing on ENSO in the future because they have different representations of the mechanics and mean state of the Pacific subtropical cell
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{deser2020isolating}}
\label{sec:org43ffddf}

\begin{itemize}
\item Main documentation for CESM1 Single Forcing Ensemble
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{dewitte2012reinterpreting}}
\label{sec:org2367605}

\section{\cite{emile2007nino}}
\label{sec:orgbb125ce}

\begin{itemize}
\item Analyzed wavelet power spectrum of ENSO variability in models forced by sunspot and orbital changes
\item Orbital changes increase long-term ENSO variability
\item It is possible that ENSO was the mechanic that allowed prehistoric solar/orbital changes to control the earth's climate
\end{itemize}

\section{\cite{graham2014effectiveness}}
\label{sec:org7d28ecd}

\begin{itemize}
\item tested how accurate the Bjerknes Stability Index is at measuring the mechanics of ENSO in a couple models
\item BJ index overestimates the importance of the Thermocline feedback.
\item BJ index assumes that terms should be linear when combined, but they actually aren't.
\end{itemize}

\section{\cite{hu2018cross}}
\label{sec:org9e81eeb}
\section{\cite{jia2019weakening}}
\label{sec:orgc220f64}
\section{{\bfseries\sffamily TODO} \cite{jimenez2016record}}
\label{sec:org422a8cd}
\section{\cite{kay2015community}}
\label{sec:org5b0d352}
\section{\cite{kestin1998time}}
\label{sec:orgffee63b}
\section{{\bfseries\sffamily TODO} \cite{kim2014response}}
\label{sec:org1250259}
\section{{\bfseries\sffamily TODO} \cite{kohyama2018weakening}}
\label{sec:orgf9af32a}
\section{\cite{levine2017impact}}
\label{sec:org4982a37}
\section{\cite{lorenz1963deterministic}}
\label{sec:org5b50a4f}
\section{{\bfseries\sffamily TODO} \cite{liu2007atmospheric}}
\label{sec:org4049c31}
\section{\cite{lubbecke2014assessing}}
\label{sec:org7b0850f}
\section{\cite{maher2018enso}}
\label{sec:orgdd094cf}
\section{\cite{nowack2017role}}
\label{sec:org9238fee}
\section{\cite{phillips2014evaluating}}
\label{sec:org830cbdd}
\section{{\bfseries\sffamily TODO} \cite{rashid2016atmospheric}}
\label{sec:orge93d755}
\section{{\bfseries\sffamily TODO} \cite{ropelewski1987global}}
\label{sec:org47cae0c}
\section{{\bfseries\sffamily TODO} \cite{son2010impact}}
\label{sec:orga87c549}
\section{{\bfseries\sffamily TODO} \cite{stevenson2010enso}}
\label{sec:org6de73f7}
\section{{\bfseries\sffamily TODO} \cite{stevenson2012will}}
\label{sec:org697dce3}
\section{\cite{stevenson2012significant}}
\label{sec:orge1990c0}
\section{\cite{stevenson2017forced}}
\label{sec:orgf634322}

\section{\cite{torrence1998practical}}
\label{sec:org2e0f412}

\begin{itemize}
\item How to use wavelets to estimate power spectrum in timeseries.
\item Uses ENSO data \emph{very niiceee}
\item Windowed Fourier Transform sucks butt because it is dependent on a time step parameter that can muck with the results depending on which value you choose.
\item A wavelet is a short \textbf{\emph{blirp}} of a wave with a mean of zero and finite amplitude/frequency and limited time domain.
\item To get an ex. Morlet Wavelet take a regular wave and multiply it by a Gaussian (normal bell curve) so that it drops off over time.
\item Will be using continuous methods, but discrete also works.
\item Use mathematical transforms to vary scale and translation of wavelet as it slides across the time series.
\item Integrate wavelet multiplied by the timeseries while varying scale and shift to generate a power spectrum.
\item Applied wavelet spectrum analysis to Nino 3 timeseries
\item strong variance in 2-8 year frequency area, but with slight changes between 1900 and 1990
\item However, results are highly dependent on which mother wavelet you choose because they all have quite different properties.
\item Trying power spectrum from a DOG (Mexican Hat) wavelet gives overall similar answer as Morelett wavelet, but it is slightly different (more detailed in time, less detailed in frequency.)
\item Use formula to pick scale limits
\item Add zeroes around the timeseries so that the wavelet equation does not misunderstand the data by thinking it is cyclical
\item Create a cone of influence to mark where the edge confusion is able to interfere with the results.
\item Make sure you convert between the wavelet scale to the Fourier period when you make your axes
\item You can also reverse the wavelet transform to get back the timeseries from the power chart if you really want to (I dont think I will).
\item Time for significance analysis!
\item take a background spectrum that serves as the null hypothesis: all spikes in the power spectrum are due to chance, the underlying signal is really random.
\item Comparing to red noise shows that the peaks of ENSO in 2-8 years are statistically significant
\item Calculate 95\% confidence interval by taking 95\% confidence \(\chi^2\) statistic and multiplying by red noise spectrum.
\item Nino3 SST wavelet power from 2-8year frequency is sometimes significantly different from red noise expectations.
\item ``The confidence interval is defined as the probability that the true wavelet power at a certain time and scale lies within a certain interval about the estimated wavelet power.''
\item \(\chi^2\) test is advantageous because it applies to a lot of situations in wavelet analysis.
\item Averaging the wavelet spectrum across the whole time range gives the overall power spectrum which can be significance tested and approximates the Fourier spectrum.
\item Smoothing/averaging increases DOF, allowing to greater significance for the peaks
\item After that, only main ENSO frequency band is shown to be statistically significant.
\item Similar to time averaging, scale averaging is sometimes a good idea
\item Wavelet analysis can be used to denoise an image/timeseries by throwing away the zones who's amplitude does not meet a certain level of significance.
\item Wavelet analysis across spatial and temporal domains when squashed by frequency allows for a great analysis of spatial and temporal variability.
\end{itemize}

\section{{\bfseries\sffamily TODO} \cite{vecchi2006weakening}}
\label{sec:orgb996a82}
\section{\cite{vega2017analysis}}
\label{sec:orga1fc1a2}
\section{{\bfseries\sffamily TODO} \cite{wang2016nino}}
\label{sec:orge694acd}
\section{{\bfseries\sffamily TODO} \cite{yeo2016role}}
\label{sec:orgeaa77a3}
\section{{\bfseries\sffamily TODO} \cite{zhang2019review}}
\label{sec:org471ee8c}
\section{{\bfseries\sffamily TODO} \cite{zheng2016intermodel}}
\label{sec:org1371712}
\section{\cite{zheng2017response}}
\label{sec:org013bb4e}

\bibliographystyle{apalike}
\bibliography{references}
\end{document}

### 2021-08-09 data/archive/Code_old/nino_contour.py

@@ -1,48 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Sun May 17 11:51:06 2020
-
-@author: bengoldman
-"""
-import netCDF4 as nc
-import numpy as np
-import matplotlib.pyplot as plt
-from eofs.standard import Eof
-from matplotlib import cm
-import cartopy.crs as ccrs
-import cartopy
-
-# %%
-f1 = nc.Dataset("/Volumes/Extreme SSD/science_research/obs_data/HadISST_sst.nc")
-
-sst1 = f1.variables["sst"][:, 59:120, 300:359]
-sst2 = f1.variables["sst"][:, 59:120, 0:100]
-sst = np.concatenate((sst1, sst2), axis=2)
-
-lats = np.flip(f1.variables["latitude"][59:120])
-lons1 = f1.variables["longitude"][300:359]
-lons2 = f1.variables["longitude"][0:100] + 360
-lons = np.concatenate((lons1, lons2), axis=0)
-
-# %%
-sst[sst == - 1e+30] = np.nan
-
-solver = Eof(sst)
-eof1 = np.flip(solver.eofsAsCorrelation(neofs=2)[1], axis=0)
-
-# %%
-
-# %%
-proj = ccrs.PlateCarree(central_longitude=180)
-
-fig = plt.figure(figsize=(15, 5))
-ax = fig.add_subplot(111, projection=proj)
-
-ax.contourf(lons, lats, eof1, cmap="coolwarm", transform=ccrs.PlateCarree(central_longitude=0))
-
-ax.coastlines()
-ax.add_feature(cartopy.feature.LAND, edgecolor='black')
-
-fig.colorbar(cm.ScalarMappable(cmap="coolwarm"), ax=ax)
-fig.suptitle("ENSO Temperature Anomaly Pattern", y=.91)

### 2021-08-09 data/archive/Code_old/openCVDP.py

@@ -1,29 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Fri May 22 11:00:38 2020
-
-@author: bengoldman
-"""
-
-import numpy as np
-import netCDF4 as nc
-import os
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy import stats
-from math import sqrt
-from random import randint
-
-
-# %%
-
-
-os.chdir('/Users/bengoldman/ENSO-amplitude/Data')
-
-ctl_cvdp = nc.Dataset("/Volumes/Extreme SSD/science_research/cvdp_data/CESM1_LENS_Coupled_Control.cvdp_data.401-2200.nc")
-ctl_amoc = ctl_cvdp.variables["amoc_timeseries_ann"][:]
-ctl_amo = ctl_cvdp.variables["amo_timeseries_mon"][:]
-
-np.savetxt("ctl_amoc.csv", ctl_amoc)
-np.savetxt("ctl_amo.csv", ctl_amo)

### 2021-08-09 data/archive/Code_old/open_nino.py

@@ -1,209 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Sun Apr 26 16:12:51 2020
-
-@author: bengoldman
-"""
-
-# %%
-import numpy as np
-import netCDF4 as nc
-import os
-
-
-lat1 = 90
-lat2 = 101
-lon1 = 152
-lon2 = 192
-
-# %%
-os.chdir("/Volumes/Extreme SSD/science_research/large_ensemble/CESM_LE/TREFHT/")
-
-# Full forcing set
-ff_set = np.zeros((2172, 34))
-print("")
-print("Full forcing")
-for i in range(2, 36):
-    if i < 10:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.cam.h0.TREFHT.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.cam.h0.TREFHT.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.cam.h0.TREFHT.208101-210012.nc" % i)
-
-        ff_set[:, i-2] = np.concatenate((
-            np.average(f1.variables["TREFHT"][:, lat1:lat2, lon1:lon2], axis=(1, 2)),
-            np.average(f2.variables["TREFHT"][:, lat1:lat2, lon1:lon2], axis=(1, 2)),
-            np.average(f3.variables["TREFHT"][:, lat1:lat2, lon1:lon2], axis=(1, 2))
-            )) - 273
-        f1.close()
-        f2.close()
-        f3.close()
-    elif i < 34:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.cam.h0.TREFHT.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.cam.h0.TREFHT.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.cam.h0.TREFHT.208101-210012.nc" % i)
-
-        ff_set[:, i-2] = np.concatenate((
-            np.average(f1.variables["TREFHT"][:, lat1:lat2, lon1:lon2], axis=(1, 2)),
[…]

### 2021-08-09 data/archive/Code_old/open_tempcep.py

@@ -1,143 +1,0 @@
-# %%
-import numpy as np
-import netCDF4 as nc
-import os
-
-os.chdir("/Volumes/Extreme SSD/LE/TEMPCEP/ff/")
-
-# %%
-ff_tempcep_set = np.zeros((34, 2172, 60, 41))
-for i in range(2, 36):
-    print(i, end="|")
-    if i < 10:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.00%s.pop.h.TEMPCEP.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPCEP.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.00%s.pop.h.TEMPCEP.208101-210012.nc" % i)
-
-        v1 = f1.variables["TEMPCEP"][:, :, :]
-        v2 = f2.variables["TEMPCEP"][:, :, :]
-        v3 = f3.variables["TEMPCEP"][:, :, :]
-
-        v1[v1.mask] = np.nan
-        v2[v2.mask] = np.nan
-        v3[v3.mask] = np.nan
-        ff_tempcep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
-
-        f1.close()
-        f2.close()
-        f3.close()
-
-    elif i < 34:
-        f1 = nc.Dataset("b.e11.B20TRC5CNBDRD.f09_g16.0%s.pop.h.TEMPCEP.192001-200512.nc" % i)
-        f2 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPCEP.200601-208012.nc" % i)
-        f3 = nc.Dataset("b.e11.BRCP85C5CNBDRD.f09_g16.0%s.pop.h.TEMPCEP.208101-210012.nc" % i)
-
-
-        v1 = f1.variables["TEMPCEP"][:, :, :]
-        v2 = f2.variables["TEMPCEP"][:, :, :]
-        v3 = f3.variables["TEMPCEP"][:, :, :]
-        
-        v1[v1.mask] = np.nan
-        v2[v2.mask] = np.nan
-        v3[v3.mask] = np.nan
-        ff_tempcep_set[i-2, :, :, :] = np.concatenate((v1, v2, v3))
-
-        f1.close()
[…]

### 2021-08-09 data/archive/Code_old/open_xsections.py

@@ -1,98 +1,0 @@
-from helpers import *
-import feather as ft
-
-# %%
-print("ff tempdt")
-ff_tempdt = get_ensemble("/Volumes/Extreme SSD/LE/TEMPDT/ff", 0, "TEMPDT")
-ff_tempdt_avg = np.average(ff_tempdt, axis=0)
-del ff_tempdt
-os.chdir("/Volumes/Extreme SSD/LE/compiled/")
-np.save("ff_tempdt_avg", ff_tempdt_avg)
-
-# %%
-print("ff tempcep")
-ff_tempcep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPCEP/ff", 0, "TEMPCEP")
-ff_tempcep_avg = np.average(ff_tempcep, axis=0)
-del ff_tempcep
-os.chdir("/Volumes/Extreme SSD/LE/compiled/")
-np.save("ff_tempcep_avg", ff_tempcep_avg)
-
-# %%
-print("ff tempwep")
-ff_tempwep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPWEP/ff", 0, "TEMPWEP")
-ff_tempwep_avg = np.average(ff_tempwep, axis=0)
-del ff_tempwep
-os.chdir("/Volumes/Extreme SSD/LE/compiled/")
-np.save("ff_tempwep_avg", ff_tempwep_avg)
-
-# %%
-print("ff tempeep")
-ff_tempeep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPEEP/ff", 0, "TEMPEEP")
-ff_tempeep_avg = np.average(ff_tempeep, axis=0)
-del ff_tempeep
-os.chdir("/Volumes/Extreme SSD/LE/compiled/")
-np.save("ff_tempeep_avg", ff_tempeep_avg)
-
-# %%
-print("ghg tempdt")
-ghg_tempdt = get_ensemble("/Volumes/Extreme SSD/LE/TEMPDT/xghg", 1, "TEMPDT")
-ghg_tempdt_avg = np.average(ghg_tempdt, axis=0)
-del ghg_tempdt
-os.chdir("/Volumes/Extreme SSD/LE/compiled/")
-np.save("ghg_tempdt_avg", ghg_tempdt_avg)
-
-# %%
-print("ghg tempcep")
-ghg_tempcep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPCEP/xghg", 1, "TEMPCEP")
[…]

### 2021-08-09 data/archive/Code_old/pptfigs.py

@@ -1,16 +1,0 @@
-from helpers import *
-
-import cartopy.crs as ccrs
-import cartopy
-import matplotlib.patches as mpatches
-
-fig = plt.figure(figsize=(4, 4))
-ax = fig.add_subplot(projection=ccrs.PlateCarree(central_longitude=180))
-ax.coastlines()
-ax.gridlines()
-ax.add_patch(mpatches.Rectangle(xy=[-170, -5], width=50, height=10, color="C0", transform=ccrs.PlateCarree()))
-ax.set_xlim((-75, 100))
-ax.set_ylim((-80, 80))
-plt.tight_layout()
-plt.savefig("../Figures/nino.pdf")
-

### 2021-08-09 data/archive/Code_old/tempcep.py

@@ -1,191 +1,0 @@
-import sys
-
-sys.path.append("./")
-from helpers import *
-from variance import ff_var_mean, ghg_var_mean, aer_var_mean
-
-# %%
-ff_tempcep_avg = np.load("../Data/ff_tempcep_average.npy")
-ff_tempcep_avg_det = ff_tempcep_avg
-
-ff_correlation = np.zeros(ff_tempcep_avg_det[0].shape)
-ff_p_vals = np.zeros(ff_tempcep_avg_det[0].shape)
-ff_var_mean_det = ff_var_mean
-signal.detrend(ff_var_mean_det[~np.isnan(ff_var_mean_det)], overwrite_data=True)
-ff_nino_ts = pd.DataFrame(ff_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
-for i in range(len(ff_tempcep_avg_det[0, :, 0])):
-    for j in range(len(ff_tempcep_avg_det[0, i, :])):
-        print(str(i) + ":" + str(j) )
-        if np.isnan(ff_tempcep_avg_det[200, i, j]):
-            r, p = (np.nan, np.nan)
-        else:
-            signal.detrend(ff_tempcep_avg_det[:, i, j], overwrite_data=True)
-            ff_tempcep_avg_det[:, i, j] = pd.DataFrame(ff_tempcep_avg_det[:, i, j]).rolling(360, center=True).mean().to_numpy()[:, 0]
-            r, p = stats.pearsonr(ff_tempcep_avg_det[300:-300, i, j], ff_nino_ts[300:-300])
-        ff_correlation[i, j], ff_p_vals[i, j] = r, p
-
-ghg_tempcep_avg = np.load("../Data/ghg_tempcep_average.npy")
-ghg_tempcep_avg_det = ghg_tempcep_avg
-
-ghg_correlation = np.zeros(ghg_tempcep_avg_det[0].shape)
-ghg_p_vals = np.zeros(ghg_tempcep_avg_det[0].shape)
-ghg_var_mean_det = ghg_var_mean
-signal.detrend(ghg_var_mean_det[~np.isnan(ghg_var_mean_det)], overwrite_data=True)
-ghg_nino_ts = pd.DataFrame(ghg_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
-for i in range(len(ghg_tempcep_avg_det[0, :, 0])):
[…]

### 2021-08-09 data/archive/Code_old/tempdt.py

@@ -1,164 +1,0 @@
-from helpers import *
-from variance import ff_var_mean, ghg_var_mean, aer_var_mean
-
-ff_tempdt_avg = np.load("../Data/ff_tempdt_average.npy")
-ff_tempdt_avg_det = ff_tempdt_avg
-
-ff_correlation = np.zeros(ff_tempdt_avg_det[0].shape)
-ff_p_vals = np.zeros(ff_tempdt_avg_det[0].shape)
-ff_var_mean_det = ff_var_mean
-signal.detrend(ff_var_mean_det[~np.isnan(ff_var_mean_det)], overwrite_data=True)
-ff_nino_ts = pd.DataFrame(ff_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
-for i in range(len(ff_tempdt_avg_det[0, :, 0])):
-    for j in range(len(ff_tempdt_avg_det[0, i, :])):
-        print(str(i) + ":" + str(j) )
-        if np.isnan(ff_tempdt_avg_det[200, i, j]):
-            r, p = (np.nan, np.nan)
-        else:
-            signal.detrend(ff_tempdt_avg_det[:, i, j], overwrite_data=True)
-            ff_tempdt_avg_det[:, i, j] = pd.DataFrame(ff_tempdt_avg_det[:, i, j]).rolling(360, center=True).mean().to_numpy()[:, 0]
-            r, p = stats.pearsonr(ff_tempdt_avg_det[300:-300, i, j], ff_nino_ts[300:-300])
-        ff_correlation[i, j], ff_p_vals[i, j] = r, p
-
-ghg_tempdt_avg = np.load("../Data/ghg_tempdt_average.npy")
-ghg_tempdt_avg_det = ghg_tempdt_avg
-
-ghg_correlation = np.zeros(ghg_tempdt_avg_det[0].shape)
-ghg_p_vals = np.zeros(ghg_tempdt_avg_det[0].shape)
-ghg_var_mean_det = ghg_var_mean
-signal.detrend(ghg_var_mean_det[~np.isnan(ghg_var_mean_det)], overwrite_data=True)
-ghg_nino_ts = pd.DataFrame(ghg_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
-for i in range(len(ghg_tempdt_avg_det[0, :, 0])):
-    for j in range(len(ghg_tempdt_avg_det[0, i, :])):
[…]

### 2021-08-09 data/archive/Code_old/tempeep.py

@@ -1,190 +1,0 @@
-
-import sys
-
-sys.path.append("./")
-from helpers import *
-from variance import ff_var_mean, ghg_var_mean, aer_var_mean
-
-ff_tempeep_avg = np.load("../Data/ff_tempeep_average.npy")
-ff_tempeep_avg_det = ff_tempeep_avg
-
-ff_correlation = np.zeros(ff_tempeep_avg_det[0].shape)
-ff_p_vals = np.zeros(ff_tempeep_avg_det[0].shape)
-ff_var_mean_det = ff_var_mean
-signal.detrend(ff_var_mean_det[~np.isnan(ff_var_mean_det)], overwrite_data=True)
-ff_nino_ts = pd.DataFrame(ff_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
-for i in range(len(ff_tempeep_avg_det[0, :, 0])):
-    for j in range(len(ff_tempeep_avg_det[0, i, :])):
-        print(str(i) + ":" + str(j) )
-        if np.isnan(ff_tempeep_avg_det[200, i, j]):
-            r, p = (np.nan, np.nan)
-        else:
-            signal.detrend(ff_tempeep_avg_det[:, i, j], overwrite_data=True)
-            ff_tempeep_avg_det[:, i, j] = pd.DataFrame(ff_tempeep_avg_det[:, i, j]).rolling(360, center=True).mean().to_numpy()[:, 0]
-            r, p = stats.pearsonr(ff_tempeep_avg_det[300:-300, i, j], ff_nino_ts[300:-300])
-        ff_correlation[i, j], ff_p_vals[i, j] = r, p
-
-ghg_tempeep_avg = np.load("../Data/ghg_tempeep_average.npy")
-ghg_tempeep_avg_det = ghg_tempeep_avg
-
-ghg_correlation = np.zeros(ghg_tempeep_avg_det[0].shape)
-ghg_p_vals = np.zeros(ghg_tempeep_avg_det[0].shape)
-ghg_var_mean_det = ghg_var_mean
-signal.detrend(ghg_var_mean_det[~np.isnan(ghg_var_mean_det)], overwrite_data=True)
-ghg_nino_ts = pd.DataFrame(ghg_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
-for i in range(len(ghg_tempeep_avg_det[0, :, 0])):
[…]

### 2021-08-09 miscellaneous/Summer2020/UHSFiles/Code/20year_variance.py

@@ -1,726 +1,0 @@
-#!/usr/bin/env python3
-# -*- coding: utf-8 -*-
-"""
-Created on Mon Apr 27 14:01:34 2020
-
-@author: bengoldman
-"""
-# %%
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy import stats, signal
-from matplotlib import gridspec
-from math import sqrt
-from random import randint
-import os
-from matplotlib import cm
-
-from helpers import *
-
-# %%
-
-os.chdir('/Users/bengoldman/ENSO-amplitude/Data')
-ctl = np.loadtxt("ctl.csv")
-ff_set = np.loadtxt("ff_set.csv")
-ff_1850 = np.loadtxt("ff_1850.csv")
-ghg_set = np.loadtxt("ghg_set.csv")
-aer_set = np.loadtxt("aer_set.csv")
-bmb_set = np.loadtxt("bmb_set.csv")
-luc_set = np.loadtxt("luc_set.csv")
-ozo_set = np.loadtxt("ozo_set.csv")
-oic_set = np.loadtxt("oic_set.csv")
-oic_1850 = np.loadtxt("oic_1850.csv")
-ctl_amoc = np.loadtxt("ctl_amoc.csv")
-ctl_amo = np.loadtxt("ctl_amo.csv")
-ff_tempdt_avg = np.load("/Users/bengoldman/ENSO-amplitude/Data/ff_tempdt_average.npy")
-lat = np.loadtxt("lat.csv")
-lon = np.loadtxt("lon.csv")
-lon1 = np.loadtxt("lon1.csv")
-depth = np.loadtxt("depth.csv")
-
-os.chdir('/Users/bengoldman/ENSO-amplitude/Figures')
-
-# %%
-ctl_var = pd.DataFrame(ctl).rolling(240, center=True).var().to_numpy()[:, 0]
-ff_var_set = pd.DataFrame(ff_set).rolling(240, center=True).var().to_numpy()
-ff_1850_var = pd.DataFrame(ff_1850).rolling(240, center=True).var().to_numpy()
-ghg_var_set = pd.DataFrame(ghg_set).rolling(240, center=True).var().to_numpy()
-aer_var_set = pd.DataFrame(aer_set).rolling(240, center=True).var().to_numpy()
[…]

### 2021-08-09 miscellaneous/Summer2020/UHSFiles/Code/control_variance.py

@@ -1,62 +1,0 @@
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy import stats, signal
-from matplotlib import gridspec
-from math import sqrt
-from random import randint
-import os
-from matplotlib import cm
-
-from helpers import *
-
-ctl_var_nan = ctl_var[~np.isnan(ctl_var)]
-
-x = np.linspace(0.7, 2, 500)
-
-ctl_var_div_20 = np.zeros((1000, 20))
-for i in range(20):
-    ctl_var_div_20[:, i] = ctl_var_nan[i*1000:(i+1)*1000]
-
-ctl_var_div_10 = np.zeros((2000, 10))
-for i in range(10):
-    ctl_var_div_10[:, i] = ctl_var_nan[i*2000:(i+1)*2000]
-
-ctl_var_div_5 = np.zeros((4000, 5))
-for i in range(5):
-    ctl_var_div_5[:, i] = ctl_var_nan[i*4000:(i+1)*4000]
-
-ctl_var_div_20_pdf = np.zeros((500, 20))
-for i in range(20):
-    ctl_var_div_20_pdf[:, i] = stats.gaussian_kde(ctl_var_div_20[:, i]).evaluate(x)
-
-ctl_var_div_10_pdf = np.zeros((500, 10))
-for i in range(10):
-    ctl_var_div_10_pdf[:, i] = stats.gaussian_kde(ctl_var_div_10[:, i]).evaluate(x)
-
-ctl_var_div_5_pdf = np.zeros((500, 5))
-for i in range(5):
-    ctl_var_div_5_pdf[:, i] = stats.gaussian_kde(ctl_var_div_5[:, i]).evaluate(x)
-
-ctl_var_pdf = stats.gaussian_kde(ctl_var_nan).evaluate(x)
-
-# %%
-gs = gridspec.GridSpec(20, 3)
-
-fig = plt.figure(figsize=(10, 10))
-
-for i in range(20):
-    ax = fig.add_subplot(gs[i, 0])
-    ax.plot(x, ctl_var_div_20_pdf[:, i])
-#    ax.set_xticklabels([])
-
-for i in range(10):
-    ax = fig.add_subplot(gs[2*i:2*i+2, 1])
-    ax.plot(x, ctl_var_div_10_pdf[:, i])
-
-for i in range(5):
-    ax = fig.add_subplot(gs[4*i:4*i+4, 2])
[…]

### 2021-08-09 miscellaneous/Summer2020/UHSFiles/Code/correlation_xsections.py

@@ -1,79 +1,0 @@
-from helpers import *
-from variance import ff_var_mean, ghg_var_mean, aer_var_mean
-
-# %%
-print("ff")
-ff_tempdt = np.load("../Data/ff_tempdt_avg.npy")
-ff_tempdt_rs, ff_tempdt_ps = correlation_detrended(ff_tempdt, ff_var_mean)
-
-# %%
-print("ff")
-ff_tempcep = np.load("../Data/ff_tempcep_avg.npy")
-ff_tempcep_rs, ff_tempcep_ps = correlation_detrended(ff_tempcep, ff_var_mean)
-
-# %%
-print("ff")
-ff_tempwep = np.load("../Data/ff_tempwep_avg.npy")
-ff_tempwep_rs, ff_tempwep_ps = correlation_detrended(ff_tempwep, ff_var_mean)
-
-# %%
-print("ff")
-ff_tempeep = np.load("../Data/ff_tempeep_avg.npy")
-ff_tempeep_rs, ff_tempeep_ps = correlation_detrended(ff_tempeep, ff_var_mean)
-
-# %%
-print("ghg")
-ghg_tempdt = np.load("../Data/ghg_tempdt_avg.npy")
-ghg_tempdt_rs, ghg_tempdt_ps = correlation_detrended(ghg_tempdt, ghg_var_mean)
-
-# %%
-print("ghg")
-ghg_tempcep = np.load("../Data/ghg_tempcep_avg.npy")
-ghg_tempcep_rs, ghg_tempcep_ps = correlation_detrended(ghg_tempcep, ghg_var_mean)
-
-# %%
-print("ghg")
-ghg_tempwep = np.load("../Data/ghg_tempwep_avg.npy")
-ghg_tempwep_rs, ghg_tempwep_ps = correlation_detrended(ghg_tempwep, ghg_var_mean)
-
-# %%
-print("ghg")
-ghg_tempeep = np.load("../Data/ghg_tempeep_avg.npy")
-ghg_tempeep_rs, ghg_tempeep_ps = correlation_detrended(ghg_tempeep, ghg_var_mean)
-
-# %%
-print("aer")
-aer_tempdt = np.load("../Data/aer_tempdt_avg.npy")
-aer_tempdt_rs, aer_tempdt_ps = correlation_detrended(aer_tempdt, aer_var_mean)
-
-# %%
-print("aer")
[…]

### 2021-08-09 miscellaneous/Summer2020/UHSFiles/Code/correlations.py

@@ -1,10 +1,0 @@
-from helpers import *
-from variance import ff_var_set
-
-# %%
-os.chdir("/Volumes/Extreme SSD/LE/TEMPDT/ff")
-ff = get_ensemble(0, "TEMPDT", 2172)
-ff_correlation = np.zeros((34, 60, 180))
-
-# %%
-correlation = list(map(lambda x, y, z: correlation_detrended(ff_var_set[x, y, z], ff_var_set), ff))

### 2021-08-09 miscellaneous/Summer2020/UHSFiles/Code/fix_data.py

@@ -1,10 +1,0 @@
-import netCDF4 as nc
-import numpy as np
-import os
-
-os.chdir("/Volumes/Extreme SSD/LE/TEMPWEP/ff")
-for file in os.listdir():
-    f = nc.Dataset(file, mode="a")
-    del f.variables["TEMPWEP"].valid_range
-    print(f.variables["TEMPWEP"])
-    f.close()

### 2021-08-09 miscellaneous/Summer2020/UHSFiles/Code/helpers.py

@@ -1,223 +1,0 @@
-import netCDF4 as nc
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy import stats, signal
-from matplotlib import gridspec
-from math import sqrt
-from random import randint
-import os
-from matplotlib import cm, colors
-import re
-
-lat = np.loadtxt("../Data/lat.csv")
-lon = np.loadtxt("../Data/lon.csv")
-lon1 = np.loadtxt("../Data/lon1.csv")
-lat1 = np.loadtxt("../Data/lat1.csv")
-depth = np.loadtxt("../Data/depth.csv")/100
-
-date_2172 = np.linspace(1920, 2100, 2172)
-date_1872 = np.linspace(1850, 2005, 1872)
-date_1032 = np.linspace(1920, 2005, 1032)
-date_1932 = np.linspace(1920, 2080, 1932)
-date_1320 = np.linspace(1920, 2029, 1320)
-date_612 = np.linspace(1955, 2005, 612)
-date_960 = np.linspace(1920, 1999, 960)
-date_1800 = np.linspace(1850, 1999, 1800)
-date_80 = np.linspace(1920, 1999, 80)
-date_ctl_20600 = np.linspace(400, 2200, 21600)
-date_ctl_1800 = np.linspace(400, 2200, 1800)
-
-
-def envelope(ax, x, y, err, col):
-    ax.plot(x, y, color=col)
-    ax.fill_between(x, y-err, y+err, alpha=.2, color=col)
-
-
-def figlett(ax, a):
-    ax.text(0.01, 0.95, a, transform=ax.transAxes, va='top')
-
-size1 = (7, 4)
-size2 = (7, 8)
-mp = np.ones((400, 3))
-mp[21:40] = [.6, .6, .6]
-mp[-40:-21] = [.6, .6, .6]
-mp[0:20, 0] = np.linspace(0, 1, 20)
-mp[0:20, 1] = np.linspace(0, 1, 20)
-mp[0:20, 2] = 1
-mp[-21:-1, 0] =1
-mp[-21:-1, 1] = np.linspace(1, 0, 20)
-mp[-21:-1, 2] = np.linspace(1, 0, 20)
-cmap = colors.ListedColormap(mp)
-
-mp = np.zeros((2, 4))
-mp[1, :] = [1, 1, 1, 1]
-mp[0, :] = [1, 1, 1, 0]
[…]

### 2021-08-09 miscellaneous/Summer2020/UHSFiles/Code/open_xsections.py

@@ -1,98 +1,0 @@
-from helpers import *
-import feather as ft
-
-# %%
-print("ff tempdt")
-ff_tempdt = get_ensemble("/Volumes/Extreme SSD/LE/TEMPDT/ff", 0, "TEMPDT")
-ff_tempdt_avg = np.average(ff_tempdt, axis=0)
-del ff_tempdt
-os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
-np.save("ff_tempdt_avg", ff_tempdt_avg)
-
-# %%
-print("ff tempcep")
-ff_tempcep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPCEP/ff", 0, "TEMPCEP")
-ff_tempcep_avg = np.average(ff_tempcep, axis=0)
-del ff_tempcep
-os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
-np.save("ff_tempcep_avg", ff_tempcep_avg)
-
-# %%
-print("ff tempwep")
-ff_tempwep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPWEP/ff", 0, "TEMPWEP")
-ff_tempwep_avg = np.average(ff_tempwep, axis=0)
-del ff_tempwep
-os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
-np.save("ff_tempwep_avg", ff_tempwep_avg)
-
-# %%
-print("ff tempeep")
-ff_tempeep = get_ensemble("/Volumes/Extreme SSD/LE/TEMPEEP/ff", 0, "TEMPEEP")
-ff_tempeep_avg = np.average(ff_tempeep, axis=0)
-del ff_tempeep
-os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
-np.save("ff_tempeep_avg", ff_tempeep_avg)
-
-# %%
-print("ghg tempdt")
-ghg_tempdt = get_ensemble("/Volumes/Extreme SSD/LE/TEMPDT/xghg", 1, "TEMPDT")
-ghg_tempdt_avg = np.average(ghg_tempdt, axis=0)
-del ghg_tempdt
-os.chdir("/Users/bengoldman/ENSO-amplitude/Data/")
-np.save("ghg_tempdt_avg", ghg_tempdt_avg)
-
-# %%
-print("ghg tempcep")
[…]

### 2021-08-09 miscellaneous/Summer2020/UHSFiles/Code/tempwep.py

@@ -1,192 +1,0 @@
-import sys
-
-sys.path.append("./")
-from helpers import *
-from variance import ff_var_mean, ghg_var_mean, aer_var_mean
-
-ff_tempwep_avg = np.load("../Data/ff_tempwep_average.npy")
-ff_tempwep_avg_det = ff_tempwep_avg
-
-ff_correlation = np.zeros(ff_tempwep_avg_det[0].shape)
-ff_p_vals = np.zeros(ff_tempwep_avg_det[0].shape)
-ff_var_mean_det = ff_var_mean
-signal.detrend(ff_var_mean_det[~np.isnan(ff_var_mean_det)], overwrite_data=True)
-ff_nino_ts = pd.DataFrame(ff_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
-for i in range(len(ff_tempwep_avg_det[0, :, 0])):
-    for j in range(len(ff_tempwep_avg_det[0, i, :])):
-        print(str(i) + ":" + str(j) )
-        if np.isnan(ff_tempwep_avg_det[200, i, j]):
-            r, p = (np.nan, np.nan)
-        else:
-            signal.detrend(ff_tempwep_avg_det[:, i, j], overwrite_data=True)
-            ff_tempwep_avg_det[:, i, j] = pd.DataFrame(ff_tempwep_avg_det[:, i, j]).rolling(360, center=True).mean().to_numpy()[:, 0]
-            r, p = stats.pearsonr(ff_tempwep_avg_det[300:-300, i, j], ff_nino_ts[300:-300])
-        ff_correlation[i, j], ff_p_vals[i, j] = r, p
-
-ghg_tempwep_avg = np.load("../Data/ghg_tempwep_average.npy")
-ghg_tempwep_avg_det = ghg_tempwep_avg
-
-ghg_correlation = np.zeros(ghg_tempwep_avg_det[0].shape)
-ghg_p_vals = np.zeros(ghg_tempwep_avg_det[0].shape)
-ghg_var_mean_det = ghg_var_mean
-signal.detrend(ghg_var_mean_det[~np.isnan(ghg_var_mean_det)], overwrite_data=True)
-ghg_nino_ts = pd.DataFrame(ghg_var_mean_det).rolling(360, center=True).mean().to_numpy()[:, 0]
-for i in range(len(ghg_tempwep_avg_det[0, :, 0])):
[…]

### 2021-08-09 miscellaneous/Summer2020/UHSFiles/Code/variance.py

@@ -1,71 +1,0 @@
-
-import numpy as np
-import pandas as pd
-import matplotlib.pyplot as plt
-from scipy import stats, signal
-from math import sqrt
-from random import randint
-import os
-
-# %%
-
-os.chdir('/Users/bengoldman/ENSO-amplitude/Data')
-ctl = np.loadtxt("ctl.csv")
-ff_set = np.loadtxt("ff_set.csv")
-ff_1850 = np.loadtxt("ff_1850.csv")
-ghg_set = np.loadtxt("ghg_set.csv")
-aer_set = np.loadtxt("aer_set.csv")
-bmb_set = np.loadtxt("bmb_set.csv")
-luc_set = np.loadtxt("luc_set.csv")
-ozo_set = np.loadtxt("ozo_set.csv")
-oic_set = np.loadtxt("oic_set.csv")
-oic_1850 = np.loadtxt("oic_1850.csv")
-ctl_amoc = np.loadtxt("ctl_amoc.csv")
-ctl_amo = np.loadtxt("ctl_amo.csv")
-ff_tempdt_avg = np.load("/Users/bengoldman/ENSO-amplitude/Data/ff_tempdt_average.npy")
-lat = np.loadtxt("lat.csv")
-lon = np.loadtxt("lon.csv")
-lon1 = np.loadtxt("lon1.csv")
-depth = np.loadtxt("depth.csv")
-
-os.chdir('/Users/bengoldman/ENSO-amplitude/Figures')
-
-# %%
-ctl_var = pd.DataFrame(ctl).rolling(240, center=True).var().to_numpy()[:, 0]
-ff_var_set = pd.DataFrame(ff_set).rolling(240, center=True).var().to_numpy()
-ff_1850_var = pd.DataFrame(ff_1850).rolling(240, center=True).var().to_numpy()
-ghg_var_set = pd.DataFrame(ghg_set).rolling(240, center=True).var().to_numpy()
-aer_var_set = pd.DataFrame(aer_set).rolling(240, center=True).var().to_numpy()
-bmb_var_set = pd.DataFrame(bmb_set).rolling(240, center=True).var().to_numpy()
-luc_var_set = pd.DataFrame(luc_set).rolling(240, center=True).var().to_numpy()
-ozo_var_set = pd.DataFrame(ozo_set).rolling(240, center=True).var().to_numpy()
[…]

### 2021-08-11 data/correlation.r
very good pictures for presentation

@@ -4,3 +4,16 @@ library(dplyr)
 library(ggplot2)
 library(WaveletComp)
 library(reshape2)
+
+ncin <- nc_open("/Volumes/Extreme SSD/DATA/stacked/CESM1/TEMP/b.e11.BRCP85C5CNBDRD.f09_g16.all.pop.h.TEMPDT.192001-210012.nc")
+tempdt <- ncvar_get(ncin, "TEMPDT", start = c(21, 1, 1, 1), count = c(121, -1, -1, 1))
+lon <- ncvar_get(ncin, "lon", start = 21, count = 121)
+depth <- ncvar_get(ncin, "depth") / 100
+date <- ncvar_get(ncin, "time") / 365
+
+woohoo <- ncvar_get(ncin, "TEMPDT", start = c(21, 1, 1, 1), count = c(1, 1, -1, -1))
+
+dimnames(tempdt) <- list(lon, depth, date)
+
+ggplot(tempdt[,,4] %>% melt, aes(Var1, log(Var2), z = value)) +
+    geom_contour_filled(bins = 16)

### 2021-08-11 data/eof_fig.r

@@ -1,40 +1,0 @@
-library(ncdf4)
-library(zoo)
-library(dplyr)
-library(ggplot2)
-library(reshape2)
-
-ncin <- nc_open("/Volumes/Extreme SSD/DATA/HadISST_sst.nc")
-
-sst <- ncvar_get(ncin, "sst")
-sst[sst == -1000] <- NA
-
-lat <- ncvar_get(ncin, "latitude")
-lon <- ncvar_get(ncin, "longitude")
-
-rownames(sst) <- lon
-colnames(sst) <- lat
-
-sst_wrap <- array(NA, dim = dim(sst))
-## lon_wrap <- rep(NA, times = length(lon))
-## lon_wrap[0:211] <- lon[150:360]
-## lon_wrap[212:360] <- lon[1:149]
-lon_wrap <- lon + 180
-sst_wrap[0:180,,] <- sst[181:360,,]
-sst_wrap[181:360,,] <- sst[1:180,,]
-rownames(sst_wrap) <- lon_wrap
-colnames(sst_wrap) <- lat
-
-## NEED TO MAKE ANOMALY PLOT!
-
-sst_example <- apply(sst_wrap[,,1356:1360], c(1, 2), mean)
-sst_example2 <- apply(sst_wrap[,,1359 + 24 :1360 + 24], c(1, 2), mean)
-
-sst_melt <- sst_example2 %>% melt(varnames = c("longitude", "latitude"))
-ggplot(sst_melt %>% filter(latitude < 50 &
-                           latitude > -50 &
-                           longitude > 50 &
-                           longitude < 300), aes(longitude, latitude, z = value)) +
-    geom_contour_filled(bins = 16)
-
-#  %>% filter(latitude < 20 & latitude > -20),

### 2021-08-11 data/intro_fig.r

@@ -1,0 +1,99 @@
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+library(scico)
+
+ncin <- nc_open("/Volumes/Extreme SSD/DATA/HadISST_sst.nc")
+
+sst <- ncvar_get(ncin, "sst")
+sst[sst == -1000] <- NA
+
+date <- ncvar_get(ncin, "time")
+date <- date/365 + 1870
+
+lat <- ncvar_get(ncin, "latitude")
+lon <- ncvar_get(ncin, "longitude")
+
+rownames(sst) <- lon
+colnames(sst) <- lat
+
+sst_wrap <- array(NA, dim = dim(sst))
+## lon_wrap <- rep(NA, times = length(lon))
+## lon_wrap[0:211] <- lon[150:360]
+## lon_wrap[212:360] <- lon[1:149]
+lon_wrap <- lon + 180
+sst_wrap[0:180,,] <- sst[181:360,,]
+sst_wrap[181:360,,] <- sst[1:180,,]
+rownames(sst_wrap) <- lon_wrap
+colnames(sst_wrap) <- lat
+
+## NEED TO MAKE ANOMALY PLOT!
+monthly_avgs <- array(NA, dim = c(360, 180, 12))
+
+for (i in c(1:12)) {
+    writeLines(as.character(i))
+    monthly_avgs[,,i] <- sst_wrap[,,seq(i, 1800 + i, 12)] %>% apply(c(1,2), mean, na.rm = TRUE)
+}
+
+rownames(monthly_avgs) <- lon_wrap
+colnames(monthly_avgs) <- lat
+
+monthly_avgs_melt <- monthly_avgs %>% melt(varnames = c("longitude", "latitude", "month"))
+
+monthly_avgs_melt$month <- as.factor(monthly_avgs_melt$month)
+# levels = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Sep", "Oct", "Nov", "Dec"))
+
+sst_anom <- array(NA, dim = dim(sst_wrap))
+months <- rep(1:12, length.out = dim(sst_wrap)[3])
+
+for(t in 1:dim(sst_wrap)[3]) {
+    sst_anom[,,t] <- sst_wrap[,,t] - monthly_avgs[,,months[t]]
+}
+dimnames(sst_anom) <- list(lon_wrap, lat, date)
+
+
+
[…]

### 2021-08-12 presentations/presentation/presentation.tex
tried to make a figure for presentation
very good pictures for presentation
new intro figure, added intro slides

@@
 \renewcommand{\bibsection}{}
+\setbeamerfont{caption}{size=\tiny}
 
@@
 
+\newcommand{\myfig}[3]{
+  \begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{figures/#1}
+    \caption{#2}
+    \label{fig:#3}
+  \end{figure}
+}
+
 \begin{document}
@@
   \begin{columns}
-    \column{0.5\textwidth}
+    \column{0.4\textwidth}
     \begin{itemize}
-    \item Here we go...
-    \item \cite{bjerknes1969atmospheric}
+    \item The earth is getting warmer. \citep{pachauri2014climate}
+    \item Climate varies on different scales.
+    \item Long-term trends and short-term noise.
     \end{itemize}
-    \column{0.5\textwidth}
-    \begin{figure}
-      \centering
-      \includegraphics[width=\textwidth]{figures/example.jpg}
-      \caption{This is a very nice figure}
-      \label{fig:this}
-    \end{figure}
+    \column{0.6\textwidth}
+    \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{this}
   \end{columns}
@@
 \begin{frame}{El Niño}
-
+  \begin{figure}
+    \begin{columns}
+      \column{0.5\textwidth}
+      \begin{itemize}
+      \item Warming and cooling of the Pacific Ocean.
+      \item Affects human societies through temperature and rainfall. \citep{ropelewski1987global}
+      \item May be affected by climate change.
+      \end{itemize}
+      \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
+      \column{0.5\textwidth}
+      \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
+    \end{columns}
+  \end{figure}
 \end{frame}
@@
 \begin{frame}{ENSO in the Future}
-  Problem: What effect does global warming have on ENSO? Can we expect ENSO to become stronger, weaker, or neither?
 \end{frame}
@@
   \bibliographystyle{apalike}
+  \fontsize{7pt}{7.2}\selectfont
   \bibliography{references.bib}

### 2021-08-12 data/intro_fig.r

@@ -5,9 +5,9 @@ library(ggplot2)
 library(reshape2)
 library(scico)
 
-ncin <- nc_open("/Volumes/Extreme SSD/DATA/HadISST_sst.nc")
+ncin1 <- nc_open("/Volumes/Extreme SSD/DATA/HadISST_sst.nc")
 
-sst <- ncvar_get(ncin, "sst")
+sst <- ncvar_get(ncin1, "sst")
 sst[sst == -1000] <- NA
 
 date <- ncvar_get(ncin, "time")
@@ -97,3 +97,32 @@ ggplot(sst_anom_compare %>% filter(latitude < 60 &
          fill = " SST \n Anomaly \n (°C)")
 
 gsave("intro_fig_2.pdf")
+
+ncin2 <- nc_open("/Volumes/Extreme SSD/DATA/BEN/air.2x2.1200.mon.anom.land.nc")
+
+temp <- ncvar_get(ncin2, "air")
+
+avg_temp <- temp %>% apply(3, mean, na.rm = TRUE)
+
+index_1 <- data.frame(date = date[6:(1698-6)], air = avg_temp %>% rollapply(width = 12, mean))
+index_5 <- data.frame(date = date[(6 * 5):(1698 - 6 * 5)], air = avg_temp %>% rollapply(width = 12 * 5, mean))
+index_10 <- data.frame(date = date[(6 * 10):(1698 - 6 * 10)], air = avg_temp %>% rollapply(width = 12 * 10, mean))
+index_20 <- data.frame(date = date[(6 * 20):(1698 - 6 * 20)], air = avg_temp %>% rollapply(width = 12 * 20, mean))
+
+reg <- lm(air ~ date, index_1, y = TRUE)
+
+reg_vec <- data.frame(date = date, air = date * reg$coefficients[2] + reg$coefficients[1])
+
+temp_df <- index_1 %>% mutate(case = "1-Year") %>%
+    bind_rows(index_10 %>% mutate(case = "10-Year"))
+    #bind_rows(reg_vec %>% mutate(case = "Linear Model")) %>%
+    # bind_rows(index_20 %>% mutate(case = "20-Year Smoothing"))
+
+ggplot(temp_df, aes(date, air, color = case)) +
+    geom_path() +
+    labs(title = "Mean Land Air Temperature Anomaly",
+         x = "Date",
+         y = "Temperature Anomaly (°C)",
+         color = "Smoothing\nWindow Size")
+
+gsave("intro_fig_3.pdf", dimensions = c(8, 6))

### 2021-08-15 articles/notes/notes.tex
paper notes and article notes

@@
-% Created 2021-07-13 Tue 22:21
+% Created 2021-08-15 Sun 14:58
 % Intended LaTeX compiler: pdflatex
@@
 \usepackage{natbib}
+\usepackage[margin=1in]{geometry}
 \setlist{noitemsep}
@@
 \section{\cite{an2017feedback}}
-\label{sec:org3655487}
+\label{sec:org87b42f5}
 
@@
 \section{\cite{bjerknes1969atmospheric}}
-\label{sec:org136b470}
+\label{sec:org7b6b0b3}
 
@@
 \section{{\bfseries\sffamily TODO} \cite{boer2000transient}}
-\label{sec:org04662d3}
+\label{sec:org8596bf6}
 
 \section{\cite{cai2018increased}}
-\label{sec:org328012a}
+\label{sec:org6b26555}
 
@@
 \section{{\bfseries\sffamily TODO} \cite{chen2015causes}}
-\label{sec:orgd3d49d0}
+\label{sec:orgc84b0ce}
 
 \section{\cite{chen2017possible}}
-\label{sec:org21b124c}
+\label{sec:org826e065}
 
@@
 \section{{\bfseries\sffamily TODO} \cite{deser2020isolating}}
-\label{sec:org43ffddf}
+\label{sec:orgd7f2af0}
 
@@
 \section{{\bfseries\sffamily TODO} \cite{dewitte2012reinterpreting}}
-\label{sec:org2367605}
+\label{sec:org1549366}
 
 \section{\cite{emile2007nino}}
-\label{sec:orgbb125ce}
+\label{sec:org5ae4263}
 
@@
 
+\section{\cite{gistemp2019giss}}
+\label{sec:org3883bf1}
+
+\begin{itemize}
+\item citation for GISSTEMP dataset
+\end{itemize}
+
 \section{\cite{graham2014effectiveness}}
-\label{sec:org7d28ecd}
+\label{sec:orga71922b}
 
@@
 \section{\cite{hu2018cross}}
-\label{sec:org9e81eeb}
+\label{sec:org025902a}
+
+\begin{itemize}
+\item Observed and model data used
+\item Strengthening in cross-equatorial winds changes ITCZ position, weakening ENSO.
+\end{itemize}
+
 \section{\cite{jia2019weakening}}
-\label{sec:orgc220f64}
+\label{sec:org43889c8}
+
+\begin{itemize}
+\item My first paper!
+\item Atlantic Niño/Niña is correlated/causes opposite ENSO form
+\item Global warming is weakening this connection, impairing prediction efforts
+\item CESM5 coupled models used
+\item Correlation coefficient
+\end{itemize}
+
 \section{{\bfseries\sffamily TODO} \cite{jimenez2016record}}
-\label{sec:org422a8cd}
-\section{\cite{kay2015community}}
-\label{sec:org5b0d352}
+\label{sec:org518f64c}
+
+\section{{\bfseries\sffamily TODO} \cite{kay2015community}}
+\label{sec:orga1c53a6}
+
+\begin{itemize}
+\item CESM1 LE citation
+\end{itemize}
+
 \section{\cite{kestin1998time}}
-\label{sec:orgffee63b}
+\label{sec:org97d4f0c}
+
+\begin{itemize}
+\item apply wavelet and Fourier analysis to ENSO in observed data
+\item 2-3 year ENSO reduced from 1920-1960
+\item Stochastic model used to simulate ENSO agrees with analysis of observed data
+\item Concluded that models are reasonably accurate for simulating decadal ENSO variability
+\end{itemize}
+
 \section{{\bfseries\sffamily TODO} \cite{kim2014response}}
-\label{sec:org1250259}
+\label{sec:org4882548}
 \section{{\bfseries\sffamily TODO} \cite{kohyama2018weakening}}
-\label{sec:orgf9af32a}
+\label{sec:orgdf9c6cc}
+
 \section{\cite{levine2017impact}}
-\label{sec:org4982a37}
+\label{sec:orgea9bd4a}
+
+\begin{itemize}
+\item Observed ICOADS data
+\item Inverse relationship between AMO index and ENSO SPB strength
+\item Coupled model simulation
+\end{itemize}
+
 \section{\cite{lorenz1963deterministic}}
-\label{sec:org5b50a4f}
+\label{sec:org9ccfa1a}
+
+\begin{itemize}
+\item Nonlinear differential equations have solutions that must be solved numerically
+\item When graphed in phase space, these solutions form attractors that never repeat themselves
+\item Many natural phenomena, including weather, are chaotic
+\item Accurate long-term weather prediction is likely impossible
+\end{itemize}
+
 \section{{\bfseries\sffamily TODO} \cite{liu2007atmospheric}}
-\label{sec:org4049c31}
+\label{sec:org771a8c8}
+
 \section{\cite{lubbecke2014assessing}}
-\label{sec:org7b0850f}
+\label{sec:orgccacdfc}
+
+\begin{itemize}
+\item Change in ENSO characteristics after 2000
+\item Lower thermocline slope reduces TH feedback, making it harder for strong ENSO to form
+\end{itemize}
+
 \section{\cite{maher2018enso}}
-\label{sec:orgdd094cf}
+\label{sec:org8070eaf}
+
+\begin{itemize}
+\item Use two different large ensembles
+\item ENSO has lots of internal variability in models
+\item 30-40 members of an ensemble are needed to detect significant changes to ENSO in the ensemble mean
+\item Increase in ENSO amplitude in worst case (RCP8.5) scenario
+\end{itemize}
+
 \section{\cite{nowack2017role}}
-\label{sec:org9238fee}
+\label{sec:org123f132}
+
+\begin{itemize}
+\item Model simulations that are forced with differing levels of greenhouse and aerosol emissions
+\item Concluded that aerosol forcing decreases greenhouse strengthening of ENSO
+\end{itemize}
+
+\section{{\bfseries\sffamily TODO} \cite{pachauri2014climate}}
+\label{sec:orgb3ac32a}
+
+\begin{itemize}
+\item IPCC report 2014
+\end{itemize}
+
 \section{\cite{phillips2014evaluating}}
-\label{sec:org830cbdd}
+\label{sec:orgc0ab115}
+
+\begin{itemize}
+\item Citation for CVDP data
+\end{itemize}
+
 \section{{\bfseries\sffamily TODO} \cite{rashid2016atmospheric}}
-\label{sec:orge93d755}
+\label{sec:org826e0bc}
 \section{{\bfseries\sffamily TODO} \cite{ropelewski1987global}}
-\label{sec:org47cae0c}
+\label{sec:orgc7c5d50}
 \section{{\bfseries\sffamily TODO} \cite{son2010impact}}
-\label{sec:orga87c549}
+\label{sec:org8390afa}
 \section{{\bfseries\sffamily TODO} \cite{stevenson2010enso}}
-\label{sec:org6de73f7}
+\label{sec:org7d57c16}
 \section{{\bfseries\sffamily TODO} \cite{stevenson2012will}}
-\label{sec:org697dce3}
+\label{sec:orge0c0888}
+
 \section{\cite{stevenson2012significant}}
-\label{sec:orge1990c0}
+\label{sec:orgc796e1b}
+
+\begin{itemize}
+\item CMIP5 dataset
+\item No significant change in ENSO amplitude for multi-model mean during 20-21st century
+\item Some significant changes in certain models
+\item Larger ensembles are needed
+\end{itemize}
+
 \section{\cite{stevenson2017forced}}
-\label{sec:orgf634322}
+\label{sec:orgcf41ef7}
+
+\begin{itemize}
+\item ENSO diversity: difference between CP and EP ENSO events
+\item CESM Last Millennium ensemble
+\item Less forced impact on ENSO amplitude, some impact on diversity
+\item Lots of natural factors have an impact too
+\end{itemize}
 
 \section{\cite{torrence1998practical}}
-\label{sec:org2e0f412}
+\label{sec:org1dc9592}
 
@@
 \section{{\bfseries\sffamily TODO} \cite{vecchi2006weakening}}
-\label{sec:orgb996a82}
+\label{sec:orgbbf1261}
+
 \section{\cite{vega2017analysis}}
-\label{sec:orga1fc1a2}
+\label{sec:org27d5f82}
+
+\begin{itemize}
+\item Control and global warming forced simulations
+\item Lots of internal variability
+\end{itemize}
+
 \section{{\bfseries\sffamily TODO} \cite{wang2016nino}}
-\label{sec:orge694acd}
+\label{sec:org937c167}
 \section{{\bfseries\sffamily TODO} \cite{yeo2016role}}
-\label{sec:orgeaa77a3}
+\label{sec:orgf9d0601}
 \section{{\bfseries\sffamily TODO} \cite{zhang2019review}}
-\label{sec:org471ee8c}
+\label{sec:orgfbc580a}
 \section{{\bfseries\sffamily TODO} \cite{zheng2016intermodel}}
-\label{sec:org1371712}
+\label{sec:orgbd6185d}
+
 \section{\cite{zheng2017response}}
-\label{sec:org013bb4e}
+\label{sec:orgcd8dd72}
+
+\begin{itemize}
+\item CESM-LE ensemble
+\item Lots of internal variability
+\item At least 15 models are necessary to separate forced from internal variability
+\end{itemize}
 
 \bibliographystyle{apalike}
-\bibliography{references}
+\bibliography{../references}
 \end{document}

### 2021-08-15 data/extract.py

@@ -1,0 +1,32 @@
+#!/usr/bin/env python
+
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import netCDF4 as nc
+
+PATH_FF = "/Volumes/Extreme SSD/DATA/stacked/CESM1/TREFHT/"
+PATH_SF = "/Volumes/Extreme SSD/DATA/stacked/CESM1SF/TREFHT/"
+
+FILE_FF = "b.e11.BRCP85C5CNBDRD.f09_g16.all.cam.h0.TREFHT.192001-210012.nc"
+FILES_SF = [
+    "b.e11.B20TRLENS_RCP85.f09_g16.xghg.all.cam.h0.TREFHT.192001-208012.nc",
+    "b.e11.B20TRLENS_RCP85.f09_g16.xaer.all.cam.h0.TREFHT.192001-208012.nc",
+    "b.e11.B20TRLENS_RCP85.f09_g16.xbmb.all.cam.h0.TREFHT.192001-202912.nc",
+    "b.e11.B20TRLENS_RCP85.f09_g16.xlulc.all.cam.h0.TREFHT.192001-202912.nc"]
+
+nino34 = [[90, 101],
+          [152, 192]]
+
+
+def extract_nino_34(file, nino34 = [[90, 101], [152, 192]]):
+    f = nc.Dataset(file)
+    nino = f.variables["TREFHT"][:,
+                                 :,
+                                 nino34[0][0]:nino34[0][1],
+                                 nino34[1][0]:nino34[1][1]].mean((2, 3))
+    return nino
+
+def get_variance(x, win = 240):
+    out = x.rolling(win, center = True).var().to_numpy()[:, 0]
+    return out

### 2021-08-15 data/varcalc.py

@@ -1,0 +1,6 @@
+#!/usr/bin/env python
+
+import numpy as np
+import pandas as pd
+import matplotlib.pyplot as plt
+import netcdf4 as nc

### 2021-08-15 syncrefs.sh

@@ -1,0 +1,14 @@
+#!/usr/bin/env bash
+set -euo pipefail
+
+source="articles/references.bib"
+
+targets=(
+    "presentations/presentation/references.bib"
+    "paperwork/paper/references.bib"
+)
+
+for t in $targets
+do
+    cp $source $t
+done

### 2021-08-16 data/correlation.r
TEMPDT correlation plot!

@@ -5,15 +5,108 @@ library(ggplot2)
 library(WaveletComp)
 library(reshape2)
 
+# subtract ensemble mean from member
+prepare <- function(x) {
+    x.mean <- x %>% apply(1, mean)
+    return(x - x.mean)
+}
+
+# calculate 20-year rolling variance
+rollvar <- function(x) {
+  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
+  return(out)
+}
+
+# load variance from ensemble file
+load_variance <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% apply(2, rollvar))
+
+se <- function(x, ...) sd(x, ...)/sqrt(length(x))
+
+# bind ensemble statistics into dataframe while preserving date offset
+bind_stats <- function(this.date, this.mean, this.se, date.start = NA, date.end = NA) {
+    if(length(this.mean) == length(this.date)) {
+        this.df <- data.frame(date = this.date, mean = this.mean, se = this.se)
+    } else {
+        if(is.na(date.end)){
+            date.end <- length(this.date)
+        }
+        if(is.na(date.start)){
+            date.start <- 0
+        }
+        this.df <- data.frame(date = this.date, mean = c(
+                                                    rep(NA, date.start),
+                                                    this.mean,
+                                                    rep(NA, (length(this.date) - date.end))),
+                              se =
+                                  c(rep(NA, date.start),
+                                    this.se,
+                                    rep(NA, (length(this.date) - date.end))))
+        writeLines(as.character(date.start))
+        writeLines(as.character(date.end))
+        return(this.df)
+    }
+}
+
+# Produce ensemble stats dataframe
+make_df <- function(this, this.date, date.start = NA, date.end = NA) {
+    this.mean <- this %>% apply(1, mean)
+    this.se <- this %>% apply(1, se)
+    writeLines(as.character(length(this.mean)))
+    this.df <- bind_stats(this.date, this.mean, this.se, date.start, date.end)
+    return(this.df)
+}
+
[…]

### 2021-08-20 paperwork/paper/notes.tex
paper outline, data analysis now confusing :(

% Created 2021-08-19 Thu 20:28
% Intended LaTeX compiler: pdflatex
\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{Benjamin Goldman}
\date{\today}
\title{Notes for Paper}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Notes for Paper},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{Introduction}
\label{sec:orge21977b}
\subsection{Rationale}
\label{sec:org22bff9d}
\subsubsection{Greenhouse emissions are increasing global temperatures (breifly state)}
\label{sec:org2d93704}
\subsubsection{ENSO is a change in the temperature of the pacific ocean with effects around the world}
\label{sec:org5775d0e}
\subsubsection{If climate change affects ENSO, then we need to find out \underline{now}}
\label{sec:org56c62ca}
\subsection{Review of literature:}
\label{sec:orge71ba68}
\subsubsection{Climate change affects ENSO}
\label{sec:orgddb4469}
\begin{enumerate}
\item Comparison of amplitude between many models with CMIP5 is mostly inconclusive
\label{sec:org0ff3528}
\item CESM-LE shows increase in amplitude that is significant
\label{sec:org409cf8d}
\end{enumerate}
\subsubsection{Role of multiple natural and artificial factors}
\label{sec:org7c2bb2f}
\begin{enumerate}
\item Greenhouse gasses may be increasing ENSO amplitude
\label{sec:org2ca497c}
\item aerosol emissions dampen GHG increase in amplitude
\label{sec:org21ed145}
\item What about everything else?
\label{sec:org5e6776a}
\item Not using a large ensemble mostly
\label{sec:org4781853}
\end{enumerate}
\subsection{Gap}
\label{sec:orga9e7bba}
\subsubsection{not enough analysis on how ghg, aer, and others affect ENSO amplitude}
\label{sec:org9c142a4}
\begin{enumerate}
\item how the two factors interact
\label{sec:org2cfc500}
\item how their effects look compared to internal variability
\label{sec:orgf657ec5}
\item What physical mechanisms mediate their effect
\label{sec:org99afee8}
\end{enumerate}
\subsubsection{Use some interesting methods: wavelet analysis}
\label{sec:org8f83908}
\subsection{Variables:}
\label{sec:org9774955}
\subsubsection{Independent variable: external forcing}
\label{sec:org2e804f0}
\begin{enumerate}
\item input climate models with observed and predicted greenhouse gas, aerosol, biomass burning, land use/cover, etc
\label{sec:org72954cd}
\item GHG:
\label{sec:org5622f92}
\begin{enumerate}
\item Observed and predicted to increase due to industrial burning
\label{sec:org6a16208}
\item Lead to higher air (and water) temperature due to greenhouse effect
\label{sec:orgda4de99}
\end{enumerate}
\item AER:
\label{sec:orgf592de7}
\begin{enumerate}
\item Aerosols produced by industry (smoke) and nature (volcanoes)
\label{sec:org9b2d67f}
\item Usually reduce global temperatures by reflecting oncoming sunlight
\label{sec:orge9600f1}
\item Location of emission may change as industry moves from US to China
\label{sec:orgca63294}
\end{enumerate}
\item BMB:
\label{sec:org2a56291}
\begin{enumerate}
\item Increased frequency of forest and grassland fires probably due to global warming
\label{sec:org5a531ca}
\item Releases CO\textsubscript{2} (GHG) and smoke (AER)
\label{sec:orgfc5ba1a}
\end{enumerate}
\item LULC:
\label{sec:orgc698b75}
\begin{enumerate}
\item Deforestation and desertification
\label{sec:orgde04da2}
\item Greenery reduces land temperature
\label{sec:org231cd88}
\item Greenery reduces CO\textsubscript{2} levels
\label{sec:org6abc24d}
\item Modifies water cycle as leaves increase evaporation
\label{sec:org2ceaf22}
\end{enumerate}
\end{enumerate}
\section{Data}
\label{sec:orgf91b064}
\subsection{CESM1}
\label{sec:orgdbe1af7}
\subsubsection{Full forcing}
\label{sec:org36a19d8}
\begin{enumerate}
\item Observed emissions levels 1920-2005
\label{sec:orgb6bc9c1}
\item RCP8.5 (worst case) emissions levels 2005-2100
\label{sec:org33be126}
\end{enumerate}
\subsubsection{Single forcing}
\label{sec:org7605e03}
\begin{enumerate}
\item All but one forcing group
\label{sec:org5ca7629}
\begin{enumerate}
\item GHG
\label{sec:org0bfc156}
\item AER
\label{sec:org1fc732e}
\item BMB
\label{sec:org31cca4f}
\item LULC
\label{sec:orgad3afc0}
\end{enumerate}
\item Observed 1920-2005
\label{sec:orgfc27146}
\item Predicted emissions 2005-2080
\label{sec:orgcd0ade9}
\end{enumerate}
\subsubsection{Control}
\label{sec:org3521a66}
\begin{enumerate}
\item 1850 fixed emissions levels
\label{sec:org78d4232}
\end{enumerate}
\subsection{CESM2}
\label{sec:org903fc5c}
\subsubsection{Higher ensemble size}
\label{sec:orgfa93ce4}
\subsubsection{Longer record length}
\label{sec:org6ef82de}
\subsubsection{Newer code}
\label{sec:org1bfcb6e}
\subsubsection{Full forcing}
\label{sec:orgad33236}
\begin{enumerate}
\item SMBB vs CMIP6 discrepancy
\label{sec:org7c1e55b}
\item Observed forcing levels 1850-2005
\label{sec:org82d9b1c}
\item Worst case emissions 2006-2100
\label{sec:org849b535}
\end{enumerate}
\subsubsection{Single Forcing}
\label{sec:org373f00c}
\begin{enumerate}
\item Similar to CESM1
\label{sec:org690c5d6}
\item Not all members are done yet
\label{sec:org8ea3d2f}
\end{enumerate}
\subsubsection{Control}
\label{sec:orgb301cb6}
\begin{enumerate}
\item Similar to CESM1
\label{sec:org735b3e3}
\end{enumerate}
\subsection{Observed data}
\label{sec:orgeb1c1df}
\subsubsection{Might mention in introduction?}
\label{sec:orga6c7ed4}
\subsubsection{Used for introductory presentation figures, not yet for actual methods}
\label{sec:org9778c23}
\section{Methods and Results}
\label{sec:org1eb5e61}
\subsection{For all ensembles and all variables:}
\label{sec:orgc5525b4}
\subsubsection{Subtract ensemble mean from each member to derive internal variability i.e. ENSO}
\label{sec:orgba6e56a}
\subsection{Niño 3.4 index}
\label{sec:orgf22a460}
\subsubsection{Mean temperature in Niño 3.4 region}
\label{sec:org7245a90}
\subsubsection{20-year rolling variance for each member}
\label{sec:org084d966}
\subsubsection{Ensemble statistics: mean and standard error (dependent on ensemble size)}
\label{sec:orgdf5efb6}
\subsubsection{Control?}
\label{sec:orgb67f98a}
\subsubsection{Figure 1: CESM1 and CESM2 FF Niño 3.4 variance ensemble means}
\label{sec:orgdb13ad6}
\subsubsection{Explanation:}
\label{sec:org1c938fa}
\begin{enumerate}
\item CESM1
\label{sec:org20d12a8}
\begin{enumerate}
\item Starts with low variance, steeply increasing from 1950-2050, dropping slightly after 2050
\label{sec:orga9e35ea}
\item Probably nonlinear response to forcing: does not keep increasing
\label{sec:org97cb690}
\begin{enumerate}
\item Are emissions slowing down post-2050?
\label{sec:org42ff186}
\end{enumerate}
\end{enumerate}
\item CESM2
\label{sec:org5378e06}
\begin{enumerate}
\item Exponential? increase in variance until 2025, then sharp decrease
\label{sec:org3dea794}
\begin{enumerate}
\item Why?
\label{sec:org487e458}
\end{enumerate}
\end{enumerate}
\end{enumerate}
\subsection{Single forcing ensembles}
\label{sec:org85701d9}
\subsubsection{SF ensembles for CESM1 and CESM2}
\label{sec:org863b2ac}
\subsubsection{All-but one}
\label{sec:org55682bb}
\subsubsection{Figure 2: ENSO amplitude index for SF and FF models}
\label{sec:org9520816}
\subsubsection{Explanation}
\label{sec:orge8bf1b3}
\begin{enumerate}
\item -BMB and -LULC means are pretty close to FF mean, signaling little impact
\label{sec:orge62b262}
\item -AER and -GHG means are mostly unchanged, signaling stronger impact
\label{sec:org778cb4f}
\end{enumerate}
\subsubsection{Bootstrap}
\label{sec:org4a237e8}
\begin{enumerate}
\item Subtract random SF member from random FF member
\label{sec:org54d0446}
\item Leaves behind isolated effect of single factor
\label{sec:orgf1b88f4}
\item aeu
\label{sec:orgbab39ac}
\end{enumerate}
\end{document}

### 2021-08-20 data/varcalc.r

@@ -18,8 +18,19 @@ rollvar <- function(x) {
   return(out)
 }
 
+rollmean <- function(x) {
+  out <- rollapply(x, width = 12, FUN = mean, align = "center", fill = NA)
+  return(out)
+}
+
+getvar <- function(x) {
+    apply(x, 2, rollvar)
+}
+
 # load variance from ensemble file
 load_variance <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% prepare %>% apply(2, rollvar))
+
+load_temp <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% apply(2, rollmean))
 
 se <- function(x, ...) sd(x, ...)/sqrt(length(x))
 
@@ -57,6 +68,14 @@ make_df <- function(this, this.date, date.start = NA, date.end = NA) {
     return(this.df)
 }
 
+make_df_manvar <- function(this, n, this.date, date.start = NA, date.end = NA) {
+    this.mean <- this %>% apply(1, mean)
+    this.se <- this %>% apply(1, se_bs, n = n)
+    writeLines(as.character(length(this.mean)))
+    this.df <- bind_stats(this.date, this.mean, this.se, date.start, date.end)
+    return(this.df)
+}
+
 # calculate standard error in bootstrap process setting
 se_bs <- function(x, n) {return(sd(x)/sqrt(n))}
 
@@ -83,22 +102,41 @@ subtract.bootstrap <- function(data1, data2, reps, date, n) {
     return(out.df)
 }
 
+subtract.bootstrap.raw <- function(data1, data2, reps) {
+    len <- dim(data1)[1]
+    num1 <- dim(data1)[2]
+    num2 <- dim(data2)[2]
+    out <- matrix(NA, ncol = reps, nrow = len)
+#    writeLines(as.character(dim(out)))
+#    writeLines(as.character(len))
+    choice1 <- sample(x = 1:num1, size = reps, replace = TRUE)
+    choice2 <- sample(x = 1:num2, size = reps, replace = TRUE)
+    for(i in 1:reps){
+#        writeLines(as.character(choice1[i]))
+#        writeLines(as.character(choice2[i]))
[…]

### 2021-08-20 presentations/presentation/presentation.tex
Presentation i use tikz!!!

@@
 \usepackage{tikz}
+\usepackage{varwidth}
+\usetikzlibrary{arrows,snakes,backgrounds}
 \usetheme{metropolis}
@@
   \begin{columns}
-    \column{0.4\textwidth}
+    \column{0.6\textwidth}
     \begin{itemize}
@@
     \item Long-term trends and short-term noise.
+    \item Forcing: any external factor that affects climate
+      \begin{itemize}
+      \item Greenhouse gasses
+      \item Aerosols (natural: volcanic ash, artificial: smoke)
+      \item Biomass burning
+      \item Land use/cover (deforestation, desertification)
       \end{itemize}
-    \column{0.6\textwidth}
+    \end{itemize}
+    \column{0.4\textwidth}
     \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{this}
@@
 \begin{frame}{Climate Simulation}
+
+  \begin{tikzpicture}[node distance = 1.5in, thick]
+
+    \node (input) [rectangle, draw] {
+      \begin{varwidth}{1.5in}
+        Independent variables
+        Ex. greenhouse gas
+      \end{varwidth}};
+
+    \node (model) [right of=input, circle, draw] {
+      \begin{varwidth}{\linewidth}
+        Computer\\Simulation
+      \end{varwidth}};
+
+    \node (output) [right of=model, rectangle, draw] {
+      \begin{varwidth}{1.4in}
+        Dependent variables
+        Model's predictions
+      \end{varwidth}};
+
+    \draw [->] (input.east) -- (model.west);
+    \draw [->] (model.east) -- (output.west);
+
+  \end{tikzpicture}
 
@@
 
-\begin{frame}{Measuring ENSO}
-
+\begin{frame}{Model Setup}
+  Input:
+  \begin{itemize}
+  \item Observed forcing levels from 1850-2005
+  \item Predicted forcing levels from 2005-2100
+  \item \textbf{Simulation ran a total of 40 times with slightly different starting point}
+  \item Control simulation with pre-1850 forcing levels
+  \end{itemize}
+  Model:
+  \begin{itemize}
+  \item CESM1 \citep{kay2015community}
+  \item CESM2 \citep{danabasoglu2020community}
+  \end{itemize}
+  Output: Sea Surface Temperature (SST)
 \end{frame}
@@
 \begin{frame}{Measuring ENSO Intensity}
+
+  \tikzstyle{process}=[circle, draw=black!50,fill=black!20, line width = 0.2mm]
+  \tikzstyle{data}=[rectangle, draw=blue!50,fill=blue!20, line width = 0.2mm]
+  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
+
+    \node [data] (input) {
+      \begin{varwidth}{1.5in}
+        External forcing
+      \end{varwidth}};
+
+    \node [process] (model) [right of=input] {
+      \begin{varwidth}{\linewidth}
+        CESM1\\and\\CESM2
+      \end{varwidth}};
+
+    \node [data] (output) [right of=model] {
+      \begin{varwidth}{1.2in}
+        Sea surface temperature (SST)
+      \end{varwidth}};
+
+    \node [data] (nino) [below of=input] {
+      \begin{varwidth}{1.4in}
+        Niño 3.4 index
+      \end{varwidth}};
+
+    \node [process] (variance) [right of=nino] {
+      \begin{varwidth}{1.4in}
+        Windowed\\variance
+      \end{varwidth}};
+
+    \node [data] (intensity) [right of=variance] {
+      \begin{varwidth}{1.4in}
+        ENSO intensity
+      \end{varwidth}};
+
+    \draw [->] (input) to (model);
+    \draw [->] (model) to (output);
+    \draw [->] (output) to [out = 270, in = 90] (nino);
+    \draw [->] (nino) to (variance);
+    \draw [->] (variance) to (intensity);
+
+  \end{tikzpicture}
 
@@
 \begin{frame}{ENSO is Becoming Stronger}
-
-\end{frame}
-
-\begin{frame}{It's not That Simple}
 
@@
 \begin{frame}{Stratification}
-
-\end{frame}
-
-\begin{frame}{Stratification in Other Ensembles}
 

### 2021-08-23 presentations/presentation/presentation.tex
lotsa presentation stuff

@@
 \usetikzlibrary{arrows,snakes,backgrounds}
-\usetheme{metropolis}
+\usetheme{ben}
 \renewcommand{\bibsection}{}
-\setbeamerfont{caption}{size=\tiny}
-
+\setbeamerfont{caption}{size=\small}
+\setbeamertemplate{caption}[numbered]
 % \begin{frame}{Example Frame With Text and Image}
@@
 
+  \tikzstyle{process}=[circle, draw=red!50,fill=red!20, line width = 0.3mm]
+  \tikzstyle{data}=[rectangle, draw=lightblue!50,fill=lightblue!20, line width = 0.3mm]
+
 \maketitle
@@
     \item Long-term trends and short-term noise.
-    \item Forcing: any external factor that affects climate
+    \item \alert{Forcing}: any external factor that affects climate.
       \begin{itemize}
@@
 
-\begin{frame}{El Niño}
+\begin{frame}{El Niño (ENSO)}
   \begin{figure}
@@
 \begin{frame}{Climate Simulation}
-
-  \begin{tikzpicture}[node distance = 1.5in, thick]
-
-    \node (input) [rectangle, draw] {
+  \begin{itemize}
+  \item Run climate simulation with predicted forcing levels as input.
+  \item \alert{Ensemble:} set of repeated simulations.
+  \end{itemize}
+  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
+    \node (input) [data] {
       \begin{varwidth}{1.5in}
-        Independent variables
-        Ex. greenhouse gas
+        Independent variables (forcing)
       \end{varwidth}};
-
-    \node (model) [right of=input, circle, draw] {
+    \node (model) [right of=input] [process] {
       \begin{varwidth}{\linewidth}
@@
       \end{varwidth}};
-
-    \node (output) [right of=model, rectangle, draw] {
+    \node (output) [right of=model] [data] {
       \begin{varwidth}{1.4in}
-        Dependent variables
-        Model's predictions
+        Dependent variables (climatology)
       \end{varwidth}};
-
-    \draw [->] (input.east) -- (model.west);
-    \draw [->] (model.east) -- (output.west);
-
+    \node (repeat) [above of=model, node distance = 1in] [process] {
+      \begin{varwidth}{0.5in}
+        Repeat
+      \end{varwidth}};
+    \draw [>->] (input) to (model);
+    \draw [>->] (model) to (output);
+    \draw [>->] (model) to [out = 50, in = 330] (repeat);
+    \draw [>->] (repeat) to [out = 210, in = 130] (model);
   \end{tikzpicture}
-
 \end{frame}
 
-\begin{frame}{ENSO in the Future}
+\begin{frame}{ENSO and Climate Change}
+  \begin{itemize}
+  \item ENSO changes over time \citep{lubbecke2014assessing}.
+  \item ENSO responds to external forcing.
+    \begin{itemize}
+    \item Correlation between ENSO strength and sunspot activity \citep{emile2007nino}.
+    \item Weakened ENSO during the Ice Age due to reduced CO$_2$ levels \citep{zhu2017reduced}.
+    \end{itemize}
+  \item Models show possible increasing ENSO activity in the future \citep{zheng2017response} and \citep{maher2018enso}.
+  \item Factors other than CO$_2$ can affect ENSO.
+    \begin{itemize}
+    \item Ozone emission may reduce ENSO activity \citep{nowack2017role}
+    \item Aerosol emission may reduce ENSO activity \citep{stevenson2017forced}
+    \end{itemize}
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{Gap and Goal}
-
-\end{frame}
-
-\begin{frame}{Research Questions}
-
+\begin{frame}{Gap and Questions}
+  Gap:
+  \begin{itemize}
+  \item Little research using a large ensemble to examine the effect of individual factors on ENSO.
+  \item Considerable disagreement between studies on whether ENSO will strengthen or weaken due to global warming
+  \end{itemize}
 \end{frame}
@@
 \section{Data and Methods}
-
-\begin{frame}{Ensembles: CESM1 and CESM2}
-
-\end{frame}
-
-\begin{frame}{Analysis Tools}
-  \begin{columns}[t]
-    \column{.33\textwidth}
-    R:
-    \begin{itemize}
-    \item ncdf4
-    \item zoo
-    \item dplyr
-    \item ggplot2
-    \item WaveletComp
-    \item reshape2
-    \end{itemize}
-    \column{.33\textwidth}
-    Python:
-    \begin{itemize}
-    \item numpy
-    \item pandas
-    \item scipy
-    \item matplotlib
-    \item netCDF4
-    \end{itemize}
-    \column{.33\textwidth}
-    Other:
-    \begin{itemize}
-    \item nco
-    \end{itemize}
-  \end{columns}
-
-\end{frame}
 
@@
 \begin{frame}{Model Setup}
-  Input:
-  \begin{itemize}
-  \item Observed forcing levels from 1850-2005
-  \item Predicted forcing levels from 2005-2100
-  \item \textbf{Simulation ran a total of 40 times with slightly different starting point}
-  \item Control simulation with pre-1850 forcing levels
-  \end{itemize}
-  Model:
   \begin{itemize}
@@
   \item CESM2 \citep{danabasoglu2020community}
+  \item Observed forcing levels from 1850-2005
+  \item Predicted forcing levels from 2005-2100
+  \item Ensembles have 40 and 50 simulations respectively
+  \item Control simulation with pre-1850 forcing levels
   \end{itemize}
-  Output: Sea Surface Temperature (SST)
 \end{frame}
@@
 \begin{frame}{Measuring ENSO Intensity}
-
-  \tikzstyle{process}=[circle, draw=black!50,fill=black!20, line width = 0.2mm]
-  \tikzstyle{data}=[rectangle, draw=blue!50,fill=blue!20, line width = 0.2mm]
   \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
-
     \node [data] (input) {
@@
       \end{varwidth}};
-
     \node [process] (model) [right of=input] {
@@
       \end{varwidth}};
-
     \node [data] (output) [right of=model] {
@@
       \end{varwidth}};
-
     \node [data] (nino) [below of=input] {
@@
       \end{varwidth}};
-
     \node [process] (variance) [right of=nino] {
@@
       \end{varwidth}};
-
     \node [data] (intensity) [right of=variance] {
@@
       \end{varwidth}};
-
     \draw [->] (input) to (model);
@@
     \draw [->] (variance) to (intensity);
-
   \end{tikzpicture}
-
 \end{frame}
@@
 \begin{frame}{Signal and Noise}
-
+  \begin{columns}
+    \column{0.3\textwidth}
+    \begin{itemize}
+      \item Butterfly effect
+      \item Use ensemble mean to remove noise.
+      \item The larger the ensemble, the less noise there is.
+    \end{itemize}
+    \column{0.7\textwidth}
+    \myfig{variance_3.pdf}{ENSO intensity for each run of the CESM1 Large Ensemble}{variance_3}
+  \end{columns}
 \end{frame}
@@
 \begin{frame}{ENSO is Becoming Stronger}
-
+  \begin{columns}
+    \column{0.5\textwidth}
+    \begin{itemize}
+    \item Increase in ENSO intensity in both ensembles.
+    \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
+      \begin{itemize}
+      \item May be caused by aerosol emissions.
+      \end{itemize}
+    \end{itemize}
+    \column{0.5\textwidth}
+    \myfig{ff_compare.pdf}{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}{ff_compare}
+  \end{columns}
 \end{frame}
@@
 
+\begin{frame}{Wavelet Analysis}
+
+\end{frame}
+
+\begin{frame}{Power Spectrum}
+
+\end{frame}
+
 \section{Conclusion}
@@
 \begin{frame}{Acknowledgments}
-
+  \begin{itemize}
+  \item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
+1852977.
+  \item Thank you to my teacher, my family, and my mentor!
+  \item Software used: R, ncdf4, zoo, dplyr, ggplot2, WaveletComp, reshape2, nco.
+  \end{itemize}
 \end{frame}
@@
   \bibliographystyle{apalike}
-  \fontsize{7pt}{7.2}\selectfont
+  \fontsize{5pt}{6}\selectfont
   \bibliography{references.bib}
@@
 
-\maketitle
-
 \end{document}

### 2021-08-24 presentations/presentation/presentation.tex
presentation looks even better

@@
 
-  \tikzstyle{process}=[circle, draw=red!50,fill=red!20, line width = 0.3mm]
-  \tikzstyle{data}=[rectangle, draw=lightblue!50,fill=lightblue!20, line width = 0.3mm]
+  \tikzstyle{process}=[circle, draw=process, fill=process!20, line width = 0.3mm]
+  \tikzstyle{data}=[rectangle, draw=data, fill=data!20, line width = 0.3mm]
 
@@
   \bibliographystyle{apalike}
-  \fontsize{5pt}{6}\selectfont
+  \fontsize{4pt}{5}\selectfont
   \bibliography{references.bib}
@@
 \end{document}
+
+%hi

### 2021-08-31 paperwork/paper/paper.tex
paper additions

--- 2021-07-08 paperwork/paper/paper.tex
@@
-% Created 2021-07-07 Wed 20:51
-% Intended LaTeX compiler: xelatex
+% Created 2021-08-31 Tue 13:01
+% Intended LaTeX compiler: pdflatex
 \documentclass[11pt]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 \usepackage{graphicx}
@@
 \usepackage{hyperref}
+\usepackage[letterpaper, margins=1in]{geometry}
 \usepackage{natbib}
-\usepackage[letterpaper, margins=1in]{geometry}
-\setlength{\parskip}{0.35em}
 \author{Benjamin Goldman}
@@
 \begin{abstract}
-
 The El Niño/Southern Oscillation (ENSO) is the dominant mode of interannual climate variability, with substantial associated global socio-economic impacts.
@@
 However, it is currently unclear what effect greenhouse gas (GHG) and industrial aerosol (AER) emissions have on ENSO, as well as what effect these factors have when combined.
-This study examined transient changes to ENSO variance under a variety of forcing scenarios using the CESM1 Large and Single-Forcing Ensembles.
+This study examined transient changes to ENSO variance under a variety of forcing scenarios using the CESM1 and CESM2 Large and Single-Forcing Ensembles.
 These multi-member ensembles span the historical record (1920-2005) and much of the 21st C (2006-2080 for GHG/AER).
@@
 We identifed significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
-
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:org3fc8d9e}
+\label{sec:orgd19dc1d}
+
 El Niño is the main mode of interanual climate variability, originating from an interaction between the atmosphere and water movement and temperature in the Pacific ocean \citep{bjerknes1969atmospheric}.
@@
 \section{Data}
-\label{sec:org0534843}
-The primary data source for this study is NCAR's Community Earth System Model Large Ensemble and Single Forcing Ensemble (CESM1-LE).
+\label{sec:orge703c0c}
+
+The primary data source for this study is NCAR's Community Earth System Model versions 1 and 2 (CESM1 and CESM2) Large Ensembles and Single Forcing Ensembles.
 The Large Ensemble contains 40 simulations of the CESM1 coupled model, forced with historic radiative forcing from 1920 to 2005 and according to the RCP 8.5 protocol from 2006 to 2100 \citep{kay2015community}.
+The CESM2 Large Ensemble consists of 100 total member, 50 of which are forced according to the CMIP6 scheme, and 50 of which receive the SMBB forcing scheme \citep{danabasoglu2020community}.
 The single forcing ensemble is a collection of sub-ensembles for various climate factors (greenhouse gas, aerosols, biomass burning, land use, ozone).
@@
 \section{Methods and Results}
-\label{sec:org6ea3bdd}
+\label{sec:orgb29ba8d}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org907e898}
-We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
+\label{sec:org97f814f}
+We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 and CESM2 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
+First, we subtract the ensemble mean from each member to isolate the internal variability for each member.
 We measure variance on 20-year centered sliding windows.
 We calculate the multi-model mean for each ensemble, as well as the ensemble standard error.
-The results of this calculation are shown in figure \ref{fig:variance_1}.
+The results of this calculation  for the CESM1 Large and Single Forcing Ensembles are shown in figure \ref{fig:variance_1}.
 
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/variance_1.pdf}
-\caption{\label{fig:variance_1}20-year windowed variance of the Niño 3.4 region SST, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc, f) fixedO3.
-Grey bars are PI control mean and 1x, 2x, 3x standard error.
-Colored bars represent ensemble standard error.}
+\includegraphics[width=.9\linewidth]{../../data/figures/cesm1.pdf}
+\caption{\label{fig:variance_1}20-year windowed variance of the Niño 3.4 index, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc. Colored bars represent ensemble standard error.}
+\end{figure}
+
+Incorporating the CESM2 Large Ensemble further supports these conclusions, as shown in figure \ref{fig:variance_2}.
+The CESM2 ensemble mean increases at a rate comparable to that of the CESM1, and its larger ensemble size allows for a smaller confidence interval.
+In contrast to the CESM1's slight decrease in ENSO intensity in the late 21\textsuperscript{st} century, the CESM2 ensemble mean decreases drastically after 2050, possibly ending below pre-1850 levels.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{../../data/figures/ff_compare.pdf}
+\caption{\label{fig:variance_2}20-year windowed variance of the Niño 3.4 index for the CESM1 and CESM2 fully-forced Large Ensembles. Colored bars represent ensemble standard error.}
 \end{figure}
@@
 The AMO has been shown to have an influence on ENSO strength and seasonal growth rate \citep{levine2017impact}.
-We filter the control Niño 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \citep{phillips2014evaluating}\}.
+We filter the control Niño 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \citep{phillips2014evaluating}.
 The Probability Distribution Function of Niño 3.4 variance is estimated for AMO/C > 1/2\(\sigma\) and < -1/-2\(\sigma\) using a Kernel Density Estimation.
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:orge41456d}
+\label{sec:org9d1d268}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/bootstrap_1.pdf}
-\caption{\label{fig:bootstrap_1}Difference between fully forced and single forced ensembles derived from the bootstrap process for a) greenhouse gas, b) aerosol emissions c) biomass burning, d) land use/cover, e) ozone; f) compares the sum of the bootstrapped ensemble means for the single forcing ensembles (blue) and the ensemble mean for the full forcing ensemble, detrended by a constant to center on zero (orange).}
+\includegraphics[width=.9\linewidth]{../../data/figures/cesm1_sf.pdf}
+\caption{\label{fig:bootstrap_1}Difference between fully forced and single forced ensembles derived from the bootstrap process for a) aerosol emissions, b) biomass burning c) greenhouse gasses, d) land use/cover}
 \end{figure}
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:orgabb8356}
+\label{sec:org91e95b0}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/tempdt.pdf}
-\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Niño 3.4 variance.
-Grey color indicates points where p<0.05.
-The white regions at 25 and 280 degrees East are Africa and South America.}
+\includegraphics[width=.9\linewidth]{../../data/figures/tempdt.pdf}
+\caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Niño 3.4 variance.}
 \end{figure}
@@
 \centering
-\includegraphics[width=.9\linewidth]{./figures/tempcep.pdf}
-\caption{\label{fig:tempcep}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.
-Grey color indicates points where p>0.05.}
+\includegraphics[width=.9\linewidth]{../../data/figures/tempcep.pdf}
+\caption{\label{fig:tempcep}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.}
 \end{figure}
 
-\section{Conclusions and Discussion}
-\label{sec:org811a462}
+\subsection{Wavelet Analysis}
+\label{sec:org7c0c948}
+Finally, we conduct a wavelet transform on the Niño 3.4 index for the CESM1 and CESM2 Large Ensembles. First, we
 
-In this study, predicted changes to ENSO amplitude are examined in the CESM1 Large Ensemble and Single Forcing Ensemble.
-The fully-forced large ensemble exhibits increased ENSO amplitude, as measured by calculating the 20-year variance of the Niño 3.4 index.
-These changes appear to be statistically significant, but need to be scrutinized further, as they are in disagreement with results of some previous studies \citep{stevenson2012significant}.
-Additionally, the fully-forced ensemble, as well as all the others contains considerable noise, as the individual members cover a wide range of variability.
-Greenhouse gas and aerosol emissions are likely the most significant contributors to these forced changes, as shown by the fact that the ensemble mean for the xGHG and xAER ensembles are the most different from the fully-forced mean.
-All the other single-forcing ensembles had insignificant differences from the fully forced ensemble, or were lacking a sample size large enough to lend credibility to their results.
-Analysis of the correlation coefficient between ENSO amplitude and Pacific ocean temperature reveals that changes to ENSO intensity are tied to changes to the subsurface layer of the tropical Pacific Ocean.
-It is still unclear what is driving this connection, as well as how external forcing such as global warming affects it.
-There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \citep{dewitte2012reinterpreting}.
-
-This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \citep{maher2018enso}.
-Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citet{stevenson2017forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
-In contrast, this study observes that the impact on ENSO amplitude of greenhouse gas and aerosols has the same sign.
-This result is also surprising given that greenhouse and aerosol emissions have been shown to have opposite effects on sea surface temperature and global circulation, with greenhouse emissions favoring a more Niño-like state with weakened Walker circulation and higher SST, and aerosol emissions counteracting these effects \citep{boer2000transient}.
-
-This study has a number of goals for future continuation.
-The most pressing of these goals is a deeper analysis of changes to Pacific Ocean structure, which will be done by examining correlation between ENSO amplitude and potential density, as well as comparisons of changes to stratification in the Pacific.
-Additionally, the methods of this study will be repeated on the CESM2 large ensemble, which has a larger ensemble size and a longer record period, as compared to CESM1.
-Analysis of the CMIP6 models are also necessary to verify these results across variations in model physics.
-Deeper statistical analyses of changes to ENSO amplitude should also be tone to further examine the likelihood and intensity of future changes to ENSO amplitude.
-Lastly, correlation between ENSO amplitude and internal variability should be further investigated using the records provided by the CVDP.
-Doing these further analyses will help to verify the results of this study, while more precisely analyzing future changes to ENSO.
-
-The results of this study will help to direct future studies on the impact of climate change on ENSO by providing one further data point supporting the conclusion that ENSO will intensify due to global warming.
-This study and others like it will help societies impacted by ENSO to prepare for intensification in the future.
-
-\section{Acknowledgements}
-\label{sec:org25124e8}
-This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
-1852977.
-Computing resources (\href{https://doi.org/10.5065/D6RX99HX}{doi:10.5065/D6RX99HX}) were provided by the Climate Simulation Laboratory at NCAR's Computational and Information Systems Laboratory, sponsored by the National Science Foundation and other agencies.
-I would like to thank Ms.
-Kimberly Fleming and the White Plains High School Science Research Program for providing support in this study.
+\section{Conclusion}
+\label{sec:org89eb66f}
+\section{Acknowledgments}
+\label{sec:org75ba429}
 

### 2021-08-31 paperwork/paper/paper.tex
paper stuff and hours

@@
-% Created 2021-08-31 Tue 13:01
+% Created 2021-08-31 Tue 13:20
 % Intended LaTeX compiler: pdflatex
@@
 \section{Introduction}
-\label{sec:orgd19dc1d}
+\label{sec:orgf2c8c6a}
 
@@
 Several modes of internal variability have been shown to modulate ENSO, including the AMO \citep{levine2017impact} and Tropical Pacific Decadal Variability (TPDV) \citep{zheng2017response}.
-An analysis of the Max Plank Institute's Grand Ensemble as well as NCAR's CESM Large Ensemble suggests that 80$\backslash$% of changes to ENSO amplitude can be attributed to internal variability, but given a large enough ensemble, significant changes in ENSO amplitude due to climate change can be detected \citep{maher2018enso}.
+An analysis of the Max Plank Institute's Grand Ensemble as well as NCAR's CESM Large Ensemble suggests that 80\% of changes to ENSO amplitude can be attributed to internal variability, but given a large enough ensemble, significant changes in ENSO amplitude due to climate change can be detected \citep{maher2018enso}.
 
@@
 \section{Data}
-\label{sec:orge703c0c}
+\label{sec:org80af315}
 
@@
 \section{Methods and Results}
-\label{sec:orgb29ba8d}
+\label{sec:org99aa203}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org97f814f}
-We estimate 20th and 21st century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 and CESM2 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
+\label{sec:org6cadc23}
+We estimate 20\textsuperscript{th} and 21\textsuperscript{st} century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 and CESM2 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
 First, we subtract the ensemble mean from each member to isolate the internal variability for each member.
@@
 The fully-forced ensemble exhibits moderate increase in variance, with the other ensembles showing less meaningful changes.
-In the fully-forced ensemble, the variance of the Niño 3.4 region increases beyond 2 standard errors of the control, increasing until the mid-21st century, and then decreasing gradually.
+In the fully-forced ensemble, the variance of the Niño 3.4 region increases beyond 2 standard errors of the control, increasing until the mid-21\textsuperscript{st} century, and then decreasing gradually.
 The excluded greenhouse, aerosol, and biomass-burning ensembles also contain a slight increase, which may not be statistically significant.
@@
 
-The fully-forced ensemble exhibits reduced variance in the mid-late 20th century, below that of the PI control.
+The fully-forced ensemble exhibits reduced variance in the mid-late 20\textsuperscript{th} century, below that of the PI control.
 The most likely explanation for this phenomenon is internal variability.
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org9d1d268}
+\label{sec:orgec2797d}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 
-In both the fully-forced scenario, and the greenhouse and aerosol only simulations, there is noticeably reduced Niño 3.4 variance in the mid-20$\backslash$\textsuperscript{th} century, below 2 standard errors of the control.
+In both the fully-forced scenario, and the greenhouse and aerosol only simulations, there is noticeably reduced Niño 3.4 variance in the mid-20\textsuperscript{th} century, below 2 standard errors of the control.
 We hypothesize that this discrepancy may be the result of anomalous initial conditions caused by internal variability of the control, as the control conditions are used to initialize all the forced runs.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:org91e95b0}
+\label{sec:org4afb475}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 \subsection{Wavelet Analysis}
-\label{sec:org7c0c948}
-Finally, we conduct a wavelet transform on the Niño 3.4 index for the CESM1 and CESM2 Large Ensembles. First, we
+\label{sec:org653115c}
+Finally, we conduct a wavelet transform on the Niño 3.4 index for the CESM1 and CESM2 Large Ensembles.
+First, we subtract the ensemble mean from each member and apply a 1-year windowed mean.
+Next, we construct a power spectrum of the Niño 3.4 index for each member following the methods of \cite{torrence1998practical} and calculate the ensemble mean of these power spectra, as displayed in figure \ref{fig:wavelet_1} for the CESM1.
+The majority of ENSO variability is contained in the 2 to 8-year frequency band.
+This band exhibits steadily increasing wavelet power until the mid 21\textsuperscript{st} century, conforming to previous results of this study.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{../../data/figures/spectrum1.pdf}
+\caption{\label{fig:wavelet_1}Ensemble mean wavelet power spectrum for the CESM1 Large Ensemble.}
+\end{figure}
 
 \section{Conclusion}
-\label{sec:org89eb66f}
+\label{sec:org0359a03}
+
+In this study, predicted changes to ENSO amplitude are examined in the CESM1 Large Ensemble and Single Forcing Ensemble.
+The fully-forced large ensemble exhibits increased ENSO amplitude, as measured by calculating the 20-year variance of the Niño 3.4 index.
+These changes appear to be statistically significant, but need to be scrutinized further, as they are in disagreement with results of some previous studies \citep{stevenson2012significant}.
+Additionally, the fully-forced ensemble, as well as all the others contains considerable noise, as the individual members cover a wide range of variability.
+Greenhouse gas and aerosol emissions are likely the most significant contributors to these forced changes, as shown by the fact that the ensemble mean for the xGHG and xAER ensembles are the most different from the fully-forced mean.
+All the other single-forcing ensembles had insignificant differences from the fully forced ensemble, or were lacking a sample size large enough to lend credibility to their results.
+Analysis of the correlation coefficient between ENSO amplitude and Pacific ocean temperature reveals that changes to ENSO intensity are tied to changes to the subsurface layer of the tropical Pacific Ocean.
+It is still unclear what is driving this connection, as well as how external forcing such as global warming affects it.
+There has been shown to be a connection between Pacific stratification and ENSO variability as higher levels of stratification result in a stronger thermocline feedback, which causes the Pacific to become less stable, making strong ENSO events more likely \citep{dewitte2012reinterpreting}.
+
+This study produced similar results as of previous studies, as it predicts an increase in ENSO amplitude, as well as overwhelming noise caused by internal variability \citep{maher2018enso}.
+Although there is limited research on the impact of individual external factors on ENSO, this study complements the results of \citet{stevenson2017forced}, who observed conflicting effects of greenhouse gas and aerosol emissions on ENSO diversity.
+In contrast, this study observes that the impact on ENSO amplitude of greenhouse gas and aerosols has the same sign.
+This result is also surprising given that greenhouse and aerosol emissions have been shown to have opposite effects on sea surface temperature and global circulation, with greenhouse emissions favoring a more Niño-like state with weakened Walker circulation and higher SST, and aerosol emissions counteracting these effects \citep{boer2000transient}.
+
+This study has a number of goals for future continuation.
+The most pressing of these goals is a deeper analysis of changes to Pacific Ocean structure, which will be done by examining correlation between ENSO amplitude and potential density, as well as comparisons of changes to stratification in the Pacific.
+Additionally, the methods of this study will be repeated on the CESM2 large ensemble, which has a larger ensemble size and a longer record period, as compared to CESM1.
+Analysis of the CMIP6 models are also necessary to verify these results across variations in model physics.
+Deeper statistical analyses of changes to ENSO amplitude should also be tone to further examine the likelihood and intensity of future changes to ENSO amplitude.
+Lastly, correlation between ENSO amplitude and internal variability should be further investigated using the records provided by the CVDP.
+Doing these further analyses will help to verify the results of this study, while more precisely analyzing future changes to ENSO.
+
+The results of this study will help to direct future studies on the impact of climate change on ENSO by providing one further data point supporting the conclusion that ENSO will intensify due to global warming.
+This study and others like it will help societies impacted by ENSO to prepare for intensification in the future.
+
 \section{Acknowledgments}
-\label{sec:org75ba429}
+\label{sec:org42fbab9}
+
+This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
+1852977.
+Computing resources (\href{https://doi.org/10.5065/D6RX99HX}{doi:10.5065/D6RX99HX}) were provided by the Climate Simulation Laboratory at NCAR's Computational and Information Systems Laboratory, sponsored by the National Science Foundation and other agencies.
+I would like to thank Ms.
+Kimberly Fleming and the White Plains High School Science Research Program for providing support in this study.
 

### 2021-09-03 competitions/sts/regeneron_essay_qs.tex
compiled assessment directory

@@
-% Created 2021-07-20 Tue 17:46
+% Created 2021-09-03 Fri 14:05
 % Intended LaTeX compiler: pdflatex
-\documentclass[11pt]{article}
+\documentclass[little]{basic}
 \usepackage[utf8]{inputenc}
@@
 \section{Research Project “Layperson’s Summary” (maximum 200 words)}
-\label{sec:org3b34335}
+\label{sec:org3c116f0}
 Summarize your project in layperson’s terms, while maintaining scientific accuracy.
@@
 
+El Niño, or ENSO, is a hugely influential climate cycle that occurs in the equatorial Pacific Ocean.
+It is a form of short-term climate change, so it causes certain years to be warmer or colder than others.
+My project shows that through greenhouse gas and aerosol emissions, human activities are causing ENSO to become stronger.
+I showed this by using data output by a climate model.
+This model uses a computer to simulate the earth's climate.
+By repeating this simulation with and without greenhouse gas, aerosols, and other factors, researchers have created a picture of what climate will be like over the 21\textsuperscript{st} century.
+I used this data to analyze changes to ENSO in the future by using established methods of measuring El Niño.
+After deducing that it was becoming stronger, I went into greater depth, analyzing what parts of ENSO are becoming stronger and why.
+It appears that global warming is allowing heat to spread between layers of the ocean more easily, which makes it easier for ENSO events to form.
+If these conclusions are correct, then we can expect more extreme weather in the future.
+
 \section{Project Benefits and Impact (maximum 200 words)}
-\label{sec:org8665bc4}
+\label{sec:org3884d1d}
 What benefits do you think your research will bring to the world, and/or to your field? What additional steps, and by whom, might be needed for this benefit to be realized?
@@
 \section{Your Potential as a Scientist, Mathematician or Engineer (maximum 250 words)}
-\label{sec:orgd6d85d8}
+\label{sec:orgd083e59}
 Address through specific and concrete examples what characteristics you have that best demonstrate your affinity and aptitude for being a good scientist.
@@
 \section{Major Scientific Question (maximum 250 words)}
-\label{sec:org3b74cef}
+\label{sec:org44dc78c}
 What is a major scientific question in your field whose answer you believe will have a significant impact on the world in the next 20 years, and why?
@@
 \section{``Tweet'' about your project! Tell us about your project in 280 characters or less.}
-\label{sec:org9998841}
+\label{sec:orge35468a}
 The Society might share this response if you are named a scholar or finalist.
 
+Models show that CO2 emissions are causing El Niño to become stronger due to global warming.
+
 \section{Greatest Accomplishment or Challenge (COMMON APP)}
-\label{sec:org228c064}
+\label{sec:org33a2be4}
 Please answer ONE of the prompts below from the Common App; we prefer that you think beyond your research project in this essay. (maximum 200 words)
@@
 \section{About You (COMMON APP) (optional)}
-\label{sec:orgbad95ff}
+\label{sec:orgb28081b}
 Some students have a background, identity, interest, or talent that is so meaningful they believe their application would be incomplete without it.
@@
 \section{COVID-19 Pandemic (adapted from COMMON APP) (optional)}
-\label{sec:org2636bf9}
+\label{sec:org89440bb}
 

### 2021-09-03 presentations/presentation/presentation.tex
compiled assessment directory

@@
 
+  \begin{columns}
+    \column{0.5\textwidth}
+    \begin{itemize}
+    \item Placeholder
+    \end{itemize}
+    \column{0.5\textwidth}
+  \myfig{cesm1_sf_3.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{cesm1_sf}
+  \end{columns}
+
 \end{frame}
@@
 
-\end{frame}
-
-\begin{frame}{Stratification}
+  \begin{columns}
+    \column{0.5\textwidth}
+    \begin{itemize}
+    \item Placeholder
+    \end{itemize}
+    \column{0.5\textwidth}
+  \myfig{tempdt.pdf}{Correlation coefficient between ENSO amplitude and ocean temperature in the fully-forced CESM1 ensemble}{tempdt}
+  \end{columns}
 
@@
 
+  \begin{columns}
+    \column{0.5\textwidth}
+    \begin{itemize}
+    \item Placeholder
+    \end{itemize}
+    \column{0.5\textwidth}
+  \myfig{spectrum1.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 ensemble}{spectrum1}
+  \end{columns}
+
 \end{frame}
@@
 
+  \begin{itemize}
+  \item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with citet:cai2018increased.
+  \item Increase is likely caused by the combined influence of greenhouse gasses and aerosols.
+  \item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
+  \end{itemize}
+
 \end{frame}
@@
 \begin{frame}{Discussion}
+
+  \begin{description}
+  \item[Application] Improve prediction ability to help people prepare for increased likelihood of extreme weather.
+  \item[Limitation] Niño 3.4 index may not be fully accurate for various models \citep{cai2018increased}.
+  \item[Next steps]
+    \begin{itemize}
+    \item Examine other variables to further analyze mediator process.
+    \item Continue wavelet analysis methods to focus on individual frequency bands.
+    \end{itemize}
+  \end{description}
 
@@
 \end{document}
-
-%hi

### 2021-10-22 competitions/sts/regeneron_essay_qs.tex
big big big stuff lotsa stuff

@@
-% Created 2021-09-03 Fri 14:05
+% Created 2021-10-19 Tue 07:53
 % Intended LaTeX compiler: pdflatex
@@
 \usepackage{graphicx}
-\usepackage{grffile}
 \usepackage{longtable}
@@
 \usepackage{amsmath}
-\usepackage{textcomp}
 \usepackage{amssymb}
@@
 \section{Research Project “Layperson’s Summary” (maximum 200 words)}
-\label{sec:org3c116f0}
+\label{sec:org77e12e2}
 Summarize your project in layperson’s terms, while maintaining scientific accuracy.
@@
 \section{Project Benefits and Impact (maximum 200 words)}
-\label{sec:org3884d1d}
+\label{sec:org1415842}
 What benefits do you think your research will bring to the world, and/or to your field? What additional steps, and by whom, might be needed for this benefit to be realized?
 
+My research will improve humans' ability to prepare for ENSO in the future.
+My prediction that ENSO will become stronger due to greenhouse gas emissions provides additional motivation towards added protection against extreme weather.
+By showing that ENSO-related extreme weather will become more common in the twenty-first century, my project will encourage the construction of defenses such as flood barriers in places that are directly affected by ENSO.
+Additionally, my study provides an additional warning that we are only beginning to feel the negative effects of industrial emissions.
+Regarding the field of climatology, this study deepens our knowledge of the interaction between global warming and ENSO.
+
+In order for the benefits of my study to be realized, my results must first be verified in other studies.
+Examinations with other models and with recorded data must be done to ensure that ENSO is most likely to become stronger.
+Additionally, examinations with higher resolution that focus on predictions for specific regions should be conducted to determine which specific cities will be most impacted and how.
+
 \section{Your Potential as a Scientist, Mathematician or Engineer (maximum 250 words)}
-\label{sec:orgd083e59}
+\label{sec:orgd5c249b}
 Address through specific and concrete examples what characteristics you have that best demonstrate your affinity and aptitude for being a good scientist.
@@
 
+My strong background and achievement in mathematics demonstrates my aptitude as a scientist, which I have applied in my project for learning complex methods.
+I have succeeded in extremely difficult math courses at White Plains High School, such as AP BC Calculus and Multivariable Calculus.
+In these classes, I picked up mathematical skills quite quickly.
+This ability also applies to my future career as a scientist because I will be able to acquire certain skills more easily.
+For example, my project required me to learn countless technical skills like from navigating the command line to programming in at least 4 new languages, and managing datasets containing hundreds of separate files.
+I demonstrated my ability to learn new skills when my mentor asked me to apply a method I complex calculus-based method called wavelet analysis.
+Consequently, I watched a recording of a difficult university lecture and derived the necessary details to implement wavelet analysis in my methods.
+
+My scientific and technical ability extends beyond my project.
+I have spent hours experimenting with my computer, learning how operating systems work, picking up programming languages, and even learning the basics of hacking (legal of course!).
+Through these activities, I have honed my technology skills while gaining a deep understanding of how the world's computing infrastructure functions.
+
+In college, I plan to focus on physics while also pursuing computer science and more scientific research.
+I am unsure of my career plans, but I can imagine myself as an industry climatologist or data scientist.
+
 \section{Major Scientific Question (maximum 250 words)}
-\label{sec:org44dc78c}
+\label{sec:org97cb04c}
 What is a major scientific question in your field whose answer you believe will have a significant impact on the world in the next 20 years, and why?
@@
 
+Climate predictions are usually grouped into three categories based on human greenhouse gas emissions: worst case, average case, and best case.
+Under the worst case predictions, environmental regulation is minimal, and industry continues with the same practices it has today.
+In the best case scenario, renewable energy becomes widespread and fossil fuel burning all but stops.
+The average case is between the best and worst cases.
+The major question in my field is which pathway is the most accurate?
+If humans continue to emit greenhouse gasses like we are doing now, then climatologists' worst predictions will come true, but if we implement strict regulations and switch to renewable energy, much of this disaster may be averted.
+The answer to this question will determine which set of predictions is most applicable for helping humans to respond to climate change.
+
+I would go about answering this question by comparing characteristics such as El Niño from climate models to those in the observed record of the climate.
+By correlating these characteristics with levels of greenhouse gas emissions in the observed and simulated climate.
+
 \section{``Tweet'' about your project! Tell us about your project in 280 characters or less.}
-\label{sec:orge35468a}
+\label{sec:orgefbfab8}
 The Society might share this response if you are named a scholar or finalist.
@@
 Models show that CO2 emissions are causing El Niño to become stronger due to global warming.
+My analysis of climate predictions shows that El Niño's strength is increasing, which could cause more extreme weather.
 
-\section{Greatest Accomplishment or Challenge (COMMON APP)}
-\label{sec:org33a2be4}
+\section{{\bfseries\sffamily TODO} Greatest Accomplishment or Challenge (COMMON APP)}
+\label{sec:orgbdcedb2}
 Please answer ONE of the prompts below from the Common App; we prefer that you think beyond your research project in this essay. (maximum 200 words)
@@
 \section{About You (COMMON APP) (optional)}
-\label{sec:orgb28081b}
+\label{sec:orgfd10795}
 Some students have a background, identity, interest, or talent that is so meaningful they believe their application would be incomplete without it.
@@
 \section{COVID-19 Pandemic (adapted from COMMON APP) (optional)}
-\label{sec:org89440bb}
+\label{sec:orgee7c2ff}
 

### 2021-10-22 feedback/presentation_main.tex
big big big stuff lotsa stuff

% Created 2021-10-19 Tue 07:57
% Intended LaTeX compiler: pdflatex
\documentclass[little]{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\setcounter{secnumdepth}{1}
\author{Benjamin Goldman}
\date{\today}
\title{Feedback for main presentation}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Feedback for main presentation},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{Feedback 1}
\label{sec:org1da6f23}
Date: \textit{<2021-09-26 Sun>}
Time: 20:20
Listener: Abby Goldman
\subsection*{Feedback}
\label{sec:org1d74604}
\begin{itemize}
\item lots of arm waving. SOOOOMUCH ARM WAVING!!! And very little breathing
\item make a list of “overhead” - make sure all concepts are well defined, in your notes, possibly in your slides
\item summary of what your many slides are — many different crunching of different data to circle around a process that we’re trying to under
\item insert heading slides, overall need a good deal more organization.
\item slide 1: break up into two or even 3?
\item Slide 2: Need to explain more generally what’s in the figure
\item Slide 3: climate simulation — is this method? If so, say so
\item Slide 4: ENSO and climate change slide — have “literature review” in heading.
\item slide about Mentor: do a better job explaining the interactive process between you and your mentor.
\item 3.4 index - whahah? I didn’t follow
\item Single forcing ensembles slide — not sure if all those arrows are working for you. Skip over single forcing setup, maybe?
\item use consistent language about the aerosols and their influence, considering that it’s a prediction
\end{itemize}

\section{Feedback 2}
\label{sec:org11974c6}
Date: \textit{<2021-09-30 Thu>}
Time: 8:20
Listener: Marcus Chung, Sean O'Donnell, Jimena Perez
\subsection*{Feedback:}
\label{sec:orgdbbe7e5}
\begin{itemize}
\item Figure text is too small
\item Caption text can be smaller to make room
\item Say some transition for the conclusion heading slide
\item Clarify difference between El Niño and La Niña
\begin{itemize}
\item overall make intro 2 more organized
\end{itemize}
\item Good understanding and well communicated
\item Are some slides too empty-looking?
\item Say fewer words because it is over time and talking too fast
\item Use note cards
\item Periods should be consistent.
\end{itemize}

\section{Feedback 3}
\label{sec:org52baac0}
Date: \textit{<2021-09-30 Thu>}
Time: 16:00
Listener: Michael Goldman
\subsection*{Feedback:}
\label{sec:orgf282c69}
\begin{itemize}
\item Make a few bullets clearer
\item Good tone
\item Slow down pacing
\item Introduction is slightly unclear
\item Use note cards
\end{itemize}

\section{Feedback 4}
\label{sec:org585437d}
Date: \textit{<2021-10-17 Sun>}
Time: 17:12
Listener: Jimena Perez
\subsection*{Feedback:}
\label{sec:orgeffe568}
\begin{itemize}
\item Make figures bigger and make captions smaller
\item Love how detailed the content is
\item Great application slide
\item Make sure you explain your results in the context of your study, think big picture
\item Intro slide 2 is great we’ll explained
\end{itemize}
\end{document}

### 2021-10-22 paperwork/paper/paper.tex
big big big stuff lotsa stuff

@@
-% Created 2021-08-31 Tue 13:20
+% Created 2021-10-14 Thu 09:16
 % Intended LaTeX compiler: pdflatex
@@
 \usepackage{hyperref}
-\usepackage[letterpaper, margins=1in]{geometry}
+\usepackage[letterpaper, margin=1in]{geometry}
 \usepackage{natbib}
@@
 \maketitle
+\tableofcontents
+
 \begin{abstract}
@@
 \section{Introduction}
-\label{sec:orgf2c8c6a}
+\label{sec:org2a498ff}
 
@@
 \section{Data}
-\label{sec:org80af315}
+\label{sec:org7ef515b}
 
@@
 \section{Methods and Results}
-\label{sec:org99aa203}
+\label{sec:org2868dc1}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org6cadc23}
+\label{sec:org65b5cbb}
 We estimate 20\textsuperscript{th} and 21\textsuperscript{st} century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 and CESM2 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:orgec2797d}
+\label{sec:org75cad93}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:org4afb475}
+\label{sec:org5c7fa6f}
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 Next, we calculated the Pearson's correlation coefficient between each gridpoint and the Niño 3.4 variance timeseries.
-The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt}, and for the central slice in figure \ref{fig:tempcep}.
+The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt}.
 
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{../../data/figures/tempcep.pdf}
-\caption{\label{fig:tempcep}Correlation coefficient between detrended and smoothed ocean temperature of the central pacific slice and detrended and smoothed Niño 3.4 variance.}
-\end{figure}
-
 \subsection{Wavelet Analysis}
-\label{sec:org653115c}
-Finally, we conduct a wavelet transform on the Niño 3.4 index for the CESM1 and CESM2 Large Ensembles.
+\label{sec:orgd4dad13}
+Finally, we conduct a wavelet analysis of the Niño 3.4 index for the CESM1 and CESM2 Large Ensembles.
+Wavelet analysis separates a timeseries into time and frequency domains, displaying the amplitude for each frequency at each time step.
+The wavelet transform works by scaling a specified ``mother wavelet'' and translating it across the series at each scaling.
+This method is superior to a windowed Fourier Transform because it has high resolution on both time and frequency domains.
 First, we subtract the ensemble mean from each member and apply a 1-year windowed mean.
-Next, we construct a power spectrum of the Niño 3.4 index for each member following the methods of \cite{torrence1998practical} and calculate the ensemble mean of these power spectra, as displayed in figure \ref{fig:wavelet_1} for the CESM1.
+Next, we construct a power spectrum of the Niño 3.4 index for each member following the methods of \cite{torrence1998practical} and calculate the ensemble mean of these power spectra, as displayed in figure \ref{fig:wavelet_2}.
 The majority of ENSO variability is contained in the 2 to 8-year frequency band.
 This band exhibits steadily increasing wavelet power until the mid 21\textsuperscript{st} century, conforming to previous results of this study.
+Both ensembles show an increase in the dominant ENSO signal, peaking in the 21\textsuperscript{st} century.
+The CESM2 ENSO amplitude peaks slightly earlier than the CESM1, in accordance with figure \ref{fig:variance_2}.
+The CESM2 exhibits a slight decrease in the dominant ENSO period during the 21\textsuperscript{st} century, with the mean power spectrum's ridgeline decreasing from a period of approximately 4 years to 2.5 years.
+It is unknown whether this period decrease is statistically robust.
+Finally, the CESM2 has a lower peak wavelet power than the CESM1, in contrast with that shown in figure \ref{fig:variance_2}.
+
+To further examine the nature of forced changes to the ENSO spectrum, we split ENSO's power spectrum into two records for each ensemble, the two groups having periods less than and greater than 3 years.
+Then, we calculate the mean power along the time axis, effectively generating records of ENSO amplitude split into two frequency groups.
+Ensemble mean and standard error are found for each time-step.
+These records are shown in figure \ref{fig:wavelet_2}.
 
@@
 \centering
-\includegraphics[width=.9\linewidth]{../../data/figures/spectrum1.pdf}
-\caption{\label{fig:wavelet_1}Ensemble mean wavelet power spectrum for the CESM1 Large Ensemble.}
+\includegraphics[width=.9\linewidth]{../../data/figures/wavelet2.pdf}
+\caption{\label{fig:wavelet_1}Ensemble mean wavelet power spectra for both large ensembles.}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[width=.9\linewidth]{../../data/figures/wavelet3.pdf}
+\caption{\label{fig:wavelet_2}Ensemble mean wavelet power records for low and high period ENSO.}
 \end{figure}
@@
 \section{Conclusion}
-\label{sec:org0359a03}
+\label{sec:org272adc7}
 
@@
 \section{Acknowledgments}
-\label{sec:org42fbab9}
+\label{sec:org9f9ff26}
 

### 2021-10-22 paperwork/paper2/paper.tex
big big big stuff lotsa stuff

--- 2021-10-22 paperwork/paper/paper.tex
@@
-% Created 2021-10-14 Thu 09:16
-% Intended LaTeX compiler: pdflatex
 \documentclass[11pt]{article}
@@
 \maketitle
-\tableofcontents
 
@@
 Due to their significance, shifts in ENSO under climate change also have the potential to substantially impact human society and natural ecosystems.
-However, it is currently unclear what effect greenhouse gas (GHG) and industrial aerosol (AER) emissions have on ENSO, as well as what effect these factors have when combined.
+However, it is currently unclear what effect greenhouse gas and industrial aerosol emissions have on ENSO, as well as what effect these factors have when combined.
 This study examined transient changes to ENSO variance under a variety of forcing scenarios using the CESM1 and CESM2 Large and Single-Forcing Ensembles.
-These multi-member ensembles span the historical record (1920-2005) and much of the 21st C (2006-2080 for GHG/AER).
+These multi-member ensembles span the historical record (1920-2005) and much of the 21st century (2006-2080 for GHG/AER).
 A 2000-year pre-industrial (PI) control simulation is used to account for model drift and 20-year running variance of the Niño 3.4 SST index is used as a proxy for ENSO variance.
@@
 We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
+Finally, we examined the wavelet power spectrume of the historical and predicted Niño 3.4 index to support the data given from the 20-year variance calculation.
 We identifed significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
@@
 \section{Introduction}
-\label{sec:org2a498ff}
 
@@
 \section{Data}
-\label{sec:org7ef515b}
 
@@
 \section{Methods and Results}
-\label{sec:org2868dc1}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org65b5cbb}
+
 We estimate 20\textsuperscript{th} and 21\textsuperscript{st} century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 and CESM2 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{../../data/figures/cesm1.pdf}
-\caption{\label{fig:variance_1}20-year windowed variance of the Niño 3.4 index, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc. Colored bars represent ensemble standard error.}
-\end{figure}
-
 Incorporating the CESM2 Large Ensemble further supports these conclusions, as shown in figure \ref{fig:variance_2}.
@@
 In contrast to the CESM1's slight decrease in ENSO intensity in the late 21\textsuperscript{st} century, the CESM2 ensemble mean decreases drastically after 2050, possibly ending below pre-1850 levels.
-
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{../../data/figures/ff_compare.pdf}
-\caption{\label{fig:variance_2}20-year windowed variance of the Niño 3.4 index for the CESM1 and CESM2 fully-forced Large Ensembles. Colored bars represent ensemble standard error.}
-\end{figure}
 
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=.5\linewidth]{../../data/figures/cesm1.pdf}
+\caption{\label{fig:variance_1}20-year windowed variance of the Niño 3.4 index, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc. Colored bars represent ensemble standard error.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=.5\linewidth]{../../data/figures/ff_compare.pdf}
+\caption{\label{fig:variance_2}20-year windowed variance of the Niño 3.4 index for the CESM1 and CESM2 fully-forced Large Ensembles. Colored bars represent ensemble standard error.}
+\end{figure}
+
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org75cad93}
+
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 
-\begin{figure}[htbp]
-\centering
-\includegraphics[width=.9\linewidth]{../../data/figures/cesm1_sf.pdf}
-\caption{\label{fig:bootstrap_1}Difference between fully forced and single forced ensembles derived from the bootstrap process for a) aerosol emissions, b) biomass burning c) greenhouse gasses, d) land use/cover}
-\end{figure}
-
 In both the fully-forced scenario, and the greenhouse and aerosol only simulations, there is noticeably reduced Niño 3.4 variance in the mid-20\textsuperscript{th} century, below 2 standard errors of the control.
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=.5\linewidth]{../../data/figures/cesm1_sf.pdf}
+\caption{\label{fig:bootstrap_1}Difference between fully forced and single forced ensembles derived from the bootstrap process for a) aerosol emissions, b) biomass burning c) greenhouse gasses, d) land use/cover}
+\end{figure}
+
 \subsection{Correlation With Ocean Temperature}
-\label{sec:org5c7fa6f}
+
 Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
@@
 
-\begin{figure}[htbp]
+\begin{figure}
 \centering
-\includegraphics[width=.9\linewidth]{../../data/figures/tempdt.pdf}
+\includegraphics[width=.5\linewidth]{../../data/figures/tempdt.pdf}
 \caption{\label{fig:tempdt}Correlation coefficient between detrended and smoothed ocean temperature averaged along the equator and detrended and smoothed Niño 3.4 variance.}
@@
 \subsection{Wavelet Analysis}
-\label{sec:orgd4dad13}
+
 Finally, we conduct a wavelet analysis of the Niño 3.4 index for the CESM1 and CESM2 Large Ensembles.
@@
 
-\begin{figure}[htbp]
+\begin{figure}
 \centering
-\includegraphics[width=.9\linewidth]{../../data/figures/wavelet2.pdf}
+\includegraphics[width=.5\linewidth]{../../data/figures/wavelet2.pdf}
 \caption{\label{fig:wavelet_1}Ensemble mean wavelet power spectra for both large ensembles.}
@@
 
-\begin{figure}[htbp]
+\begin{figure}
 \centering
-\includegraphics[width=.9\linewidth]{../../data/figures/wavelet3.pdf}
+\includegraphics[width=.5\linewidth]{../../data/figures/wavelet3.pdf}
 \caption{\label{fig:wavelet_2}Ensemble mean wavelet power records for low and high period ENSO.}
@@
 \section{Conclusion}
-\label{sec:org272adc7}
 
@@
 \section{Acknowledgments}
-\label{sec:org9f9ff26}
 

### 2021-10-22 presentations/elevator/elevator.tex
big big big stuff lotsa stuff

% Created 2021-10-06 Wed 09:19
% Intended LaTeX compiler: pdflatex
\documentclass{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{textcomp}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\author{Ben Goldman}
\professor{Ms. Fleming}
\class{Science Research}
\author{Benjamin Goldman}
\date{\today}
\title{Elevator Speech}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Elevator Speech},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle
Hello, my name is Ben Goldman, and I am a Junior in the Science Research Program at White Plains High School. I am doing research on climate science, and I have found that El Niño, a crucial climate cycle, could intensify in the future due to global warming. El Nino, or ENSO has a great impact on human society, by regulating droughts and flooding all over the world. Research on the impact of climate change on ENSO is important to protect the human societies affected by it. My methodology involves using sets of climate models to determine changes to ENSO intensity. These models use an input of predicted greenhouse gas levels, and then output various aspects of the climate, such as sea temperature. By measuring intensity of El Niño in the model set’s output, I have predicted an increase in ENSO intensity in the mid-21st century, meaning that El Niño could be stronger in the future. Thank you for listening.
\end{document}

### 2021-10-22 presentations/presentation/notes.tex
big big big stuff lotsa stuff

% Created 2021-10-19 Tue 07:55
% Intended LaTeX compiler: pdflatex
\documentclass[little]{basic}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{wrapfig}
\usepackage{rotating}
\usepackage[normalem]{ulem}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{capt-of}
\usepackage{hyperref}
\setcounter{secnumdepth}{1}
\author{Benjamin Goldman}
\date{\today}
\title{Presentation Notes}
\hypersetup{
 pdfauthor={Benjamin Goldman},
 pdftitle={Presentation Notes},
 pdfkeywords={},
 pdfsubject={},
 pdfcreator={Emacs 27.2 (Org mode 9.5)}, 
 pdflang={English}}
\begin{document}

\maketitle

\section{General:}
\label{sec:orgbf39085}
\begin{itemize}
\item Remember to say meaning of axes and colors
\item Talk slower
\item Less detail in results, more in intro?
\end{itemize}

\section{Title}
\label{sec:orgd30528c}
\section{Introduction}
\label{sec:orgc222b61}
\subsection*{Climate Change and Variability}
\label{sec:orga02a357}
\begin{itemize}
\item Climate change: long term trends in temp, etc (fig: red line)
\item Climate variability: short (a few years) change in climate (fig: blue line)
\begin{itemize}
\item May be cyclical or random
\end{itemize}
\end{itemize}
\subsection*{Climate Forcing}
\label{sec:orga059bc9}
\begin{itemize}
\item External factors that affect climate change and/or variability
\item List factors, ghg, aer, bmb, lulc
\item Sources:
\begin{itemize}
\item ghg: industry, livestock
\item aer: industry, volcanoes (smoke, dust, sulphites?)
\item bmb: wildfires
\item lulc: deforestation, agriculture, desertification
\end{itemize}
\item Greenhouse effect
\begin{itemize}
\item gasses and particulates affect atmospheric chemistry and sunlight reflection/absorption
\item ghg absorbs ``blanket'' trapping heat (fig: orange arrows)
\item aer reflects in upper atmosphere blocking heat out. (fig: yellow arrows)
\item bmb, lulc affect reflection, absorption on surface (fig: yellow arrows)
\end{itemize}
\end{itemize}
\subsection*{El Niño (ENSO)}
\label{sec:org2234521}
\begin{itemize}
\item Temperature of the pacific ocean
\item Cold -> La Niña
\item Hot -> El Niño
\item Entire cycle: ENSO (El Niño/Southern Oscillation)
\item Affects humans: hot year, dry year, cold year, wet year
\item Figure: temperature differences between strong La Niña year and strong El Niño year
\begin{itemize}
\item Blue=colder, red=warmer
\item Point out California hot for El Niño (wildfires)
\end{itemize}
\end{itemize}
\subsection*{Method: Climate Simulation}
\label{sec:org8ab0f5f}
\begin{itemize}
\item Main way of making predictions
\item Predictions of forcing levels are fed to computers
\item Computers simulate climate on a grid of data containing temperature and much more
\item Predictions are usually run many times
\item My contain biases but are quite well tested
\end{itemize}
\subsection*{Review of Literature}
\label{sec:orga43d7c2}
\begin{itemize}
\item Slide is notes
\end{itemize}
\subsection*{Gap}
\label{sec:orgf383e67}
\begin{itemize}
\item Slide is notes
\end{itemize}
\subsection*{Questions}
\label{sec:org5727808}
\begin{itemize}
\item What, why, how
\item Slide is notes
\end{itemize}
\section{Data, Methods, and Results}
\label{sec:org815aceb}
\subsection*{Methods Overview}
\label{sec:orge5d6cab}
\begin{itemize}
\item Slide is notes
\end{itemize}
\subsection*{Role of Mentor and Student}
\label{sec:org1650fef}
\begin{itemize}
\item Slide is notes
\end{itemize}
\subsection*{Model Setup}
\label{sec:orgb994e4d}
\begin{itemize}
\item Slide is notes
\end{itemize}
\subsection*{Measuring ENSO Intensity}
\label{sec:org40dcc5e}
\begin{itemize}
\item Make sure you talk about what Niño 3.4 index is: number that represents how strong El Niño is at each time
\item Windowed variance calculates amount of variability of the Niño 3.4 index ie how intense the ENSO cycle is
\end{itemize}
\subsection*{ENSO is Becoming Stronger}
\label{sec:orgcdec18c}
\subsection*{Influence of Aerosols and Greenhouse Gasses}
\label{sec:org3b80de0}
\subsection*{Correlation With Ocean Temperature}
\label{sec:org0736ced}
\subsection*{Wavelet Analysis}
\label{sec:org2f294d1}
\section{Conclusion}
\label{sec:orgd03482b}
\subsection*{Conclusions and Discussion}
\label{sec:org4e13d7c}
\subsection*{Application, Limitation, and Next Steps}
\label{sec:org0486ffb}
\subsection*{Acknowledgments}
\label{sec:org4ae1490}
\subsection*{References}
\label{sec:orgcee0089}
\end{document}

### 2021-10-22 presentations/presentation/presentation.tex
big big big stuff lotsa stuff

--- 2021-08-24 presentations/presentation/presentation.tex
@@
 \documentclass{beamer}
+
 \title{The Impact of Anthropogenic Forcing on ENSO Amplitude}
@@
 \date{\today}
+
 \usepackage{natbib}
@@
 \usepackage{varwidth}
+
+\usetheme{ben}
+\setbeamersize{description width=0.5cm}
+
 \usetikzlibrary{arrows,snakes,backgrounds}
-\usetheme{ben}
-\renewcommand{\bibsection}{}
-\setbeamerfont{caption}{size=\small}
-\setbeamertemplate{caption}[numbered]
-% \begin{frame}{Example Frame With Text and Image}
-%   \begin{columns}
-%     \column{0.5\textwidth}
-%     \begin{itemize}
-%     \item Here we go...
-%     \end{itemize}
-%     \column{0.5\textwidth}
-%     \begin{figure}
-%       \centering
-%       \includegraphics[width=\textwidth]{figures/example.jpg}
-%       \caption{This is a very nice figure}
-%       \label{fig:this}
-%     \end{figure}
-%   \end{columns}
-% \end{frame}
+\tikzstyle{process}=[rectangle, draw=process, fill=process!20, line width = 0.3mm]
+\tikzstyle{data}=[rectangle, draw=data, fill=data!20, line width = 0.3mm]
+\tikzstyle{tight}=[node distance = 0.5in]
+\tikzstyle{loose}=[node distance = 2in]
+\definecolor{process}{HTML}{3d5f8f}
+\definecolor{data}{HTML}{8f6d3d}
 
@@
 }
+\renewcommand{\bibsection}{}
 
 \begin{document}
-
-  \tikzstyle{process}=[circle, draw=process, fill=process!20, line width = 0.3mm]
-  \tikzstyle{data}=[rectangle, draw=data, fill=data!20, line width = 0.3mm]
 
@@
 
-\begin{frame}{Climate Change}
+\begin{frame}{Climate Change and Variability}
+  \begin{columns}
+    \column{0.4\textwidth}
+    \begin{itemize}
+    \item Global warming
+    \item Long-term trends vs short-term randomness
+    \end{itemize}
+    \column{0.6\textwidth}
+    \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{this}
+  \end{columns}
+\end{frame}
 
+\begin{frame}{Climate forcing}
   \begin{columns}
+    \column{0.5\textwidth}
+    \begin{itemize}
+    \item \alert{Forcing}: any external factor that affects climate.
+      \begin{description}
+      \item[\alert{GHG}] Greenhouse gasses
+      \item[\alert{AER}] Aerosols (natural: volcanic ash, artificial: smoke)
+      \item[\alert{BMB}] Biomass burning
+      \item[\alert{LULC}] Land use/cover (deforestation, desertification)
+      \end{description}
+    \end{itemize}
     \column{0.6\textwidth}
-    \begin{itemize}
-    \item The earth is getting warmer. \citep{pachauri2014climate}
-    \item Climate varies on different scales.
-    \item Long-term trends and short-term noise.
-    \item \alert{Forcing}: any external factor that affects climate.
-      \begin{itemize}
-      \item Greenhouse gasses
-      \item Aerosols (natural: volcanic ash, artificial: smoke)
-      \item Biomass burning
-      \item Land use/cover (deforestation, desertification)
-      \end{itemize}
-    \end{itemize}
-    \column{0.4\textwidth}
-    \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{this}
+    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect.}{this}
   \end{columns}
@@
 
-\begin{frame}{Climate Simulation}
+\begin{frame}{Method: Climate Simulation}
   \begin{itemize}
   \item Run climate simulation with predicted forcing levels as input.
-  \item \alert{Ensemble:} set of repeated simulations.
+  \item \alert{Ensemble:} set of repeated simulations of the same model.
   \end{itemize}
@@
       \end{varwidth}};
-    \draw [>->] (input) to (model);
-    \draw [>->] (model) to (output);
-    \draw [>->] (model) to [out = 50, in = 330] (repeat);
-    \draw [>->] (repeat) to [out = 210, in = 130] (model);
+    \draw [->] (input) to (model);
+    \draw [->] (model) to (output);
+    \draw [->] (model) to [out = 50, in = 330] (repeat);
+    \draw [->] (repeat) to [out = 210, in = 130] (model);
   \end{tikzpicture}
@@
 
-\begin{frame}{ENSO and Climate Change}
+\begin{frame}{Review of Literature}
   \begin{itemize}
-  \item ENSO changes over time \citep{lubbecke2014assessing}.
+  \item ENSO's properties observed vary across different decades. \citep{lubbecke2014assessing}.
   \item ENSO responds to external forcing.
@@
     \begin{itemize}
-    \item Ozone emission may reduce ENSO activity \citep{nowack2017role}
-    \item Aerosol emission may reduce ENSO activity \citep{stevenson2017forced}
+    \item Ozone emissions reduce ENSO activity \citep{nowack2017role}.
+    \item Aerosol emissions modify ENSO geographical center \citep{stevenson2017forced}.
     \end{itemize}
@@
 
-\begin{frame}{Gap and Questions}
-  Gap:
+\begin{frame}{Gap}
   \begin{itemize}
@@
 
-\section{Data and Methods}
+\begin{frame}{Questions}
+  \begin{description}
+  \item[What?] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
+  \item[Why?] Is the predicted increase (or decrease) due to human activities?
+  \item[How?] What processes are causing greenhouse gasses and aerosols to affect ENSO?
+  \end{description}
+\end{frame}
+
+\section{Data, Methods, and Results}
+
+\begin{frame}{Methods Overview}
+  \begin{itemize}
+  \item Precollected predictions of sea surface temperature from climate models.
+  \item Calculate ENSO intensity in model output.
+  \item Use single forcing ensembles to estimate contributions each forcing item.
+  \item Plot correlation between ENSO intensity and ocean temperature to examine relationship between heat transfer, forcing, and ENSO.
+  \item Use wavelet analysis to analyze changes to ENSO at different frequencies.
+  \end{itemize}
+\end{frame}
 
@@
     \column{.5\textwidth}
+    Student:
+    \begin{itemize}
+    \item Analyze raw data on computer
+    \item Produce graphics for analysis and publication
+    \item Write documentation
+    \item Identify key features of results
+    \end{itemize}
+    \column{.5\textwidth}
     Mentor:
     \begin{itemize}
-    \item Suggest future methods
-    \item Conduct parallel analysis to complement student work
-    \item Provide raw precollected data
-    \item Interpret data produced by student
     \item Review student writing
-    \end{itemize}
-    \column{.5\textwidth}
-    Student:
-    \begin{itemize}
-    \item Analyze data on computer
-    \item Produce graphics for analysis and publication
-    \item Write documentation
-    \item Suggest interpretations of data
+    \item Interpret results in the context of climatology
+    \item Conduct parallel analysis
+    \item Provide raw data from facility
     \end{itemize}
@@
   \begin{itemize}
-  \item CESM1 \citep{kay2015community}
-  \item CESM2 \citep{danabasoglu2020community}
+  \item CESM1 \citep{kay2015community} and CESM2 \citep{danabasoglu2020community}
   \item Observed forcing levels from 1850-2005
@@
   \item Control simulation with pre-1850 forcing levels
+  \item Single forcing ensembles that represent influence of single factor
   \end{itemize}
@@
 
-\begin{frame}{Signal and Noise}
-  \begin{columns}
-    \column{0.3\textwidth}
-    \begin{itemize}
-      \item Butterfly effect
-      \item Use ensemble mean to remove noise.
-      \item The larger the ensemble, the less noise there is.
-    \end{itemize}
-    \column{0.7\textwidth}
-    \myfig{variance_3.pdf}{ENSO intensity for each run of the CESM1 Large Ensemble}{variance_3}
-  \end{columns}
-\end{frame}
-
 \begin{frame}{ENSO is Becoming Stronger}
@@
 
-\begin{frame}{Single Forcing Ensembles}
-
-\end{frame}
-
 \begin{frame}{Influence of Aerosols and Greenhouse Gasses}
-
+  \begin{columns}
+    \column{0.65\textwidth}
+    \begin{itemize}
+    \item Influence of each factor on ENSO amplitude.
+    \item Increased variance due to greenhouse gas emissions.
+    \item Somewhat increased variance from aerosol emissions, but not linear.
+    \item Inconclusive results from biomass burning and land use forcing
+    \end{itemize}
+    \alert{Takeaway:} Human activities are triggering predicted strengthening of ENSO.
+    \column{0.5\textwidth}
+    \myfig{cesm1_sf.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{cesm1_sf}
+  \end{columns}
 \end{frame}
@@
 \begin{frame}{Correlation With Ocean Temperature}
-
-\end{frame}
-
-\begin{frame}{Stratification}
-
+  \begin{columns}
+    \column{0.65\textwidth}
+    \begin{itemize}
+    \item Correlation coefficient between ocean temperature and ENSO amplitude.
+    \item Negative coefficient in subsurface layer.
+    \item Positive coefficient in surface layer.
+    \item Suggests that ocean stratification may be mediating global warming influence on ENSO.
+    \item Difference in heating modifies mechanics of ENSO cycle.
+    \end{itemize}
+    \column{0.5\textwidth}
+    \myfig{tempdt.pdf}{Correlation coefficient between ENSO amplitude and ocean temperature in equatorial cross-section in the fully-forced CESM1 ensemble}{tempdt}
+  \end{columns}
 \end{frame}
@@
 \begin{frame}{Wavelet Analysis}
-
-\end{frame}
-
-\begin{frame}{Power Spectrum}
-
+  \begin{columns}
+    \column{0.6\textwidth}
+    \begin{itemize}
+    \item Separate ENSO record into frequency over time.
+    \item Increase in power in late 21\textsuperscript{st} century agrees with previous results.
+    \item CESM2 shows a slight ``speeding up'' of ENSO as period decreases in late 21\textsuperscript{st} century.
+    \end{itemize}
+    \column{0.6\textwidth}
+    \myfig{wavelet2.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 and CESM2 ensembles}{wavelet2}
+  \end{columns}
 \end{frame}
@@
 
-\begin{frame}{Conclusions}
-
+\begin{frame}{Conclusions and Discussion}
+  \begin{itemize}
+  \item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
+  \item Variance increase is likely caused by the combined influence of greenhouse gasses and aerosols.
+  \item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
+  \end{itemize}
 \end{frame}
 
-\begin{frame}{Discussion}
-
+\begin{frame}{Application, Limitation, and Next Steps}
+  \begin{description}
+  \item[\alert{Application}] Improve prediction ability to help people prepare for increased likelihood of extreme weather.
+  \item[\alert{Limitation}] Niño 3.4 index may not be fully accurate for various models \citep{cai2018increased}. Also, CESM may contain biases and is not completely accurate.
+  \item[\alert{Next steps}] Examine other variables to further analyze mediator process, continue wavelet analysis methods to focus on individual frequency bands.
+  \end{description}
 \end{frame}
@@
 
-%hi
+% hi

### 2021-10-22 data/analyzeVariance.r

@@ -1,4 +1,4 @@
-# Imports
+                                        # Imports
 library(ncdf4)
 library(zoo)
 library(dplyr)
@@ -6,7 +6,7 @@ library(ggplot2)
 library(WaveletComp)
 library(reshape2)
 
-# save a plot my way
+                                        # save a plot my way
 gsave <- function(name, plot = last_plot(), dimensions = c(8, 6), path = "./figures/") {
     ggsave(name, path = path, width = dimensions[1], height = dimensions[2])
 }
@@ -27,33 +27,37 @@ ff_compare <- ff1_df %>% mutate(case = "CESM1") %>%
     bind_rows(ff2_df %>% mutate(case = "CESM2"))
 
 ggplot(cesm_1, aes(y = mean, x = date, color = case, fill = case, ymin = (mean - 2*se), ymax = (mean + 2*se)), show.legend = FALSE) +
-  geom_line() +
-  geom_ribbon(alpha = m, color = NA) +
-  facet_wrap(vars(case), dir = "v") +
-  theme(legend.position = "none") +
-  xlim(xlim1) +
-  labs(title = "CESM1 20-Year Variance", x = "Date", y = "Variance", color = "Case")
+    geom_line() +
+    geom_ribbon(alpha = m, color = NA) +
+    facet_wrap(vars(case), dir = "v") +
+    theme(legend.position = "none") +
+    xlim(xlim1) +
+    labs(title = "CESM1 20-Year Variance", x = "Date", y = "Variance", color = "Case") +
+    theme(text = element_text(size = 20))
 gsave("cesm1.pdf")
 
 ggplot(cesm_1, aes(y = mean, x = date, color = case, fill = case, ymin = (mean - 2*se), ymax = (mean + 2*se))) +
-  geom_line() +
-  geom_ribbon(alpha = l, color = NA) +
-  xlim(xlim1) +
-  labs(title = "CESM1 20-Year Variance", x = "Date", y = "Variance", color = "Case", fill = "Case")
+    geom_line() +
+    geom_ribbon(alpha = l, color = NA) +
+    xlim(xlim1) +
+    labs(title = "CESM1 20-Year Variance", x = "Date", y = "Variance", color = "Case", fill = "Case") +
+    theme(text = element_text(size = 20))
 gsave("cesm1_2.pdf")
 
[…]

### 2021-10-22 data/correlation.r

@@ -1,3 +1,4 @@
+library(pracma)
 library(ncdf4)
 library(zoo)
 library(dplyr)
@@ -5,12 +6,7 @@ library(ggplot2)
 library(WaveletComp)
 library(reshape2)
 library(abind)
-
-# subtract ensemble mean from member
-prepare <- function(x) {
-    x.mean <- x %>% apply(1, mean)
-    return(x - x.mean)
-}
+library(scales)
 
 # calculate 20-year rolling variance
 rollvar <- function(x) {
@@ -18,8 +14,22 @@ rollvar <- function(x) {
   return(out)
 }
 
+rollmean <- function(x) {
+  out <- rollapply(x, width = 12, FUN = mean, align = "center", fill = NA)
+  return(out)
+}
+
+getvar <- function(x) {
+    apply(x, 2, rollvar)
+}
+
+prepare <- function(x) {
+    x.mean <- x %>% apply(1, mean)
+    return(x - x.mean)
+}
+
 # load variance from ensemble file
-load_variance <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% apply(2, rollvar))
+load_temp <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% apply(2, rollmean))
 
 se <- function(x, ...) sd(x, ...)/sqrt(length(x))
 
@@ -65,46 +75,44 @@ na_cor <- function(x, y, ...) {
     }
 }
 
-ncin <- nc_open("/Volumes/Extreme SSD/DATA/stacked/CESM1/TEMP/b.e11.BRCP85C5CNBDRD.f09_g16.all.pop.h.TEMPDT.192001-210012.nc")
+date1 <- seq(1920, 2100, length.out = 2172)
+
+ff1 <- load_temp("./data/CESM1/ff.csv")
+ff1_df <- make_df(ff1 %>% prepare () %>% getvar(), date1)
+
+ncin <- nc_open("/Volumes/Extreme-SSD/DATA/stacked/CESM1/TEMP/b.e11.BRCP85C5CNBDRD.f09_g16.all.pop.h.TEMPDT.192001-210012.average.nc")
+
 lon <- ncvar_get(ncin, "lon", start = 21, count = 121)
 depth <- ncvar_get(ncin, "depth") / 100
 date <- ncvar_get(ncin, "time") / 365
 
-tempdt <- ncvar_get(ncin, "TEMPDT", start = c(21, 1, 1, 1), count = c(121, -1, -1, 40), verbose = TRUE)
[…]

### 2021-10-22 data/correlation_old.r

@@ -1,0 +1,163 @@
+library(ncdf4)
+library(zoo)
+library(dplyr)
+library(ggplot2)
+library(WaveletComp)
+library(reshape2)
+library(abind)
+
+# subtract ensemble mean from member
+prepare <- function(x) {
+    x.mean <- x %>% apply(1, mean)
+    return(x - x.mean)
+}
+
+# calculate 20-year rolling variance
+rollvar <- function(x) {
+  out <- rollapply(x, width = 240, FUN = var, align = "center", fill = NA)
+  return(out)
+}
+
+# load variance from ensemble file
+load_variance <- function(file) (read.table(file, sep = ",") %>% as.matrix %>% apply(2, rollvar))
+
+se <- function(x, ...) sd(x, ...)/sqrt(length(x))
+
+# bind ensemble statistics into dataframe while preserving date offset
+bind_stats <- function(this.date, this.mean, this.se, date.start = NA, date.end = NA) {
+    if(length(this.mean) == length(this.date)) {
+        this.df <- data.frame(date = this.date, mean = this.mean, se = this.se)
+    } else {
+        if(is.na(date.end)){
+            date.end <- length(this.date)
+        }
+        if(is.na(date.start)){
+            date.start <- 0
+        }
+        this.df <- data.frame(date = this.date, mean = c(
+                                                    rep(NA, date.start),
+                                                    this.mean,
+                                                    rep(NA, (length(this.date) - date.end))),
+                              se =
+                                  c(rep(NA, date.start),
+                                    this.se,
+                                    rep(NA, (length(this.date) - date.end))))
+        writeLines(as.character(date.start))
+        writeLines(as.character(date.end))
+        return(this.df)
+    }
+}
+
+# Produce ensemble stats dataframe
+make_df <- function(this, this.date, date.start = NA, date.end = NA) {
+    this.mean <- this %>% apply(1, mean)
+    this.se <- this %>% apply(1, se)
+    writeLines(as.character(length(this.mean)))
[…]

### 2021-10-22 data/intro_fig.r

@@ -5,9 +5,9 @@ library(ggplot2)
 library(reshape2)
 library(scico)
 
-ncin1 <- nc_open("/Volumes/Extreme SSD/DATA/HadISST_sst.nc")
+ncin <- nc_open("/Volumes/Extreme-SSD/DATA/HadISST_sst.nc")
 
-sst <- ncvar_get(ncin1, "sst")
+sst <- ncvar_get(ncin, "sst")
 sst[sst == -1000] <- NA
 
 date <- ncvar_get(ncin, "time")
@@ -80,25 +80,27 @@ ggplot(sst_anom_compare %>% filter(latitude < 60 &
     labs(title = "Pacific SST Anomalies Comparison",
          x = "Longitude",
          y = "Latitude",
-         fill = " SST \n Anomaly \n (°C)")
+         fill = " SST \n Anomaly \n (°C)") +
+    theme(text = element_text(size = 20))
 
-gsave("intro_fig.pdf")
+gsave("intro_fig.pdf", dimensions = c(8, 8))
 
-ggplot(sst_anom_compare %>% filter(latitude < 60 &
-                           latitude > -60 &
-                           longitude > 50 &
-                           longitude < 350), aes(longitude, latitude, z = value)) +
-    geom_contour_filled(bins = 20) +
-    facet_wrap(vars(case), nrow = 2, ncol = 1) +
-    scale_fill_scico_d(palette = "vik") +
-    labs(title = "Pacific SST Anomalies Comparison",
-         x = "Longitude",
-         y = "Latitude",
-         fill = " SST \n Anomaly \n (°C)")
+## ggplot(sst_anom_compare %>% filter(latitude < 60 &
+##                            latitude > -60 &
+##                            longitude > 50 &
+##                            longitude < 350), aes(longitude, latitude, z = value)) +
+##     geom_contour_filled(bins = 20) +
+##     facet_wrap(vars(case), nrow = 2, ncol = 1) +
+##     scale_fill_scico_d(palette = "vik") +
+##     labs(title = "Pacific SST Anomalies Comparison",
+##          x = "Longitude",
+##          y = "Latitude",
+##          fill = " SST \n Anomaly \n (°C)") +
+##     theme(text = element_text(size = 20))
 
-gsave("intro_fig_2.pdf")
+## gsave("intro_fig_2.pdf")
 
[…]

### 2021-10-22 data/varcalc.r

@@ -161,17 +161,16 @@ ff_compare$case <- as.factor(ff_compare$case)
 # Bootstrap process to generate sf timeseries
 reps <- 100
 
-ghg1_sf <- subtract.bootstrap.raw(ff1, ghg1, reps) %>% getvar() %>% make_df_manvar(20, date2, date.start = 840)
-aer1_sf <- subtract.bootstrap.raw(ff1, aer1, reps) %>% getvar() %>% make_df_manvar(20, date2, date.start = 840)
-bmb1_sf <- subtract.bootstrap.raw(ff1, bmb1, reps) %>% getvar() %>% make_df_manvar(15, date2, date.start = 840)
-luc1_sf <- subtract.bootstrap.raw(ff1, luc1, reps) %>% getvar() %>% make_df_manvar(5, date2, date.start = 840)
+ghg1_sf <- subtract.bootstrap.raw(ff1 %>% getvar(), ghg1 %>% getvar(), reps) %>% make_df_manvar(20, date2, date.start = 840)
+aer1_sf <- subtract.bootstrap.raw(ff1 %>% getvar(), aer1 %>% getvar(), reps) %>% make_df_manvar(20, date2, date.start = 840)
+bmb1_sf <- subtract.bootstrap.raw(ff1 %>% getvar(), bmb1 %>% getvar(), reps) %>% make_df_manvar(15, date2, date.start = 840)
+luc1_sf <- subtract.bootstrap.raw(ff1 %>% getvar(), luc1 %>% getvar(), reps) %>% make_df_manvar(5, date2, date.start = 840)
 
 # bootstrap ensemble statistics in single dataframe
-cesm1_sf <- data.frame(date = ff1_df$date, mean = ff1_df$mean, se = ff1_df$se, case = "FF") %>%
-    bind_rows(ghg1_sf %>% mutate(case = "GHG")) %>%
-    bind_rows(aer1_sf %>% mutate(case = "AER")) %>%
-    bind_rows(bmb1_sf %>% mutate(case = "BMB")) %>%
-    bind_rows(luc1_sf %>% mutate(case = "LUC"))
+cesm1_sf <- ghg1_sf %>% mutate(case = "GHG") %>%
+        bind_rows(aer1_sf %>% mutate(case = "AER")) %>%
+        bind_rows(bmb1_sf %>% mutate(case = "BMB")) %>%
+        bind_rows(luc1_sf %>% mutate(case = "LUC"))
[…]

### 2021-10-22 data/wavelets2.r

@@ -1,4 +1,4 @@
-# Imports
+                                        # Imports
 library(ncdf4)
 library(zoo)
 library(dplyr)
@@ -12,55 +12,115 @@ prepare <- function(x) {
     return(x - x.mean)
 }
 
-ff <- read.table("./data/CESM1/ff.csv", sep = ",") %>% prepare() %>% rollapply(width = 12, FUN = mean, align = "center") %>% as.matrix()
-ghg <- read.table("./data/CESM1/ghg.csv", sep = ",") %>% prepare() %>% rollapply(width = 12, FUN = mean, align = "center")
-aer <- read.table("./data/CESM1/aer.csv", sep = ",") %>% prepare() %>% rollapply(width = 12, FUN = mean, align = "center")
-bmb <- read.table("./data/CESM1/bmb.csv", sep = ",") %>% prepare() %>% rollapply(width = 12, FUN = mean, align = "center")
-luc <- read.table("./data/CESM1/luc.csv", sep = ",") %>% prepare() %>% rollapply(width = 12, FUN = mean, align = "center")
+se <- function(x, ...) sd(x, ...)/sqrt(length(x))
 
-
-ff_mean <- ff %>% apply(1, mean) %>% as.data.frame()
-
-ff_w_raw <- ff %>% apply(2, WaveletTransform, dt = 1/12, lowerPeriod = 1)
-
-ff_w <- array(NA, dim = c(119, 2161, 40))
-for(r in 1:40) {
-    writeLines(as.character(r))
-    ff_w[, , r] <- ff_w_raw[[r]]$Power
+wavelet_ensemble <- function(x) {
+    wv_raw <- x %>% apply(2, WaveletTransform, dt = 1/12, lowerPeriod = 1)
+    nr <- wv_raw[[1]]$nr
+    t <- dim(x)[1]
+    n <- dim(x)[2]
+    pwr <- array(NA, dim = c(nr, t, n))
+    writeLines(as.character(dim(pwr)))
+    for(r in 1:n) {
+        writeLines(as.character(r))
+        pwr[,,r] <- wv_raw[[r]]$Power
+    }
+    return(pwr)
 }
 
-mydate <- seq(1920, 2099, length.out = 2161)
-period <- ff_w_raw[[1]]$Period
-nr <- ff_w_raw[[1]]$nr
+bind_stats <- function(this.date, this.mean, this.se, date.start = NA, date.end = NA) {
+    if(length(this.mean) == length(this.date)) {
[…]

### 2021-10-31 paperwork/abstract/abstract.tex
more stuff woohoo

@@
-% Created 2021-05-10 Mon 19:08
+% Created 2021-10-28 Thu 18:11
 % Intended LaTeX compiler: pdflatex
@@
 \usepackage{hyperref}
+\usepackage[margin=1in]{geometry}
 \date{\today}
@@
 \hypersetup{
- pdfauthor={John Doe},
+ pdfauthor={Benjamin Goldman},
  pdftitle={Abstract},
@@
 \begin{center}
+\textbf{The Impact of Industrial Emissions on ENSO Amplitude}
 
-{\bf The Impact of Industrial Emissions on ENSO Amplitude}
-{\bf Benjamin Goldman}
-
+\textbf{Benjamin Goldman}
 \end{center}
@@
 The El Niño/Southern Oscillation (ENSO) is the dominant mode of interannual climate variability, with substantial associated global socio-economic impacts.
-Due to their significance, shifts in ENSO under climate change also have the potential to substantial impact human society and natural ecosystems.
-However, it is currently unclear what effect greenhouse gas (GHG) and industrial aerosol (AER) emissions will have on ENSO and even less clear what effect a combination of these factors might have when changing in tandem.
-This study examines transient changes to ENSO variance under a variety of forcing scenarios using the CESM1 Large and Single-Forcing Ensembles.
-These multi-member ensembles span the historical record (1920-2005) and much of the 21st C (2006-2080 for GHG/AER).
+Due to their significance, shifts in ENSO under climate change also have the potential to substantially impact human society and natural ecosystems.
+However, it is currently unclear what effect greenhouse gas and industrial aerosol emissions have on ENSO, as well as what effect these factors have when combined.
+This study examined transient changes to ENSO variance under a variety of forcing scenarios using the CESM1 and CESM2 Large and Single-Forcing Ensembles.
+These multi-member ensembles span the historical record (1920-2005) and much of the 21st century (2006-2080 for GHG/AER).
 A 2000-year pre-industrial (PI) control simulation is used to account for model drift and 20-year running variance of the Niño 3.4 SST index is used as a proxy for ENSO variance.
-The ensemble mean and standard error of each ensemble is calculated, to estimate the statistical significance of simulated changes.
-We identify significant increases in variance under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
-Computing detrended and smoothed correlation between 20-year variance of the Niño 3.4 index and ocean temperature in the tropical Pacific Ocean reveals that stratification caused by global warming is connected with increased ENSO amplitude.
+The ensemble mean and standard error of each ensemble was calculated, while the Probability Density Function (PDF) is computed for the PI control simulation to estimate the statistical significance of simulated changes.
+We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
+Finally, we examined the wavelet power spectrum of the historical and predicted Niño 3.4 index to support the data given from the 20-year variance calculation.
+We identified significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
+
+\begin{center}
+\textbf{Presentations}
+\end{center}
+
+\textbf{2020:} \emph{Virtual Poster Board presenter at the White Plains High School Science Research Symposium (June)}
+
+\textbf{2021:} \emph{Virtual Powerpoint presenter at the Westchester Science and Engineering Fair (March)}
+
+\begin{center}
+\textbf{Awards}
+\end{center}
+
+\textbf{2021:} \emph{American Meteorological Society Award at the Westchester Science and Engineering Fair}
+
+\textbf{2021:} \emph{4\textsuperscript{th} Place in Earth Science at the Westchester Science and Engineering Fair}
+
+\begin{center}
+\textbf{Mentor Information}
+
+Dr. John Fasullo
+
+National Center for Atmospheric Research
+
+Boulder, Colorado
+\end{center}
 \end{document}

### 2021-10-31 paperwork/paper/paper.tex
more stuff woohoo

@@
 We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
-We identifed significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
+Finally, we examined the wavelet power spectrum of the historical and predicted Niño 3.4 index to support the data given from the 20-year variance calculation.
+We identified significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
 \end{abstract}
@@
 The Probability Distribution Function of Niño 3.4 variance is estimated for AMO/C > 1/2\(\sigma\) and < -1/-2\(\sigma\) using a Kernel Density Estimation.
-No meaningful differences were found in the distribution of ENSO intensity under any of these conditions (See Suplementary information figure 1).
+No meaningful differences were found in the distribution of ENSO intensity under any of these conditions (See Supplementary information figure 1).
+
+\begin{figure}
+\centering
+\includegraphics[width=.5\linewidth]{../../data/figures/cesm1.pdf}
+\caption{\label{fig:variance_1}20-year windowed variance of the Niño 3.4 index, for ensembles a) full-forcing, b) xghg, c) xaer, d) xbmb, e) xlulc. Colored bars represent ensemble standard error.}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=.5\linewidth]{../../data/figures/ff_compare.pdf}
+\caption{\label{fig:variance_2}20-year windowed variance of the Niño 3.4 index for the CESM1 and CESM2 fully-forced Large Ensembles. Colored bars represent ensemble standard error.}
+\end{figure}
 
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org75cad93}
+
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=.5\linewidth]{../../data/figures/cesm1_sf.pdf}
+\caption{\label{fig:bootstrap_1}Difference between fully forced and single forced ensembles derived from the bootstrap process for a) aerosol emissions, b) biomass burning c) greenhouse gasses, d) land use/cover}
+\end{figure}
+
 \subsection{Correlation With Ocean Temperature}
-\label{sec:org5c7fa6f}
-Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
+
+Next, we analyzed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
 To do this we used 4 slices of ocean temperature from the fully-forced, xghg, and xaer ensembles, including a slice averaged along the equator and slices through the western, central, and eastern Pacific basins.
-We linearly detrend and smooth with a 30-year windowed mean the timeseries at each gridpoint and the Niño 3.4 variance.
-Next, we calculated the Pearson's correlation coefficient between each gridpoint and the Niño 3.4 variance timeseries.
+We linearly detrend and smooth with a 30-year windowed mean the timeseries at each grid-point and the Niño 3.4 variance.
+Next, we calculated the Pearson's correlation coefficient between each grid-point and the Niño 3.4 variance timeseries.
 The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt}.

### 2021-10-31 paperwork/paper_old/paper.tex
more stuff woohoo

--- 2021-10-22 paperwork/paper/paper.tex
@@
-% Created 2021-10-14 Thu 09:16
+% Created 2021-10-28 Thu 09:33
 % Intended LaTeX compiler: pdflatex
@@
 \maketitle
-\tableofcontents
-
 \begin{abstract}
@@
 We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
-We identifed significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
+We identified significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:org2a498ff}
+\label{sec:orgfd64463}
 
@@
 \section{Data}
-\label{sec:org7ef515b}
+\label{sec:orgf0dfd76}
 
@@
 \section{Methods and Results}
-\label{sec:org2868dc1}
+\label{sec:org9a01f67}
 
 \subsection{Niño 3.4 Variance}
-\label{sec:org65b5cbb}
+\label{sec:org4dc417f}
 We estimate 20\textsuperscript{th} and 21\textsuperscript{st} century ENSO amplitude for each model in the fully-forced, single-forced, and PI control CESM1 and CESM2 ensembles using the variance of the Niño 3.4 region of the Pacific ocean (5N-5S, 170W-120W).
@@
 The AMO has been shown to have an influence on ENSO strength and seasonal growth rate \citep{levine2017impact}.
-We filter the control Niño 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diognostics Package (CVDP) \citep{phillips2014evaluating}.
+We filter the control Niño 3.4 variance data according to the strength of the AMO and AMOC, using records of AMO/C strength in the Climate Variability Diagnostics Package (CVDP) \citep{phillips2014evaluating}.
 The Probability Distribution Function of Niño 3.4 variance is estimated for AMO/C > 1/2\(\sigma\) and < -1/-2\(\sigma\) using a Kernel Density Estimation.
-No meaningful differences were found in the distribution of ENSO intensity under any of these conditions (See Suplementary information figure 1).
+No meaningful differences were found in the distribution of ENSO intensity under any of these conditions (See Supplementary information figure 1).
 
 \subsection{Single Forcing Scenarios and the Bootstrap Process}
-\label{sec:org75cad93}
+\label{sec:org2956178}
 We analyze the role of individual factors using the CESM Single Forcing Ensembles.
@@
 \subsection{Correlation With Ocean Temperature}
-\label{sec:org5c7fa6f}
-Next, we analyzeed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
+\label{sec:org1985d36}
+Next, we analyzed the correlation between ENSO amplitude and changes to ocean structure in the CESM1 ensembles.
 To do this we used 4 slices of ocean temperature from the fully-forced, xghg, and xaer ensembles, including a slice averaged along the equator and slices through the western, central, and eastern Pacific basins.
-We linearly detrend and smooth with a 30-year windowed mean the timeseries at each gridpoint and the Niño 3.4 variance.
-Next, we calculated the Pearson's correlation coefficient between each gridpoint and the Niño 3.4 variance timeseries.
+We linearly detrend and smooth with a 30-year windowed mean the timeseries at each grid-point and the Niño 3.4 variance.
+Next, we calculated the Pearson's correlation coefficient between each grid-point and the Niño 3.4 variance timeseries.
 The correlation coefficients for the equatorial slice are shown in figure \ref{fig:tempdt}.
@@
 \subsection{Wavelet Analysis}
-\label{sec:orgd4dad13}
+\label{sec:orgcd04aa3}
 Finally, we conduct a wavelet analysis of the Niño 3.4 index for the CESM1 and CESM2 Large Ensembles.
@@
 \section{Conclusion}
-\label{sec:org272adc7}
+\label{sec:orgf6b6554}
 
@@
 \section{Acknowledgments}
-\label{sec:org9f9ff26}
+\label{sec:orgb2d1109}
 

### 2021-10-31 presentations/presentation/presentation.tex
more stuff woohoo

@@
 \usepackage{varwidth}
+\usepackage{hyperref}
 
@@
     \column{0.6\textwidth}
-    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect.}{this}
+    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect. \url{https://www.coolaustralia.org/the-greenhouse-effect-secondary}}{this}
   \end{columns}
@@
     \item Increase in power in late 21\textsuperscript{st} century agrees with previous results.
-    \item CESM2 shows a slight ``speeding up'' of ENSO as period decreases in late 21\textsuperscript{st} century.
     \end{itemize}

### 2021-11-03 presentations/presentation/notes.tex
woohoo presentation changes i think

@@
-% Created 2021-10-19 Tue 07:55
+% Created 2021-10-31 Sun 14:05
 % Intended LaTeX compiler: pdflatex
@@
 \section{General:}
-\label{sec:orgbf39085}
+\label{sec:orgf4efe11}
 \begin{itemize}
@@
 \section{Title}
-\label{sec:orgd30528c}
+\label{sec:org197abb8}
 \section{Introduction}
-\label{sec:orgc222b61}
+\label{sec:org7e341be}
 \subsection*{Climate Change and Variability}
-\label{sec:orga02a357}
+\label{sec:orgeb7a331}
 \begin{itemize}
@@
 \subsection*{Climate Forcing}
-\label{sec:orga059bc9}
+\label{sec:org08234e7}
 \begin{itemize}
@@
 \subsection*{El Niño (ENSO)}
-\label{sec:org2234521}
+\label{sec:org0ec8ade}
 \begin{itemize}
@@
 \subsection*{Method: Climate Simulation}
-\label{sec:org8ab0f5f}
+\label{sec:org902ab02}
 \begin{itemize}
@@
 \subsection*{Review of Literature}
-\label{sec:orga43d7c2}
+\label{sec:orgf798249}
 \begin{itemize}
@@
 \subsection*{Gap}
-\label{sec:orgf383e67}
+\label{sec:org177c1d6}
 \begin{itemize}
@@
 \subsection*{Questions}
-\label{sec:org5727808}
+\label{sec:org1ab33ee}
 \begin{itemize}
@@
 \section{Data, Methods, and Results}
-\label{sec:org815aceb}
+\label{sec:org84e14f6}
 \subsection*{Methods Overview}
-\label{sec:orge5d6cab}
+\label{sec:org86e31d8}
 \begin{itemize}
@@
 \subsection*{Role of Mentor and Student}
-\label{sec:org1650fef}
+\label{sec:org2e78332}
 \begin{itemize}
@@
 \subsection*{Model Setup}
-\label{sec:orgb994e4d}
+\label{sec:org0f6ba67}
 \begin{itemize}
@@
 \subsection*{Measuring ENSO Intensity}
-\label{sec:org40dcc5e}
+\label{sec:orge16e5c1}
 \begin{itemize}
@@
 \subsection*{ENSO is Becoming Stronger}
-\label{sec:orgcdec18c}
+\label{sec:orgb07762a}
+\begin{itemize}
+\item Graph axes and colors CESM1 and CESM2
+\item Both have increase in ENSO intensity
+\item Slowdown/decrease after 2050
+\begin{itemize}
+\item Aerosol emissions
+\end{itemize}
+\end{itemize}
 \subsection*{Influence of Aerosols and Greenhouse Gasses}
-\label{sec:org3b80de0}
+\label{sec:org2b52160}
+\begin{itemize}
+\item Data source: single forcing ensembles
+\item Explain figure axes, subplots
+\item Biggest, constant increase in GHG
+\item Significant changes in AER, but not one direction
+\item Insignificant changes for others
+\item Takeaway
+\end{itemize}
 \subsection*{Correlation With Ocean Temperature}
-\label{sec:org0736ced}
+\label{sec:orgbdaf9d4}
+\begin{itemize}
+\item Calculated correlation coefficient with ocean temp. and ENSO amplitude
+\item CESM1 only so far
+\item Plot:
+\begin{itemize}
+\item Cross section of pacific along equator, x:lon, y:depth, color:correlation
+\item Positive coefficient in surface layer, negative coefficient in subsurface layer
+\item Stratification
+\item More work necessary
+\end{itemize}
+\end{itemize}
 \subsection*{Wavelet Analysis}
-\label{sec:org2f294d1}
+\label{sec:orga32dcde}
+\begin{itemize}
+\item WA = mathematical process that takes a signal and derives changes to each frequency over time
+\item Plot:
+\begin{itemize}
+\item axes, color, subplots
+\item Confirm results from previous steps
+\end{itemize}
+\end{itemize}
 \section{Conclusion}
-\label{sec:orgd03482b}
+\label{sec:org46c3af5}
 \subsection*{Conclusions and Discussion}
-\label{sec:org4e13d7c}
+\label{sec:org10adbad}
+\begin{itemize}
+\item Slide is notes
+\end{itemize}
 \subsection*{Application, Limitation, and Next Steps}
-\label{sec:org0486ffb}
+\label{sec:org10fbd6a}
+\begin{itemize}
+\item Slide is notes
+\end{itemize}
 \subsection*{Acknowledgments}
-\label{sec:org4ae1490}
+\label{sec:orgf4b1b34}
+\begin{itemize}
+\item Thank you to NCAR for running the climate models
+\item Thank you to my teacher, family, and mentor
+\item Thank you to the software makers
+\end{itemize}
 \subsection*{References}
-\label{sec:orgcee0089}
+\label{sec:orgd9329c4}
 \end{document}

### 2021-11-03 presentations/presentation/presentation.tex
woohoo presentation changes i think

@@
 \maketitle
-
-\section{Introduction}
 
@@
 
-\begin{frame}{Method: Climate Simulation}
-  \begin{itemize}
-  \item Run climate simulation with predicted forcing levels as input.
-  \item \alert{Ensemble:} set of repeated simulations of the same model.
-  \end{itemize}
-  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
-    \node (input) [data] {
-      \begin{varwidth}{1.5in}
-        Independent variables (forcing)
-      \end{varwidth}};
-    \node (model) [right of=input] [process] {
-      \begin{varwidth}{\linewidth}
-        Computer\\Simulation
-      \end{varwidth}};
-    \node (output) [right of=model] [data] {
-      \begin{varwidth}{1.4in}
-        Dependent variables (climatology)
-      \end{varwidth}};
-    \node (repeat) [above of=model, node distance = 1in] [process] {
-      \begin{varwidth}{0.5in}
-        Repeat
-      \end{varwidth}};
-    \draw [->] (input) to (model);
-    \draw [->] (model) to (output);
-    \draw [->] (model) to [out = 50, in = 330] (repeat);
-    \draw [->] (repeat) to [out = 210, in = 130] (model);
-  \end{tikzpicture}
-\end{frame}
-
 \begin{frame}{Review of Literature}
@@
   \item Models show possible increasing ENSO activity in the future \citep{zheng2017response} and \citep{maher2018enso}.
-  \item Factors other than CO$_2$ can affect ENSO.
+  \item Factors other than CO\textsubscript{2} can affect ENSO.
     \begin{itemize}
@@
 \end{frame}
-
-\section{Data, Methods, and Results}
 
@@
 
-\section{Conclusion}
-
 \begin{frame}{Conclusions and Discussion}

### 2021-11-08 presentations/presentation/presentation.tex
fixed some figs, edited presentation

@@
   \begin{columns}
-    \column{0.5\textwidth}
+    \column{0.4\textwidth}
     \begin{itemize}
@@
     \end{itemize}
-    \column{0.5\textwidth}
+    \column{0.6\textwidth}
     \myfig{ff_compare.pdf}{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}{ff_compare}
@@
     \item Somewhat increased variance from aerosol emissions, but not linear.
-    \item Inconclusive results from biomass burning and land use forcing
     \end{itemize}
@@
     \column{0.5\textwidth}
-    \myfig{cesm1_sf.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{cesm1_sf}
+    \myfig{cesm1_sf_4.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{cesm1_sf_4}
   \end{columns}
@@
   \begin{columns}
-    \column{0.6\textwidth}
+    \column{0.5\textwidth}
     \begin{itemize}
-    \item Separate ENSO record into frequency over time.
+    \item Separate ENSO record into changes in period over time.
     \item Increase in power in late 21\textsuperscript{st} century agrees with previous results.
+    \item In CESM1, increase in ENSO intensity is mainly strengthening of longer-period cycle.
+    \item In CESM2, longer-period ENSO weakens after 2025.
     \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{wavelet2.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 and CESM2 ensembles}{wavelet2}
+    \column{0.5\textwidth}
+    \myfig{wavelet3.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 and CESM2 ensembles}{wavelet2}
   \end{columns}
+\end{frame}
+
+% SPLIT UP CONCLUSIONS DISCUSSION
+
+\begin{frame}{Discussion}
+  \begin{itemize}
+  \item Rising greenhouse gas levels increase Pacific Ocean stratification, strengthening ENSO cycle.
+  \item Aerosol influence nonlinear because aerosol levels are not purely increasing.
+  \item Stronger ENSO will lead to greater climate variability and extreme weather.
+  \item
+  \end{itemize}
 \end{frame}

### 2021-11-08 data/analyzeVariance.r

@@ -97,3 +97,12 @@ ggplot(cesm1_sf %>% filter(case != "LUC"), aes(y = mean, x = date, color = case,
     labs(title = "CESM1 20-Year Variance", subtitle = "Single Forcing Net Impact", x = "Date", y = "Variance", color = "Case", fill = "Case") +
     theme(text = element_text(size = 20))
 gsave("cesm1_sf_3.pdf")
+
+ggplot(cesm1_sf %>% filter(case == "AER" | case == "GHG"), aes(y = mean, x = date, color = case, fill = case, ymin = (mean - 2 * se), ymax = (mean + 2 * se))) +
+    facet_wrap(vars(case), dir = "v") +
+    geom_line() +
+    geom_ribbon(alpha = l, color = NA) +
+    xlim(xlim1) +
+    labs(title = "CESM1 20-Year Variance", subtitle = "Single Forcing Net Impact", x = "Date", y = "Variance", color = "Case", fill = "Case") +
+    theme(text = element_text(size = 20))
+gsave("cesm1_sf_4.pdf")

### 2021-11-08 data/wavelets2.r

@@ -112,15 +112,15 @@ ff1_high <- ff1_pwr[34:61, ,] %>% apply(c(2, 3), mean) %>% make_df(mydate1, date
 ff2_low <- ff2_pwr[21:33, ,] %>% apply(c(2, 3), mean) %>% make_df(mydate2, date.start = 840)
 ff2_high <- ff2_pwr[34:61, ,] %>% apply(c(2, 3), mean) %>% make_df(mydate2, date.start = 840)
 
-ff_stuff <- ff1_low %>% as.data.frame() %>% mutate(ensemble = "CESM1", group = "p < 3") %>%
-    bind_rows(ff1_high %>% as.data.frame() %>% mutate(ensemble = "CESM1", group = "p > 3")) %>%
-    bind_rows(ff2_low %>% as.data.frame() %>% mutate(ensemble = "CESM2", group = "p < 3")) %>%
-    bind_rows(ff2_high %>% as.data.frame() %>% mutate(ensemble = "CESM2", group = "p > 3"))
+ff_stuff <- ff1_low %>% as.data.frame() %>% mutate(ensemble = "CESM1", group = "T < 3") %>%
+    bind_rows(ff1_high %>% as.data.frame() %>% mutate(ensemble = "CESM1", group = "T > 3")) %>%
+    bind_rows(ff2_low %>% as.data.frame() %>% mutate(ensemble = "CESM2", group = "T < 3")) %>%
+    bind_rows(ff2_high %>% as.data.frame() %>% mutate(ensemble = "CESM2", group = "T > 3"))
 
 ggplot(ff_stuff, aes(y = mean, x = date, color = group, fill = group, ymin = (mean - 2 * se), ymax = (mean + 2 * se))) +
     geom_line() +
     geom_ribbon(alpha = l, color = NA) +
     facet_wrap(vars(ensemble), dir = "v") +
-    labs(title = "Mean Wavelet Power", x = "Date", y = "Power", color = "Case", fill = "Case") +
+    labs(title = "Mean Wavelet Power", x = "Date", y = "Power", color = "Period", fill = "Period") +
     theme(text = element_text(size = 20))
 gsave("wavelet3.pdf")

### 2021-12-15 paperwork/abstract/abstract.tex
wooohooooo here goes

@@
-% Created 2021-10-28 Thu 18:11
+% Created 2021-11-30 Tue 09:32
 % Intended LaTeX compiler: pdflatex
@@
 The ensemble mean and standard error of each ensemble was calculated, while the Probability Density Function (PDF) is computed for the PI control simulation to estimate the statistical significance of simulated changes.
-We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
-Finally, we examined the wavelet power spectrum of the historical and predicted Niño 3.4 index to support the data given from the 20-year variance calculation.
-We identified significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
+I calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
+Finally, I examined the wavelet power spectrum of the historical and predicted Niño 3.4 index to support the data given from the 20-year variance calculation.
+I identified significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
 

### 2021-12-15 paperwork/paper/paper.tex
wooohooooo here goes

@@
 The ensemble mean and standard error of each ensemble was calculated, while the Probability Density Function (PDF) is computed for the PI control simulation to estimate the statistical significance of simulated changes.
-We calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
-Finally, we examined the wavelet power spectrum of the historical and predicted Niño 3.4 index to support the data given from the 20-year variance calculation.
-We identified significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come.
+I calculated the correlation coefficient between ocean temperature in the equatorial Pacific and Niño 3.4 under various forcing conditions, concluding that Pacific stratification likely is tied to changes to ENSO amplitude.
+Finally, I examined the wavelet power spectrum of the historical and predicted Niño 3.4 index to support the data given from the 20-year variance calculation.
+I identified significant increases in variance of the Niño 3.4 index under full-forcing conditions during the historical record and in the future simulations and attribute these mainly to changes in GHG, with the potential emergence of AER-driven increases in the decades to come, some of which may conflict with GHG-driven changes.
 \end{abstract}

### 2021-12-15 paperwork/researchplan/plan.tex
removed archive to separate database, reorganized some stuff
wooohooooo here goes

--- 2021-05-23 paperwork/ResearchPlan2/plan.tex
@@
-% Created 2021-05-21 Fri 08:44
+% Created 2021-12-06 Mon 09:29
 % Intended LaTeX compiler: pdflatex
@@
 \usepackage{natbib}
+\usepackage[letterpaper, margin=1in]{geometry}
 \author{Benjamin Goldman}
@@
 \section{Rationale}
-\label{sec:org38aeab9}
+\label{sec:orge884475}
 
@@
 Past methods will be repeated on this dataset and results will be compared to those from the CESM1, with the hope that the CESM2 results will provide insight into those descrepancies that appear in the CESM1 output.
-Additionally, deeper analysis of the role of air currents and ocean structure will be conducted for each ensemble to further clarify the processes that mediate the aformentioned forced effect on ENSO amplitude.
+Additionally, deeper analysis of the role of air currents and ocean structure will be conducted for each ensemble to further clarify the processes that mediate the aforementioned forced effect on ENSO amplitude.
 
 \section{Research Goals}
-\label{sec:orgc5ea83c}
+\label{sec:org3779bf2}
 \begin{itemize}
@@
 \section{Methods}
-\label{sec:org42d33f3}
+\label{sec:org4b357da}
 \subsection{Materials and Data}
-\label{sec:orgb6a9935}
+\label{sec:orgab15497}
 \begin{itemize}
@@
 \subsection{Procedures}
-\label{sec:org1d14759}
+\label{sec:org0ae14af}
 \emph{Methods are subject to change in response to results}
 \begin{itemize}
-\item Prodedures from previous version of this study will be repeated using the Python and R languages.
-\item ENSO amplitude will be measured in each simulation by calculating the 20-year running variance of the Niño 3.4 index.
-\item A bootstrap process will be used to isolate the influence of single forcings, comparing their contributions in the context of global warming.
-\item Correlation coefficient between Niño 3.4 variance and sea temperature will be calculated for each simulation.
-\item A mixed-layer heat budget analysis will be used to further clarify the role of heat transfer in determining changes to ENSO under global warming.
+\item Wavelet transform will be conducted on the Niño 3.4 index of each member of the CESM1 and CESM2 ensembles.
+\item The mean wavelet power spectrum will be calculated and graphed
+\item The mean power spectrum will be divided into two frequency bands, that with a period greater than 3 years, and one with a period less than three years.
+\item The ensemble mean and standard error will be calculated for both frequency bands
 \end{itemize}
@@
 \subsection{Risk and Safety}
-\label{sec:org7d700d5}
+\label{sec:orgc939962}
 No risk/safety issues relevant as all methods are digital.
@@
 \section{Role of Mentor and Student}
-\label{sec:org841d5d2}
+\label{sec:org7d01d61}
 \begin{center}
@@
 \hline
-Perform all calculations & Discuss and interperet results & Review results and methods\\
+Perform all calculations & Discuss and interpret results & Review results and methods\\
 Produce plots and documents &  & Provide large datasets\\
+ &  & Check student work\\
 \end{tabular}

### 2021-12-15 presentations/board/left.tex
wooohooooo here goes

--- 2021-11-03 presentations/presentation/presentation.tex
@@
 \usepackage{hyperref}
-
-\usetheme{ben}
-\setbeamersize{description width=0.5cm}
+\usepackage[scale=1.4]{beamerposter}
+\geometry{papersize={20in,48in}}
+\usepackage{lipsum}
 
 \usetikzlibrary{arrows,snakes,backgrounds}
+\usetikzlibrary{positioning}
 \tikzstyle{process}=[rectangle, draw=process, fill=process!20, line width = 0.3mm]
@@
 
-\newcommand{\myfig}[3]{
+\usetheme{ben}
+
+\newcommand{\myfig}[4]{
   \begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{figures/#1}
+    \includegraphics[width=#3\textwidth]{figures/#1}
     \caption{#2}
-    \label{fig:#3}
+    \label{fig:#4}
   \end{figure}
@@
 
-\maketitle
+\begin{frame}
 
-\begin{frame}{Climate Change and Variability}
+  \begin{block}{Climate Change and Variability}
+    \begin{figure}
       \begin{columns}
@@
         \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{this}
+        \caption{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}
+        \column{0.4\textwidth}
+        \includegraphics[width=1.0\textwidth]{figures/intro_fig_3.pdf}
       \end{columns}
-\end{frame}
+    \end{figure}
+  \end{block}
+  \vfill
 
-\begin{frame}{Climate forcing}
+  \begin{block}{Climate Forcing}
     \begin{columns}
-    \column{0.5\textwidth}
-    \begin{itemize}
-    \item \alert{Forcing}: any external factor that affects climate.
+      \column{0.4\linewidth}
+      \alert{Forcing}: any external factor that affects climate.
       \begin{description}
@@
       \end{description}
-    \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect. \url{https://www.coolaustralia.org/the-greenhouse-effect-secondary}}{this}
+      \column{0.4\linewidth}
+      \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect. \url{https://www.coolaustralia.org/the-greenhouse-effect-secondary}}{1.0}{this}
     \end{columns}
-\end{frame}
+  \end{block}
+  \vfill
 
-\begin{frame}{El Niño (ENSO)}
+  \begin{block}{El Niño (ENSO)}
     \begin{figure}
       \begin{columns}
-      \column{0.5\textwidth}
+        \column{0.4\textwidth}
         \begin{itemize}
@@
         \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
-      \column{0.5\textwidth}
-      \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
+        \column{0.4\textwidth}
+        \includegraphics[width=1.0\textwidth]{figures/intro_fig.pdf}
       \end{columns}
     \end{figure}
-\end{frame}
+  \end{block}
+  \vfill
 
-\begin{frame}{Review of Literature}
+  \begin{block}{Review of Literature}
     \begin{itemize}
@@
     \item ENSO responds to external forcing.
-    \begin{itemize}
-    \item Correlation between ENSO strength and sunspot activity \citep{emile2007nino}.
     \item Weakened ENSO during the Ice Age due to reduced CO$_2$ levels \citep{zhu2017reduced}.
+    \item Models show possible increasing ENSO activity in the future \citep{zheng2017response} and \citep{maher2018enso}.
     \end{itemize}
-  \item Models show possible increasing ENSO activity in the future \citep{zheng2017response} and \citep{maher2018enso}.
-  \item Factors other than CO\textsubscript{2} can affect ENSO.
-    \begin{itemize}
-    \item Ozone emissions reduce ENSO activity \citep{nowack2017role}.
-    \item Aerosol emissions modify ENSO geographical center \citep{stevenson2017forced}.
-    \end{itemize}
-  \end{itemize}
-\end{frame}
+  \end{block}
+  \vfill
 
-\begin{frame}{Gap}
+  \begin{block}{Gap and Questions}
     \begin{itemize}
@@
     \end{itemize}
-\end{frame}
-
-\begin{frame}{Questions}
     \begin{description}
@@
     \end{description}
-\end{frame}
+  \end{block}
+  \vfill
 
-\begin{frame}{Methods Overview}
-  \begin{itemize}
-  \item Precollected predictions of sea surface temperature from climate models.
-  \item Calculate ENSO intensity in model output.
-  \item Use single forcing ensembles to estimate contributions each forcing item.
-  \item Plot correlation between ENSO intensity and ocean temperature to examine relationship between heat transfer, forcing, and ENSO.
-  \item Use wavelet analysis to analyze changes to ENSO at different frequencies.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Role of Mentor and Student}
-  \begin{columns}[t]
-    \column{.5\textwidth}
-    Student:
-    \begin{itemize}
-    \item Analyze raw data on computer
-    \item Produce graphics for analysis and publication
-    \item Write documentation
-    \item Identify key features of results
-    \end{itemize}
-    \column{.5\textwidth}
-    Mentor:
-    \begin{itemize}
-    \item Review student writing
-    \item Interpret results in the context of climatology
-    \item Conduct parallel analysis
-    \item Provide raw data from facility
-    \end{itemize}
-  \end{columns}
-
-\end{frame}
-
-\begin{frame}{Model Setup}
+  \begin{block}{Model Setup (Data)}
     \begin{itemize}
     \item CESM1 \citep{kay2015community} and CESM2 \citep{danabasoglu2020community}
-  \item Observed forcing levels from 1850-2005
-  \item Predicted forcing levels from 2005-2100
-  \item Ensembles have 40 and 50 simulations respectively
-  \item Control simulation with pre-1850 forcing levels
-  \item Single forcing ensembles that represent influence of single factor
+    \item Predicts climate over 21st century with global warming.
+    \item 40-50 simulations per ensemble.
+    \item Control simulation with pre-1850 forcing levels.
+    \item Single forcing ensembles that represent influence of single factor.
     \end{itemize}
-\end{frame}
+  \end{block}
+  \vfill
 
-\begin{frame}{Measuring ENSO Intensity}
-  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
+  \begin{block}{Measuring ENSO Intensity}
+    \begin{center}
+      \begin{tikzpicture}[node distance = 2.0in and 0.2\linewidth, line width = 1.5mm]
         \node [data] (input) {
-      \begin{varwidth}{1.5in}
+          \begin{varwidth}{4.0in}
             External forcing
           \end{varwidth}};
-    \node [process] (model) [right of=input] {
-      \begin{varwidth}{\linewidth}
+        \node [process] (model) [right=of input] {
+          \begin{varwidth}{4.0in}
             CESM1\\and\\CESM2
           \end{varwidth}};
-    \node [data] (output) [right of=model] {
-      \begin{varwidth}{1.2in}
+        \node [data] (output) [right=of model] {
+          \begin{varwidth}{4.0in}
             Sea surface temperature (SST)
           \end{varwidth}};
-    \node [data] (nino) [below of=input] {
-      \begin{varwidth}{1.4in}
-        Niño 3.4 index
+        \node [data] (nino) [below=of input] {
+          \begin{varwidth}{4.0in}
+            Niño 4.0 index
           \end{varwidth}};
-    \node [process] (variance) [right of=nino] {
-      \begin{varwidth}{1.4in}
+        \node [process] (variance) [right=of nino] {
+          \begin{varwidth}{4.0in}
             Windowed\\variance
           \end{varwidth}};
-    \node [data] (intensity) [right of=variance] {
-      \begin{varwidth}{1.4in}
+        \node [data] (intensity) [right=of variance] {
+          \begin{varwidth}{4.0in}
             ENSO intensity
@@
         \draw [->] (model) to (output);
-    \draw [->] (output) to [out = 270, in = 90] (nino);
+        \draw [->] (output) to [out = 210, in = 30] (nino);
         \draw [->] (nino) to (variance);
@@
       \end{tikzpicture}
-\end{frame}
+    \end{center}
+  \end{block}
 
-\begin{frame}{ENSO is Becoming Stronger}
-  \begin{columns}
-    \column{0.5\textwidth}
-    \begin{itemize}
-    \item Increase in ENSO intensity in both ensembles.
-    \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
-      \begin{itemize}
-      \item May be caused by aerosol emissions.
-      \end{itemize}
-    \end{itemize}
-    \column{0.5\textwidth}
-    \myfig{ff_compare.pdf}{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}{ff_compare}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Influence of Aerosols and Greenhouse Gasses}
-  \begin{columns}
-    \column{0.65\textwidth}
-    \begin{itemize}
-    \item Influence of each factor on ENSO amplitude.
-    \item Increased variance due to greenhouse gas emissions.
-    \item Somewhat increased variance from aerosol emissions, but not linear.
-    \item Inconclusive results from biomass burning and land use forcing
-    \end{itemize}
-    \alert{Takeaway:} Human activities are triggering predicted strengthening of ENSO.
-    \column{0.5\textwidth}
-    \myfig{cesm1_sf.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{cesm1_sf}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Correlation With Ocean Temperature}
-  \begin{columns}
-    \column{0.65\textwidth}
-    \begin{itemize}
-    \item Correlation coefficient between ocean temperature and ENSO amplitude.
-    \item Negative coefficient in subsurface layer.
-    \item Positive coefficient in surface layer.
-    \item Suggests that ocean stratification may be mediating global warming influence on ENSO.
-    \item Difference in heating modifies mechanics of ENSO cycle.
-    \end{itemize}
-    \column{0.5\textwidth}
-    \myfig{tempdt.pdf}{Correlation coefficient between ENSO amplitude and ocean temperature in equatorial cross-section in the fully-forced CESM1 ensemble}{tempdt}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Wavelet Analysis}
-  \begin{columns}
-    \column{0.6\textwidth}
-    \begin{itemize}
-    \item Separate ENSO record into frequency over time.
-    \item Increase in power in late 21\textsuperscript{st} century agrees with previous results.
-    \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{wavelet2.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 and CESM2 ensembles}{wavelet2}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Conclusions and Discussion}
-  \begin{itemize}
-  \item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
-  \item Variance increase is likely caused by the combined influence of greenhouse gasses and aerosols.
-  \item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Application, Limitation, and Next Steps}
-  \begin{description}
-  \item[\alert{Application}] Improve prediction ability to help people prepare for increased likelihood of extreme weather.
-  \item[\alert{Limitation}] Niño 3.4 index may not be fully accurate for various models \citep{cai2018increased}. Also, CESM may contain biases and is not completely accurate.
-  \item[\alert{Next steps}] Examine other variables to further analyze mediator process, continue wavelet analysis methods to focus on individual frequency bands.
-  \end{description}
-\end{frame}
-
-\begin{frame}{Acknowledgments}
-  \begin{itemize}
-  \item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
-1852977.
-  \item Thank you to my teacher, my family, and my mentor!
-  \item Software used: R, ncdf4, zoo, dplyr, ggplot2, WaveletComp, reshape2, nco.
-  \end{itemize}
 \end{frame}
@@
   \bibliographystyle{apalike}
-  \fontsize{4pt}{5}\selectfont
   \bibliography{references.bib}
@@
 \end{document}
-
-% hi

### 2021-12-15 presentations/board/middle.tex
wooohooooo here goes

--- 2021-11-03 presentations/presentation/presentation.tex
@@
 \usepackage{hyperref}
-
-\usetheme{ben}
-\setbeamersize{description width=0.5cm}
+\usepackage[scale=1.4]{beamerposter}
+\geometry{papersize={32in,48in}}
+\usepackage{lipsum}
 
 \usetikzlibrary{arrows,snakes,backgrounds}
+\usetikzlibrary{positioning}
 \tikzstyle{process}=[rectangle, draw=process, fill=process!20, line width = 0.3mm]
@@
 
-\newcommand{\myfig}[3]{
+\usetheme{ben}
+
+\newcommand{\myfig}[4]{
   \begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{figures/#1}
+    \includegraphics[width=#3\textwidth]{figures/#1}
     \caption{#2}
-    \label{fig:#3}
+    \label{fig:#4}
   \end{figure}
 }
+
 \renewcommand{\bibsection}{}
@@
 \begin{document}
+\begin{frame}{The Impact of Anthropogenic Forcing on ENSO Amplitude}{Ben Goldman}
 
-\maketitle
-
-\begin{frame}{Climate Change and Variability}
+  \begin{block}{ENSO is Becoming Stronger (Past year results)}
     \begin{columns}
@@
       \begin{itemize}
-    \item Global warming
-    \item Long-term trends vs short-term randomness
-    \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{this}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Climate forcing}
-  \begin{columns}
-    \column{0.5\textwidth}
-    \begin{itemize}
-    \item \alert{Forcing}: any external factor that affects climate.
-      \begin{description}
-      \item[\alert{GHG}] Greenhouse gasses
-      \item[\alert{AER}] Aerosols (natural: volcanic ash, artificial: smoke)
-      \item[\alert{BMB}] Biomass burning
-      \item[\alert{LULC}] Land use/cover (deforestation, desertification)
-      \end{description}
-    \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect. \url{https://www.coolaustralia.org/the-greenhouse-effect-secondary}}{this}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{El Niño (ENSO)}
-  \begin{figure}
-    \begin{columns}
-      \column{0.5\textwidth}
-      \begin{itemize}
-      \item Warming and cooling of the Pacific Ocean.
-      \item Affects human societies through temperature and rainfall. \citep{ropelewski1987global}
-      \item May be affected by climate change.
-      \end{itemize}
-      \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
-      \column{0.5\textwidth}
-      \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
-    \end{columns}
-  \end{figure}
-\end{frame}
-
-\begin{frame}{Review of Literature}
-  \begin{itemize}
-  \item ENSO's properties observed vary across different decades. \citep{lubbecke2014assessing}.
-  \item ENSO responds to external forcing.
-    \begin{itemize}
-    \item Correlation between ENSO strength and sunspot activity \citep{emile2007nino}.
-    \item Weakened ENSO during the Ice Age due to reduced CO$_2$ levels \citep{zhu2017reduced}.
-    \end{itemize}
-  \item Models show possible increasing ENSO activity in the future \citep{zheng2017response} and \citep{maher2018enso}.
-  \item Factors other than CO\textsubscript{2} can affect ENSO.
-    \begin{itemize}
-    \item Ozone emissions reduce ENSO activity \citep{nowack2017role}.
-    \item Aerosol emissions modify ENSO geographical center \citep{stevenson2017forced}.
-    \end{itemize}
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Gap}
-  \begin{itemize}
-  \item Little research using a large ensemble to examine the effect of individual factors on ENSO.
-  \item Considerable disagreement between studies on whether ENSO will strengthen or weaken due to global warming
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Questions}
-  \begin{description}
-  \item[What?] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
-  \item[Why?] Is the predicted increase (or decrease) due to human activities?
-  \item[How?] What processes are causing greenhouse gasses and aerosols to affect ENSO?
-  \end{description}
-\end{frame}
-
-\begin{frame}{Methods Overview}
-  \begin{itemize}
-  \item Precollected predictions of sea surface temperature from climate models.
-  \item Calculate ENSO intensity in model output.
-  \item Use single forcing ensembles to estimate contributions each forcing item.
-  \item Plot correlation between ENSO intensity and ocean temperature to examine relationship between heat transfer, forcing, and ENSO.
-  \item Use wavelet analysis to analyze changes to ENSO at different frequencies.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Role of Mentor and Student}
-  \begin{columns}[t]
-    \column{.5\textwidth}
-    Student:
-    \begin{itemize}
-    \item Analyze raw data on computer
-    \item Produce graphics for analysis and publication
-    \item Write documentation
-    \item Identify key features of results
-    \end{itemize}
-    \column{.5\textwidth}
-    Mentor:
-    \begin{itemize}
-    \item Review student writing
-    \item Interpret results in the context of climatology
-    \item Conduct parallel analysis
-    \item Provide raw data from facility
-    \end{itemize}
-  \end{columns}
-
-\end{frame}
-
-\begin{frame}{Model Setup}
-  \begin{itemize}
-  \item CESM1 \citep{kay2015community} and CESM2 \citep{danabasoglu2020community}
-  \item Observed forcing levels from 1850-2005
-  \item Predicted forcing levels from 2005-2100
-  \item Ensembles have 40 and 50 simulations respectively
-  \item Control simulation with pre-1850 forcing levels
-  \item Single forcing ensembles that represent influence of single factor
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Measuring ENSO Intensity}
-  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
-    \node [data] (input) {
-      \begin{varwidth}{1.5in}
-        External forcing
-      \end{varwidth}};
-    \node [process] (model) [right of=input] {
-      \begin{varwidth}{\linewidth}
-        CESM1\\and\\CESM2
-      \end{varwidth}};
-    \node [data] (output) [right of=model] {
-      \begin{varwidth}{1.2in}
-        Sea surface temperature (SST)
-      \end{varwidth}};
-    \node [data] (nino) [below of=input] {
-      \begin{varwidth}{1.4in}
-        Niño 3.4 index
-      \end{varwidth}};
-    \node [process] (variance) [right of=nino] {
-      \begin{varwidth}{1.4in}
-        Windowed\\variance
-      \end{varwidth}};
-    \node [data] (intensity) [right of=variance] {
-      \begin{varwidth}{1.4in}
-        ENSO intensity
-      \end{varwidth}};
-    \draw [->] (input) to (model);
-    \draw [->] (model) to (output);
-    \draw [->] (output) to [out = 270, in = 90] (nino);
-    \draw [->] (nino) to (variance);
-    \draw [->] (variance) to (intensity);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}{ENSO is Becoming Stronger}
-  \begin{columns}
-    \column{0.5\textwidth}
-    \begin{itemize}
       \item Increase in ENSO intensity in both ensembles.
       \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
-      \begin{itemize}
-      \item May be caused by aerosol emissions.
-      \end{itemize}
       \end{itemize}
       \column{0.5\textwidth}
-    \myfig{ff_compare.pdf}{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}{ff_compare}
+      \myfig{ff_compare.pdf}{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}{1.0}{ff_compare}
     \end{columns}
-\end{frame}
+  \end{block}
 
-\begin{frame}{Influence of Aerosols and Greenhouse Gasses}
+  \vfill
+
+  \begin{block}{Influence of Aerosols and Greenhouse Gasses (Past year results)}
     \begin{columns}
-    \column{0.65\textwidth}
+      \column{0.4\textwidth}
       \begin{itemize}
@@
       \item Somewhat increased variance from aerosol emissions, but not linear.
-    \item Inconclusive results from biomass burning and land use forcing
       \end{itemize}
@@
       \column{0.5\textwidth}
-    \myfig{cesm1_sf.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{cesm1_sf}
+      \myfig{cesm1_sf_4.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{1.0}{cesm1_sf_4}
     \end{columns}
-\end{frame}
+  \end{block}
 
-\begin{frame}{Correlation With Ocean Temperature}
+  \vfill
+
+  \begin{block}{Correlation With Ocean Temperature (Current results)}
     \begin{columns}
-    \column{0.65\textwidth}
+      \column{0.4\textwidth}
       \begin{itemize}
@@
       \column{0.5\textwidth}
-    \myfig{tempdt.pdf}{Correlation coefficient between ENSO amplitude and ocean temperature in equatorial cross-section in the fully-forced CESM1 ensemble}{tempdt}
+      \myfig{tempdt.pdf}{Correlation coefficient between ENSO amplitude and ocean temperature in equatorial cross-section in the fully-forced CESM1 ensemble}{1.0}{tempdt}
     \end{columns}
+  \end{block}
+
+  \vfill
+
 \end{frame}
-
-\begin{frame}{Wavelet Analysis}
-  \begin{columns}
-    \column{0.6\textwidth}
-    \begin{itemize}
-    \item Separate ENSO record into frequency over time.
-    \item Increase in power in late 21\textsuperscript{st} century agrees with previous results.
-    \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{wavelet2.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 and CESM2 ensembles}{wavelet2}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Conclusions and Discussion}
-  \begin{itemize}
-  \item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
-  \item Variance increase is likely caused by the combined influence of greenhouse gasses and aerosols.
-  \item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Application, Limitation, and Next Steps}
-  \begin{description}
-  \item[\alert{Application}] Improve prediction ability to help people prepare for increased likelihood of extreme weather.
-  \item[\alert{Limitation}] Niño 3.4 index may not be fully accurate for various models \citep{cai2018increased}. Also, CESM may contain biases and is not completely accurate.
-  \item[\alert{Next steps}] Examine other variables to further analyze mediator process, continue wavelet analysis methods to focus on individual frequency bands.
-  \end{description}
-\end{frame}
-
-\begin{frame}{Acknowledgments}
-  \begin{itemize}
-  \item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
-1852977.
-  \item Thank you to my teacher, my family, and my mentor!
-  \item Software used: R, ncdf4, zoo, dplyr, ggplot2, WaveletComp, reshape2, nco.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{References}
-  \bibliographystyle{apalike}
-  \fontsize{4pt}{5}\selectfont
-  \bibliography{references.bib}
-\end{frame}
-
 \end{document}
-
-% hi

### 2021-12-15 presentations/board/right.tex
wooohooooo here goes

--- 2021-11-08 presentations/presentation/presentation.tex
@@
 \usepackage{hyperref}
+\usepackage[scale=1.4]{beamerposter}
+\geometry{papersize={20in,48in}}
+\usepackage{lipsum}
 
 \usetheme{ben}
-\setbeamersize{description width=0.5cm}
 
-\usetikzlibrary{arrows,snakes,backgrounds}
-\tikzstyle{process}=[rectangle, draw=process, fill=process!20, line width = 0.3mm]
-\tikzstyle{data}=[rectangle, draw=data, fill=data!20, line width = 0.3mm]
-\tikzstyle{tight}=[node distance = 0.5in]
-\tikzstyle{loose}=[node distance = 2in]
-\definecolor{process}{HTML}{3d5f8f}
-\definecolor{data}{HTML}{8f6d3d}
-
-\newcommand{\myfig}[3]{
+\newcommand{\myfig}[4]{
   \begin{figure}
     \centering
-    \includegraphics[width=\textwidth]{figures/#1}
+    \includegraphics[width=#3\textwidth]{figures/#1}
     \caption{#2}
-    \label{fig:#3}
+    \label{fig:#4}
   \end{figure}
 }
+
 \renewcommand{\bibsection}{}
@@
 
-\maketitle
+\nocite{gistemp2019giss}
+\nocite{lenssen2019improvements}
+\nocite{ropelewski1987global}
+\nocite{rayner2003global}
+\nocite{lubbecke2014assessing}
+\nocite{zhu2017reduced}
+\nocite{zheng2017response}
+\nocite{maher2018enso}
+\nocite{kay2015community}
+\nocite{danabasoglu2020community}
 
-\begin{frame}{Climate Change and Variability}
-  \begin{columns}
-    \column{0.4\textwidth}
+\begin{frame}
+
+  \begin{block}{Wavelet Analysis (Current results)}
     \begin{itemize}
-    \item Global warming
-    \item Long-term trends vs short-term randomness
+    \item Separate ENSO record into changes in period over time.
+    \item Increase in power in late 21\textsuperscript{st} century agrees with previous results.
+    \item In CESM1, increase in ENSO intensity is mainly strengthening of longer-period cycle.
+    \item In CESM2, longer-period ENSO weakens after 2025.
     \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{this}
-  \end{columns}
-\end{frame}
+    \myfig{wavelet3.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 and CESM2 ensembles}{0.8}{wavelet2}
+  \end{block}
 
-\begin{frame}{Climate forcing}
-  \begin{columns}
-    \column{0.5\textwidth}
+  \vfill
+
+  \begin{block}{Discussion}
     \begin{itemize}
-    \item \alert{Forcing}: any external factor that affects climate.
-      \begin{description}
-      \item[\alert{GHG}] Greenhouse gasses
-      \item[\alert{AER}] Aerosols (natural: volcanic ash, artificial: smoke)
-      \item[\alert{BMB}] Biomass burning
-      \item[\alert{LULC}] Land use/cover (deforestation, desertification)
-      \end{description}
+    \item Rising greenhouse gas levels increase Pacific Ocean stratification, strengthening ENSO cycle.
+    \item Aerosol influence is nonlinear because aerosol levels are not purely increasing.
+    \item Stronger ENSO may lead to greater temperature variability and extreme weather.
+    \item CESM1 and CESM2 conflict in their prediction of the changes to ENSO's frequency.
     \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect. \url{https://www.coolaustralia.org/the-greenhouse-effect-secondary}}{this}
-  \end{columns}
-\end{frame}
+  \end{block}
 
-\begin{frame}{El Niño (ENSO)}
-  \begin{figure}
-    \begin{columns}
-      \column{0.5\textwidth}
+  \vfill
+
+  \begin{block}{Limitations and Applications}
+    Limitations:
     \begin{itemize}
-      \item Warming and cooling of the Pacific Ocean.
-      \item Affects human societies through temperature and rainfall. \citep{ropelewski1987global}
-      \item May be affected by climate change.
+    \item Niño 3.4 index shown to be inaccurate for some models \citep{cai2018increased}.
+    \item CESM may contain biases.
     \end{itemize}
-      \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
-      \column{0.5\textwidth}
-      \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
-    \end{columns}
-  \end{figure}
-\end{frame}
+    Application: to improve our ability to predict ENSO and help people prepare for increased likelihood of extreme weather.
+  \end{block}
 
-\begin{frame}{Review of Literature}
+  \vfill
+
+  \begin{block}{Acknowledgments}
     \begin{itemize}
-  \item ENSO's properties observed vary across different decades. \citep{lubbecke2014assessing}.
-  \item ENSO responds to external forcing.
-    \begin{itemize}
-    \item Correlation between ENSO strength and sunspot activity \citep{emile2007nino}.
-    \item Weakened ENSO during the Ice Age due to reduced CO$_2$ levels \citep{zhu2017reduced}.
+    \item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
+1852977.
+    \item Thank you to my teacher, my family, and my mentor!
+    \item Software used: R, ncdf4, zoo, dplyr, ggplot2, WaveletComp, reshape2, nco.
     \end{itemize}
-  \item Models show possible increasing ENSO activity in the future \citep{zheng2017response} and \citep{maher2018enso}.
-  \item Factors other than CO\textsubscript{2} can affect ENSO.
-    \begin{itemize}
-    \item Ozone emissions reduce ENSO activity \citep{nowack2017role}.
-    \item Aerosol emissions modify ENSO geographical center \citep{stevenson2017forced}.
-    \end{itemize}
-  \end{itemize}
-\end{frame}
+  \end{block}
 
-\begin{frame}{Gap}
-  \begin{itemize}
-  \item Little research using a large ensemble to examine the effect of individual factors on ENSO.
-  \item Considerable disagreement between studies on whether ENSO will strengthen or weaken due to global warming
-  \end{itemize}
-\end{frame}
+  \vfill
 
-\begin{frame}{Questions}
-  \begin{description}
-  \item[What?] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
-  \item[Why?] Is the predicted increase (or decrease) due to human activities?
-  \item[How?] What processes are causing greenhouse gasses and aerosols to affect ENSO?
-  \end{description}
-\end{frame}
-
-\begin{frame}{Methods Overview}
-  \begin{itemize}
-  \item Precollected predictions of sea surface temperature from climate models.
-  \item Calculate ENSO intensity in model output.
-  \item Use single forcing ensembles to estimate contributions each forcing item.
-  \item Plot correlation between ENSO intensity and ocean temperature to examine relationship between heat transfer, forcing, and ENSO.
-  \item Use wavelet analysis to analyze changes to ENSO at different frequencies.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Role of Mentor and Student}
+  \begin{block}{Role of Mentor and Student}
     \begin{columns}[t]
-    \column{.5\textwidth}
+      \column{.4\textwidth}
       Student:
@@
       \end{itemize}
-    \column{.5\textwidth}
+      \column{.4\textwidth}
       Mentor:
@@
     \end{columns}
+  \end{block}
+
+  \vfill
+
+  \begin{block}{References}
+    \bibliographystyle{apalike}
+    \scriptsize{\bibliography{references.bib}}
+  \end{block}
+
+  \vfill
 
 \end{frame}
-
-\begin{frame}{Model Setup}
-  \begin{itemize}
-  \item CESM1 \citep{kay2015community} and CESM2 \citep{danabasoglu2020community}
-  \item Observed forcing levels from 1850-2005
-  \item Predicted forcing levels from 2005-2100
-  \item Ensembles have 40 and 50 simulations respectively
-  \item Control simulation with pre-1850 forcing levels
-  \item Single forcing ensembles that represent influence of single factor
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Measuring ENSO Intensity}
-  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
-    \node [data] (input) {
-      \begin{varwidth}{1.5in}
-        External forcing
-      \end{varwidth}};
-    \node [process] (model) [right of=input] {
-      \begin{varwidth}{\linewidth}
-        CESM1\\and\\CESM2
-      \end{varwidth}};
-    \node [data] (output) [right of=model] {
-      \begin{varwidth}{1.2in}
-        Sea surface temperature (SST)
-      \end{varwidth}};
-    \node [data] (nino) [below of=input] {
-      \begin{varwidth}{1.4in}
-        Niño 3.4 index
-      \end{varwidth}};
-    \node [process] (variance) [right of=nino] {
-      \begin{varwidth}{1.4in}
-        Windowed\\variance
-      \end{varwidth}};
-    \node [data] (intensity) [right of=variance] {
-      \begin{varwidth}{1.4in}
-        ENSO intensity
-      \end{varwidth}};
-    \draw [->] (input) to (model);
-    \draw [->] (model) to (output);
-    \draw [->] (output) to [out = 270, in = 90] (nino);
-    \draw [->] (nino) to (variance);
-    \draw [->] (variance) to (intensity);
-  \end{tikzpicture}
-\end{frame}
-
-\begin{frame}{ENSO is Becoming Stronger}
-  \begin{columns}
-    \column{0.4\textwidth}
-    \begin{itemize}
-    \item Increase in ENSO intensity in both ensembles.
-    \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
-      \begin{itemize}
-      \item May be caused by aerosol emissions.
-      \end{itemize}
-    \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{ff_compare.pdf}{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}{ff_compare}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Influence of Aerosols and Greenhouse Gasses}
-  \begin{columns}
-    \column{0.65\textwidth}
-    \begin{itemize}
-    \item Influence of each factor on ENSO amplitude.
-    \item Increased variance due to greenhouse gas emissions.
-    \item Somewhat increased variance from aerosol emissions, but not linear.
-    \end{itemize}
-    \alert{Takeaway:} Human activities are triggering predicted strengthening of ENSO.
-    \column{0.5\textwidth}
-    \myfig{cesm1_sf_4.pdf}{Influence of GHG, AER, and BMB forcing on ENSO amplitude in CESM1}{cesm1_sf_4}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Correlation With Ocean Temperature}
-  \begin{columns}
-    \column{0.65\textwidth}
-    \begin{itemize}
-    \item Correlation coefficient between ocean temperature and ENSO amplitude.
-    \item Negative coefficient in subsurface layer.
-    \item Positive coefficient in surface layer.
-    \item Suggests that ocean stratification may be mediating global warming influence on ENSO.
-    \item Difference in heating modifies mechanics of ENSO cycle.
-    \end{itemize}
-    \column{0.5\textwidth}
-    \myfig{tempdt.pdf}{Correlation coefficient between ENSO amplitude and ocean temperature in equatorial cross-section in the fully-forced CESM1 ensemble}{tempdt}
-  \end{columns}
-\end{frame}
-
-\begin{frame}{Wavelet Analysis}
-  \begin{columns}
-    \column{0.5\textwidth}
-    \begin{itemize}
-    \item Separate ENSO record into changes in period over time.
-    \item Increase in power in late 21\textsuperscript{st} century agrees with previous results.
-    \item In CESM1, increase in ENSO intensity is mainly strengthening of longer-period cycle.
-    \item In CESM2, longer-period ENSO weakens after 2025.
-    \end{itemize}
-    \column{0.5\textwidth}
-    \myfig{wavelet3.pdf}{Wavelet power spectrum for the Niño 3.4 index in the fully-forced CESM1 and CESM2 ensembles}{wavelet2}
-  \end{columns}
-\end{frame}
-
-% SPLIT UP CONCLUSIONS DISCUSSION
-
-\begin{frame}{Discussion}
-  \begin{itemize}
-  \item Rising greenhouse gas levels increase Pacific Ocean stratification, strengthening ENSO cycle.
-  \item Aerosol influence nonlinear because aerosol levels are not purely increasing.
-  \item Stronger ENSO will lead to greater climate variability and extreme weather.
-  \item
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Conclusions and Discussion}
-  \begin{itemize}
-  \item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
-  \item Variance increase is likely caused by the combined influence of greenhouse gasses and aerosols.
-  \item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Application, Limitation, and Next Steps}
-  \begin{description}
-  \item[\alert{Application}] Improve prediction ability to help people prepare for increased likelihood of extreme weather.
-  \item[\alert{Limitation}] Niño 3.4 index may not be fully accurate for various models \citep{cai2018increased}. Also, CESM may contain biases and is not completely accurate.
-  \item[\alert{Next steps}] Examine other variables to further analyze mediator process, continue wavelet analysis methods to focus on individual frequency bands.
-  \end{description}
-\end{frame}
-
-\begin{frame}{Acknowledgments}
-  \begin{itemize}
-  \item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
-1852977.
-  \item Thank you to my teacher, my family, and my mentor!
-  \item Software used: R, ncdf4, zoo, dplyr, ggplot2, WaveletComp, reshape2, nco.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{References}
-  \bibliographystyle{apalike}
-  \fontsize{4pt}{5}\selectfont
-  \bibliography{references.bib}
-\end{frame}
-
 \end{document}
-
-% hi

### 2021-12-15 presentations/presentation/presentation.tex
wooohooooo here goes

@@
     \column{0.6\textwidth}
-    \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{this}
+    \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{woohoo}
   \end{columns}
@@
 
-\begin{frame}{Methods Overview}
-  \begin{itemize}
-  \item Precollected predictions of sea surface temperature from climate models.
-  \item Calculate ENSO intensity in model output.
-  \item Use single forcing ensembles to estimate contributions each forcing item.
-  \item Plot correlation between ENSO intensity and ocean temperature to examine relationship between heat transfer, forcing, and ENSO.
-  \item Use wavelet analysis to analyze changes to ENSO at different frequencies.
-  \end{itemize}
-\end{frame}
+% \begin{frame}{Methods Overview}
+%   \begin{itemize}
+%   \item Precollected predictions of sea surface temperature from climate models.
+%   \item Calculate ENSO intensity in model output.
+%   \item Use single forcing ensembles to estimate contributions each forcing item.
+%   \item Plot correlation between ENSO intensity and ocean temperature to examine relationship between heat transfer, forcing, and ENSO.
+%   \item Use wavelet analysis to analyze changes to ENSO at different frequencies.
+%   \end{itemize}
+% \end{frame}
 
-\begin{frame}{Role of Mentor and Student}
-  \begin{columns}[t]
-    \column{.5\textwidth}
-    Student:
-    \begin{itemize}
-    \item Analyze raw data on computer
-    \item Produce graphics for analysis and publication
-    \item Write documentation
-    \item Identify key features of results
-    \end{itemize}
-    \column{.5\textwidth}
-    Mentor:
-    \begin{itemize}
-    \item Review student writing
-    \item Interpret results in the context of climatology
-    \item Conduct parallel analysis
-    \item Provide raw data from facility
-    \end{itemize}
-  \end{columns}
-
-\end{frame}
+% \begin{frame}{Role of Mentor and Student}
+%   \begin{columns}[t]
+%     \column{.5\textwidth}
+%     Student:
+%     \begin{itemize}
+%     \item Analyze raw data on computer
+%     \item Produce graphics for analysis and publication
+%     \item Write documentation
+%     \item Identify key features of results
+%     \end{itemize}
+%     \column{.5\textwidth}
+%     Mentor:
+%     \begin{itemize}
+%     \item Review student writing
+%     \item Interpret results in the context of climatology
+%     \item Conduct parallel analysis
+%     \item Provide raw data from facility
+%     \end{itemize}
+%   \end{columns}
+% \end{frame}
 
@@
     \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
-      \begin{itemize}
-      \item May be caused by aerosol emissions.
-      \end{itemize}
     \end{itemize}
@@
   \item Rising greenhouse gas levels increase Pacific Ocean stratification, strengthening ENSO cycle.
-  \item Aerosol influence nonlinear because aerosol levels are not purely increasing.
-  \item Stronger ENSO will lead to greater climate variability and extreme weather.
-  \item
+  \item Aerosol influence is nonlinear because aerosol levels are not purely increasing.
+  \item Stronger ENSO may lead to greater temperature variability and extreme weather.
+  \item CESM1 and CESM2 conflict in their prediction of the changes to ENSO's frequency.
   \end{itemize}
@@
 
-\begin{frame}{Conclusions and Discussion}
+\begin{frame}{Limitations and Applications}
+  Limitations:
   \begin{itemize}
-  \item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
-  \item Variance increase is likely caused by the combined influence of greenhouse gasses and aerosols.
-  \item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
+  \item Niño 3.4 index shown to be inaccurate for some models \citep{cai2018increased}.
+  \item CESM may contain biases.
   \end{itemize}
+  Application: to improve our ability to predict ENSO and help people prepare for increased likelihood of extreme weather.
 \end{frame}
+% \begin{frame}{Limitations}
+%   \begin{itemize}
+%   \item Niño 3.4 index is inaccurate for certain models.
+%   \item Only one climate model used in analysis.
+%   \item Climate models are all imperfect.
+%   \end{itemize}
+% \end{frame}
 
-\begin{frame}{Application, Limitation, and Next Steps}
-  \begin{description}
-  \item[\alert{Application}] Improve prediction ability to help people prepare for increased likelihood of extreme weather.
-  \item[\alert{Limitation}] Niño 3.4 index may not be fully accurate for various models \citep{cai2018increased}. Also, CESM may contain biases and is not completely accurate.
-  \item[\alert{Next steps}] Examine other variables to further analyze mediator process, continue wavelet analysis methods to focus on individual frequency bands.
-  \end{description}
-\end{frame}
+% \begin{frame}{Future Research}
+%   \begin{itemize}
+%   \item Conduct similar analysis on large ensembles of other climate models.
+%   \end{itemize}
+% \end{frame}
+
+% \begin{frame}{Conclusions and Discussion}
+%   \begin{itemize}
+%   \item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
+%   \item Variance increase is likely caused by the combined influence of greenhouse gasses and aerosols.
+%   \item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
+%   \end{itemize}
+% \end{frame}
+%
+% \begin{frame}{Application, Limitation, and Next Steps}
+%   \begin{description}
+%   \item[\alert{Application}] Improve prediction ability to help people prepare for increased likelihood of extreme weather.
+%   \item[\alert{Limitation}] Niño 3.4 index may not be fully accurate for various models \citep{cai2018increased}. Also, CESM may contain biases and is not completely accurate.
+%   \item[\alert{Next steps}] Examine other variables to further analyze mediator process, continue wavelet analysis methods to focus on individual frequency bands.
+%   \end{description}
+% \end{frame}
 

### 2021-12-15 data/correlation.r

@@ -106,7 +106,7 @@ dimnames(corrs) <- list(lon, depth)
 
 ggplot(corrs[,1:38] %>% melt, aes(Var1, Var2, z = value)) +
     geom_contour_filled() +
-    labs(title = "Correlation Coefficient Between Niño 3.4 Variance and Ocean Temperature",
+    labs(title = "Correlation Coefficient",
          x = "Longitude",
          y = "Depth (m)",
          fill = "Correlation\nCoefficient") +

### 2021-12-21 presentations/presentation/presentation.tex
some changes lol

--- 2021-11-08 presentations/presentation/presentation.tex
@@
   \item ENSO responds to external forcing.
-    \begin{itemize}
-    \item Correlation between ENSO strength and sunspot activity \citep{emile2007nino}.
   \item Weakened ENSO during the Ice Age due to reduced CO$_2$ levels \citep{zhu2017reduced}.
-    \end{itemize}
   \item Models show possible increasing ENSO activity in the future \citep{zheng2017response} and \citep{maher2018enso}.
-  \item Factors other than CO\textsubscript{2} can affect ENSO.
-    \begin{itemize}
-    \item Ozone emissions reduce ENSO activity \citep{nowack2017role}.
-    \item Aerosol emissions modify ENSO geographical center \citep{stevenson2017forced}.
-    \end{itemize}
   \end{itemize}
@@
 
-\begin{frame}{Gap}
+\begin{frame}{Gap and Questions}
   \begin{itemize}
@@
   \end{itemize}
-\end{frame}
-
-\begin{frame}{Questions}
   \begin{description}
@@
 
-\begin{frame}{Methods Overview}
-  \begin{itemize}
-  \item Precollected predictions of sea surface temperature from climate models.
-  \item Calculate ENSO intensity in model output.
-  \item Use single forcing ensembles to estimate contributions each forcing item.
-  \item Plot correlation between ENSO intensity and ocean temperature to examine relationship between heat transfer, forcing, and ENSO.
-  \item Use wavelet analysis to analyze changes to ENSO at different frequencies.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Role of Mentor and Student}
-  \begin{columns}[t]
-    \column{.5\textwidth}
-    Student:
-    \begin{itemize}
-    \item Analyze raw data on computer
-    \item Produce graphics for analysis and publication
-    \item Write documentation
-    \item Identify key features of results
-    \end{itemize}
-    \column{.5\textwidth}
-    Mentor:
-    \begin{itemize}
-    \item Review student writing
-    \item Interpret results in the context of climatology
-    \item Conduct parallel analysis
-    \item Provide raw data from facility
-    \end{itemize}
-  \end{columns}
-
-\end{frame}
-
-\begin{frame}{Model Setup}
+\begin{frame}{Model Setup (Data)}
   \begin{itemize}
   \item CESM1 \citep{kay2015community} and CESM2 \citep{danabasoglu2020community}
-  \item Observed forcing levels from 1850-2005
-  \item Predicted forcing levels from 2005-2100
-  \item Ensembles have 40 and 50 simulations respectively
-  \item Control simulation with pre-1850 forcing levels
-  \item Single forcing ensembles that represent influence of single factor
+  \item Predicts climate over 21st century with global warming.
+  \item 40-50 simulations per ensemble.
+  \item Control simulation with pre-1850 forcing levels.
+  \item Single forcing ensembles that represent influence of single factor.
   \end{itemize}
@@
 
-\begin{frame}{Conclusions and Discussion}
+\begin{frame}{Limitations and Applications}
+  Limitations:
   \begin{itemize}
-  \item There is likely to be an increase in ENSO strength over the next 100 years. Agrees with \citet{cai2018increased}.
-  \item Variance increase is likely caused by the combined influence of greenhouse gasses and aerosols.
-  \item Global warming increases ENSO intensity by warming upper layers of the Pacific faster than central layers.
+  \item Niño 3.4 index shown to be inaccurate for some models \citep{cai2018increased}.
+  \item CESM may contain biases.
   \end{itemize}
-\end{frame}
-
-\begin{frame}{Application, Limitation, and Next Steps}
-  \begin{description}
-  \item[\alert{Application}] Improve prediction ability to help people prepare for increased likelihood of extreme weather.
-  \item[\alert{Limitation}] Niño 3.4 index may not be fully accurate for various models \citep{cai2018increased}. Also, CESM may contain biases and is not completely accurate.
-  \item[\alert{Next steps}] Examine other variables to further analyze mediator process, continue wavelet analysis methods to focus on individual frequency bands.
-  \end{description}
+  Application: to improve our ability to predict ENSO and help people prepare for increased likelihood of extreme weather.
 \end{frame}
@@
 
+\begin{frame}{Role of Mentor and Student}
+  \begin{columns}[t]
+    \column{.4\textwidth}
+    Student:
+    \begin{itemize}
+    \item Analyze raw data on computer
+    \item Produce graphics for analysis and publication
+    \item Write documentation
+    \item Identify key features of results
+    \end{itemize}
+    \column{.4\textwidth}
+    Mentor:
+    \begin{itemize}
+    \item Review student writing
+    \item Interpret results in the context of climatology
+    \item Conduct parallel analysis
+    \item Provide raw data from facility
+    \end{itemize}
+  \end{columns}
+\end{frame}
+
 \begin{frame}{References}

### 2022-01-18 feedback/presentation_main.tex
stuff lol

@@
-% Created 2021-10-19 Tue 07:57
+% Created 2022-01-06 Thu 19:00
 % Intended LaTeX compiler: pdflatex
@@
 \section{Feedback 1}
-\label{sec:org1da6f23}
+\label{sec:orgb0ce4b1}
 Date: \textit{<2021-09-26 Sun>}
@@
 \subsection*{Feedback}
-\label{sec:org1d74604}
+\label{sec:orgeae9fd0}
 \begin{itemize}
@@
 \section{Feedback 2}
-\label{sec:org11974c6}
+\label{sec:org3a9f9f0}
 Date: \textit{<2021-09-30 Thu>}
@@
 \subsection*{Feedback:}
-\label{sec:orgdbbe7e5}
+\label{sec:orgd187f3b}
 \begin{itemize}
@@
 \section{Feedback 3}
-\label{sec:org52baac0}
+\label{sec:org304f650}
 Date: \textit{<2021-09-30 Thu>}
@@
 \subsection*{Feedback:}
-\label{sec:orgf282c69}
+\label{sec:org2526349}
 \begin{itemize}
@@
 \section{Feedback 4}
-\label{sec:org585437d}
+\label{sec:org661561a}
 Date: \textit{<2021-10-17 Sun>}
@@
 \subsection*{Feedback:}
-\label{sec:orgeffe568}
+\label{sec:orgfb0e433}
 \begin{itemize}
@@
 \end{itemize}
+
+\section{Feedback 5}
+\label{sec:org4ee2cd0}
+
+Date: \textit{<2022-01-06 Thu>}
+Time: 09:20
+Listener: Jimena Perez
+\subsection*{Feedback:}
+\label{sec:orgcae29f9}
+\begin{itemize}
+\item Put Student vs mentor role earlier in presentation
+\item Slow down speech
+\item Write a script for wavelet slide
+\item Don't admit you don't understand it
+\item No monospace for urls
+\item Clean up global warming slide
+\end{itemize}
 \end{document}

### 2022-01-18 presentations/presentation/presentation.tex
stuff lol

@@
       \begin{itemize}
+      \item ENSO = El Niño/Southern Oscillation
       \item Warming and cooling of the Pacific Ocean.
@@
 \begin{frame}{Gap and Questions}
+  \alert{Gap}
   \begin{itemize}
@@
   \end{itemize}
+  \alert{Questions}
   \begin{description}
-  \item[What?] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
-  \item[Why?] Is the predicted increase (or decrease) due to human activities?
-  \item[How?] What processes are causing greenhouse gasses and aerosols to affect ENSO?
+  \item[\alert{What?}] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
+  \item[\alert{Why?}] Is the predicted increase (or decrease) due to human activities?
+  \item[\alert{How?}] What processes are causing greenhouse gasses and aerosols to affect ENSO?
   \end{description}
@@
 
+\section{Methods and Results}
+
 \begin{frame}{Model Setup (Data)}
   \begin{itemize}
-  \item CESM1 \citep{kay2015community} and CESM2 \citep{danabasoglu2020community}
+  \item Community Earth System Model (CESM) Versions 1 and 2 \citep{kay2015community} \citep{danabasoglu2020community}.
   \item Predicts climate over 21st century with global warming.
@@
   \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
-    \node [data] (input) {
-      \begin{varwidth}{1.5in}
-        External forcing
+    \node [process] (input) {
+      \begin{varwidth}{1.2in}
+        External forcing records and predictions
       \end{varwidth}};
     \node [process] (model) [right of=input] {
-      \begin{varwidth}{\linewidth}
-        CESM1\\and\\CESM2
+      \begin{varwidth}{1.2in}
+        CESM1 and CESM2 climate models
       \end{varwidth}};
-    \node [data] (output) [right of=model] {
+    \node [process] (output) [right of=model] {
       \begin{varwidth}{1.2in}
@@
       \end{varwidth}};
-    \node [data] (nino) [below of=input] {
-      \begin{varwidth}{1.4in}
-        Niño 3.4 index
+    \node [process] (nino) [below of=input] {
+      \begin{varwidth}{1.2in}
+        Mean tropical Pacific temperature
       \end{varwidth}};
     \node [process] (variance) [right of=nino] {
-      \begin{varwidth}{1.4in}
-        Windowed\\variance
+      \begin{varwidth}{1.0in}
+        Windowed variance: Mean variance over every 20 years.
       \end{varwidth}};
-    \node [data] (intensity) [right of=variance] {
-      \begin{varwidth}{1.4in}
-        ENSO intensity
+    \node [process] (intensity) [right of=variance] {
+      \begin{varwidth}{1.2in}
+        Mean and standard error
       \end{varwidth}};
@@
   \begin{columns}
-    \column{0.4\textwidth}
+    \column{0.5\textwidth}
     \begin{itemize}
-    \item Increase in ENSO intensity in both ensembles.
+    \item Increase in ENSO intensity in both ensembles. (Exceeds 2 standard errors)
     \item Increase slows down in CESM1 and decreases in CESM2 after around 2050.
     \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{ff_compare.pdf}{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}{ff_compare}
+    \column{0.5\textwidth}
+    \begin{figure}
+      \begin{tikzpicture}
+        \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=\textwidth]{figures/ff_compare.pdf}};
+        \coordinate (a) at (0.5, 1.5);
+        \node [red] (b) at (1.0, 1.0) {
+          \begin{varwidth}{0.6in}
+            Increasing variance
+          \end{varwidth}};
+        \coordinate (c) at (3.0, 1.5) {};
+        \draw[red, |-|] (a) to (c);
+
+        \coordinate (d) at (3.2, 1.5);
+        \node [red] (e) at (3.5, 1.0) {
+          \begin{varwidth}{0.6in}
+            Decreasing variance
+          \end{varwidth}};
+        \coordinate (f) at (4.3, 1.5) {};
+        \draw[red, |-|] (d) to (f);
+      \end{tikzpicture}
+      \caption{ENSO intensity ensemble mean and standard error for CESM1 and CESM2}
+      \label{ff_compare}
+    \end{figure}
   \end{columns}
@@
     \item Separate ENSO record into changes in period over time.
-    \item Increase in power in late 21\textsuperscript{st} century agrees with previous results.
     \item In CESM1, increase in ENSO intensity is mainly strengthening of longer-period cycle.
     \item In CESM2, longer-period ENSO weakens after 2025.
+    \item Indicates that longer frequency bands are more susceptible to climate change.
     \end{itemize}
@@
   \item CESM may contain biases.
+  \item Models are only an approximation of the Earth's actual climate.
   \end{itemize}

### 2022-01-24 presentations/presentation/notes.tex
stuff lol

@@
-% Created 2021-10-31 Sun 14:05
+% Created 2022-01-18 Tue 20:19
 % Intended LaTeX compiler: pdflatex
@@
 \section{General:}
-\label{sec:orgf4efe11}
+\label{sec:org2299927}
 \begin{itemize}
@@
 \section{Title}
-\label{sec:org197abb8}
+\label{sec:org95ba26f}
 \section{Introduction}
-\label{sec:org7e341be}
+\label{sec:orgee84b8b}
 \subsection*{Climate Change and Variability}
-\label{sec:orgeb7a331}
+\label{sec:org68534c1}
 \begin{itemize}
@@
 \subsection*{Climate Forcing}
-\label{sec:org08234e7}
+\label{sec:orga3b6ea1}
 \begin{itemize}
@@
 \subsection*{El Niño (ENSO)}
-\label{sec:org0ec8ade}
+\label{sec:org3e6158d}
 \begin{itemize}
@@
 \subsection*{Method: Climate Simulation}
-\label{sec:org902ab02}
+\label{sec:orgc27209d}
 \begin{itemize}
@@
 \subsection*{Review of Literature}
-\label{sec:orgf798249}
+\label{sec:org4631399}
 \begin{itemize}
@@
 \subsection*{Gap}
-\label{sec:org177c1d6}
+\label{sec:org64f4a3a}
 \begin{itemize}
@@
 \subsection*{Questions}
-\label{sec:org1ab33ee}
+\label{sec:org41ce2a5}
 \begin{itemize}
@@
 \section{Data, Methods, and Results}
-\label{sec:org84e14f6}
+\label{sec:org4f1184c}
 \subsection*{Methods Overview}
-\label{sec:org86e31d8}
+\label{sec:orgea6d725}
 \begin{itemize}
@@
 \subsection*{Role of Mentor and Student}
-\label{sec:org2e78332}
+\label{sec:org996c977}
 \begin{itemize}
@@
 \subsection*{Model Setup}
-\label{sec:org0f6ba67}
+\label{sec:org8ffde0b}
 \begin{itemize}
@@
 \subsection*{Measuring ENSO Intensity}
-\label{sec:orge16e5c1}
+\label{sec:org8f32f98}
 \begin{itemize}
@@
 \subsection*{ENSO is Becoming Stronger}
-\label{sec:orgb07762a}
+\label{sec:org5844385}
 \begin{itemize}
@@
 \subsection*{Influence of Aerosols and Greenhouse Gasses}
-\label{sec:org2b52160}
+\label{sec:org80018e5}
 \begin{itemize}
@@
 \subsection*{Correlation With Ocean Temperature}
-\label{sec:orgbdaf9d4}
+\label{sec:org403aa1c}
 \begin{itemize}
@@
 \subsection*{Wavelet Analysis}
-\label{sec:orga32dcde}
+\label{sec:orgb0f8166}
 \begin{itemize}
@@
 \end{itemize}
+
+I conducted a wavelet analysis to identify changes to specific frequency bands of ENSO.
+One way of thinking about wavelet analysis is using a radio to listen to a variety of frequencies, all at once, except with data rather than radio waves.
+So figure 7 shows, the power, similar to loudness, of the ENSO signal where the period is greater than three years and less than three years, for both the CESM1 and CESM2.
+Looking at CESM1, much of the increase in power over the 20th century is in the longer-period band, meaning that much of the increase in ENSO intensity is happening in the lower frequencies.
+Looking at CESM2, one interesting feature is that the decreasing strength we saw earlier is mostly happening in the longer frequency.
+These results suggest that longer frequency ENSO is more succeptible to climate change than ENSO with a faster cycle.
+
 \section{Conclusion}
-\label{sec:org46c3af5}
+\label{sec:org92a7dae}
 \subsection*{Conclusions and Discussion}
-\label{sec:org10adbad}
+\label{sec:orge52295e}
 \begin{itemize}
@@
 \subsection*{Application, Limitation, and Next Steps}
-\label{sec:org10fbd6a}
+\label{sec:org0c817e6}
 \begin{itemize}
@@
 \subsection*{Acknowledgments}
-\label{sec:orgf4b1b34}
+\label{sec:orgeaf66eb}
 \begin{itemize}
@@
 \subsection*{References}
-\label{sec:orgd9329c4}
+\label{sec:org0630c81}
 \end{document}

### 2022-01-24 presentations/presentation/presentation.tex
stuff lol

@@
 
-\begin{frame}{Climate Change and Variability}
-  \begin{columns}
-    \column{0.4\textwidth}
-    \begin{itemize}
-    \item Global warming
-    \item Long-term trends vs short-term randomness
-    \end{itemize}
-    \column{0.6\textwidth}
-    \myfig{intro_fig_3.pdf}{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}{woohoo}
-  \end{columns}
+\begin{frame}{Climate Change: Global Warming vs. Climate Variability}
+  \begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{figures/intro_fig_3.pdf}
+    \caption{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}
+    \label{fig:woohoo}
+  \end{figure}
 \end{frame}
 
-\begin{frame}{Climate forcing}
+\begin{frame}{Greenhouse Gasses and Friends}
   \begin{columns}
@@
 
-\begin{frame}{El Niño (ENSO)}
+\begin{frame}{El Niño, La Niña, and ENSO}
   \begin{figure}
@@
       \item ENSO = El Niño/Southern Oscillation
+      \item El Niño is the warm phase, and La Niña is the cool phase.
       \item Warming and cooling of the Pacific Ocean.
       \item Affects human societies through temperature and rainfall. \citep{ropelewski1987global}
-      \item May be affected by climate change.
       \end{itemize}
       \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
-      \column{0.5\textwidth}
+      \column{0.6\textwidth}
       \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
@@
   \item ENSO's properties observed vary across different decades. \citep{lubbecke2014assessing}.
-  \item ENSO responds to external forcing.
   \item Weakened ENSO during the Ice Age due to reduced CO$_2$ levels \citep{zhu2017reduced}.
@@
   \alert{Questions}
-  \begin{description}
-  \item[\alert{What?}] Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
-  \item[\alert{Why?}] Is the predicted increase (or decrease) due to human activities?
-  \item[\alert{How?}] What processes are causing greenhouse gasses and aerosols to affect ENSO?
-  \end{description}
+  \begin{enumerate}
+  \item Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
+  \item Is the predicted increase (or decrease) due to human activities?
+  \end{enumerate}
 \end{frame}
@@
 
-\begin{frame}{Model Setup (Data)}
+\begin{frame}{Data: Community Earth System Model}
   \begin{itemize}
@@
   \item Predicts climate over 21st century with global warming.
-  \item 40-50 simulations per ensemble.
-  \item Control simulation with pre-1850 forcing levels.
+  \item Ensemble: collection of multiple simulations.
   \item Single forcing ensembles that represent influence of single factor.
@@
 
-\begin{frame}{Correlation With Ocean Temperature}
-  \begin{columns}
-    \column{0.65\textwidth}
-    \begin{itemize}
-    \item Correlation coefficient between ocean temperature and ENSO amplitude.
-    \item Negative coefficient in subsurface layer.
-    \item Positive coefficient in surface layer.
-    \item Suggests that ocean stratification may be mediating global warming influence on ENSO.
-    \item Difference in heating modifies mechanics of ENSO cycle.
-    \end{itemize}
-    \column{0.5\textwidth}
-    \myfig{tempdt.pdf}{Correlation coefficient between ENSO amplitude and ocean temperature in equatorial cross-section in the fully-forced CESM1 ensemble}{tempdt}
-  \end{columns}
-\end{frame}
+% \begin{frame}{Correlation With Ocean Temperature}
+%   \begin{columns}
+%     \column{0.65\textwidth}
+%     \begin{itemize}
+%     \item Correlation coefficient between ocean temperature and ENSO amplitude.
+%     \item Negative coefficient in subsurface layer.
+%     \item Positive coefficient in surface layer.
+%     \item Suggests that ocean stratification may be mediating global warming influence on ENSO.
+%     \item Difference in heating modifies mechanics of ENSO cycle.
+%     \end{itemize}
+%     \column{0.5\textwidth}
+%     \myfig{tempdt.pdf}{Correlation coefficient between ENSO amplitude and ocean temperature in equatorial cross-section in the fully-forced CESM1 ensemble}{tempdt}
+%   \end{columns}
+% \end{frame}
 
@@
   \begin{itemize}
-  \item Rising greenhouse gas levels increase Pacific Ocean stratification, strengthening ENSO cycle.
+  \item Rising greenhouse gas levels strengthen ENSO cycle.
   \item Aerosol influence is nonlinear because aerosol levels are not purely increasing.
   \item Stronger ENSO may lead to greater temperature variability and extreme weather.
-  \item CESM1 and CESM2 conflict in their prediction of the changes to ENSO's frequency.
+  \item External forcing affects lower frequency ENSO more.
   \end{itemize}
@@
   \begin{columns}[t]
-    \column{.4\textwidth}
+    \column{.5\textwidth}
     Student:
@@
     \end{itemize}
-    \column{.4\textwidth}
+    \column{.5\textwidth}
     Mentor:

### 2022-03-02 feedback/presentation_main.tex
stuff!

@@
-% Created 2022-01-06 Thu 19:00
+% Created 2022-01-27 Thu 09:32
 % Intended LaTeX compiler: pdflatex
@@
 \section{Feedback 1}
-\label{sec:orgb0ce4b1}
+\label{sec:orge5d83f9}
 Date: \textit{<2021-09-26 Sun>}
@@
 \subsection*{Feedback}
-\label{sec:orgeae9fd0}
+\label{sec:org59ea01a}
 \begin{itemize}
@@
 \section{Feedback 2}
-\label{sec:org3a9f9f0}
+\label{sec:orgf550f9c}
 Date: \textit{<2021-09-30 Thu>}
@@
 \subsection*{Feedback:}
-\label{sec:orgd187f3b}
+\label{sec:org46618e2}
 \begin{itemize}
@@
 \section{Feedback 3}
-\label{sec:org304f650}
+\label{sec:org4fbc4a2}
 Date: \textit{<2021-09-30 Thu>}
@@
 \subsection*{Feedback:}
-\label{sec:org2526349}
+\label{sec:orgb1ba02d}
 \begin{itemize}
@@
 \section{Feedback 4}
-\label{sec:org661561a}
+\label{sec:org94d9adb}
 Date: \textit{<2021-10-17 Sun>}
@@
 \subsection*{Feedback:}
-\label{sec:orgfb0e433}
+\label{sec:orgff9e78c}
 \begin{itemize}
@@
 \section{Feedback 5}
-\label{sec:org4ee2cd0}
+\label{sec:org71846e3}
 
@@
 \subsection*{Feedback:}
-\label{sec:orgcae29f9}
+\label{sec:orga0d61f9}
 \begin{itemize}
@@
 \end{itemize}
+
+\section{Feedback 6}
+\label{sec:orge523970}
+
+Date: \textit{<2022-01-15 Sat>}
+Time: 09:20
+Listener: Abby Goldman
+\subsection*{Feedback}
+\label{sec:org2833856}
+\begin{itemize}
+\item Make some notes to follow
+\item Couple slides were unclear
+\item Speak slower and less nervous
+\item Good introduction
+\end{itemize}
+
+\section{Feedback 7}
+\label{sec:org3149d17}
+Date: \textit{<2022-01-20 Thu>}
+Time: 9:25
+Listener: Jimena Perez
+\subsection*{Feedback:}
+\label{sec:org43a5e7b}
+\begin{itemize}
+\item Love how the intro slides flow, the flow chart in the methodology slide becomes easier to understand every time you present it
+\item Maintain speaking pace, describe implications of methodology while you're going through methodolgy.
+\end{itemize}
+
+\section{Feedback 8}
+\label{sec:org6818510}
+Date: \textit{<2022-01-25 Tue>}
+Time: 9:25
+Listener: Sean O'Donnell
+\subsection*{Feedback:}
+\label{sec:orge3ca3f6}
+\begin{itemize}
+\item Don't have to say ziggy zaggy line on 2nd slide (say red line or blue line)
+\item Don't think you described figure 2
+\item Liked explanation of figure 3 and El Nino/La nina in general
+\item Yes, def mention ENSO first before EL Nino and La nina in introduction
+\item Liked simplicity of powerpoint design, made it very easy to follow
+\item Maybe mention CESM before stating research questions so people can understand its background (may not be necessary though)?
+\item Strong explanation of methodology (liked the flowchart)
+\item Liked the addition of ``increasing variance'' and decreasing variance`` to Figure 4
+\item What's difference between CESM1 and CESM 2 (may have missed this in presentation)?
+\item Yes, def note software used before acknowledgements (don't have to describe each database)
+\end{itemize}
 \end{document}

### 2022-03-02 presentations/presentation/presentation.tex
Presentation is 16:10, not 4:3
stuff!

@@
-\documentclass{beamer}
+\documentclass[aspectratio=1610]{beamer}
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{figures/intro_fig_3.pdf}
+    \includegraphics[width=0.9\textheight]{figures/intro_fig_3.pdf}
     \caption{Global mean land air temperature in GISSTEMP 4 dataset. \citep{gistemp2019giss} and \citep{lenssen2019improvements}}
@@
 
-\begin{frame}{Greenhouse Gasses and Friends}
+\begin{frame}{Climate Forcing}
   \begin{columns}
@@
     \column{0.6\textwidth}
-    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect. \url{https://www.coolaustralia.org/the-greenhouse-effect-secondary}}{this}
+    \myfig{greenhouse_Effect.jpg}{Factors that contribute to the greenhouse effect.}{this}
   \end{columns}
@@
       \caption{Comparison of SST anomaly between 1975 La Niña event and 1997 El Niño event in HadISST 1 dataset. \citep{rayner2003global}}
-      \column{0.6\textwidth}
+      \column{0.5\textwidth}
       \includegraphics[width = \textwidth]{figures/intro_fig.pdf}
@@
 \begin{frame}{Gap and Questions}
-  \alert{Gap}
+  \alert{Research Gap}
   \begin{itemize}
@@
   \begin{enumerate}
-  \item Do the CESM1 and CESM2 predict increased or decreased ENSO intensity in the future?
+  \item Do the CESM1 and CESM2 climate models predict increased or decreased ENSO intensity in the future?
   \item Is the predicted increase (or decrease) due to human activities?
@@
   \item Single forcing ensembles that represent influence of single factor.
+  \item Software used: R, ncdf4, zoo, dplyr, ggplot2, WaveletComp, reshape2, nco.
   \end{itemize}
@@
 \begin{frame}{Measuring ENSO Intensity}
-  \begin{tikzpicture}[node distance = 1.5in, line width = 0.5mm]
+  \begin{tikzpicture}[node distance = .37\textwidth, line width = 0.5mm]
     \node [process] (input) {
@@
         \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=\textwidth]{figures/ff_compare.pdf}};
-        \coordinate (a) at (0.5, 1.5);
-        \node [red] (b) at (1.0, 1.0) {
+        \coordinate (a) at (1.0, 1.5);
+        \node [red] (b) at (1.3, 1.0) {
           \begin{varwidth}{0.6in}
@@
           \end{varwidth}};
-        \coordinate (c) at (3.0, 1.5) {};
+        \coordinate (c) at (3.8, 1.5) {};
         \draw[red, |-|] (a) to (c);
 
-        \coordinate (d) at (3.2, 1.5);
-        \node [red] (e) at (3.5, 1.0) {
+        \coordinate (d) at (4.0, 1.5);
+        \node [red] (e) at (4.3, 1.0) {
           \begin{varwidth}{0.6in}
@@
           \end{varwidth}};
-        \coordinate (f) at (4.3, 1.5) {};
+        \coordinate (f) at (5.0, 1.5) {};
         \draw[red, |-|] (d) to (f);
@@
   \begin{columns}
-    \column{0.65\textwidth}
+    \column{0.5\textwidth}
     \begin{itemize}
@@
 
-% SPLIT UP CONCLUSIONS DISCUSSION
-
-\begin{frame}{Discussion}
-  \begin{itemize}
-  \item Rising greenhouse gas levels strengthen ENSO cycle.
-  \item Aerosol influence is nonlinear because aerosol levels are not purely increasing.
-  \item Stronger ENSO may lead to greater temperature variability and extreme weather.
-  \item External forcing affects lower frequency ENSO more.
-  \end{itemize}
-\end{frame}
-
-\begin{frame}{Limitations and Applications}
-  Limitations:
-  \begin{itemize}
-  \item Niño 3.4 index shown to be inaccurate for some models \citep{cai2018increased}.
-  \item CESM may contain biases.
-  \item Models are only an approximation of the Earth's actual climate.
-  \end{itemize}
-  Application: to improve our ability to predict ENSO and help people prepare for increased likelihood of extreme weather.
-\end{frame}
-
-\begin{frame}{Acknowledgments}
-  \begin{itemize}
-  \item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
-1852977.
-  \item Thank you to my teacher, my family, and my mentor!
-  \item Software used: R, ncdf4, zoo, dplyr, ggplot2, WaveletComp, reshape2, nco.
-  \end{itemize}
-\end{frame}
-
 \begin{frame}{Role of Mentor and Student}
@@
 
+\begin{frame}{Conclusion and Discussion}
+  \begin{enumerate}
+  \item Do the CESM1 and CESM2 climate models predict increased or decreased ENSO intensity in the future?
+  \item Is the predicted increase (or decrease) due to human activities?
+  \end{enumerate}
+  \begin{enumerate}
+  \item Yes, ENSO variance increases significantly over the 21$^{st}$ century
+  \item Yes, the major factor contributing to this increase is greenhouse gasses, with aerosol emissions playing a secondary role.
+  \end{enumerate}
+\end{frame}
+
+\begin{frame}{Limitations and Applications}
+  Limitations:
+  \begin{itemize}
+  \item Niño 3.4 index shown to be inaccurate for some models \citep{cai2018increased}.
+  \item CESM may contain biases.
+  \item Models are only an approximation of the Earth's actual climate.
+  \end{itemize}
+  Application: to improve our ability to predict ENSO and help people prepare for increased likelihood of extreme weather.
+\end{frame}
+
+\begin{frame}{Acknowledgments}
+  \begin{itemize}
+  \item This material is based upon work supported by the National Center for Atmospheric Research, which is a major facility sponsored by the National Science Foundation under Cooperative Agreement No.
+1852977.
+  \item Thank you to my teacher, my family, and my mentor!
+  \end{itemize}
+\end{frame}
+
 \begin{frame}{References}
   \bibliographystyle{apalike}
-  \fontsize{4pt}{5}\selectfont
+  \fontsize{6pt}{5}\selectfont
   \bibliography{references.bib}
+  \url{https://www.coolaustralia.org/the-greenhouse-effect-secondary}
 \end{frame}

### 2022-04-06 feedback/presentation_main.tex
yay

@@
-% Created 2022-01-27 Thu 09:32
+% Created 2022-03-21 Mon 09:23
 % Intended LaTeX compiler: pdflatex
@@
 \section{Feedback 1}
-\label{sec:orge5d83f9}
+\label{sec:org3d23bd9}
 Date: \textit{<2021-09-26 Sun>}
@@
 \subsection*{Feedback}
-\label{sec:org59ea01a}
+\label{sec:org6215fef}
 \begin{itemize}
@@
 \section{Feedback 2}
-\label{sec:orgf550f9c}
+\label{sec:orgd8b9b86}
 Date: \textit{<2021-09-30 Thu>}
@@
 \subsection*{Feedback:}
-\label{sec:org46618e2}
+\label{sec:orga4a15c1}
 \begin{itemize}
@@
 \section{Feedback 3}
-\label{sec:org4fbc4a2}
+\label{sec:org01fa312}
 Date: \textit{<2021-09-30 Thu>}
@@
 \subsection*{Feedback:}
-\label{sec:orgb1ba02d}
+\label{sec:org3e66130}
 \begin{itemize}
@@
 \section{Feedback 4}
-\label{sec:org94d9adb}
+\label{sec:orgd2fd241}
 Date: \textit{<2021-10-17 Sun>}
@@
 \subsection*{Feedback:}
-\label{sec:orgff9e78c}
+\label{sec:org826e6ec}
 \begin{itemize}
@@
 \section{Feedback 5}
-\label{sec:org71846e3}
+\label{sec:orgad8b6bb}
 
@@
 \subsection*{Feedback:}
-\label{sec:orga0d61f9}
+\label{sec:org3b02608}
 \begin{itemize}
@@
 \section{Feedback 6}
-\label{sec:orge523970}
+\label{sec:org068e467}
 
@@
 \subsection*{Feedback}
-\label{sec:org2833856}
+\label{sec:org044ac2b}
 \begin{itemize}
@@
 \section{Feedback 7}
-\label{sec:org3149d17}
+\label{sec:orgba0123a}
 Date: \textit{<2022-01-20 Thu>}
@@
 \subsection*{Feedback:}
-\label{sec:org43a5e7b}
+\label{sec:org6749e2b}
 \begin{itemize}
@@
 \section{Feedback 8}
-\label{sec:org6818510}
+\label{sec:orgf10f8be}
 Date: \textit{<2022-01-25 Tue>}
@@
 \subsection*{Feedback:}
-\label{sec:orge3ca3f6}
+\label{sec:orgd778e46}
 \begin{itemize}
@@
 \end{itemize}
+
+\section{Feedback 9}
+\label{sec:org9a0e673}
+Date: \textit{<2022-03-16 Wed>}
+Time: 9:25
+Listener: Giovanni
+\subsection*{Feedback:}
+\label{sec:org0947ba6}
+\begin{itemize}
+\item Good overall
+\item Slow down
+\item Weird spacing
+\end{itemize}
+
+\section{Feedback 10}
+\label{sec:orgfe25489}
+Date: \textit{<2022-03-17 Thu>}
+Time: 9:25
+Listener: Nurashau
+\subsection*{Feedback:}
+\label{sec:org28862e7}
+\begin{itemize}
+\item Slow down
+\item Weird spacing
+\item Mispronouncing ``statistically''
+\end{itemize}
+
+\section{Feedback 11}
+\label{sec:org0efdbe0}
+Date: \textit{<2022-03-18 Fri>}
+Time: 9:25
+Listener: Abby Goldman
+\subsection*{Feedback:}
+\label{sec:orgf9a6d41}
+\begin{itemize}
+\item Slow down
+\item Improve hook and application to be a little more specific
+\end{itemize}
 \end{document}

### 2022-04-06 paperwork/abstract/abstract.tex
yay
stuff

@@
 
+\textbf{2022:} \emph{Regional presenter at the Westchester Junior Science and Hummanities Fair}
+
+\textbf{2022:} \emph{Virtual Powerpoint presenter at the Westchester Science and Engineering Fair (March)}
+
 \begin{center}
@@
 \end{center}
+
+\textbf{2022:} \emph{American Meteorological Society Award at the Westchester Science and Engineering Fair}
 
